### 2023-02-24 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@mkenworthy}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{showyourwork}
+
+\usepackage{astro_bib_macro} %shortcuts of journals defined in this file
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
+    see if we can get this to update....and can er add the macros safely?
+    Come friendly bombs and fall on Slough it isn't fit for humans now....
     Duis eget orci sit amet orci dignissim rutrum.

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-02-24 src/tex/ms.tex
@@
-% Define document class
-\documentclass[twocolumn]{aastex631}
-\usepackage{showyourwork}
+\RequirePackage{cmap}
 
-% Begin!
+\documentclass[
+    usenatbib,
+]{mnras}
+
+\usepackage[T1]{fontenc}
+
+\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
+
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{mathrsfs}
+
+\usepackage{xurl}
+\urlstyle{tt}
+
+\usepackage{booktabs}
+\usepackage{graphicx}
+\usepackage[
+    % colorlinks=true, 
+    % allcolors=blue,
+]{hyperref}
+
+\usepackage{libertine}
+\usepackage[libertine]{newtxmath}
+\usepackage[scaled=0.76]{beramono}
+\usepackage{microtype}
+\usepackage{csquotes}
+\usepackage[capitalise]{cleveref}
+\usepackage[dvipsnames, x11names]{xcolor}
+\usepackage{xspace}
+
+\usepackage{standalone}
+\usepackage{tikz}
+\usetikzlibrary{
+    positioning,
+    shapes,
+}
+
+\usepackage{siunitx}
+\sisetup{
+    range-units = single,
+    range-phrase = {--},
+    separate-uncertainty = true,
+    multi-part-units = brackets,
+    product-units = brackets,
+    detect-weight = true,
+    per-mode = symbol,
+    mode = text,
+}
+\DeclareSIUnit\parsec{pc}
+\DeclareSIUnit\au{au}
+
+%REMOVE LATER
+\usepackage{soul}
+
+\usepackage{astro_bib_macro} %shortcuts of journals defined in this file
+
+\newcommand{\todo}[1]{\textcolor{red}{[#1]}}
+
+% Define macros
+\newcommand{\IWA}{\ensuremath{\mathrm{IWA}}}
+
+% Define the colors of the \rainbows
+\definecolor{C0}{HTML}{ff0000}
+\definecolor{C1}{HTML}{ff6d38}
+\definecolor{C2}{HTML}{ecc86f}
+\definecolor{C3}{HTML}{a4f89f}
+\definecolor{C4}{HTML}{5af8c8}
+\definecolor{C5}{HTML}{12c8e6}
+\definecolor{C6}{HTML}{386df9}
+\definecolor{C7}{HTML}{8000ff}
+
+\newcommand{\rainbows}{%
+    \textcolor{C0}{r}%
+    \textcolor{C1}{a}%
+    \textcolor{C2}{i}%
+    \textcolor{C3}{n}%
+    \textcolor{C4}{b}%
+    \textcolor{C5}{o}%
+    \textcolor{C6}{w}%
+    \textcolor{C7}{s}%
+    \xspace
+}
+
+% Chasing rainbows with the Habitable Worlds Observatory
+\title{Chasing \rainbows with the Habitable Worlds Observatory}
+%Scattering phase angle probed by directly imaged planets, or
+%Don't ask us about contrast
+
+% Authors are in alphabetical order:
+\author[Sophia R. Vaughan et al.]{%
+    Sophia R. Vaughan$^{1,}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
+    Kim Bott$^{2}$,
+    Sarah L. Casewell$^{3}$,
+    Nicolas B. Cowan$^{4}$,
+    David Doelman$^{5}$,\newauthor
+    Timothy D. Gebhard$^{6}$,
+    Matthew Kenworthy$^{5}$,
+    Johan Mazoyer$^{7}$,
+    Maxwell A. Millar-Blanchaer$^{8}$
+    \newauthor [remaining workshop participants in alphabetical order] \newauthor 
+    \\
+    $^{1}$ Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK\\
+    $^{2}$ Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\
+    $^{3}$ Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK\\
+    $^{4}$ Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada\\
+    $^{5}$ Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
+    $^{6}$ Max Planck Institute for Intelligent Systems, Max Planck Ring 4, 72076 Tübingen, Germany \& ETH Zurich, Switzerland \\
+    $^{7}$ LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-92195 Meudon, France \\
+    $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA 
+}
+\date{Draft version: \today}
+
+\pagestyle{plain}
+
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+\maketitle
 
-% Author list
-\author{@mkenworthy}
+\begin{abstract}
+NASA recently announced the Habitable Worlds Observatory, which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures. 
+%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
+After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties.
+Surface liquid water is the definition of planetary habitability.
+Photometric and polarimetric phase variations might betray the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects.
+These phenomena tend to occur at scattering phase angles close to new or full phase, whereas direct imaging missions are optimised to detect planets near quadrature.
+The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle.
+Planets with relatively face-on orbits exhibit a limited range of phases.
+For planets that orbit close to edge-on, on the other hand, the accessible phase angles are limited by obscuration by the coronagraph.
+We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena.
+Any exo-Earths visible to the Habitable Worlds Observatory could be searched for the Rayleigh scattering peak, which occurs near quadrature.
+The glint signature at scattering phases of 50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems.
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.
 
-% Abstract with filler text
-\begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
 \end{abstract}
 
-% Main body with filler text
+\begin{keywords}
+planets and satellites: terrestrial planets -- instrumentation: high angular resolution -- planets and satellites: atmospheres
+\end{keywords}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\begin{figure*}%[t]
+   \centering
+   \includegraphics[width=\linewidth]{Fig1IWAPol.png}
+   \caption{BOTT PLOT this is a holder until Kim updates the figure}
+    \label{fig:bottplot}
+\end{figure*}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\textcolor{red}{introduce motivation, specs for habitable worlds observatory and the document we got the target list from.}
 
+\textcolor{blue}{Habitable worlds bit}
+
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
+This assertion requires an instrument with a coronograph capable of high contrast imaging in the optical-near infrared.
+Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (HWO).
+The precursor technology recommended by the Decadal Survey also lists "direct imaging to probe polarized ocean glint on terrestrial planets" as one of the priority capabilities (\citet[Box E.1 in][]{decadal}) for ground and space-based observatories.
+The performance and precise characteristics of this telescope (on-axis or off-axis, exact diameter, type of segmentation, type of coronagraph) are still to be determined.
+The development will be heavily influenced by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
+The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \cite{Stark2019_exoplanetyield}.
+
+Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \SIrange{200}{2000}{\nano\meter} instrument for exoplanet characterisation, with a similar coronagraph operating between \num{450} and \SI{1800}{\nano\meter} planned for HabEx.
+POLLUX, a proposed UV instrument for the LUVOIR study could also allow the detection of polarised light from hot Jupiters at \SI{300}{\nano\meter} \citep{Bouret2018_pollux}.
+However, the challenges associated with the design of a coronagraph in UV are enormous, and it is unlikely that the HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
+
+%We ignore contrast in this study
+%OWA -> should go in discussions
+
+\textcolor{blue}{Science motivation below: Kim}\\
+\textbf{Why Polarization curves worth looking at [Kim]}
+%compare to features in nonpol (strength rainbow); current capabilties?
+
+ Phase curves of planets are more complex than simple Lambertian curves. 
+ The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
+Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citet[see, for example,][]{kopparla2018}.
+In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
+Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
+
+ In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
+Notably, in the polarized phase curve the rainbow may be an especially strong feature and the maximum reflectance signal (from Rayleigh scattering) occurs
+
+%Polarimetry provides the possibility to detect molecules, hazes, clouds and different surfaces  as they all produce unique polarization signatures which cannot be disentagled by using broadband flux alone.
+Polarimetry measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
+Polarimetry has previously played an important role in the understanding of solar system planet atmospheres, for instance, polarimetric measurements of Venus’ atmosphere suggested that the upper atmosphere of Venus is dominated by a haze of liquid sulphuric acid droplets \citep{Venus74}, which has been confirmed by in situ observations.
+
+\textbf{Why rainbow worth it [Kim]} %calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
+
+%The rai
+
+\textbf{Why glint [Kim]}
+% probe of surface conditions; habitability
+
+%\cite{2006astro.ph.10518M} first noted the potential for photometric and polarimetric scattered light phase to detect oceans on terrestrial exoplanets.
+--> No it was talked about previous to this, per the paper cited here in fact...
+"Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint.
+Schneider (2006) had similar ideas." -KB
+
+%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
+They noted that surface oceans would be detectable because they are a) less reflective near full phase,
+% not following what you mean here -KB
+%b) more reflective at crescent phases, and c) exhibit a peak in polarization at crescent phases.
+\cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.
+Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
+\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to simulations; they found .
+%\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths.
+This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+\cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
+\cite{2022A&A...664A.172T}
+
+%cloud coverage paper in polarisation :
+%\cite{rossi17} computed flux and polarisation of reflected starlight for different types of cloud cover on Earth-like planets and determined different types of cloud cover can be distinguished from each other.
+%Alternative options or papers here....
+Kim can write this tomorrow , needs other Stam, Karalidi, Kopparla, maybe Gordon, need to reference rainbows in Bailey paper.
+Look at citations in Gordon paper (our bib)...
+he has a really thorough lit review --KB
+
+\textbf{In this paper we...}
+% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.
+
+% We do not consider the particulars of specific planet scenarios as that work is completed by foward models
+
+%We do not separate Stokes parameters as we are interested in where the net signal is achievable with IWA and OWA.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section{Plots to Include}
+%\label{sec:plots}
+%\begin{enumerate}
+%    \item \st{Cartoon defining various phase angles: orbital phase, scattering phase, $\beta$ ? (Sophia)}
+%    \item Bott plot: annotated polarized and unpolarized phase curves for ocean world and cloudy worlds (Kim)
+%    \item \st{3x3 cartoon showing effect of inclination and IWA compared to orbit (Sophia)}
+%    \item Cumulative distribution functions, both simple (circular, edge-on), bit more realistic (circular, but range of inclinations), and full-on (range of inclinations and eccentricities).
+Max \& Matthew
+%    \item Scatter plot of stellar effective temperature vs system distance (Timmy)
+%\end{enumerate}
+
+\section{Methods/simulations}
+\subsection{Habitable Worlds Observatory and its target star list}
+%\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
+The HWO is still in the early stages of development, and the exact telescope and instrument design is undecided. 
+In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
+
+HWO will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions.
+The IWA of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations.
+In this work, we investigate the scattering phase coverage obtainable for several choices for the IWA of the observatory.
+We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor.
+
+For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{700}{\nano\meter}$.
+
+\begin{table}
+    \centering
+    \caption{
+        Conversion between $\lambda / D$ and mas for different wavelengths.
+\textbf{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function.
+Also update to use the only wavelength we choose}
+    }
+    \label{tab:IWA_OWA}
+    \begin{tabular}{ c c c c } 
+    \toprule
+     & $\lambda/D$ & mas (at \SI{600}{\nano\meter}) & mas (at \SI{700}{\nano\meter}) \\
+    \midrule
+    \midrule
+    IWA & 1 & 21 & 24\\
+    IWA & 2 & 42 & 48\\
+    IWA & 3 & 63 & 72\\
+    IWA & 4& 84& 96\\
+    \midrule
+    OWA & 32 & 660 & 770 \\
+    OWA & 64 & 1320 & 1540\\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+%\subsection{Stellar sample and assumed planet parameters} 
+We use the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
+\footnote{\url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
+This list comprises $\sim$160 stars, the majority of which are Sun-like dwarfs; 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
+The target list is constructed assuming a maximum planet magnitude of $R_c = 31$ and a contrast floor of \num{2.5e-11}. %with an occurrence rate of
+%rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
+The corresponding adopted
+habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the
+bolometric luminosity normalized to the Sun. This range of orbital separations
+corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}.
+
+The authors of the report determined that these are the nearby stars (maximum distance \SI{25}{\parsec}) for which exo-Earths would be the most observable for an imaging survey of habitable zones with a \SI{6}{\meter} telescope.
+
+\subsection{Maximum scattering phase angle coverage}
+\label{sec:Delta_phi}
+
+The coronagraph, while necessary for observing faint exoplanets, can obscure part of the planet's orbit. 
+As demonstrated in \cref{fig:orb-grid}, this can prevent the planet being detected at high and low scattering angles. 
+In this section, we derive the minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible for a given orbit of semi-major $a$ (in \si{\au}), line of sight inclination $i$, star distance $d_*$ in parsecs (\si{\parsec}) and for a given inner working angle of the instrument \IWA (in milliarcseconds).
+We define the inclination $i$, such that $i=0$ corresponds to a face-on orbit.
+
+\begin{figure*}%[t]
+   \centering
+   \includegraphics[width=0.99\textwidth]{orb-grid.pdf}
+   \caption{
+        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle. 
+        If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
+        For orbital inclinations close to edge-on (bottom panels) the planet goes through all phases, but the scattering phase angles near 0$^\circ$ and 180$^\circ$ are not accessible due to obscuration by the coronagraph mask; for circular orbits, the accessible scattering phases are $\phi \in 90^\circ \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+    }
+    \label{fig:orb-grid}
+\end{figure*}
+
+In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), $\phi_\mathrm{min}$ and $\phi_\mathrm{max}$ are symmetric about 90 degrees (). 
+We define the angle $\Delta \phi$ such that: 
+\begin{equation}
+    \label{eq:Delta_phi}
+    \Delta \phi 
+    = \phi_\mathrm{max} - \phi_\mathrm{min}
+    =  2(\phi_\mathrm{max} - 90) 
+    =  2(\phi_\mathrm{min} + 90)
+\end{equation}
+
+For a given system with inclination $i$, we have identified 2 regimes, shown in \cref{fig:orb-grid}:
+
+(1)~If the inclination is face-on enough such that the coronagraph mask does not cover any part of the projected orbit on-sky, then the maximum scattering phase angle coverage depends solely on the disk inclination ;
+
+(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the IWA and orbit semi-major axis.
+In this section, we derive the analytical formula for this case in the circular orbit case.
+
+\begin{figure}%[t]
+    \centering
+    % \includegraphics[width=\linewidth]{figures/scattering_angles.pdf}
+    \includestandalone{figures/scattering-phase-angle}
+    \caption{
+        The angular separation, $\varepsilon$, of a planet orbiting with semi-major-axis $a$ about a star a distance $d_*$ from the observer. 
+        The scattering phase angle is $\varphi$. 
+        The smallest accessible scattering phase angle occurs when the angular separation equals the coronagraph inner working angle: $\varphi_{\rm min} = {\rm IWA} \cdot d_*/a$.
+    }
+    \label{fig:scattering_angles}
+\end{figure}
+
+From \cref{fig:scattering_angles}, we deduce that for all points on the orbit, the scattering angle $\varphi$ satisfies:
+\begin{equation}
+    \sin(\varphi) = \frac{r_\mathrm{proj}}{a}
+\end{equation}
+with $r_\mathrm{proj}$ the on-sky projected planet distance (in \si{\au}), $a$ the distance to the star (constant in this section, in \si{\au}).
+
+$\varepsilon$ is the angular separation of the planet at a given point of the orbit ($r_\mathrm{proj} = \delta d_*$). 
+We access the minimum scattering angle when the projected distance reaches the inner working angle separation $\delta = \mathrm{IWA}$. 
+Using \cref{eq:Delta_phi}, we deduce:
+
+\begin{equation}
+    \cos\left(\dfrac{\Delta \phi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a}
+\end{equation}
+
+We now deduce the 2 regime equation for all cases (circular orbit): 
+\begin{equation}
+\label{eq:Delta_phi_max}
+    \Delta \phi = 
+    \begin{cases}
+        2 \cdot i & \textrm{for} \cos(i) > \dfrac{\mathrm{IWA}\cdot d_* }{a}
+  \\ 
+        2 \cdot \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right)  & \textrm{for} \cos(i) < \dfrac{\mathrm{IWA}\cdot d_* }{a}
+    \end{cases}\,.
+\end{equation}
+
+\subsection{Edge-on circular orbits}
+\label{sec:edge-on}
+We initially performed a simple test considering only edge-on circular orbits.
+This gives an upper limit for circular orbits for the visible scattering phase range as a function of inclination.
+Fig.
+\ref{fig:scatterplot} shows $\Delta \phi$ for the systems in the target list assuming a $3 \lambda / D$, at $\SI{600}{\nano\meter}$, coronograph.
+This optimistic case indicates that it will be possible to observe Rayleigh scattering for all targets and that rainbows will be accessible in some cases.
+
+\begin{figure*}
+    \centering
+    \includegraphics{scatterplot.pdf}
+    \caption{
+        Scatter plot at 3 $\lambda/D$ for the target sample showing stellar effective temperature and stellar distance. 
+        The size of the points represents the angular separation of the star and planet in milliarcseconds as presented in the target list.
+The colour of the points shows the atmospheric phenomenon that can be detected with darker colors including all lighter (yellow) color phenomenon.
+Thus dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
+    }
+    \label{fig:scatterplot}
+\end{figure*}
+
+\subsection{Assumed planet parameters}
+
+As shown in \cref{sec:Delta_phi}, $\Delta \phi$ depends on the angular separation, inclination and eccentricity of the planets orbit. 
+We use the angular separation for an Earth-Like planet with the same instellation as Earth which is provided in the input star list. 
+The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
+
+\subsection{Making Contrast Curves}
+\textbf{Kim and David}
+%
+
+% \subsection{Monte Carlo Simulations}
+% \label{sec:montecarlo}
+% For each star on the HabWorlds target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
+The reflected flux and polarisation are not constant
+
+\section{Results}
+
+\subsection{Circular randomly inclined orbits}
+\label{sec:circular}
+For each star on the HabWorlds target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
+The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$.
+We then use Eqn \ref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
+Figure blah shows the cumulative distribution for the most extreme scattering phase angles accessible for the HabWorlds target list, depending on the choice of inner working angle.
+
+\subsection{Eccentric randomly inclined orbits}
+\label{sec:eccentric}
+Introduce full orbits \\
+
+We repeat the monte-carlo simulation from section \ref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
+\citet{Guimond_2019} and references therein).
+
+We randomly drew 1000 eccentricities from the beta distribution for each star and determined the orbits.
+We then filtered on IWA for multiples of 1, 2 3 and 4 $\lambda / D$, before determining which targets we would be able to determine whether ocean glint, rainbows or Rayleigh peak are detectable for.
+
+Yarn ball plot \cref{fig:ball-o-yarn} \\
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{madness.pdf}  
+    \caption{
+        Eccentric orbits for the stellar sample $\IWA$ of 1, 2, 3 and 4~$\lambda / D$ corresponding to 21, 42, 63 and 84 mas.
+        The habitable zone is identified in each panel. 
+    }
+    \label{fig:ball-o-yarn}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.6\textwidth]{features_plot_peak.png}  
+    \caption{
+        Number of systems vs inner working angle
+    }
+    \label{fig:accessible_phase_angles}
+\end{figure*}
+
+full version cumulative distribution \\
+- discussion of differences \\
+
+\subsection{The glory of Alpha Cen}
+\label{sec:ealpha-cen}
+\textbf{Glory science, what it tells us (Kim)}
+% Kim will write para on science of glory (what tells us, vaguely why appears here)
+
+\subsection{Contrast Curves}
+\label{sec:contrast}
+
+\section{Discussion}
+What did we determine?
+Can we actually detect this stuff for anything in this sample?
+Do we need to not something about things like A stars which are not included in the sample - they were rejected by the list creators - do we need to comment on this?
+
+\section*{Software Data Availability}
+%Maybe we could upload the code that we used for our analysis / plots somewhere?
+%Just for full transparency / reproducibility?
+The NASA ExEP target list containing the data used for these simulations is available online.% 
+\footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
+This work has made use of \textsf{numpy}
+ \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy},\footnote{\url{https://www.astropy.org}} a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
+
+\section*{Acknowledgements}
+The workshop on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, Netherlands.
+This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).
+
+This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France.
+
+SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
+SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
+TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
+Specific grant funding and any other specific requirements
+KMB acknowledges support from NASA Habitable
+Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
+
+% Add references
+\bibliographystyle{mnras}
 \bibliography{bib}

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
-
+\usepackage{showyourwork}
 \usepackage{amsmath}
@@
 \usepackage[capitalise]{cleveref}
-\usepackage[dvipsnames, x11names]{xcolor}
-\usepackage{xspace}
+%\usepackage[dvipsnames, x11names]{xcolor}
+%\usepackage{xspace}
 
@@
    \centering
-   \includegraphics[width=\linewidth]{Fig1IWAPol.png}
+   \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
    \caption{BOTT PLOT this is a holder until Kim updates the figure}
@@
    \centering
-   \includegraphics[width=0.99\textwidth]{orb-grid.pdf}
+   \includegraphics[width=0.99\textwidth]{figures/orb-grid.pdf}
    \caption{
@@
 
-\begin{figure}%[t]
-    \centering
+%\begin{figure}%[t]
+%    \centering
     % \includegraphics[width=\linewidth]{figures/scattering_angles.pdf}
-    \includestandalone{figures/scattering-phase-angle}
-    \caption{
-        The angular separation, $\varepsilon$, of a planet orbiting with semi-major-axis $a$ about a star a distance $d_*$ from the observer. 
-        The scattering phase angle is $\varphi$. 
-        The smallest accessible scattering phase angle occurs when the angular separation equals the coronagraph inner working angle: $\varphi_{\rm min} = {\rm IWA} \cdot d_*/a$.
-    }
-    \label{fig:scattering_angles}
-\end{figure}
+%    \includestandalone{figures/scattering-phase-angle}
+%    \caption{
+%        The angular separation, $\varepsilon$, of a planet orbiting with semi-major-axis $a$ about a star a distance $d_*$ from the observer. 
+%        The scattering phase angle is $\varphi$. 
+%        The smallest accessible scattering phase angle occurs when the angular separation equals the coronagraph inner working angle: $\varphi_{\rm min} = {\rm IWA} \cdot d_*/a$.
+%    }
+%    \label{fig:scattering_angles}
+%\end{figure}
 
@@
     \centering
-    \includegraphics{scatterplot.pdf}
+    \includegraphics{figures/scatterplot.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=\textwidth]{madness.pdf}  
+    \includegraphics[width=\textwidth]{figures/madness.pdf}  
     \caption{
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{features_plot_peak.png}  
+    \includegraphics[width=0.6\textwidth]{figures/features_plot_peak.png}  
     \caption{

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \textcolor{C7}{s}%
-    \xspace
+    % \xspace
 }
@@
 % Chasing rainbows with the Habitable Worlds Observatory
-\title{Chasing \rainbows with the Habitable Worlds Observatory}
+\title{Chasing \rainbows \hspace{1pt} with the Habitable Worlds Observatory}
 %Scattering phase angle probed by directly imaged planets, or
@@
     Sophia R. Vaughan$^{1,}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
-    Kim Bott$^{2}$,
+    Kimberly Bott$^{2,9}$,
     Sarah L. Casewell$^{3}$,
@@
     $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA 
+    $^{9}$ NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA
 }
@@
 After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties.
-Surface liquid water is the definition of planetary habitability.
-Photometric and polarimetric phase variations might betray the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects.
-These phenomena tend to occur at scattering phase angles close to new or full phase, whereas direct imaging missions are optimised to detect planets near quadrature.
+% JMa I've seen someone commented my text. I'm ok with that but I think it would be important to have a 1 sentence description of the 
+Surface liquid water is central to the definition of planetary habitability.
+Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects.
+Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest.
 The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle.
-Planets with relatively face-on orbits exhibit a limited range of phases.
-For planets that orbit close to edge-on, on the other hand, the accessible phase angles are limited by obscuration by the coronagraph.
+Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph.
 We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena.
-Any exo-Earths visible to the Habitable Worlds Observatory could be searched for the Rayleigh scattering peak, which occurs near quadrature.
-The glint signature at scattering phases of 50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems.
+We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak.
+The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems.
 The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.
@@
 \section{Methods/simulations}
-\subsection{Habitable Worlds Observatory and its target star list}
+\subsection{HWO and Stellar Sample}
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
@@
 
-From \cref{fig:scattering_angles}, we deduce that for all points on the orbit, the scattering angle $\varphi$ satisfies:
+%From \cref{fig:scattering_angles}, we deduce that for all points on the orbit, the scattering angle $\varphi$ satisfies:
 \begin{equation}
@@
 % For each star on the HabWorlds target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
-The reflected flux and polarisation are not constant
+The reflected flux and polarisation are not constant as a function of scattering phase angle.
+This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
+These are modelled using ...
 
@@
 \label{sec:eccentric}
-Introduce full orbits \\
-
-We repeat the monte-carlo simulation from section \ref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
+We repeat the Monte Carlo simulation from section \ref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
 \citet{Guimond_2019} and references therein).
-
 We randomly drew 1000 eccentricities from the beta distribution for each star and determined the orbits.
-We then filtered on IWA for multiples of 1, 2 3 and 4 $\lambda / D$, before determining which targets we would be able to determine whether ocean glint, rainbows or Rayleigh peak are detectable for.
-
-Yarn ball plot \cref{fig:ball-o-yarn} \\
+We then filtered on $\IWA$ for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows or the Rayleigh peak occur.
+Figure \ref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's.
+Each plot has been scaled such that the habitable zone occupies the same radius hence the \IWA's vary in apparent size compared to the orbit.
 

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory, which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures. 
+NASA recently announced the Habitable Worlds Observatory, a coronagraphic mission which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures.
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
 After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties.
-% JMa I've seen someone commented my text. I'm ok with that but I think it would be important to have a 1 sentence description of the 
 Surface liquid water is central to the definition of planetary habitability.

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\textbf{Is there a plot for this?}
+
 \subsection{Eccentric randomly inclined orbits}
@@
     \caption{
-        Eccentric orbits for the stellar sample $\IWA$ of 1, 2, 3 and 4~$\lambda / D$ corresponding to 21, 42, 63 and 84 mas.
-        The habitable zone is identified in each panel. 
+        A example of the eccentric orbits generated for the stellar sample.
+The orbits are scaled by the Earth-equivalent flux distance.
+The shaded region indicates \IWA's of 1, 2, 3 and 4~$\lambda / D$ corresponding to 21, 42, 63 and 84 mas.
+The \IWA can significantly affect the range of scattering phases observable with each orbit.
     }
@@
 \end{figure*}
-
-full version cumulative distribution \\
-- discussion of differences \\
 

### 2023-02-24 src/scripts/create-scatterplot.py
[showyourwork] overleaf sync

@@ -1,0 +1,180 @@
+"""
+Create a color-coded scatter plot the the target list.
+"""
+
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+import paths
+from typing import Any
+
+from matplotlib.colorbar import Colorbar
+from matplotlib.lines import Line2D
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+print(mpl.__version__)
+
+# -----------------------------------------------------------------------------
+# AUXILIARY FUNCTIONS
+# -----------------------------------------------------------------------------
+
+def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
+    """
+    Auxiliary function to set the fontsize of all elements in a plot.
+    """
+
+    for item in (
+        [ax.title, ax.xaxis.label, ax.yaxis.label]
+        + ax.get_xticklabels()
+        + ax.get_yticklabels()
+    ):
+        item.set_fontsize(fontsize)
+
+
+# -----------------------------------------------------------------------------
+# MAIN CODE
+# -----------------------------------------------------------------------------
+
+if __name__ == '__main__':
+    
+    # -------------------------------------------------------------------------
+    # Load the data
+    # -------------------------------------------------------------------------
+
+    # Load the target list
+    file_path = paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv'
+    data = pd.read_csv(file_path, header=1)
+
+    # Load the beta values
+    file_path = paths.data / 'beta_max_3_lambda_over_d.csv'
+    beta_values = np.loadtxt(file_path)
+
+    # -------------------------------------------------------------------------
+    # Create the scatter plot
+    # -------------------------------------------------------------------------
+
+    # Define a function to map a quantity (e.g., angular separation)
+    # onto a marker size
+    def quantity_to_ms(q: float) -> float:
+        return 5 * (1 + q / 500)
+
+    # Define a mapping of quantity (e.g., beta value) to a color
[…]

### 2023-02-24 src/scripts/create-scatterplot.py
[showyourwork] overleaf sync

@@ -65,7 +65,7 @@ if __name__ == '__main__':
 
     # Define a mapping of quantity (e.g., beta value) to a color
     norm = mpl.colors.Normalize(vmin=0, vmax=90)
-    cmap = mpl.colormaps['viridis_r']
+    cmap = mpl.cm.get_cmap('viridis_r')
 
     # Define a function to map a quantity (e.g., beta value) onto a color
     def quantity_to_color(q: float) -> Any:

### 2023-02-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Chasing rainbows with the Habitable Worlds Observatory
-\title{Chasing \rainbows \hspace{1pt} with the Habitable Worlds Observatory}
+\title{Chasing \rainbows{} with the Habitable Worlds Observatory}
 %Scattering phase angle probed by directly imaged planets, or
@@
     $^{7}$ LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-92195 Meudon, France \\
-    $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA 
+    $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA \\
     $^{9}$ NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA
@@
 Surface liquid water is central to the definition of planetary habitability.
+%
 Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects.
@@
 The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
+%
 This assertion requires an instrument with a coronograph capable of high contrast imaging in the optical-near infrared.
+%
 Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (HWO).
-The precursor technology recommended by the Decadal Survey also lists "direct imaging to probe polarized ocean glint on terrestrial planets" as one of the priority capabilities (\citet[Box E.1 in][]{decadal}) for ground and space-based observatories.
+%
+The precursor technology recommended by the Decadal Survey also lists ``direct imaging to probe polarized ocean glint on terrestrial planets'' as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
+%
 The performance and precise characteristics of this telescope (on-axis or off-axis, exact diameter, type of segmentation, type of coronagraph) are still to be determined.
+%
 The development will be heavily influenced by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
-The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \cite{Stark2019_exoplanetyield}.
+%
+The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
 
 Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \SIrange{200}{2000}{\nano\meter} instrument for exoplanet characterisation, with a similar coronagraph operating between \num{450} and \SI{1800}{\nano\meter} planned for HabEx.
+%
 POLLUX, a proposed UV instrument for the LUVOIR study could also allow the detection of polarised light from hot Jupiters at \SI{300}{\nano\meter} \citep{Bouret2018_pollux}.
+%
 However, the challenges associated with the design of a coronagraph in UV are enormous, and it is unlikely that the HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
@@
  Phase curves of planets are more complex than simple Lambertian curves.
+ %
  The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
-Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citet[see, for example,][]{kopparla2018}.
+ %
+ Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
+ %
  In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
+ %
  Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
@@
     \label{fig:scatterplot}
+    \script{create-scatterplot.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/madness.pdf}  
+    \includegraphics[width=\textwidth]{figures/ball_of_yarn.pdf}  
     \caption{
@@
     \label{fig:ball-o-yarn}
+    \script{make_yarn_plot.py}
 \end{figure*}

### 2023-02-24 src/scripts/make_yarn_plot.py

@@ -1,0 +1,106 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Circle
+import paths
+from phases import *
+from samplers import *
+
+np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
+
+# this will make a plot with the number of orbits per panel and side of plots along one side
+plotit = 1 # set to 0 to skip plot generation
+norbs_plot = 10
+nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
+
+tyb = dict(color='black', fontsize=14)
+
+def draw_iwa(ax, iwa):
+	from matplotlib.patches import Circle
+	for r in iwa:
+		ax.add_patch(Circle((0,0),radius=r,alpha=0.2))
+
+# number of planet sims per star
+nplanets = 10
+
+# number of epochs to sample along the orbit of the planet
+norbitsample = 1000
+
+# IWA ranges
+iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax at
+
+# read in the stars
+from astropy.io import ascii
+
+fname = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
+t = ascii.read(paths.data / fname ,header_start = 1, data_start = 2)
+
+nstars = len(t)
+print(f'# read in {nstars} stars from {fname}')
+
+# set up output arrays
+betamin = np.zeros((nplanets,nstars,iwa.size))
+betamax = np.zeros((nplanets,nstars,iwa.size))
+
+if (plotit):
+	fig, axes = plt.subplots(nstars_plot, nstars_plot, figsize=(10,10))
+	ax = axes.flatten()
+	fig.suptitle(f'Inner Working Angles = {iwa} mas')
+	# remove the x and y ticks
+	for a in ax:
+		a.set_xticks([])
+		a.set_yticks([])
+		a.set_aspect('equal')
+
+	fig.subplots_adjust(wspace=0, hspace=0)
+
+cax = 0 # counter for the grid of subplots - this increases by 1 for each star AND also counts the star number for output array
+
+for l in tqdm(t): # loop over all the stars in the table
+	star = l['hip_id']    		# name
[…]

### 2023-02-24 src/scripts/samplers.py

@@ -1,0 +1,87 @@
+# from sampling-inclinations-and-eccentricities.ipynb
+#
+# Tim Gebhard
+from scipy.stats import beta
+
+import numpy as np
+
+def sample_cos_i(n: int = 1) -> np.array:
+    """
+    Sample cos(inclination) uniformly from [0, 1].
+    """
+    return np.random.uniform(0, 1, n)
+
+def sample_i(n: int = 1) -> np.array:
+    """
+    Sample `n` inclinations such that cos(i) is uniform in [0, 1].
+    The resulting values are returned in *degree*.
+    """
+    return np.rad2deg(np.arccos(sample_cos_i(n)))
+
+
+def sample_e(n: int, a: float = 0.867, b: float = 3.03) -> np.ndarray:
+    """
+    Sample `n` eccentricities `e` from a beta distribution with shape
+    parameters `a` and `b`. The default values for `a` and `b` are
+    taken from:
+    
+        Guimond and Cowan (2019): arXiv:1903.06184
+    
+    We are truncating `e` to [0, 1] to exclude hyperbolic orbits.
+    """
+    
+    return np.clip(beta.rvs(a=a, b=b, size=n), 0, 1)
+
+
+if __name__ == "__main__":
+
+	import matplotlib.pyplot as plt
+	n = 100_000
+
+	# Sample inclinations i uniformly in cos(i)
+	inclinations = sample_i(n)
+
+	# Plot a histogram of the inclinations
+	fig, ax = plt.subplots(figsize=(9, 3))
+	ax.hist(
+	    inclinations, 
+	    bins=np.linspace(0, 90, 90),
+	    histtype='step',
+	    lw=2,
+	    density=True,
+	)
+	ax.set_xlim(0, 90)
+	ax.set_xlabel('Inclination (deg)')
+	ax.set_ylabel('Density')
+
+	plt.show()
+
+	n = 100_000
+
+	# Sample eccentricities from beta distribution
+	eccentricities = sample_e(n)
+
+	# Plot a histogram of the inclinations
+	fig, ax = plt.subplots(figsize=(9, 3))
+	ax.hist(
+	    eccentricities, 
+	    bins=np.linspace(0, 1, 100),
+	    histtype='step',
+	    lw=2,
+	    density=True,
+	    zorder=1,
+	)
+
+	# Plot the beta PDF
[…]

### 2023-02-24 src/scripts/make_yarn_plot.py
[showyourwork] overleaf sync

@@ -55,7 +55,7 @@ if (plotit):
 
 cax = 0 # counter for the grid of subplots - this increases by 1 for each star AND also counts the star number for output array
 
-for l in tqdm(t): # loop over all the stars in the table
+for l in t: # loop over all the stars in the table
 	star = l['hip_id']    		# name
 	d = l['sy_dist']      		# parsecs
 	Vmag = l['sy_vmag']   		# V magnitude

### 2023-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \DeclareSIUnit\au{au}
+\DeclareSIUnit\mas{mas}
 
@@
 \newcommand{\IWA}{\ensuremath{\mathrm{IWA}}}
+\newcommand{\hwo}{HabWorlds}
 
@@
 %
-Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (HWO).
+Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
 %
@@
 %
-However, the challenges associated with the design of a coronagraph in UV are enormous, and it is unlikely that the HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
+However, the challenges associated with the design of a coronagraph in UV are enormous, and it is unlikely that \hwo will include a UV high-contrast instrument sensitive to Earth-like planets.
 
@@
 \section{Methods/simulations}
-\subsection{HWO and Stellar Sample}
+\subsection{\hwo\ and Stellar Sample}
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-The HWO is still in the early stages of development, and the exact telescope and instrument design is undecided. 
+\hwo\ is still in the early stages of development, and the exact telescope and instrument design is undecided. 
 In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
 
-HWO will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions.
-The IWA of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations.
-In this work, we investigate the scattering phase coverage obtainable for several choices for the IWA of the observatory.
+\hwo\ will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions.
+The \IWA\ of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations.
+In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory.
 We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor.
@@
 %rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
-The corresponding adopted
-habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the
-bolometric luminosity normalized to the Sun. This range of orbital separations
-corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}.
+The corresponding adopted habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the
+bolometric luminosity normalized to the Sun. 
+This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}.
 
@@
 As demonstrated in \cref{fig:orb-grid}, this can prevent the planet being detected at high and low scattering angles. 
-In this section, we derive the minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible for a given orbit of semi-major $a$ (in \si{\au}), line of sight inclination $i$, star distance $d_*$ in parsecs (\si{\parsec}) and for a given inner working angle of the instrument \IWA (in milliarcseconds).
+In this section, we derive the minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible for a given orbit of semi-major $a$ (in \si{\au}), line of sight inclination $i$, star distance $d_*$ in parsecs (\si{\parsec}) and for a given inner working angle of the instrument \IWA\ (in milliarcseconds).
 We define the inclination $i$, such that $i=0$ corresponds to a face-on orbit.
@@
 
-(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the IWA and orbit semi-major axis.
+(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the \IWA\ and orbit semi-major axis.
 In this section, we derive the analytical formula for this case in the circular orbit case.
@@
 This gives an upper limit for circular orbits for the visible scattering phase range as a function of inclination. 
-Fig.
-\ref{fig:scatterplot} shows $\Delta \phi$ for the systems in the target list assuming a $3 \lambda / D$, at $\SI{600}{\nano\meter}$, coronograph.
-This optimistic case indicates that it will be possible to observe Rayleigh scattering for all targets and that rainbows will be accessible in some cases.
+\Cref{fig:scatterplot} shows $\Delta \phi$ for the systems in the target list assuming a $3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$. 
+This optimistic case indicates that it will be possible to observe Rayleigh scattering for all targets, and that rainbows will be accessible in some cases.
 
@@
     \caption{
-        Scatter plot at 3 $\lambda/D$ for the target sample showing stellar effective temperature and stellar distance. 
+        \todo{Are those the results we obtained for circular, edge-on orbits? 
+        We should clarify this also in the figure caption.}
+        \todo{The figure is currently broken because showyourwork is using matplotlib 3.4 --- I (Timmy) can debug this if it turns out we cannot use a newer version of matplotlib, but if updating to 3.7 is okay, I will save myself the efforts.}
+        Scatter plot at 3 $\lambda/D$ for the target sample, showing stellar effective temperature and stellar distance. 
         The size of the points represents the angular separation of the star and planet in milliarcseconds as presented in the target list. 
-The colour of the points shows the atmospheric phenomenon that can be detected with darker colors including all lighter (yellow) color phenomenon.
-Thus dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
+        The colour of the points shows the atmospheric phenomenon that can be detected with darker colours, including all lighter (yellow) colour phenomenon. 
+        Thus, dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
     }
@@
 
-As shown in \cref{sec:Delta_phi}, $\Delta \phi$ depends on the angular separation, inclination and eccentricity of the planets orbit. 
-We use the angular separation for an Earth-Like planet with the same instellation as Earth which is provided in the input star list. 
+As shown in \cref{sec:Delta_phi}, $\Delta \phi$ depends on the angular separation, inclination, and eccentricity of the planet's orbit. 
+We use the angular separation for an Earth-Like planet with the same instellation as Earth, which is provided in the input star list. 
 The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
@@
 \label{sec:circular}
-For each star on the HabWorlds target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
+For each star on the \hwo target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
 The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$.
-We then use Eqn \ref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
-Figure blah shows the cumulative distribution for the most extreme scattering phase angles accessible for the HabWorlds target list, depending on the choice of inner working angle.
+We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet. 
+\todo{Figure X} shows the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo target list, depending on the choice of inner working angle.
 
@@
 \label{sec:eccentric}
-We repeat the Monte Carlo simulation from section \ref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
+We repeat the Monte Carlo simulation from \cref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
 \citet{Guimond_2019} and references therein).
 We randomly drew 1000 eccentricities from the beta distribution for each star and determined the orbits. 
-We then filtered on $\IWA$ for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows or the Rayleigh peak occur.
-Figure \ref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's.
-Each plot has been scaled such that the habitable zone occupies the same radius hence the \IWA's vary in apparent size compared to the orbit.
+We then filtered on \IWA for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's. 
+Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
 
@@
     \caption{
-        A example of the eccentric orbits generated for the stellar sample.
+        An example of the eccentric orbits generated for the stellar sample. %Are we putting all of these somewhere?  Its prob about 7 pages worth in total?
         The orbits are scaled by the Earth-equivalent flux distance. 
-The shaded region indicates \IWA's of 1, 2, 3 and 4~$\lambda / D$ corresponding to 21, 42, 63 and 84 mas.
-The \IWA can significantly affect the range of scattering phases observable with each orbit.
+        The shaded region indicates \IWA's of 1, 2, 3 and 4\,$\lambda / D$ corresponding to 21, 42, 63 and \SI{84}{\mas}.
+        The \IWA\ can significantly affect the range of scattering phases observable with each orbit.
     }
     \label{fig:ball-o-yarn}
-    \script{make_yarn_plot.py}
+    \script{make_yarn_plot.py} % most awesome command in this paper.  This knitter approves.
 \end{figure*}

### 2023-02-27 src/scripts/orbfunc.py
[showyourwork] overleaf sync

@@ -1,0 +1,74 @@
+import numpy as np
+
+from scipy.spatial.transform import Rotation as R
+from scipy.interpolate import interp1d
+
+import astropy.units as u
+
+###########################################################################################
+
+def phase(time, params):
+
+        phases = ( ( (time - params['time periapsis passage']).jd ) /params['orbital period'].to(u.day).value)%1
+        phases = np.array(phases).flatten()
+        phases = convert_ma_to_ta(phases, params)
+        
+        return phases
+        
+###########################################################################################
+
+def convert_ma_to_ta(ma, params):
+
+        phases = np.arange(0,1,0.00001)
+        f = interp1d(mean_anomaly(phases, params), phases, bounds_error=False, fill_value='extrapolate')
+
+        return f(ma)
+
+###########################################################################################
+
+def mean_anomaly(true_anomaly, params):
+
+        e = params['eccentricity']
+        ta = true_anomaly*np.pi*2
+        ma = 2*np.arctan(np.sqrt((1-e)/(1+e))*np.tan(ta/2)) + [2*np.pi if p>np.pi else 0 for p in ta] - e*(1-e**2)*np.sin(ta)/(1+e*np.cos(ta))
+
+        return ma / (np.pi*2)
+
+###########################################################################################
+
+def orbit_radius(time, params):
+
+        phases = phase(time, params)
+        r = ( params['semi-major axis']*(1 - params['eccentricity']**2)/(1 + params['eccentricity']*np.cos(2 *np.pi*phases)) ).to(u.au)
+
+        return np.array(r).flatten() * u.au
+
+###########################################################################################
+
+def coordinates3D(coords, time, params):
+
+        aperi = -(params['argument of periapsis'] + params['periapsis precession']*(time - params['time periapsis passage']) ).to(u.rad).value
+        inc = np.zeros(time.shape) + (np.pi/2 - params['inclination'].to(u.rad).value)
+        lan = np.zeros(time.shape) - params['longditude of ascending node'].to(u.rad).value
+
+        rotation_matrix = R.from_euler('ZXZ', np.array([lan, inc, aperi]).T)
+
+        return rotation_matrix.apply(coords)
+
+###########################################################################################
+
+def xyz_position(time, params):
+
+        phases = phase( time, params )
+        r = orbit_radius( time, params )
+
+        x = -r*np.cos(2 *np.pi*phases)
+        y = r*np.sin(2 *np.pi*phases)
+
[…]

### 2023-02-27 src/scripts/plot-orb-grid.py

@@ -1,0 +1,201 @@
+import numpy as np
+
+import astropy.units as u
+from astropy.time import Time
+
+import paths
+
+import matplotlib.pyplot as plt
+from matplotlib.patches import Arrow
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+from mpl_toolkits import mplot3d
+
+from orbfunc import xyz_position, phase
+
+##########################################################################################
+
+
+
+params = {
+    'time periapsis passage':Time('2021-09-29T00:00:00', format='fits'), 
+    'orbital period':1*u.day,
+    'semi-major axis':1*u.au,
+    'eccentricity':0.0,
+    'argument of periapsis':np.pi*0/180*u.rad,
+    'periapsis precession':0*u.rad/u.s,
+    'inclination':np.pi*75/180*u.rad,
+    'longditude of ascending node':np.pi*90/180*u.rad
+}
+
+time = params['time periapsis passage'] - 0.5*u.day + np.arange(0,1,0.001)*u.day
+
+##########################################################################################
+
+def setup_plot(fig, i, oplane=False, losplane=False):
+
+    ax = fig.add_subplot(4, 3, i, projection='3d')
+    ax.set_proj_type('ortho')
+    
+    # plot orbit
+    ax.plot3D(xyz[:,0], xyz[:,1], xyz[:,2], 'k-', lw=2)
+
+    # plane of los
+    if losplane==True:
+        lossurfx, lossurfy = np.meshgrid(np.linspace(-0.65,0.65,10),np.linspace(-0.65,0.65,10))
+        x1 = ax.plot_surface(lossurfx, lossurfy, np.zeros((10,10)), color='g', edgecolor=(0,0,0,0.15), alpha=0.1, label='Plane of Nodes')
+        
+        ### Trying to catch a version dependent error in matplotlib, may not work for all versions ###
+        try:
+            x1._facecolors2d = x1._facecolor3d
+            x1._edgecolors2d = x1._edgecolor3d
+        except:
+            try:
[…]

### 2023-02-27 src/scripts/create-scatterplot.py
[showyourwork] overleaf sync

@@ -65,7 +65,7 @@ if __name__ == '__main__':
 
     # Define a mapping of quantity (e.g., beta value) to a color
     norm = mpl.colors.Normalize(vmin=0, vmax=90)
-    cmap = mpl.cm.get_cmap('viridis_r')
+    cmap = mpl.colormaps['viridis_r']
 
     # Define a function to map a quantity (e.g., beta value) onto a color
     def quantity_to_color(q: float) -> Any:

### 2023-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \label{fig:orb-grid}
+    \script{plot-orb-grid.py}
 \end{figure*}
@@
         We should clarify this also in the figure caption.}
-        \todo{The figure is currently broken because showyourwork is using matplotlib 3.4 --- I (Timmy) can debug this if it turns out we cannot use a newer version of matplotlib, but if updating to 3.7 is okay, I will save myself the efforts.}
         Scatter plot at 3 $\lambda/D$ for the target sample, showing stellar effective temperature and stellar distance. 

### 2023-02-27 src/scripts/create-scatterplot.py
Reformat and clean up code slightly

@@ -5,7 +5,7 @@ Create a color-coded scatter plot the the target list.
 # -----------------------------------------------------------------------------
 # IMPORTS
 # -----------------------------------------------------------------------------
-import paths
+
 from typing import Any
 
 from matplotlib.colorbar import Colorbar
@@ -17,7 +17,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
-print(mpl.__version__)
+import paths
+
 
 # -----------------------------------------------------------------------------
 # AUXILIARY FUNCTIONS
@@ -40,23 +41,31 @@ def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
 # MAIN CODE
 # -----------------------------------------------------------------------------
 
-if __name__ == '__main__':
-    
+if __name__ == "__main__":
+
     # -------------------------------------------------------------------------
     # Load the data
     # -------------------------------------------------------------------------
 
+    print("\nCREATE SCATTERPLOT\n")
+
+    print("Loading data...", end=" ", flush=True)
+
     # Load the target list
-    file_path = paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv'
+    file_path = paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv"
     data = pd.read_csv(file_path, header=1)
 
     # Load the beta values
-    file_path = paths.data / 'beta_max_3_lambda_over_d.csv'
+    file_path = paths.data / "beta_max_3_lambda_over_d.csv"
     beta_values = np.loadtxt(file_path)
+
+    print("Done!")
 
     # -------------------------------------------------------------------------
     # Create the scatter plot
     # -------------------------------------------------------------------------
+
+    print("Creating scatter plot...", end=" ", flush=True)
 
     # Define a function to map a quantity (e.g., angular separation)
     # onto a marker size
@@ -65,7 +74,7 @@ if __name__ == '__main__':
 
     # Define a mapping of quantity (e.g., beta value) to a color
     norm = mpl.colors.Normalize(vmin=0, vmax=90)
-    cmap = mpl.colormaps['viridis_r']
+    cmap = mpl.colormaps["viridis_r"]
 
     # Define a function to map a quantity (e.g., beta value) onto a color
     def quantity_to_color(q: float) -> Any:
[…]

### 2023-02-27 src/scripts/make_yarn_plot.py
Adjust figure size to match other plots in paper

@@ -12,12 +12,14 @@ plotit = 1 # set to 0 to skip plot generation
 norbs_plot = 10
 nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
 
-tyb = dict(color='black', fontsize=14)
+tyb = dict(color='black', fontsize=10)
+
 
 def draw_iwa(ax, iwa):
 	from matplotlib.patches import Circle
 	for r in iwa:
-		ax.add_patch(Circle((0,0),radius=r,alpha=0.2))
+		ax.add_patch(Circle((0, 0), radius=r, alpha=0.2))
+
 
 # number of planet sims per star
 nplanets = 10
@@ -41,17 +43,25 @@ print(f'# read in {nstars} stars from {fname}')
 betamin = np.zeros((nplanets,nstars,iwa.size))
 betamax = np.zeros((nplanets,nstars,iwa.size))
 
-if (plotit):
-	fig, axes = plt.subplots(nstars_plot, nstars_plot, figsize=(10,10))
+if plotit:
+
+	pad_inches = 0.025
+	fig, axes = plt.subplots(
+		nstars_plot,
+		nstars_plot,
+		figsize=(7 - 2 * pad_inches, 7 - 2 * pad_inches)
+	)
+
 	ax = axes.flatten()
-	fig.suptitle(f'Inner Working Angles = {iwa} mas')
+	# fig.suptitle(f'Inner Working Angles = {iwa} mas')
+
 	# remove the x and y ticks
 	for a in ax:
 		a.set_xticks([])
 		a.set_yticks([])
 		a.set_aspect('equal')
 
-	fig.subplots_adjust(wspace=0, hspace=0)
+	# fig.subplots_adjust(wspace=0, hspace=0)
 
 cax = 0 # counter for the grid of subplots - this increases by 1 for each star AND also counts the star number for output array
 
@@ -63,18 +73,18 @@ for l in t: # loop over all the stars in the table
 	hz =  l['st_eei_angsep'] 	# Habitable zone distance (mas)
 
 	if (plotit) and (cax < (nstars_plot*nstars_plot)):
-		ax[cax].set_xlim(-1.5*hz, 1.5*hz)
-		ax[cax].set_ylim(-1.5*hz, 1.5*hz)
-		draw_iwa(ax[cax],iwa)
[…]

### 2023-02-27 src/scripts/make_yarn_plot.py
Adjust font weight and size

@@ -12,7 +12,7 @@ plotit = 1 # set to 0 to skip plot generation
 norbs_plot = 10
 nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
 
-tyb = dict(color='black', fontsize=10)
+tyb = dict(color='black', fontsize=6)
 
 
 def draw_iwa(ax, iwa):
@@ -76,7 +76,7 @@ for l in t: # loop over all the stars in the table
 		ax[cax].set_xlim(-2*hz, 2*hz)
 		ax[cax].set_ylim(-2*hz, 2*hz)
 		draw_iwa(ax[cax], iwa)
-		ax[cax].text(0.05, 0.95, star, ha='left', va='top', transform=ax[cax].transAxes, fontsize=10)
+		ax[cax].text(0.05, 0.95, star, ha='left', va='top', transform=ax[cax].transAxes, fontsize=6, weight='bold')
 		ax[cax].text(0.05, 0.05, f'HZ={hz:.1f} mas', ha='left', va='bottom', transform=ax[cax].transAxes, **tyb)
 
 

### 2023-02-27 src/scripts/make_yarn_plot.py
Adjust code style to reduce complaints from linter

@@ -8,9 +8,9 @@ from samplers import *
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
 # this will make a plot with the number of orbits per panel and side of plots along one side
-plotit = 1 # set to 0 to skip plot generation
+plotit = 1  # set to 0 to skip plot generation
 norbs_plot = 10
-nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
+nstars_plot = 4  # you will make a plot with nstars_plot by nstars_plot panels
 
 tyb = dict(color='black', fontsize=6)
 
@@ -28,20 +28,21 @@ nplanets = 10
 norbitsample = 1000
 
 # IWA ranges
-iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax at
+# inner working angles (in mas) to compute betamax at
+iwa = np.linspace(20, 120, 5)
 
 # read in the stars
 from astropy.io import ascii
 
 fname = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
-t = ascii.read(paths.data / fname ,header_start = 1, data_start = 2)
+t = ascii.read(paths.data / fname, header_start=1, data_start=2)
 
 nstars = len(t)
 print(f'# read in {nstars} stars from {fname}')
 
 # set up output arrays
-betamin = np.zeros((nplanets,nstars,iwa.size))
-betamax = np.zeros((nplanets,nstars,iwa.size))
+betamin = np.zeros((nplanets, nstars, iwa.size))
+betamax = np.zeros((nplanets, nstars, iwa.size))
 
 if plotit:
 
@@ -70,7 +71,7 @@ for l in t: # loop over all the stars in the table
 	d = l['sy_dist']      		# parsecs
 	Vmag = l['sy_vmag']   		# V magnitude
 	mass = l['st_mass']   		# stellar mass
-	hz =  l['st_eei_angsep'] 	# Habitable zone distance (mas)
+	hz = l['st_eei_angsep'] 	# Habitable zone distance (mas)
 
 	if (plotit) and (cax < (nstars_plot*nstars_plot)):
 		ax[cax].set_xlim(-2*hz, 2*hz)
@@ -93,20 +94,20 @@ for l in t: # loop over all the stars in the table
 		#
[…]

### 2023-02-27 src/scripts/make_yarn_plot.py
Update plot style, add $d_*$ to each panel
The new plot style attempts to be a bit more colorblind-friendly. Feel
free to undo this if you don't like it!

@@ -1,9 +1,30 @@
 import numpy as np
+import matplotlib as mpl
 import matplotlib.pyplot as plt
 from matplotlib.patches import Circle
 import paths
 from phases import *
 from samplers import *
+
+
+# Define colorblind-friendly colors (CBF)
+# Source: https://arxiv.org/abs/2107.02270
+CBF_COLORS = [
+	"#3F90DA",  # blue
+	"#FFA90E",  # orange
+	"#BD1F01",  # red
+	"#94A4A2",  # gray
+	"#832DB6",  # purple
+	"#A96B59",  # brown
+	"#E76300",  # orange
+	"#B9AC70",  # tan
+	"#717581",  # gray
+	"#92DADD",  # light blue
+]
+
+# Use colorblind-friendly colors as default color cycle
+mpl.rcParams['axes.prop_cycle'] = mpl.cycler(color=CBF_COLORS)
+
 
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
@@ -18,7 +39,13 @@ tyb = dict(color='black', fontsize=6)
 def draw_iwa(ax, iwa):
 	from matplotlib.patches import Circle
 	for r in iwa:
-		ax.add_patch(Circle((0, 0), radius=r, alpha=0.2))
+		# ax.add_patch(Circle((0, 0), radius=r, alpha=0.2))
+		ax.add_patch(
+			Circle(
+				(0, 0), radius=r, facecolor='none', edgecolor='k', ls='--',
+				lw=0.5, alpha=1.0
+			)
+		)
 
 
 # number of planet sims per star
@@ -73,13 +100,22 @@ for l in t: # loop over all the stars in the table
 	mass = l['st_mass']   		# stellar mass
 	hz = l['st_eei_angsep'] 	# Habitable zone distance (mas)
 
-	if (plotit) and (cax < (nstars_plot*nstars_plot)):
+	# Set options for text boxes
+	textbox_props = dict(
+		boxstyle='square,pad=0.25',
+		# linewidth=0.5,
+		facecolor='white',
+		edgecolor='none',
+		alpha=0.8,
+	)
+
+	if plotit and (cax < (nstars_plot * nstars_plot)):
 		ax[cax].set_xlim(-2*hz, 2*hz)
 		ax[cax].set_ylim(-2*hz, 2*hz)
 		draw_iwa(ax[cax], iwa)
[…]

### 2023-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \caption{
-        Conversion between $\lambda / D$ and mas for different wavelengths.
-\textbf{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function.
-Also update to use the only wavelength we choose}
+        Conversion between $\lambda / D$ and \si{\mas} for different wavelengths, assuming $D = \SI{6}{\meter}$. 
+        \todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
     }
@@
     \midrule
-    IWA & 1 & 21 & 24\\
-    IWA & 2 & 42 & 48\\
-    IWA & 3 & 63 & 72\\
-    IWA & 4& 84& 96\\
+    IWA & 1 & 20.63 & 24.06 \\
+    IWA & 2 & 41.25 & 48.13 \\
+    IWA & 3 & 61.88 & 72.19 \\
+    IWA & 4 & 82.51 & 96.26 \\
     \midrule
-    OWA & 32 & 660 & 770 \\
-    OWA & 64 & 1320 & 1540\\
+    OWA & 32 &  660.05 & 770.06 \\
+    OWA & 64 & 1320.09 & 1540.11 \\
     \bottomrule
@@
    \caption{
-        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle. 
+        \textbf{I think show your work may have borked this?
+The vertical arrow and the one at the top look like they got squished?} The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle.
         If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
@@
 \label{sec:circular}
-For each star on the \hwo target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
+For each star on the \hwo\ target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
 The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$.
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/ball_of_yarn.pdf}  
+    \includegraphics{figures/ball_of_yarn.pdf}  
     \caption{
-        An example of the eccentric orbits generated for the stellar sample. %Are we putting all of these somewhere?  Its prob about 7 pages worth in total?
+        %Are we putting all of these somewhere?  Its prob about 7 pages worth in total?
+        Random examples of the eccentric orbits generated for the stellar sample.
         The orbits are scaled by the Earth-equivalent flux distance. 
-        The shaded region indicates \IWA's of 1, 2, 3 and 4\,$\lambda / D$ corresponding to 21, 42, 63 and \SI{84}{\mas}.
-        The \IWA\ can significantly affect the range of scattering phases observable with each orbit.
+        The concentric circles marked by the dashed lines indicate inner working angles of 1, 2, 3, and 4\,$\lambda / D$, corresponding to 20.6, 41.3, 61.9, and \SI{82.5}{\mas}, respectively (assuming $\lambda = \SI{600}{\nano\meter}$ and $D = \SI{6}{\meter}$).
+        The figure illustrates that the \IWA\ can significantly affect the range of scattering phases observable with each orbit.
     }
@@
 
-\subsection{The glory of Alpha Cen}
+\subsection{The glory of Alpha Cen}%Halelujah
 \label{sec:ealpha-cen}
@@
 Do we need to not something about things like A stars which are not included in the sample - they were rejected by the list creators - do we need to comment on this?
+- we decided nope.
 
-\section*{Software Data Availability}
-%Maybe we could upload the code that we used for our analysis / plots somewhere?
-%Just for full transparency / reproducibility?
+\section*{Software/Data Availability}
+
 The NASA ExEP target list containing the data used for these simulations is available online.% 

### 2023-02-27 src/scripts/plot-orb-grid.py
[showyourwork] overleaf sync

@@ -1,9 +1,9 @@
 import numpy as np
 
+#from scipy.interpolate import interp1d
+#import astropy.constants as c
 import astropy.units as u
 from astropy.time import Time
-
-import paths
 
 import matplotlib.pyplot as plt
 from matplotlib.patches import Arrow
@@ -35,6 +35,7 @@ def setup_plot(fig, i, oplane=False, losplane=False):
 
     ax = fig.add_subplot(4, 3, i, projection='3d')
     ax.set_proj_type('ortho')
+    ax.margins(-0.499)
     
     # plot orbit
     ax.plot3D(xyz[:,0], xyz[:,1], xyz[:,2], 'k-', lw=2)
@@ -82,8 +83,8 @@ def setup_plot(fig, i, oplane=False, losplane=False):
     # mark points
     ax.plot3D([0],[0],[0], 'C1*', label='Star', markersize=15)
     
-    ax.set_xlim(0.65,-0.65)
-    ax.set_ylim(-0.65,0.65)
+    ax.set_xlim(0.7,-0.7)
+    ax.set_ylim(-0.7,0.75)
     ax.set_zlim(-0.2,0.2)
     ax.axis('off')
     
@@ -93,7 +94,7 @@ def setup_plot(fig, i, oplane=False, losplane=False):
     
 xyz = xyz_position(time, params).to(u.au).value
 
-fig = plt.figure(figsize=(20,10))
+fig = plt.figure(figsize=(15,20)) #plt.figaspect(1)*2)
 
 index_an = np.argmin(abs(phase(time, params)-params['argument of periapsis'].to(u.rad).value/(2*np.pi)))
 pend = np.arctan2(xyz[index_an,1], -xyz[index_an,0])
@@ -108,7 +109,7 @@ inc = np.pi/2 - params['inclination'].to(u.rad).value
 ax.plot3D(di*np.cos(np.linspace(0,inc,100))*np.sin(pend)+xyz[index_an,0], di*np.cos(np.linspace(0,inc,100))*np.cos(pend)+xyz[index_an,1], di*np.sin(np.linspace(0,inc,100)), 'r-')
[…]

### 2023-02-27 src/scripts/plot-orb-grid.py
[showyourwork] overleaf sync

@@ -1,5 +1,5 @@
 import numpy as np
-
+import paths
 #from scipy.interpolate import interp1d
 #import astropy.constants as c
 import astropy.units as u
@@ -199,4 +199,4 @@ ax.plot3D([0],[0],[0], 'ko', label='Mask', markersize=200, alpha=0.3)
 ax.view_init(elev=-40, azim=0)
 
 plt.subplots_adjust(wspace=0.0, hspace=-0.4)
-plt.savefig('./orb-grid.pdf')
+plt.savefig(paths.figures / 'orb-grid.pdf')

### 2023-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure*}%[t]
+\begin{figure}%[t]
    \centering
-   \includegraphics[width=0.99\textwidth]{figures/orb-grid.pdf}
+   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
    \caption{
-        \textbf{I think show your work may have borked this?
-The vertical arrow and the one at the top look like they got squished?} The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle.
+        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle. 
         If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
@@
     \script{plot-orb-grid.py}
-\end{figure*}
+\end{figure}
 

### 2023-02-28 src/scripts/utils.py
Add utils.py

@@ -1,0 +1,42 @@
+"""
+General utility functions (e.g., for plotting).
+"""
+
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+
+import matplotlib.pyplot as plt
+
+
+# -----------------------------------------------------------------------------
+# DEFINITIONS
+# -----------------------------------------------------------------------------
+
+# Define colorblind-friendly colors (CBF)
+# Source: https://arxiv.org/abs/2107.02270
+CBF_COLORS = [
+    "#3F90DA",  # blue
+    "#FFA90E",  # orange
+    "#BD1F01",  # red
+    "#94A4A2",  # gray
+    "#832DB6",  # purple
+    "#A96B59",  # brown
+    "#E76300",  # orange
+    "#B9AC70",  # tan
+    "#717581",  # gray
+    "#92DADD",  # light blue
+]
+
+
+def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
+    """
+    Auxiliary function to set the fontsize of all elements in a plot.
+    """
+
+    for item in (
+        [ax.title, ax.xaxis.label, ax.yaxis.label]
+        + ax.get_xticklabels()
+        + ax.get_yticklabels()
+    ):
+        item.set_fontsize(fontsize)

### 2023-02-28 src/scripts/create-scatterplot.py
Move auxiliary functions to utils.py

@@ -17,24 +17,9 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
+from utils import set_fontsize
+
 import paths
-
-
-# -----------------------------------------------------------------------------
-# AUXILIARY FUNCTIONS
-# -----------------------------------------------------------------------------
-
-def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
-    """
-    Auxiliary function to set the fontsize of all elements in a plot.
-    """
-
-    for item in (
-        [ax.title, ax.xaxis.label, ax.yaxis.label]
-        + ax.get_xticklabels()
-        + ax.get_yticklabels()
-    ):
-        item.set_fontsize(fontsize)
 
 
 # -----------------------------------------------------------------------------

### 2023-02-28 src/scripts/samplers.py
Fix type hints and formatting

@@ -1,17 +1,28 @@
-# from sampling-inclinations-and-eccentricities.ipynb
-#
-# Tim Gebhard
+"""
+Utilities for sampling orbital parameters (inclination / eccentricity).
+"""
+
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+
 from scipy.stats import beta
 
 import numpy as np
 
-def sample_cos_i(n: int = 1) -> np.array:
+
+# -----------------------------------------------------------------------------
+# DEFINITIONS
+# -----------------------------------------------------------------------------
+
+def sample_cos_i(n: int = 1) -> np.ndarray:
     """
     Sample cos(inclination) uniformly from [0, 1].
     """
     return np.random.uniform(0, 1, n)
 
-def sample_i(n: int = 1) -> np.array:
+
+def sample_i(n: int = 1) -> np.ndarray:
     """
     Sample `n` inclinations such that cos(i) is uniform in [0, 1].
     The resulting values are returned in *degree*.
@@ -24,64 +35,78 @@ def sample_e(n: int, a: float = 0.867, b: float = 3.03) -> np.ndarray:
     Sample `n` eccentricities `e` from a beta distribution with shape
     parameters `a` and `b`. The default values for `a` and `b` are
     taken from:
-    
+
         Guimond and Cowan (2019): arXiv:1903.06184
-    
+
     We are truncating `e` to [0, 1] to exclude hyperbolic orbits.
     """
-    
+
     return np.clip(beta.rvs(a=a, b=b, size=n), 0, 1)
 
 
+# -----------------------------------------------------------------------------
+# MAIN CODE (TESTS)
+# -----------------------------------------------------------------------------
+
 if __name__ == "__main__":
 
-	import matplotlib.pyplot as plt
-	n = 100_000
+    import matplotlib.pyplot as plt
 
-	# Sample inclinations i uniformly in cos(i)
-	inclinations = sample_i(n)
+    # -------------------------------------------------------------------------
+    # Test the sampling of inclinations
+    # -------------------------------------------------------------------------
 
-	# Plot a histogram of the inclinations
[…]

### 2023-02-28 src/scripts/make_yarn_plot.py
Reformat file
This was only to improve the readability of the code (and stop my linter
from complaining); the functionality should remain unchanged!

@@ -1,176 +1,288 @@
-import numpy as np
+"""
+Create the yarn ball plot with random eccentric orbits.
+"""
+
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+
+import time
+
+from astropy.io import ascii
+from matplotlib.patches import Circle
+from tqdm.auto import tqdm
+
+import astropy.units as u
 import matplotlib as mpl
 import matplotlib.pyplot as plt
-from matplotlib.patches import Circle
+import numpy as np
+
+from phases import kep3d
+from samplers import sample_e, sample_i
+from utils import CBF_COLORS
+
 import paths
 
-from phases import *
-from samplers import *
 
+# -----------------------------------------------------------------------------
+# DEFINITIONS
+# -----------------------------------------------------------------------------
 
-# Define colorblind-friendly colors (CBF)
-# Source: https://arxiv.org/abs/2107.02270
-CBF_COLORS = [
-	"#3F90DA",  # blue
-	"#FFA90E",  # orange
-	"#BD1F01",  # red
-	"#94A4A2",  # gray
-	"#832DB6",  # purple
-	"#A96B59",  # brown
-	"#E76300",  # orange
-	"#B9AC70",  # tan
-	"#717581",  # gray
-	"#92DADD",  # light blue
-]
+def draw_iwa(ax: plt.Axes, iwas: np.ndarray) -> None:
+    """
+    Auxiliary function to draw inner working angles on a given axis.
+    """
 
-# Use colorblind-friendly colors as default color cycle
-mpl.rcParams['axes.prop_cycle'] = mpl.cycler(color=CBF_COLORS)
-
-
-np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
-
-# this will make a plot with the number of orbits per panel and side of plots along one side
-plotit = 1  # set to 0 to skip plot generation
-norbs_plot = 10
-nstars_plot = 4  # you will make a plot with nstars_plot by nstars_plot panels
-
-tyb = dict(color='black', fontsize=6)
-
-
-def draw_iwa(ax, iwa):
-	from matplotlib.patches import Circle
-	for r in iwa:
-		# ax.add_patch(Circle((0, 0), radius=r, alpha=0.2))
-		ax.add_patch(
-			Circle(
-				(0, 0), radius=r, facecolor='none', edgecolor='k', ls='--',
-				lw=0.5, alpha=1.0
-			)
-		)
[…]

### 2023-02-28 src/tex/tikz/annotated-orbit.tex
Add TikZ plots

\documentclass[10pt, crop, margin=0.5mm, dvipsnames]{standalone}

\usepackage{libertine}
\usepackage[libertine]{newtxmath}

\usepackage{tikz}
\usetikzlibrary{
    calc,
    positioning,
    shapes,
}

\usepackage{xcolor}

\begin{document}
\begin{tikzpicture}

    % Define command for the "eye" symbol
    \newcommand*\lateraleye{%
       \scalebox{0.15}{
    \tikzset{every picture/.style={line width=0.75pt}} 
    \begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
    \draw  [line width=1.5]  (300,100.33) .. controls (326,122) and (352,135) .. (378,139.33) .. controls (352,143.67) and (326,156.67) .. (300,178.33) ;
    \draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (308.94,116.33) .. controls (313.87,116.33) and (317.86,125.51) .. (317.85,136.83) .. controls (317.84,148.15) and (313.84,157.33) .. (308.91,157.33) .. controls (303.99,157.32) and (300,148.14) .. (300.01,136.82) .. controls (300.02,125.5) and (304.02,116.32) .. (308.94,116.33) -- cycle ;
    \draw  [draw opacity=0][line width=1.5]  (314.84,166.6) .. controls (311.87,164.64) and (309.14,162.18) .. (306.76,159.24) .. controls (295.12,144.82) and (296.6,124.33) .. (310.07,113.45) .. controls (311.48,112.32) and (312.96,111.33) .. (314.5,110.49) -- (331.14,139.55) -- cycle ; \draw  [line width=1.5]  (314.84,166.6) .. controls (311.87,164.64) and (309.14,162.18) .. (306.76,159.24) .. controls (295.12,144.82) and (296.6,124.33) .. (310.07,113.45) .. controls (311.48,112.32) and (312.96,111.33) .. (314.5,110.49) ;
    \draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (304.43,124.2) .. controls (306.09,124.25) and (307.32,128.01) .. (307.18,132.6) .. controls (307.05,137.19) and (305.59,140.88) .. (303.93,140.83) .. controls (302.27,140.78) and (301.03,137.02) .. (301.17,132.43) .. controls (301.31,127.83) and (302.76,124.15) .. (304.43,124.2) -- cycle ;
    \end{tikzpicture}
    }\,}

    % Command for arc with defined center
    \def\centerarc[#1](#2)(#3:#4:#5){ \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }

    % Command for planet dayside
    % Arguments: center, radius, color, rotation (-180,...,180)
    \newcommand\dayside[4]
    {%
      \ifnum #4 > 0
        \pgfmathsetmacro\lb{ #2}            % left  arc, horizontal axis
        \pgfmathsetmacro\rb{ #2*(90-#4)/90} % right arc, horizontal axis
      \else
        \pgfmathsetmacro\lb{-#2*(90+#4)/90} % left  arc, horizontal axis
        \pgfmathsetmacro\rb{-#2}            % right arc, horizontal axis  
      \fi
      \draw[thick,#3,fill=white] #1 circle (#2);
      \fill[#3,opacity=0.75] ($#1+(0,#2)$) arc (90:270:\lb cm and #2 cm)
                                          arc (270:90:\rb cm and #2 cm);
    }

    % Define color scheme
    \definecolor{rayleigh}{HTML}{dae319}
    \definecolor{glint}{HTML}{35b779}
    \definecolor{rainbow}{HTML}{31688e}
    \definecolor{other}{HTML}{440154}

    % Background (to center stuff)
    \draw [fill=white, draw=none] (-3.3, 0) rectangle (3.3, 0);

    % Plot what's blocked by the coronagraph
    \draw [fill=black!10, draw=none] (-1, -4) rectangle (1, 4);
    \node [align=center, black!30, font=\small] at (0, 4.5) {Blocked by\\ the coronagraph};

    % Define radius
    \def\radius{2.5}

    % Color-code the orbit
    \centerarc [other, line width=2mm](0, 0)(90:70:\radius);
    \centerarc [other, line width=2mm](0, 0)(90:110:\radius);
    \node [other, anchor=south, font=\small, rotate=80-90] at (80:\radius+0.25) {Glory};

    \centerarc [rainbow, line width=2mm](0, 0)(70:30:\radius);
    \centerarc [rainbow, line width=2mm](0, 0)(110:150:\radius);
    \node [rainbow, anchor=south, font=\small, rotate=50-90] at (50:\radius+0.25) {Rainbows};

    \centerarc [rayleigh, line width=2mm](0, 0)(30:-35:\radius);
    \centerarc [rayleigh, line width=2mm](0, 0)(150:215:\radius);
    \node [rayleigh, anchor=south, font=\small, rotate=0-90] at (0:\radius+0.25)  {Rayleigh};

    \centerarc [glint, line width=2mm](0, 0)(-35:-70:\radius);
    \centerarc [glint, line width=2mm](0, 0)(215:250:\radius);
    \node [glint, anchor=north, font=\small, rotate=-52.5+90] at (-52.5:\radius+0.25) {Glint};

    \centerarc [other, line width=2mm](0, 0)(-70:-90:\radius);
    \centerarc [other, line width=2mm](0, 0)(250:270:\radius);
    \node [other, anchor=north, font=\small, rotate=-80+90] at (-80:\radius+0.25) {Other};

    % Plot the orbit
    \draw [dotted, black, semithick] (0, 0) circle (\radius);

    % Star, planet position, and observer
    \node [star, fill=Dandelion, star points=5, inner sep=1mm] (star) (star) at (0, 0) {};
    \node [Dandelion, above=1mm of star] {Star};
    \node [] (planet) at (135:\radius) {};
    \node [black, left=1mm of planet.west] {planet};
    \node [circle, anchor=center, align=center, rotate=-90, inner sep=0mm] (observer) at (0, -6) {\lateraleye};
    \node [black, anchor=east, xshift=-2mm] at (observer) {Observer};

    % Plot some lines
    \draw [shorten <= 1mm, shorten >= 3mm, densely dashed] (star) -- node [pos=0.45, font=\scriptsize, below=0mm] {semi-major axis} (0:\radius);
    \draw [shorten >= 2mm, shorten <= 2mm, gray] (star) -- (planet);
    \draw [shorten >= 1mm, shorten <= 2mm, gray] (star) -- (observer);
    \begin{scope}[yshift=-5cm]
        \draw [white, thick] (0, 0) -- (0, 0.1);
        \draw [gray] (-0.2, -0.1) -- (0.2, 0.1);
        \draw [gray] (-0.2, 0) -- (0.2, 0.2);
    \end{scope}

    % Plot planets (daysides)
    \foreach\i in {-180,-135,...,180} {
        \dayside{(\i+90:\radius)}{0.2}{black}{\i};
    }

    % Plot the scattering angle in red
    \draw [draw=red, thick, shorten >= 1mm, shorten <= 1mm, <->] ($(0, 0)+({1*cos(135)},{1*sin(135)})$) arc (135:270:1);
    \node [red] at (202.5:0.75) {$\varphi$};

\end{tikzpicture}
\end{document}

### 2023-03-02 src/scripts/plot_contrasts.py
[showyourwork] overleaf sync

@@ -1,0 +1,147 @@
+from typing import Any
+
+from matplotlib.colorbar import Colorbar
+from matplotlib.lines import Line2D
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from scipy import interpolate
+
+import paths
+
+# -----------------------------------------------------------------------------
+# AUXILIARY FUNCTIONS
+# -----------------------------------------------------------------------------
+
+def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
+    """
+    Auxiliary function to set the fontsize of all elements in a plot.
+    """
+
+    for item in (
+        [ax.title, ax.xaxis.label, ax.yaxis.label]
+        + ax.get_xticklabels()
+        + ax.get_yticklabels()
+    ):
+        item.set_fontsize(fontsize)
+
+def quantity_to_ms(q: float) -> float:
+    return 5 * (1 + q / 500)
+
+
+# Define a function to map a quantity (e.g., beta value) onto a color
+def quantity_to_color(q: float) -> Any:
+    return cmap(norm(q))
+
+
+df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')
+
+hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
+contrasts =np.array(df['Contrast'].values[1:],dtype=float)
+names = np.array(df['CommonID'].values[1:],dtype=str) 
+
+data_trees = np.loadtxt(paths.data / 'Reflected_light_curves_Trees_Stam_2019.txt')
+
+
+P_670 = interpolate.interp1d(data_trees[0],data_trees[2])
+F_670 = interpolate.interp1d(data_trees[0],data_trees[1])
+
+orbital_phase = np.linspace(0,2*np.pi,200)
+
+inclination = np.radians(63)
+
+alpha = np.arccos(np.cos(orbital_phase)*np.sin(inclination))
+
+# Define a mapping of quantity (e.g., beta value) to a color
+norm = mpl.colors.Normalize(vmin=0, vmax=90)
+cmap = mpl.cm.get_cmap('viridis_r')
+
+
+names_array = np.array(['Rigil Kentaurus','Ran' ,'Lalande 21185'])
[…]

### 2023-03-02 src/tex/ms.tex
Add TikZ plots
[showyourwork]

@@
 \usepackage[T1]{fontenc}
+\usepackage[dvipsnames, x11names]{xcolor}
 
@@
 \usepackage[capitalise]{cleveref}
-%\usepackage[dvipsnames, x11names]{xcolor}
 %\usepackage{xspace}
@@
 \usetikzlibrary{
+    calc,
     positioning,
@@
     Nicolas B. Cowan$^{4}$,
-    David Doelman$^{5}$,\newauthor
+    David S. Doelman$^{5}$,\newauthor
     Timothy D. Gebhard$^{6}$,
@@
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak.
-The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in XX systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in YY systems.
+The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in \textbf{XX} systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in \textbf{YY} systems.
 The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.
@@
 We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
-
-For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{700}{\nano\meter}$.
+For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.
 
@@
 %
+A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, i.e.
+the relative flux of the reflected light from the planet divided by the stellar flux.
+While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.
+This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects. 
+Here, we aim to present some simplified examples of the impact of the scattering angle dependent contrast on the detectability of rainbow and glint features. 
+We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of Earth-like planet \citep{ treesstam2019,trees2022}.\\
+The NASA Exoplanet Exploration Program’s Mission Star
+List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
+The contrast is calculated as 
+\begin{equation}
+C = F_p/F_* = p \phi (\alpha) (R_p/a)^2,
+\end{equation}
+where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$
+is the planet’s radius, and $a$ is the separation of the planet from its star. 
+For their calculations they made the simplified assumptions of circular orbits, a geometric albedo of $p=0.2$, and that the scattering phase function, $\phi (\alpha)$, can be described by a simple Lambertian reflectance phase function, e.g.
+isotropic scattering.
+These calculations give a good order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
+Therefore, in this work we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
+We note that we do not aim to accurately calculate the absolute scaling of the contrast as this depends on many factors and this would require to redo some of the work done for the star list.\\
 
-% \subsection{Monte Carlo Simulations}
-% \label{sec:montecarlo}
-% For each star on the HabWorlds target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis.
-The reflected flux and polarisation are not constant as a function of scattering phase angle.
-This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
-These are modelled using ...
+We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of 7m/s at 670 nm.
+The 670 nm is close to the center of the Rc band for Vega (642 nm). 
+We normalize the reflected light curve at the value of 90 degrees (quadrature) and multiply it with the contrast.
+In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of 60 and 90 degrees.
+Together they give the reflected light contrast as function of separation for a full orbit.
+We multiply this flux with the degree of polarisation and retrieve the polarised contrast.
+
+See Figure~\ref{fig:contrasts}.
+
+\begin{figure*}%[t]
+   \centering
+   \includegraphics[width=0.99\textwidth]{figures/Contrast_vs_separation_inclination_angle_63.pdf}
+   \caption{
+    Contrast plots for three stars.
+    }
+    \label{fig:contrasts}
+    \script{plot_contrasts.py}
+\end{figure*}
 
@@
 
+\begin{figure}
+    \centering
+    \includestandalone{tikz/annotated-orbit}
+    \caption{
+        Annotated orbit --- let's see if this works.
+    }
+    \label{fig:annotated-orbit}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includestandalone{tikz/scattering-angle}
+    \caption{
+        Scattering angle --- let's see if this works.
+    }
+    \label{fig:scattering-angle}
+\end{figure}
+
 \subsection{The glory of Alpha Cen}%Halelujah

### 2023-03-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in \textbf{XX} systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in \textbf{YY} systems.
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with IWA.
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Methods/simulations}
@@
 \hwo\ is still in the early stages of development, and the exact telescope and instrument design is undecided. 
+%
 In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
-
-\hwo\ will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions.
+%
+In addition, \hwo\ will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
+%
 The \IWA\ of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
-In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
-We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
+%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
+%We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
 For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.
@@
 %\subsection{Stellar sample and assumed planet parameters} 
-We use the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
+\hwo is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
 \footnote{\url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
+%
 This list comprises $\sim$160 stars, the majority of which are Sun-like dwarfs; 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
-The target list is constructed assuming a maximum planet magnitude of $R_c = 31$ and a contrast floor of \num{2.5e-11}. %with an occurrence rate of
-%rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
-The corresponding adopted habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the
-bolometric luminosity normalized to the Sun. 
+%
+The target list is constructed assuming a maximum planet magnitude of $R_c = 31$ and a contrast floor of \num{2.5e-11}. 
+%with an occurrence rate of rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
+The corresponding adopted habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the bolometric luminosity normalized to the Sun.
+%
 This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}.
@@
 The coronagraph, while necessary for observing faint exoplanets, can obscure part of the planet's orbit. 
-As demonstrated in \cref{fig:orb-grid}, this can prevent the planet being detected at high and low scattering angles. 
+%
+As demonstrated in \cref{fig:annotated-orbit}, this can prevent the planet being detected at high and low scattering angles. 
+%
 In this section, we derive the minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible for a given orbit of semi-major $a$ (in \si{\au}), line of sight inclination $i$, star distance $d_*$ in parsecs (\si{\parsec}) and for a given inner working angle of the instrument \IWA\ (in milliarcseconds).
+%
 We define the inclination $i$, such that $i=0$ corresponds to a face-on orbit.
+
+\begin{figure}
+    \centering
+    \includestandalone{tikz/annotated-orbit}
+    \caption{
+        Annotated orbit --- let's see if this works.
+    }
+    \label{fig:annotated-orbit}
+\end{figure}
 
@@
     \centering
-    \includestandalone{tikz/annotated-orbit}
-    \caption{
-        Annotated orbit --- let's see if this works.
-    }
-    \label{fig:annotated-orbit}
-\end{figure}
-
-\begin{figure}
-    \centering
     \includestandalone{tikz/scattering-angle}

### 2023-03-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\todo}[1]{\textcolor{red}{[#1]}}
+\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}}
 
@@
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak.
-The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in \textbf{XX} systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in \textbf{YY} systems.
+The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$20 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$55 systems, assuming a 60 mas inner working angle (\IWA).
 The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-\textcolor{red}{introduce motivation, specs for habitable worlds observatory and the document we got the target list from.}
+%\textcolor{red}{introduce motivation, specs for habitable worlds observatory and the document we got the target list from.}
 
-\textcolor{blue}{Habitable worlds bit}
+Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be remotely detectable on an exoplanet if it is present at the surface.
+Hence the search for life on exoplanets typically focuses on habitable planets.
+To first order, the presence of surface liquid water can be predicted by its instellation---the flux it receives from its star.
+The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \cite{kasting93}.
+The HZ is popular among planet hunters because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
+Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
+
+\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
+%\textbf{Why glint [Kim]}
+% probe of surface conditions; habitability
+\cite{2001Natur.412..885F} first noted the potential for time-resolved photometry and polarimetry to identify water oceans on the surface of an exoplanet.
+\textcolor{blue}{A sentence about rotational mapping here.} \cite{2006astro.ph.10518M} further quantified how photometric and polarimetric scattered light phase variations might betray the presence of oceans on terrestrial exoplanets.
+%--> No it was talked about previous to this, per the paper cited here in fact...
+"Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint.
+Schneider (2006) had similar ideas." -KB
+\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
+They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
+\cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.
+Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
+\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
+\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths.
+This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+\cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
+\cite{2022A&A...664A.172T}...
+
+%cloud coverage paper in polarisation :
+%\cite{rossi17} computed flux and polarisation of reflected starlight for different types of cloud cover on Earth-like planets and determined different types of cloud cover can be distinguished from each other.
+%Alternative options or papers here....
+Kim can write this tomorrow , needs other Stam, Karalidi, Kopparla, maybe Gordon, need to reference rainbows in Bailey paper.
+Look at citations in Gordon paper (our bib)...
+he has a really thorough lit review --KB
 
@@
 
-\textbf{Why glint [Kim]}
-% probe of surface conditions; habitability
-
-%\cite{2006astro.ph.10518M} first noted the potential for photometric and polarimetric scattered light phase to detect oceans on terrestrial exoplanets.
---> No it was talked about previous to this, per the paper cited here in fact...
-"Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint.
-Schneider (2006) had similar ideas." -KB
-
-%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
-They noted that surface oceans would be detectable because they are a) less reflective near full phase,
-% not following what you mean here -KB
-%b) more reflective at crescent phases, and c) exhibit a peak in polarization at crescent phases.
-\cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.
-Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
-\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to simulations; they found .
-%\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths.
-This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
-\cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
-\cite{2022A&A...664A.172T}
-
-%cloud coverage paper in polarisation :
-%\cite{rossi17} computed flux and polarisation of reflected starlight for different types of cloud cover on Earth-like planets and determined different types of cloud cover can be distinguished from each other.
-%Alternative options or papers here....
-Kim can write this tomorrow , needs other Stam, Karalidi, Kopparla, maybe Gordon, need to reference rainbows in Bailey paper.
-Look at citations in Gordon paper (our bib)...
-he has a really thorough lit review --KB
-
 \textbf{In this paper we...}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Observing scattering phenomena}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Methods/simulations}
 \subsection{\hwo\ and Stellar Sample}
@@
 %
-In addition, \hwo\ will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
+In addition, the observatory will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
 %
@@
 
-%\subsection{Stellar sample and assumed planet parameters} 
-\hwo is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
+\hwo\ is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
 \footnote{\url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Maximum scattering phase angle coverage}
@@
 
-The coronagraph, while necessary for observing faint exoplanets, can obscure part of the planet's orbit. 
+The coronagraph can obscure part of the planet's orbit which, as demonstrated in \cref{fig:annotated-orbit}, can prevent the planet from being observed at high and low scattering angles. 
 %
-As demonstrated in \cref{fig:annotated-orbit}, this can prevent the planet being detected at high and low scattering angles. 
-%
-In this section, we derive the minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible for a given orbit of semi-major $a$ (in \si{\au}), line of sight inclination $i$, star distance $d_*$ in parsecs (\si{\parsec}) and for a given inner working angle of the instrument \IWA\ (in milliarcseconds).
-%
-We define the inclination $i$, such that $i=0$ corresponds to a face-on orbit.
+The range of scattering phase angles that can be observed will depend on the \IWA\ of the coronagraph as well as the shape and orientation of the exoplanet's orbit.
 
@@
     \caption{
-        Annotated orbit --- let's see if this works.
+        The coronagraph can obscure the exoplanet at high and low scattering phase angles.
+This can prevent observations of scattering phenomena which could be used to indicate the habitability and atmospheric composition of the exoplanet.
+\todo{Add why scattering is cool}
     }
@@
 
-\begin{figure}%[t]
-   \centering
-   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
-   \caption{
-        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle. 
-        If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
-        For orbital inclinations close to edge-on (bottom panels) the planet goes through all phases, but the scattering phase angles near 0$^\circ$ and 180$^\circ$ are not accessible due to obscuration by the coronagraph mask; for circular orbits, the accessible scattering phases are $\phi \in 90^\circ \pm \arcsin({\rm IWA}\cdot d_*/a)$.
-    }
-    \label{fig:orb-grid}
-    \script{plot-orb-grid.py}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), $\phi_\mathrm{min}$ and $\phi_\mathrm{max}$ are symmetric about 90 degrees (). 
+\subsubsection{Circular edge-on orbits}
+
+In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible are symmetric about 90 degrees ().
+%
 We define the angle $\Delta \phi$ such that: 
@@
 
-For a given system with inclination $i$, we have identified 2 regimes, shown in \cref{fig:orb-grid}:
-
-(1)~If the inclination is face-on enough such that the coronagraph mask does not cover any part of the projected orbit on-sky, then the maximum scattering phase angle coverage depends solely on the disk inclination ;
-
-(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the \IWA\ and orbit semi-major axis.
-In this section, we derive the analytical formula for this case in the circular orbit case.
-
-%\begin{figure}%[t]
-%    \centering
-    % \includegraphics[width=\linewidth]{figures/scattering_angles.pdf}
-%    \includestandalone{figures/scattering-phase-angle}
-%    \caption{
-%        The angular separation, $\varepsilon$, of a planet orbiting with semi-major-axis $a$ about a star a distance $d_*$ from the observer. 
-%        The scattering phase angle is $\varphi$. 
-%        The smallest accessible scattering phase angle occurs when the angular separation equals the coronagraph inner working angle: $\varphi_{\rm min} = {\rm IWA} \cdot d_*/a$.
-%    }
-%    \label{fig:scattering_angles}
-%\end{figure}
-
-%From \cref{fig:scattering_angles}, we deduce that for all points on the orbit, the scattering angle $\varphi$ satisfies:
+The on-sky projected planet distance, $r_\mathrm{proj}$ (in \si{\au}) is shown in \cref{fig:scattering-angle} assuming a non-eccentric circular orbit with a semi-major $a$ (in \si{\au}) at a distance $d_*$ (in \si{\parsec}) from the observer.
+%
+The scattering phase angle $\phi$ is this set up follows the equation,
 \begin{equation}
-    \sin(\varphi) = \frac{r_\mathrm{proj}}{a}
+    \sin(\phi) = \frac{r_\mathrm{proj}}{a}
 \end{equation}
-with $r_\mathrm{proj}$ the on-sky projected planet distance (in \si{\au}), $a$ the distance to the star (constant in this section, in \si{\au}).
-
-$\varepsilon$ is the angular separation of the planet at a given point of the orbit ($r_\mathrm{proj} = \delta d_*$). 
+%
+Using small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the angular separation on-sky.
+%
 We access the minimum scattering angle when the projected distance reaches the inner working angle separation $\delta = \mathrm{IWA}$. 
+%
 Using \cref{eq:Delta_phi}, we deduce:
@@
 \begin{equation}
+    \label{eq:scattering_angle}
     \cos\left(\dfrac{\Delta \phi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a}
@@
 
-We now deduce the 2 regime equation for all cases (circular orbit): 
-\begin{equation}
-\label{eq:Delta_phi_max}
-    \Delta \phi = 
-    \begin{cases}
-        2 \cdot i & \textrm{for} \cos(i) > \dfrac{\mathrm{IWA}\cdot d_* }{a}
-  \\ 
-        2 \cdot \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right)  & \textrm{for} \cos(i) < \dfrac{\mathrm{IWA}\cdot d_* }{a}
-    \end{cases}\,.
-\end{equation}
+\begin{figure}
+    \centering
+    \includestandalone{tikz/scattering-angle}
+    \caption{
+        The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major $a$ at a distance $d_*$.
+The scattering phase angle, $\phi$, is the angle between the line of sight to the star and the line connecting the star and planet.
+$\delta$ is the on-sky angular separation as viewed by the observer.
+\todo{should the red angle be $\phi$?}
+    }
+    \label{fig:scattering-angle}
+\end{figure}
 
-\subsection{Edge-on circular orbits}
-\label{sec:edge-on}
-We initially performed a simple test considering only edge-on circular orbits. 
-This gives an upper limit for circular orbits for the visible scattering phase range as a function of inclination. 
 \Cref{fig:scatterplot} shows $\Delta \phi$ for the systems in the target list assuming a $3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$. 
-This optimistic case indicates that it will be possible to observe Rayleigh scattering for all targets, and that rainbows will be accessible in some cases.
+%
+This assumes each planet has a semi-major axis such that it receives an Earth-equivalent instellation, which is provided in the input star list.
+%
+For edge-on circular orbits, it will be possible to observe Rayleigh scattering for all targets and rainbows will be accessible in some cases.
 
@@
 
-\subsection{Assumed planet parameters}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-As shown in \cref{sec:Delta_phi}, $\Delta \phi$ depends on the angular separation, inclination, and eccentricity of the planet's orbit. 
-We use the angular separation for an Earth-Like planet with the same instellation as Earth, which is provided in the input star list. 
-The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
+\subsubsection{Circular inclined orbits}
 
-\subsection{Making Contrast Curves}
+Most systems will not be viewed as edge on from Earth.
+%
+For a given system with inclination $i$, we have identified 2 regimes, which are shown in \cref{fig:orb-grid}:
+
+(1)~If the inclination is face-on enough such that the coronagraph mask does not cover any part of the projected orbit on-sky, then the maximum scattering phase angle coverage depends solely on the disk inclination ;
+
+(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the \IWA\ and orbit semi-major axis.
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
+   \caption{
+        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle. 
+        If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
+        For orbital inclinations close to edge-on (bottom panels) the planet goes through all phases, but the scattering phase angles near 0$^\circ$ and 180$^\circ$ are not accessible due to obscuration by the coronagraph mask; for circular orbits, the accessible scattering phases are $\phi \in 90^\circ \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+    }
+    \label{fig:orb-grid}
+    \script{plot-orb-grid.py}
+\end{figure}
+
+Using \cref{eq:scattering_angle}, we deduce the 2 regime equation for inclined circular orbits: 
+\begin{equation}
+\label{eq:Delta_phi_max}
+    \Delta \phi = 
+    \begin{cases}
+        2 \cdot i & \textrm{for} \cos(i) > \dfrac{\mathrm{IWA}\cdot d_* }{a}
+  \\ 
+        2 \cdot \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right)  & \textrm{for} \cos(i) < \dfrac{\mathrm{IWA}\cdot d_* }{a}
+    \end{cases}\,.
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Eccentric orbits}
+
+\todo{how was the angle coverage calculated in the monte carlo?}
+
+\timmy{I think so far, the MC code doesn't compute angle coverage. I came up with this quick and dirty sketch to help us get started:}
+
+\includegraphics[width=\linewidth]{figures/ellipse.pdf}
+
+\timmy{If you agree that this makes sense, I could try to add this to the yarn ball script by Matt?}
+
+%The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Simulating contrast curves}
 \textbf{Kim and David}
@@
 the relative flux of the reflected light from the planet divided by the stellar flux.
+%
 While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.
+%
 This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
+%
 Here, we aim to present some simplified examples of the impact of the scattering angle dependent contrast on the detectability of rainbow and glint features. 
-We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of Earth-like planet \citep{ treesstam2019,trees2022}.\\
-The NASA Exoplanet Exploration Program’s Mission Star
-List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
+%
+We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of Earth-like planet \citep{ treesstam2019,trees2022}.
+
+The NASA Exoplanet Exploration Program’s Mission Star List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
+%
 The contrast is calculated as 
@@
 \end{equation}
-where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$
-is the planet’s radius, and $a$ is the separation of the planet from its star. 
+where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
+%
 For their calculations they made the simplified assumptions of circular orbits, a geometric albedo of $p=0.2$, and that the scattering phase function, $\phi (\alpha)$, can be described by a simple Lambertian reflectance phase function, e.g.
 isotropic scattering.
+%
 These calculations give a good order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
+%
 Therefore, in this work we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
-We note that we do not aim to accurately calculate the absolute scaling of the contrast as this depends on many factors and this would require to redo some of the work done for the star list.\\
+%
+We note that we do not aim to accurately calculate the absolute scaling of the contrast as this depends on many factors and would have required some of the work for the star list to be redone.
 
-We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of 7m/s at 670 nm.
-The 670 nm is close to the center of the Rc band for Vega (642 nm). 
+We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
+%
+The $\SI{670}{\nano\meter}$ is close to the center of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
+%
 We normalize the reflected light curve at the value of 90 degrees (quadrature) and multiply it with the contrast.
-In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of 60 and 90 degrees.
+%
+In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of $60$ and $90$ degrees.
+\todo{Only the 60 degree plot is here at the moment}
+%
 Together they give the reflected light contrast as function of separation for a full orbit.
+%
 We multiply this flux with the degree of polarisation and retrieve the polarised contrast.
-
-See Figure~\ref{fig:contrasts}.
+%
+The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G dwarf $\alpha$ Cen A, the K dwarf $\epsilon$ Eri and the M dwarf Lalande 21185.
 
@@
    \caption{
-    Contrast plots for three stars.
+    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of $60^\circ$ for the stars $\alpha$ Cen A, $\epsilon$ Eri and Lalande 21185.
+The solid line indicates the contrast in unpolarised light with the contrast at quadrature marked by a solid dot.
+The polarised component is indicated by the colored dots for which the color represents the scattering phase angle from quadrature.
+The light grey points show the quadrature contrasts of the other targets in the star list and the dashed lines indicate $1,2$ and $3$ times the \IWA\ for \hwo.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Results}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \label{sec:circular}
+
 For each star on the \hwo\ target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
+%
 The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$. 
+%
 We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
-\todo{Figure X} shows the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo target list, depending on the choice of inner working angle.
+%
+\todo{Figure X, where the plot for this?} shows the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo\ target list, depending on the choice of inner working angle.
 
-\textbf{Is there a plot for this?}
+\todo{what does this plot show, how many systems could we see?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \citet{Guimond_2019} and references therein).
+%
 We randomly drew 1000 eccentricities from the beta distribution for each star and determined the orbits. 
-We then filtered on \IWA for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+%
+We then filtered on \IWA\ for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+%
 \Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's. 
-Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
+%
+%Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
+
+\todo{Is there a plot for this? How many systems could we see? Why is it different to non-eccentric cases?}
+
+\timmy{No plot / results yet; see above.}
 
@@
 
-\begin{figure}
-    \centering
-    \includestandalone{tikz/scattering-angle}
-    \caption{
-        Scattering angle --- let's see if this works.
-    }
-    \label{fig:scattering-angle}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection{Contrast Curves}
-\label{sec:contrast}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Discussion}
-What did we determine?
-Can we actually detect this stuff for anything in this sample?
-Do we need to not something about things like A stars which are not included in the sample - they were rejected by the list creators - do we need to comment on this?
-- we decided nope.
+%\subsection{Contrast Curves}
+%\label{sec:contrast}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusions}
+%What did we determine? 
+%Can we actually detect this stuff for anything in this sample? 
+%Do we need to not something about things like A stars which are not included in the sample - they were rejected by the list creators - do we need to comment on this?  - we decided nope.
+
+The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040s \todo{check}. 
+%
+The instrument design and capabilities are still under development and thus can be driven by the science goals of the mission.
+%
+In this work, we investigated the possibility of using \hwo\ to detect different scattering phenomena which can be used to indicate the habitability of an exoplanet. \todo{more of why this is cool}
+
+Coronagraphy is a rapidly advancing field and it is hard to predict what the capabilities of \hwo\ will be.
+%
+We primarily focused on the range of scattering phase angles and therefore which scattering phenomena could potentially be observed given the limitation of the coronagraphic \IWA.
+%
+Assuming a 60 mas \IWA, all of the systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, ocean glint would be accessible in $\sim$20 systems, and the rainbows of water clouds in $\sim$55 systems.
+%
+The number of systems increases with decreasing \IWA\ and therefore the occurrence rate of Earth-like exoplanets will drive the choice of coronagraphic \IWA\ for this mission.
+%
+While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
+%
+This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph.
 

### 2023-03-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\todo}[1]{\textcolor{red}{[#1]}}
-\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}}
+\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}} % amazing
 
@@
     \caption{
-        \todo{Are those the results we obtained for circular, edge-on orbits? 
-        We should clarify this also in the figure caption.}
+        \timmy{I believe those should be the results for circular, edge-on orbits, but let's double-check with Max.}
         Scatter plot at 3 $\lambda/D$ for the target sample, showing stellar effective temperature and stellar distance. 
@@
 
-\includegraphics[width=\linewidth]{figures/ellipse.pdf}
-
-\timmy{If you agree that this makes sense, I could try to add this to the yarn ball script by Matt?}
+\begin{figure}
+   \centering
+   \includegraphics[width=0.99\columnwidth]{figures/ellipse.pdf}
+   \caption{
+  Ellipse drawingIf you agree that this makes sense, I could try to add this to the yarn ball script by Matt?    }
+    \label{fig:ellipse}
+%    \script{plot-orb-grid.py}
+\end{figure}
 
@@
 
+\todo{Sophia: what is features\_plot\_peak.png showing?}
+
+\timmy{Good question --- maybe Matt or Max know how the data for this was generated? (It's not part of the showyourwork yet)}
+
 \begin{figure*}

### 2023-03-06 src/scripts/plot_features_vs_iwa.py

@@ -1,0 +1,213 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import interpolate as si
+import paths
+
+CBF_COLORS = [
+    "#5790fc",  # blue
+    "#f89c20",  # orange
+    "#e42536",  # red
+    "#964a8b",  # purple
+    "#9c9ca1",  # gray
+    "#7a21dd",  # purple
+]
+import matplotlib as mpl
+mpl.rcParams['axes.prop_cycle'] = mpl.cycler(color=CBF_COLORS)
+
+def make_features_plot_multi(betas,hab_zones,features,iwa_list = [1,2,3,4],eta_earth=0.24,d=6,
+                       nbins=20,wavelength=700e-9,figure_filename="nplanets_vs_iwa.png",
+                       save_fig=False,cmap_name='Spectral'):
+    '''
+    A function to make a multi-panneled plot of four different scattering features
+    Showing the numner of accessible systems for the 
+    low end, middle and upper end of the phase angle ranges of each feature, as a function of IWA. 
+
+    betas - the maximum scattering angle parameter, here defined as beta = phase_angle_max - 90
+    iwa_list - list of IWAs you want to use, shape [n_iwas]
+    features - A dictionary of scattering features that includes the lower, middle and upper phase angle ranges (in degrees)
+    hab_zones - list of habitable zone SMAs in mas, shape [n_systems]
+    eta_earth - the assumption on the number of systems have interesting planets (basically a scale factor for the y-axis), scalar
+    wavelength - wavelength you want to consider the IWA for in m
+    d - diameter of your detelescope in m
+    cmap_name - The colormap you want to use to color the lines
+    n_bins - the number of bins to use when making hisotgram bins for the cdfs, scalar
+    savefig - whether to save the figure or not, boolean
+    filename - the filename of the figure you want to save, string
+    '''
+
+    # if ax is None:
+    #     fig, axis = plt.subplots(1,1)
+    # else: 
+    #     axis=ax
+
+    fig,axes = plt.subplots(2,2,figsize=(12,5),sharex=True,sharey=True)
+    axes = axes.flatten()
+
+    for k,feature in enumerate(features):
+        lows = []
+        meds = []
+        highs = []
+
+        for i in range(len(iwa_list)):
+        
[…]

### 2023-03-06 src/scripts/plot_n_versus_phase.py

@@ -1,0 +1,177 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import paths
+
+CBF_COLORS = [
+    "#5790fc",  # blue
+    "#f89c20",  # orange
+    "#e42536",  # red
+    "#964a8b",  # purple
+    "#9c9ca1",  # gray
+    "#7a21dd",  # purple
+]
+import matplotlib as mpl
+mpl.rcParams['axes.prop_cycle'] = mpl.cycler(color=CBF_COLORS)
+
+def make_dual_cdf_plot(beta_maxes,betas,iwa_list,hab_zones,label_locations,rotations,ax=None,eta_earth=0.24,
+                       n_bins=20,savefig=False,filename="accessible_phase_angles.png"):
+    '''
+    A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle
+    for a given inner working ange. 
+
+    beta_maxes - An array of maximum scattering angles, generated by the dynamical simulations, shape [n_simulations,n_systems,n_iwas]
+    betas - An array of beta values, generated for circular orbits, shape [n_simulations,n_systems,n_iwas]
+    iwa_list - list of IWAs you want to use, shape [n_iwas]
+    hab_zones - list of habitable zone SMAs in mas, shape [n_systems]
+    label_locations - a list of locations for each iwa, shape [n_iwas,2]
+    rotations - a list of angles defining the label rotations for each iwa, shape [n_iwas]
+    ax - a matplotlib axes on which to plot
+    eta_earth - the assumption on the number of systems have interesting planets (basically a scale factor for the y-axis), scalar
+    n_bins - the number of bins to use when making hisotgram bins for the cdfs, scalar
+    savefig - whether to save the figure or not, boolean
+    filename - the filename of the figure you want to save, string
+    '''
+
+    #If an axis isn't provided, then make one. 
+    if ax is None:
+        # fig, axis = plt.subplots(1,1,figsize=(10,4))
+        fig, axis = plt.subplots(1,1,figsize=(10,4))
+    else: 
+        axis=ax
+
+    #Cycle through each IWA to plot a line. 
+    for i,iwa in enumerate(iwa_list):
+        #### First the dynamical betas.
+        # (Although we're calling them betas, really they're the maximum scattering angles)
+        
+        # Getting data of the histogram bins to build the CDF
[…]

### 2023-03-07 src/scripts/run_simulation.py
added elliptical sim code

@@ -1,0 +1,146 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Circle
+
+from phases import *
+from samplers import *
+
+from tqdm import tqdm
+import time
+
+np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
+
+# this will make a plot with the number of orbits per panel and side of plots along one side
+plotit = 0 # set to 0 to skip plot generation
+norbs_plot = 10
+nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
+
+tyb = dict(color='black', fontsize=14)
+
+def draw_iwa(ax, iwa):
+	from matplotlib.patches import Circle
+	for r in iwa:
+		ax.add_patch(Circle((0,0),radius=r,alpha=0.2))
+
+# number of planet sims per star
+nplanets = 1000
+
+# number of epochs to sample along the orbit of the planet
+norbitsample = 1000
+
+# IWA ranges
+iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax at
+
+# read in the stars
+from astropy.io import ascii
+
+fname = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
+t = ascii.read(fname ,header_start = 1, data_start = 2)
+
+nstars = len(t)
+print(f'# read in {nstars} stars from {fname}')
+
+# set up output arrays
+betamin = np.zeros((nplanets,nstars,iwa.size))
+betamax = np.zeros((nplanets,nstars,iwa.size))
+
+if (plotit):
+	fig, axes = plt.subplots(nstars_plot, nstars_plot, figsize=(10,10))
+	ax = axes.flatten()
+	fig.suptitle(f'Inner Working Angles = {iwa} mas')
+	# remove the x and y ticks
+	for a in ax:
+		a.set_xticks([])
+		a.set_yticks([])
+		a.set_aspect('equal')
+
+	fig.subplots_adjust(wspace=0, hspace=0)
+
+cax = 0 # counter for the grid of subplots - this increases by 1 for each star AND also counts the star number for output array
+
+for l in tqdm(t): # loop over all the stars in the table
[…]

### 2023-03-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
    \caption{
-  Ellipse drawingIf you agree that this makes sense, I could try to add this to the yarn ball script by Matt?    }
+  Ellipse drawingIf you agree that this makes sense, I could try to add this to the yarn ball script by Matt?
+\todo{sophia: I have posted code in the slack to calculate beta min-max for randomly orientated eccentric orbits.
+I don't know how to add it to the code we already have.}}
     \label{fig:ellipse}
@@
 
+%\begin{figure*}
+%    \centering
+%    \includegraphics[width=0.6\textwidth]{figures/features_plot_peak.png}  
+%    \caption{
+%        Number of systems vs inner working angle
+%    }
+%    \label{fig:accessible_phase_angles}
+%\end{figure*}
+
 \begin{figure*}
     \centering
-    \includegraphics[width=0.6\textwidth]{figures/features_plot_peak.png}  
+    \includegraphics[width=0.95\textwidth]{figures/nplanets_vs_iwa.png}  
     \caption{
@@
     \label{fig:accessible_phase_angles}
+    \script{plot_features_vs_iwa.py}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
+    \caption{
+        The big kahuna
+    }
+    \label{fig:betaallofit}
+    \script{plot_n_versus_phase.py}
 \end{figure*}

### 2023-03-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=0.99\columnwidth]{figures/ellipse.pdf}
-   \caption{
-  Ellipse drawingIf you agree that this makes sense, I could try to add this to the yarn ball script by Matt?
-\todo{sophia: I have posted code in the slack to calculate beta min-max for randomly orientated eccentric orbits.
-I don't know how to add it to the code we already have.}}
-    \label{fig:ellipse}
-%    \script{plot-orb-grid.py}
-\end{figure}
-
 %The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
@@
 %
-We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of Earth-like planet \citep{ treesstam2019,trees2022}.
+We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of an Earth-like planet \citep{ treesstam2019,trees2022}.
 
@@
 %
-Therefore, in this work we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
+Therefore, in this work, we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
 %
-We note that we do not aim to accurately calculate the absolute scaling of the contrast as this depends on many factors and would have required some of the work for the star list to be redone.
+We note that we do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some of the work for the star list to be redone.
 
@@
 %
-The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G dwarf $\alpha$ Cen A, the K dwarf $\epsilon$ Eri and the M dwarf Lalande 21185.
+The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri and the M~dwarf Lalande~21185.
 

### 2023-03-08 src/scripts/hwo.py

@@ -1,0 +1,6 @@
+d = 6.0 # diameter in m
+wavelength = 600e-9 # central wavelength in m
+
+if __name__ == "__main__":
+	print(f'Telescope diameter is {d:3.1f} m')	
+	print(f'Central wavelength is {wavelength*1e9:5.1f} nm')

### 2023-03-08 src/scripts/make_beta_max_3_lambda_over_d_csv.py

@@ -1,0 +1,14 @@
+import pandas as pd
+import numpy as np
+import paths
+from hwo import *
+
+# telescope d and wlen are imported from hwo file to make consistent for all the codes
+
+# givea a nan for the first element because it tries to calculate 
+# arccos of 1.032, presumably this is alpha or proxima cen
+
+df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')
+hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
+beta_max_3ld =  np.degrees(np.arccos(3*wavelength/d*206265/(hab_zones/1000)))
+np.savetxt(paths.data / "beta_max_3_lambda_over_d.csv",beta_max_3ld,delimiter=",")

### 2023-03-08 src/scripts/plot_features_vs_iwa.py
[showyourwork] overleaf sync

@@ -3,6 +3,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 from scipy import interpolate as si
 import paths
+from hwo.py import *
 
 CBF_COLORS = [
     "#5790fc",  # blue
@@ -184,9 +185,9 @@ if __name__ == "__main__":
     #The angular separation of the Earth Equivalent Instellation
     hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
 
-    ##Some assumed parameters 
-    d = 6 #telescope diameter
-    wavelength = 600e-9
+    ##Some assumed parameters (now imported from hwo.py)
+ #   d = 6 #telescope diameter
+ #   wavelength = 600e-9
 
     ## Read in the maximum and minimum scattering angles as 
     ## calculated by the dynamical simulations

### 2023-03-08 src/scripts/plot_features_vs_iwa.py
[showyourwork] overleaf sync

@@ -3,7 +3,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 from scipy import interpolate as si
 import paths
-from hwo.py import *
+from hwo import *
 
 CBF_COLORS = [
     "#5790fc",  # blue

### 2023-03-08 src/scripts/plot_n_versus_phase.py
[showyourwork] overleaf sync

@@ -2,6 +2,7 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import paths
+from hwo import *
 
 CBF_COLORS = [
     "#5790fc",  # blue
@@ -111,9 +112,9 @@ if __name__ == "__main__":
     #The angular separation of the Earth Equivalent Instellation
     hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
 
-    ##Some Assumed parameters 
-    d = 6 #telescope diameter
-    wavelength = 600e-9
+    ##Some Assumed parameters (now imported from hwo.py)
+ #   d = 6 #telescope diameter
+ #   wavelength = 600e-9
 
     ## Read in the maximum and minimum scattering angles as 
     ## calculated by the dynamical simulations

### 2023-03-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
+%\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
 %\textbf{Why glint [Kim]}
@@
 \cite{2001Natur.412..885F} first noted the potential for time-resolved photometry and polarimetry to identify water oceans on the surface of an exoplanet.
-\textcolor{blue}{A sentence about rotational mapping here.} \cite{2006astro.ph.10518M} further quantified how photometric and polarimetric scattered light phase variations might betray the presence of oceans on terrestrial exoplanets.
+Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but specular reflection, a.k.a.
+glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}.
+\cite{2006astro.ph.10518M}  quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.
 %--> No it was talked about previous to this, per the paper cited here in fact...
@@
 They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
+
+In practice, the clouds complicate the detection of ocean glint because they obscure the underlying surface and they are themselves non-Lambertian scatterers.
 \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.

### 2023-03-10 src/scripts/run_simulation.py
made run_simulation with command line options

@@ -7,6 +7,13 @@ from samplers import *
 
 from tqdm import tqdm
 import time
+import paths
+
+import sys 
+
+if (len(sys.argv) != 2):
+	print('need one command line argument (iwa, iwa2 or iwa3). Stopping...')
+	quit()
 
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
@@ -28,17 +35,27 @@ nplanets = 1000
 # number of epochs to sample along the orbit of the planet
 norbitsample = 1000
 
-# IWA ranges
-iwa = np.array([21, 42, 63, 84]) # iwa.py
-iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax at iwa2.npy
-iwa = np.linspace(20,120,20) # iwa3.npy
+# IWA ranges as selected for three possible iwa
 
-
+op = sys.argv[1]
+if op == 'iwa':
+	iwa = np.array([21, 42, 63, 84])
+	tag = ''
+elif op == 'iwa2':
+	iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax
+	tag = '2'
+elif op == 'iwa3':
+	iwa = np.linspace(20,120,20)
+	tag = '3'
+else:
+	print('not a valid command line argument. Stopping...')
+	quit()
+print(f'option selected was {op}')
 # read in the stars
 from astropy.io import ascii
 
 fname = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
-t = ascii.read(fname ,header_start = 1, data_start = 2)
+t = ascii.read(paths.data / fname ,header_start = 1, data_start = 2)
 
 nstars = len(t)
 print(f'# read in {nstars} stars from {fname}')
@@ -136,14 +153,16 @@ for l in tqdm(t): # loop over all the stars in the table
 
 	cax+=1 # increase figure number
 
-np.save('iwa2.npy',iwa)
-np.save('betamax2.npy',betamax)
-np.save('betamin2.npy',betamin)
-
 if plotit:
 	plt.draw()
 	plt.savefig('ball_of_yarn.png',dpi=100)
 	plt.savefig('ball_of_yarn.pdf')
 	plt.show()
+else:
+	print(f'saving the data with command line option {op}')
+	np.save(paths.data / f'iwa{tag}.npy',iwa)
+	np.save(paths.data / f'betamax{tag}.npy',betamax)
+	np.save(paths.data / f'betamin{tag}.npy',betamin)
+
 
 	

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In practice, the clouds complicate the detection of ocean glint because they obscure the underlying surface and they are themselves non-Lambertian scatterers.
+In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and are themselves non-Lambertian scatterers.
 \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.
@@
 \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
-\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths.
+\citet{Zugger_2011} note that ocean glint is best observed in near infrared opacity windows because of the reduced Rayleigh scattering at long wavelengths.
 This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
-\cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
-\cite{2022A&A...664A.172T}...
+Most recently, \cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
+The bottom line is that phase-dependent reflected light photometry and polarization of a directly-imaged exoplanet could establish its habitability.
 

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be remotely detectable on an exoplanet if it is present at the surface.
+Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be detectable on an exoplanet if it is present at the surface.
 Hence the search for life on exoplanets typically focuses on habitable planets.
-To first order, the presence of surface liquid water can be predicted by its instellation---the flux it receives from its star.
-The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \cite{kasting93}.
-The HZ is popular among planet hunters because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
+To first order, the presence of surface liquid water can be predicted by the flux it receives from its star.
+The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \citep{kasting93}.
+The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
 Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
@@
 Most recently, \cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
-The bottom line is that phase-dependent reflected light photometry and polarization of a directly-imaged exoplanet could establish its habitability.
+The bottom line is that a direct imaging mission sensitive to the phase-dependent reflected light photometry and polarization of an  exoplanet could establish its habitability.
 

### 2023-03-10 src/scripts/plot_features_vs_iwa.py

@@ -193,9 +193,13 @@ if __name__ == "__main__":
     ## calculated by the dynamical simulations
     ## The files with the 3 suffix include the wider range of 
     ## iwa that we want for this step
-    phase_max=np.load(paths.data / "betamax3.npy")
-    phase_min=np.load(paths.data / "betamin3.npy")
-    iwa_list=np.load(paths.data / "iwa3.npy")
+    iwa = np.load('iwa_all3.npz')
+    phase_max = iwa['betamax']
+    phase_min = iwa['betamin']
+    iwa_list =  iwa['iwa']
+#    phase_max=np.load(paths.data / "betamax3.npy")
+#    phase_min=np.load(paths.data / "betamin3.npy")
+#    iwa_list=np.load(paths.data / "iwa3.npy")
     # Note: We pass in beta to the plotting function here, based on phase_max
     # Beta is an intermediate parameter used here. 
     # The maximum scattering angle is equal to 90+beta degrees

### 2023-03-10 src/scripts/plot_n_versus_phase.py

@@ -119,9 +119,13 @@ if __name__ == "__main__":
     ## Read in the maximum and minimum scattering angles as 
     ## calculated by the dynamical simulations
     ## This set of files only has 4 IWAs. 
-    phase_max=np.load(paths.data / "betamax.npy")
-    phase_min=np.load(paths.data / "betamin.npy")
-    iwa_list=np.load(paths.data / "iwa.npy")
+    iwa = np.load('iwa_all.npz')
+    phase_max = iwa['betamax']
+    phase_min = iwa['betamin']
+    iwa_list =  iwa['iwa']
+#    phase_max=np.load(paths.data / "betamax.npy")
+#    phase_min=np.load(paths.data / "betamin.npy")
+#    iwa_list=np.load(paths.data / "iwa.npy")
     #Note: We only pass phase_max to the plotting routine. 
     # For a given system phase_max and phase_min might differ (with elliptical orbits), 
     # but the distributions mirror each other (about 90 degrees) when you average over enough orbits

### 2023-03-10 src/scripts/run_simulation.py

@@ -160,9 +160,11 @@ if plotit:
 	plt.show()
 else:
 	print(f'saving the data with command line option {op}')
-	np.save(paths.data / f'iwa{tag}.npy',iwa)
-	np.save(paths.data / f'betamax{tag}.npy',betamax)
-	np.save(paths.data / f'betamin{tag}.npy',betamin)
+#	np.save(paths.data / f'iwa{tag}.npy',iwa)
+#	np.save(paths.data / f'betamax{tag}.npy',betamax)
+#	np.save(paths.data / f'betamin{tag}.npy',betamin)
+
+	np.savez_compressed(paths.data / f'iwa_all{tag}.npy', iwa=iwa, betamax=betamax, betamin=betamin)
 
 
 	

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 To first order, the presence of surface liquid water can be predicted by the flux it receives from its star.
-The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \citep{kasting93}.
+The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which an Earth twin could maintain surface liquid water \citep{kasting93}.
 The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.

### 2023-03-10 src/scripts/plot_features_vs_iwa.py

@@ -193,7 +193,7 @@ if __name__ == "__main__":
     ## calculated by the dynamical simulations
     ## The files with the 3 suffix include the wider range of 
     ## iwa that we want for this step
-    iwa = np.load('iwa_all3.npz')
+    iwa = np.load(paths.data / 'iwa_all3.npz')
     phase_max = iwa['betamax']
     phase_min = iwa['betamin']
     iwa_list =  iwa['iwa']

### 2023-03-10 src/scripts/plot_n_versus_phase.py

@@ -119,7 +119,7 @@ if __name__ == "__main__":
     ## Read in the maximum and minimum scattering angles as 
     ## calculated by the dynamical simulations
     ## This set of files only has 4 IWAs. 
-    iwa = np.load('iwa_all.npz')
+    iwa = np.load(paths.data / 'iwa_all.npz')
     phase_max = iwa['betamax']
     phase_min = iwa['betamin']
     iwa_list =  iwa['iwa']

### 2023-03-10 src/scripts/run_simulation.py

@@ -164,7 +164,7 @@ else:
 #	np.save(paths.data / f'betamax{tag}.npy',betamax)
 #	np.save(paths.data / f'betamin{tag}.npy',betamin)
 
-	np.savez_compressed(paths.data / f'iwa_all{tag}.npy', iwa=iwa, betamax=betamax, betamin=betamin)
+	np.savez_compressed(paths.data / f'iwa_all{tag}.npz', iwa=iwa, betamax=betamax, betamin=betamin)
 
 
 	

### 2023-03-10 src/scripts/run_simulation.py
[showyourwork] overleaf sync

@@ -9,11 +9,11 @@ from tqdm import tqdm
 import time
 import paths
 
-import sys 
+#import sys 
 
-if (len(sys.argv) != 2):
-	print('need one command line argument (iwa, iwa2 or iwa3). Stopping...')
-	quit()
+#if (len(sys.argv) != 2):
+#	print('need one command line argument (iwa, iwa2 or iwa3). Stopping...')
+#	quit()
 
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
@@ -36,8 +36,8 @@ nplanets = 1000
 norbitsample = 1000
 
 # IWA ranges as selected for three possible iwa
-
-op = sys.argv[1]
+op = snakemake.params['whichsim']
+#op = sys.argv[1]
 if op == 'iwa':
 	iwa = np.array([21, 42, 63, 84])
 	tag = ''

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The bottom line is that a direct imaging mission sensitive to the phase-dependent reflected light photometry and polarization of an  exoplanet could establish its habitability.
+\textcolor{blue}{Tighten the above two paragraphs and combine with the following.}
+
+\textcolor{blue}{Science motivation below: Kim}\\
+\textbf{Why Polarization curves worth looking at [Kim]}
+%compare to features in nonpol (strength rainbow); current capabilties?
+
+ Phase curves of planets are more complex than simple Lambertian curves.
+ %
+ The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
+ %
+ Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
+ %
+ In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
+ %
+ Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
+
+ In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
+Notably, in the polarized phase curve the rainbow may be an especially strong feature and the maximum reflectance signal (from Rayleigh scattering) occurs
+
+%Polarimetry provides the possibility to detect molecules, hazes, clouds and different surfaces  as they all produce unique polarization signatures which cannot be disentagled by using broadband flux alone.
+Polarimetry measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
+Polarimetry has previously played an important role in the understanding of solar system planet atmospheres, for instance, polarimetric measurements of Venus’ atmosphere suggested that the upper atmosphere of Venus is dominated by a haze of liquid sulphuric acid droplets \citep{Venus74}, which has been confirmed by in situ observations.
+
+\textbf{Why rainbow worth it [Kim]} %calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
+
+%The rai
 
@@
 %OWA -> should go in discussions
-
-\textcolor{blue}{Science motivation below: Kim}\\
-\textbf{Why Polarization curves worth looking at [Kim]}
-%compare to features in nonpol (strength rainbow); current capabilties?
-
- Phase curves of planets are more complex than simple Lambertian curves.
- %
- The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
- %
- Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
- %
- In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
- %
- Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
-
- In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
-Notably, in the polarized phase curve the rainbow may be an especially strong feature and the maximum reflectance signal (from Rayleigh scattering) occurs
-
-%Polarimetry provides the possibility to detect molecules, hazes, clouds and different surfaces  as they all produce unique polarization signatures which cannot be disentagled by using broadband flux alone.
-Polarimetry measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
-Polarimetry has previously played an important role in the understanding of solar system planet atmospheres, for instance, polarimetric measurements of Venus’ atmosphere suggested that the upper atmosphere of Venus is dominated by a haze of liquid sulphuric acid droplets \citep{Venus74}, which has been confirmed by in situ observations.
-
-\textbf{Why rainbow worth it [Kim]} %calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
-
-%The rai
 
@@
 
-In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), minimum ($\phi_\mathrm{min}$) and maximum ($\phi_\mathrm{max}$) scattering phase angles accessible are symmetric about 90 degrees ().
+In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) scattering phase angles accessible are symmetric about 90 degrees.
 %
@@
     \label{eq:scattering_angle}
-    \cos\left(\dfrac{\Delta \phi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a}
+    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a}
 \end{equation}
@@
         The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major $a$ at a distance $d_*$.
-The scattering phase angle, $\phi$, is the angle between the line of sight to the star and the line connecting the star and planet.
-$\delta$ is the on-sky angular separation as viewed by the observer.
-\todo{should the red angle be $\phi$?}
+The on-sky angular separation as viewed by the observer is $\delta$.
+The scattering phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.
     }

### 2023-03-10 src/scripts/run_simulation.py

@@ -9,28 +9,10 @@ from tqdm import tqdm
 import time
 import paths
 
-#import sys 
-
-#if (len(sys.argv) != 2):
-#	print('need one command line argument (iwa, iwa2 or iwa3). Stopping...')
-#	quit()
-
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
-# this will make a plot with the number of orbits per panel and side of plots along one side
-plotit = 0 # set to 0 to skip plot generation
-norbs_plot = 10
-nstars_plot = 4 # you will make a plot with nstars_plot by nstars_plot panels
-
-tyb = dict(color='black', fontsize=14)
-
-def draw_iwa(ax, iwa):
-	from matplotlib.patches import Circle
-	for r in iwa:
-		ax.add_patch(Circle((0,0),radius=r,alpha=0.2))
-
 # number of planet sims per star
-nplanets = 1000
+nplanets = 20
 
 # number of epochs to sample along the orbit of the planet
 norbitsample = 1000
@@ -64,18 +46,6 @@ print(f'# read in {nstars} stars from {fname}')
 betamin = np.zeros((nplanets,nstars,iwa.size))
 betamax = np.zeros((nplanets,nstars,iwa.size))
 
-if (plotit):
-	fig, axes = plt.subplots(nstars_plot, nstars_plot, figsize=(10,10))
-	ax = axes.flatten()
-	fig.suptitle(f'Inner Working Angles = {iwa} mas')
-	# remove the x and y ticks
-	for a in ax:
-		a.set_xticks([])
-		a.set_yticks([])
-		a.set_aspect('equal')
-
-	fig.subplots_adjust(wspace=0, hspace=0)
-
 cax = 0 # counter for the grid of subplots - this increases by 1 for each star AND also counts the star number for output array
 
 for l in tqdm(t): # loop over all the stars in the table
@@ -84,16 +54,6 @@ for l in tqdm(t): # loop over all the stars in the table
 	Vmag = l['sy_vmag']   		# V magnitude
 	mass = l['st_mass']   		# stellar mass
 	hz =  l['st_eei_angsep'] 	# Habitable zone distance (mas)
-
[…]

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-We define the angle $\Delta \phi$ such that: 
+We define the angle $\Delta \varphi$ such that: 
 \begin{equation}
     \label{eq:Delta_phi}
-    \Delta \phi 
-    = \phi_\mathrm{max} - \phi_\mathrm{min}
-    =  2(\phi_\mathrm{max} - 90) 
-    =  2(\phi_\mathrm{min} + 90)
+    \Delta \varphi 
+    = \varphi_\mathrm{max} - \varphi_\mathrm{min}
+    =  2(\varphi_\mathrm{max} - 90) 
+    =  2(\varphi_\mathrm{min} + 90)
 \end{equation}
@@
 \begin{equation}
-    \sin(\phi) = \frac{r_\mathrm{proj}}{a}
+    \sin(\varphi) = \frac{r_\mathrm{proj}}{a}
 \end{equation}
@@
 
-The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040s \todo{check}. 
+The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040's \todo{check}. 
 %

### 2023-03-10 src/scripts/run_simulation.py

@@ -12,7 +12,7 @@ import paths
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
 # number of planet sims per star
-nplanets = 20
+nplanets = 1000
 
 # number of epochs to sample along the orbit of the planet
 norbitsample = 1000

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\textbf{In this paper we...}
+%\textbf{In this paper we...}
+In this paper we quantify the accessible scattering phase angles for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars.
+In \S 2.1 we describe the target list, in \S 2.2 we derive expressions for the range of scattering phase angles in a few different limits, in \S 2.3 we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4 we outline how the planet/star contrast changes as a function of orbital phase.
+We present and discuss our results in \S 3 and conclude in \S 4.
+%the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles
 % We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.
@@
 
-\section{Observing scattering phenomena}
+%\section{Observing scattering phenomena}
+\section{Methods}
 
@@
 This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}. 
+The authors of the report determined that these are the nearby stars (maximum distance \SI{25}{\parsec}) for which exo-Earths would be the most observable for an imaging survey of habitable zones with a \SI{6}{\meter} telescope.
 
-The authors of the report determined that these are the nearby stars (maximum distance \SI{25}{\parsec}) for which exo-Earths would be the most observable for an imaging survey of habitable zones with a \SI{6}{\meter} telescope.
+In what follows, we consider hypothetical Earth-size planets orbiting at the right distance from each star to receive the same flux as the Earth receives from the Sun.
+By construction, such planets should be observable at quadrature phase by HabWorlds.
 
@@
 %
-The scattering phase angle $\phi$ is this set up follows the equation,
+The scattering phase angle $\varphi$ is this set up follows the equation,
 \begin{equation}
@@
 
-\Cref{fig:scatterplot} shows $\Delta \phi$ for the systems in the target list assuming a $3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$. 
+\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the HabWorlds target list assuming an IWA of 62 mas ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
 %
-This assumes each planet has a semi-major axis such that it receives an Earth-equivalent instellation, which is provided in the input star list.
+%This assumes each planet has a semi-major axis such that it receives an Earth-equivalent instellation, which is provided in the input star list.
 %
@@
 
-(1)~If the inclination is face-on enough such that the coronagraph mask does not cover any part of the projected orbit on-sky, then the maximum scattering phase angle coverage depends solely on the disk inclination ;
+(1)~If the inclination is relatively face-on, then the coronagraph mask does not obscure any part of the sky-projected orbit, and the maximum scattering phase angle coverage depends solely on the disk inclination ;
 
-(2)~When parts of the disk (minimum and maximum scattering phase angles) are hidden by the focal plane mask, the scattering phase angle coverage does not depend on the inclination and only on the \IWA\ and orbit semi-major axis.
+(2)~When parts of the orbit  are hidden by the focal plane mask, the scattering phase angle coverage only depends on the \IWA\ and orbital semi-major axis.
 
-\begin{figure}
+\begin{figure}[htb]
    \centering
@@
 
-Using \cref{eq:scattering_angle}, we deduce the 2 regime equation for inclined circular orbits: 
+Using \cref{eq:scattering_angle}, we deduce the equation for inclined circular orbits: 
 \begin{equation}
 \label{eq:Delta_phi_max}
-    \Delta \phi = 
+    \Delta \varphi = 
     \begin{cases}
@@
 
-\subsubsection{Eccentric orbits}
+%\subsubsection{Eccentric orbits}
+
+%The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Monte Carlo Simulations}
+
+\subsubsection{Circular randomly inclined orbits}
+\label{sec:circular}
+
+For each star on the \hwo\ target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
+%
+The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$. 
+%
+We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet. 
+The solid lines in Figure \ref{fig:betaallofit} show the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo\ target list, depending on the choice of inner working angle.
+
+\todo{what does this plot show, how many systems could we see?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
+    \caption{
+        The big kahuna
+    }
+    \label{fig:betaallofit}
+    \script{plot_n_versus_phase.py}
+\end{figure*}
+
+\subsubsection{Eccentric randomly inclined orbits}
+\label{sec:eccentric}
+We repeat the Monte Carlo simulation from \cref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
+%
+We randomly drew 1000 random orbits for each star and determined the sky-projected orbits following, e.g., \cite{2010exop.book...15M}.
+Eccentricities were drawn from the beta distribution and inclinations were uniform in $\cos i$.
+%
+We then filtered on \IWA's of 1, 2, 3 and 4 times $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+%
+\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's.
+The dashed lines in Figure \ref{fig:betaallofit} show the cumulative distribution for the most extreme scattering phase angles accessible, when accounting for orbital eccentricity.
+Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph IWA and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
 
@@
 
-%The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
 
@@
 
-\subsection{Circular randomly inclined orbits}
-\label{sec:circular}
-
-For each star on the \hwo\ target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
-%
-The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$. 
-%
-We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
-%
-\todo{Figure X, where the plot for this?} shows the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo\ target list, depending on the choice of inner working angle.
-
-\todo{what does this plot show, how many systems could we see?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Eccentric randomly inclined orbits}
-\label{sec:eccentric}
-We repeat the Monte Carlo simulation from \cref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ (cf.
-\citet{Guimond_2019} and references therein).
-%
-We randomly drew 1000 eccentricities from the beta distribution for each star and determined the orbits. 
-%
-We then filtered on \IWA\ for multiples of 1, 2, 3 and 4 $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
-%
-\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's. 
-%
-%Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
-
 \todo{Is there a plot for this? How many systems could we see? Why is it different to non-eccentric cases?}
@@
     \script{plot_features_vs_iwa.py}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
-    \caption{
-        The big kahuna
-    }
-    \label{fig:betaallofit}
-    \script{plot_n_versus_phase.py}
 \end{figure*}

### 2023-03-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be detectable on an exoplanet if it is present at the surface.
-Hence the search for life on exoplanets typically focuses on habitable planets.
-To first order, the presence of surface liquid water can be predicted by the flux it receives from its star.
-The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which an Earth twin could maintain surface liquid water \citep{kasting93}.
+Surface liquid water is closely related to planetary habitability: water is required for life as we know it, and life on exoplanets will most likely on be detectable if it is present at the surface.
+
+To first order, the presence of surface liquid water can be predicted by the flux a planet receives from its star.
+The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which an Earth-like planet could maintain surface liquid water \citep{kasting93}.
 The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
-Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
+%Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
 
 %\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
+
 %\textbf{Why glint [Kim]}
 % probe of surface conditions; habitability
-\cite{2001Natur.412..885F} first noted the potential for time-resolved photometry and polarimetry to identify water oceans on the surface of an exoplanet.
+
+\cite{2001Natur.412..885F} noted the potential for time-resolved photometry to identify water oceans on the surface of an exoplanet, which was further explored by \citet{williams2004} and \citet{giados2006}.
 Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but specular reflection, a.k.a.
 glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}.
-\cite{2006astro.ph.10518M}  quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.
-%--> No it was talked about previous to this, per the paper cited here in fact...
-"Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint.
-Schneider (2006) had similar ideas." -KB
+\citet{lustig2019} explored mapping exoplanets from their ocean glint in polarized light.
+
+In polarized light the glint appears at similar phase angles but its peak can be extended further particularly in the presence of atmospheric winds \citep{kopparla2018}.
+%\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
+\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean.
 \cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
@@
 
+%"Stam & Hovenier (2006) independently examined the observability of the polarized signatures of an Earth-like extrasolar planet, including Rayleigh scattering and sea-surface glint.
+Schneider (2006) had similar ideas." -KB
+
+%% MARKER FOR WHERE KIM LEFT OFF PLEASE LEAVE %%%
+
+% Move to Clouds?
 In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and are themselves non-Lambertian scatterers.
@@
 The bottom line is that a direct imaging mission sensitive to the phase-dependent reflected light photometry and polarization of an  exoplanet could establish its habitability.
-\textcolor{blue}{Tighten the above two paragraphs and combine with the following.}
 
@@
 %Polarimetry provides the possibility to detect molecules, hazes, clouds and different surfaces  as they all produce unique polarization signatures which cannot be disentagled by using broadband flux alone.
-Polarimetry measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
+Polarized measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
 Polarimetry has previously played an important role in the understanding of solar system planet atmospheres, for instance, polarimetric measurements of Venus’ atmosphere suggested that the upper atmosphere of Venus is dominated by a haze of liquid sulphuric acid droplets \citep{Venus74}, which has been confirmed by in situ observations.
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[t]
     \centering
@@
         The big kahuna
+        \timmy{What are the dashed vs. solid lines?}
     }

### 2023-03-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\subsection{Ocean Glint}
 \cite{2001Natur.412..885F} noted the potential for time-resolved photometry to identify water oceans on the surface of an exoplanet, which was further explored by \citet{williams2004} and \citet{giados2006}.
@@
 
+% NEEDS Paragraph on mapping connection to tectonics
+
+% NEEDS KOPPARLA (prob refer to , BERDYUGINA, ROSSI?
+(cloud?), HU?, added...
+blend cloud hinderance into argument for context to lead into rainbows, mention Lietzow and jeremy there., maybe some more mention of the Earthshine observations Karalidi, Gordon, etc
+
 %"Stam & Hovenier (2006) independently examined the observability of the polarized signatures of an Earth-like extrasolar planet, including Rayleigh scattering and sea-surface glint.
@@
 
-%% MARKER FOR WHERE KIM LEFT OFF PLEASE LEAVE %%%
+% Move to Clouds?
+In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint.
+\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
+\cite{Zugger_2010} simulated photometric and polarized phase variations for Earth-like planets including the effects of partial cloud coverage in suppressing the Rayleigh scattering and ocean glint polarization.
+Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
+To suppress the effects of Rayleigh scattering, glint observations can be made at longer wavelengths
+\citep{Zugger_2011}.
 
-% Move to Clouds?
-In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and are themselves non-Lambertian scatterers.
-\cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves.
-Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
-\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
-\citet{Zugger_2011} note that ocean glint is best observed in near infrared opacity windows because of the reduced Rayleigh scattering at long wavelengths.
-This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
-Most recently, \cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
-The bottom line is that a direct imaging mission sensitive to the phase-dependent reflected light photometry and polarization of an  exoplanet could establish its habitability.
+%%% KIM's BOOKMARK FOR HERLSELF, PLEASE DO NOT DELETE %%%
 
-\textcolor{blue}{Science motivation below: Kim}\\
-\textbf{Why Polarization curves worth looking at [Kim]}
-%compare to features in nonpol (strength rainbow); current capabilties?
+\subsection{Polarized Phase Curves}
+%compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,
+
+% add latitude albedo degeneracy break here \citep{2012ApJ...752L...3C}
+
+%This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+
+ %\cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
 
@@
 
-\textbf{Why rainbow worth it [Kim]} %calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
+\subsection{Rainbows}
+%calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
 
@@
         \timmy{What are the dashed vs. solid lines?}
+        \timmy{How did this plot aggregate the simulations for each planet? mean? median? min/max?
+        }
     }
@@
 We randomly drew 1000 random orbits for each star and determined the sky-projected orbits following, e.g., \cite{2010exop.book...15M}.
-Eccentricities were drawn from the beta distribution and inclinations were uniform in $\cos i$.
+Eccentricities were drawn from the beta distribution, and inclinations were uniform in $\cos i$.
 %
@@
 \todo{how was the angle coverage calculated in the monte carlo?}
-
-\timmy{I think so far, the MC code doesn't compute angle coverage. I came up with this quick and dirty sketch to help us get started:}
 
@@
 
-\timmy{Good question --- maybe Matt or Max know how the data for this was generated? (It's not part of the showyourwork yet)}
+\timmy{Good question --- maybe Matt or Max know how the data for this was generated?
+Matt: yes, it's all in showyourwork right now.
+I ran 1000 sims for each target with a random eccentric orbit, then pulled out the betamax and betamin angles for each of the 1000 realisations.
+Max's plots encode the completeness as a function of phase angle, showing that you get about 10 percent more phenomena.
+}
 

### 2023-03-14 src/scripts/plot_contrasts.py

@@ -51,7 +51,7 @@ F_670 = interpolate.interp1d(data_trees[0],data_trees[1])
 
 orbital_phase = np.linspace(0,2*np.pi,200)
 
-inclination = np.radians(63)
+inclination = np.radians(90)
 
 alpha = np.arccos(np.cos(orbital_phase)*np.sin(inclination))
 

### 2023-03-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of $60$ and $90$ degrees.
-\todo{Only the 60 degree plot is here at the moment}
+In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given an inclination of $90$ degrees.
 %
@@
    \centering
-   \includegraphics[width=0.99\textwidth]{figures/Contrast_vs_separation_inclination_angle_63.pdf}
+   \includegraphics[width=0.99\textwidth]{figures/Contrast_vs_separation_inclination_angle_90.pdf}
    \caption{
-    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of $60^\circ$ for the stars $\alpha$ Cen A, $\epsilon$ Eri and Lalande 21185.
+    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of $90^\circ$ for the stars $\alpha$ Cen A, $\epsilon$ Eri and Lalande 21185.
 The solid line indicates the contrast in unpolarised light with the contrast at quadrature marked by a solid dot.

### 2023-03-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Sophia R. Vaughan$^{1,}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
-    Kimberly Bott$^{2,9}$,
+    Kimberly Bott$^{2,28}$,
     Sarah L. Casewell$^{3}$,
@@
     Johan Mazoyer$^{7}$,
-    Maxwell A. Millar-Blanchaer$^{8}$
-    \newauthor [remaining workshop participants in alphabetical order] \newauthor 
+    Maxwell A. Millar-Blanchaer$^{8}$,\newauthor 
+    Olivier Absil$^{16}$,
+    Lisa Altinier$^{13}$,
+    Pierre Baudoz$^{7}$,
+    Ruslan Belikov$^{24}$,
+    Alexis Bidot$^{9}$,\newauthor 
+    Markus Johannes Bonse$^{6}$,
+    Bernhard Brandl$^{5}$,
+    Alexis Carlotti$^{9}$,
+    Elodie Choquet$^{13}$,
+    Niyati Desai$^{17}$,\newauthor
+    Kevin Fogarty$^{24}$,
+    Jules Fowler$^{12}$
+    Yann Gutierrez$^{7,14,22}$,
+    Olivier Guyon$^{15,25,26,27}$,
+    Sebastiaan Y. Haffert,$^{15}$, \newauthor
+    Olivier Herscovici-Schiller$^{14}$,
+    Adrien Hours$^{9}$,
+    Roser Juanola-Parramon $^{20,21}$,
+    Elina Kleisioti $^{5}$, \newauthor
+    Lorenzo König$^{16}$,
+    Mariya Krasteva$^{23}$,
+    Iva Laginja$^{7}$,
+    Rico Landman$^{5}$,
+    Lucie Leboulleux$^{9}$,
+    David Mouillet$^{9}$,\newauthor
+    Mamadou N’Diaye$^{10}$,
+    Emiel H. Por$^{18}$,
+    Laurent Pueyo$^{18}$,
+    Frans Snik$^{5}$,
+    Dirk van Dam$^{5}$,
+    Kyle van Gorkom, $^{15}$,\newauthor
+    Maaike van Kooten$^{11}$
+    \newauthor 
     \\
@@
     $^{6}$ Max Planck Institute for Intelligent Systems, Max Planck Ring 4, 72076 Tübingen, Germany \& ETH Zurich, Switzerland \\
-    $^{7}$ LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-92195 Meudon, France \\
+    $^{7}$ LESIA, Observatoire de Paris, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e de Paris, F-92195 Meudon, France \\
     $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA \\
-    $^{9}$ NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA
-}
+$^{9}$ Univ. Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\
+$^{10}$ Universit\'e C\^ote d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange. France \\
+$^{11}$ National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C. Canada \\
+$^{12}$ University of California, Santa Cruz  \\
+$^{13}$ Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\
+$^{14}$ DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\
+$^{15}$ Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, Arizona \\
+$^{16}$ STAR Institute, Universit\'e de Li\`ege, All\'ee du Six Ao\^ut 19c, 4000 Li\`ege, Belgium \\
+$^{17}$ Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\
+$^{18}$ Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\
+$^{19}$ SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\
+$^{20}$ NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771,United States \\
+$^{21}$ University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, United States \\
+$^{22}$ DOTA, ONERA, 92322 Ch\^atillon, France \\
+$^{23}$ European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\
+$^{24}$ NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\
+$^{25}$ Subaru Telescope, NAOJ, United States \\
+$^{26}$ College of Optical Sciences, University of Arizona, Tucson, AZ 87521, United States \\
+$^{27}$ Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\
+$^{28}$ NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA}
 \date{Draft version: \today}
@@
 Surface liquid water is closely related to planetary habitability: water is required for life as we know it, and life on exoplanets will most likely on be detectable if it is present at the surface. 
-
 To first order, the presence of surface liquid water can be predicted by the flux a planet receives from its star.
@@
 The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
+We would therefore like to establish the presence of liquid water on the surface of an exoplanet before trying to interpret its atmospheric biosignatures.
 %Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
@@
 \subsection{Ocean Glint}
-\cite{2001Natur.412..885F} noted the potential for time-resolved photometry to identify water oceans on the surface of an exoplanet, which was further explored by \citet{williams2004} and \citet{giados2006}.
-Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but specular reflection, a.k.a.
+\cite{2001Natur.412..885F} noted the potential for time-resolved reflected light photometry to identify water oceans on the surface of an exoplanet.
+\cite{2008Icar..195..927W} performed simulations of scattered light from directly imaged terrestrial exoplanets.
+%to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
+They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
+%Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but
+Specular reflection, a.k.a.
 glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}.
-\citet{lustig2019} explored mapping exoplanets from their ocean glint in polarized light.
+%\citet{lustig2019} explored mapping exoplanets at crescent phases using ocean glint.
 
@@
 \citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean.
-\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
+%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
 They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.

### 2023-03-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
-They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
+They showed that oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
 %Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but
 Specular reflection, a.k.a.
-glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}.
+glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., Titan's methane lakes;][]{2010GeoRL..37.7104S}.
 %\citet{lustig2019} explored mapping exoplanets at crescent phases using ocean glint.
-
-In polarized light the glint appears at similar phase angles but its peak can be extended further particularly in the presence of atmospheric winds \citep{kopparla2018}.
+Specular reflection appears at similar phase angles in both unpolarized and polarized light and its peak orbital phase can be extended by atmospheric winds, which produce waves and hence increase the size of the glint spot \citep{2011Icar..211..722B,kopparla2018}.
 %\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
-\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean.
+\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean, proving that the method works in practice for an unresolved world.
 %\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
@@
 \subsection{Polarized Phase Curves}
+\textcolor{blue}{Recap all of polarized phase curve work, including stuff not about ocean glint. Introduce the polarization signals of Rayleigh scattering, rainbow and glory.}
 %compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,
@@
 
-\subsection{Rainbows}
+%\subsection{Rainbows}
 %calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
@@
 
+\subsection{Habitable Worlds Observatory}
 The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
@@
 %\textbf{In this paper we...}
+\subsection{Outline}
 In this paper we quantify the accessible scattering phase angles for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars.
@@
 %We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
+%From the HabWorlds report.
+Inner Working Angle: The IWA defines the region near the star that cannot be accessed for direct imaging due to a coronagraphic mask, starshade obscuration, or an interferometric null.
+It depends on the architecture of the telescope, the starlight suppression system used, and the observation wavelength.
+While a nominal value could be defined by making assumptions about the design for Habitable Worlds Observatory, we choose instead to derive the IWA from the star list itself and Astro2020’s requirement that ~100 cumulative habitable zones be surveyed by direct imaging.
+As shown in the main table, an IWA near ~70 mas will be needed to access ~100 cumulative habitable zones, and would need to be achieved at all wavelengths of interest for spectral characterization.
+For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
+
 For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.

### 2023-03-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Surface liquid water is closely related to planetary habitability: water is required for life as we know it, and life on exoplanets will most likely on be detectable if it is present at the surface. 
+Surface liquid water is closely related to planetary habitability: water is required for life as we know it, and life on exoplanets will most likely on be detectable if water is present at the surface.
 To first order, the presence of surface liquid water can be predicted by the flux a planet receives from its star.
@@
 \citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean, proving that the method works in practice for an unresolved world.
+
+\noindent {\color{Green}<Modifications by Victor Trees>}
+
 %\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
@@
 % Move to Clouds?
-In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint.
+In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint {\color{Green} in the total (polarized plus unpolarized) reflected flux}.
 \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
-\cite{Zugger_2010} simulated photometric and polarized phase variations for Earth-like planets including the effects of partial cloud coverage in suppressing the Rayleigh scattering and ocean glint polarization.
-Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
-To suppress the effects of Rayleigh scattering, glint observations can be made at longer wavelengths
-\citep{Zugger_2011}.
+
+\cite{Zugger_2010} simulated photometric and polarized phase variations {\color{Green} in the visible wavelength range} for Earth-like planets including the effects of %partial cloud coverage
+{\color{Green} Rayleigh scattering and a horizontally homogeneous cloud cover.
+They concluded that detection of an ocean glint signature in polarization is severely limited by the presence of clouds.} %Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they %again
+%find that oceans are detectable and distinguishable from other sources of polarization. 
+{\color{Green} \citet{treesstam2019} simulated the total and polarized fluxes and the degree of polarization of light reflected by ocean planets with a horizontally inhomogeneous (i.e.
+patchy) cloud cover.
+They found that an alternation of a cloud-free and cloud-covered glint, as the planet rotates and its cloud cover changes in time, can result in strong fluctuations in the degree of polarization phase curves.}
+
+{\color{Green}In the near-infrared, the Rayleigh scattering optical thickness of the exoplanet's atmosphere is smaller than in the visible wavelength range, which improves the visibility of the glint \citep{Zugger_2011}.
+Combining degree of polarization phase curves of ocean planets at short wavelengths and long wavelengths results in an intersection of those phase curves, as shown by \citet{treesstam2019}.
+This color change in polarized light, when the planet orbits its parent star, is not expected for cloudy and cloud-free dry planets, and can therefore help to distinguish between an exo-ocean and forward scattering clouds.}
+
+%Moreover, Trees & Stam (2019) combined the phase curves at various wavelengths between 350 and 865 nm, to show that the glint increasingly reddens the degree of polariza- tion and polarized flux with increasing phase angle.
+This color change in polarized light does not happen for dry planets and can thus identify an exo-ocean, also in the mean signal of heav- ily cloudy ocean planets with random patchy cloud patterns, and when surface pressures are higher than on Earth.
+
+%The spectra of Stam (2008) and the phase curves of Zugger et al.
+(2011, 2010) are for horizontally homogeneous planets: the local atmosphere-surface systems do not vary in the latitu- dinal and longitudinal directions.
+The spectra of partly cloudy planets were mimicked by weighted sums of completely cloudy and completely cloud-free planets, which results in fading of the ocean signatures.
+Zugger et al.
+(2010, 2011) therefore concluded that the detectability of an exo-ocean is strongly limited by the presence of clouds.
+
+%Trees & Stam (2019), however, simulated the total and polarized fluxes and the degree of polarization of light reflected by planets with rough ocean surfaces and with horizon- tally inhomogeneous clouds, and found that the glint appearing and hiding behind patchy clouds results in a variability of the phase curves.
+
+%Moreover, Trees & Stam (2019) combined the phase curves at various wavelengths between 350 and 865 nm, to show that the glint increasingly reddens the degree of polariza- tion and polarized flux with increasing phase angle.
+This color change in polarized light does not happen for dry planets and can thus identify an exo-ocean, also in the mean signal of heav- ily cloudy ocean planets with random patchy cloud patterns, and when surface pressures are higher than on Earth.
 
@@
 
+{\color{Green} Victor Trees: Do you mean:
+%
+\begin{equation}
+    \Delta \varphi 
+    = \varphi_\mathrm{max} - \varphi_\mathrm{min}
+    =  2(\varphi_\mathrm{max} - 90^\circ) 
+    =  2(90^\circ-\varphi_\mathrm{min} ) \notag
+\end{equation}
+%
+}
+
 The on-sky projected planet distance, $r_\mathrm{proj}$ (in \si{\au}) is shown in \cref{fig:scattering-angle} assuming a non-eccentric circular orbit with a semi-major $a$ (in \si{\au}) at a distance $d_*$ (in \si{\parsec}) from the observer.
 %
-The scattering phase angle $\varphi$ is this set up follows the equation,
+The scattering phase angle $\varphi$ in this set up follows the equation,
 \begin{equation}
@@
 Using \cref{eq:Delta_phi}, we deduce:
-
+%
 \begin{equation}
@@
 \end{equation}
-
+ %
 \begin{figure}
@@
 %
-While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.
+While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.https://www.overleaf.com/project/63f87efaa38311f6a4b2a551
 %

### 2023-03-29 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-03-16 src/tex/ms.tex
@@
 
-\subsection{Ocean Glint}
+\subsection{Ocean Glint, Cloud Glory and Rainbows}
 \cite{2001Natur.412..885F} noted the potential for time-resolved reflected light photometry to identify water oceans on the surface of an exoplanet.
@@
 \citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean, proving that the method works in practice for an unresolved world.
+
+%\noindent {\color{Green}<Modifications by Victor Trees>}
+
 %\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
@@
 % Move to Clouds?
-In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint.
+In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint in the total  reflected flux (polarized plus unpolarized).
 \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
-\cite{Zugger_2010} simulated photometric and polarized phase variations for Earth-like planets including the effects of partial cloud coverage in suppressing the Rayleigh scattering and ocean glint polarization.
-Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization.
-To suppress the effects of Rayleigh scattering, glint observations can be made at longer wavelengths
-\citep{Zugger_2011}.
+
+\cite{Zugger_2010} simulated photometric and polarized phase variations at visible wavelengths for Earth-like planets including the effects of %partial cloud coverage
+Rayleigh scattering and a horizontally homogeneous cloud cover.
+They concluded that detection of an ocean glint signature in polarization is severely limited by the presence of clouds.
+%Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they %again
+%find that oceans are detectable and distinguishable from other sources of polarization. 
+%{\color{Green} 
+\citet{treesstam2019} simulated the total and polarized fluxes and the degree of polarization of light reflected by ocean planets with patchy clouds.
+They found that the alternation of cloud-free and cloud-covered glint, as the planet rotates and its cloud cover changes in time, can result in strong fluctuations in the degree of polarization, hence resulting in noisier phase curves.
+Nonetheless, polarization phase curves of ocean planets at optical and near infrared wavelengths intersect \citep{treesstam2019}, providing a possible avenue to distinguish between an exo-ocean and forward scattering clouds.%}%}
+
+%{\color{Green}
+The Rayleigh scattering optical thickness of a planet's atmosphere is smaller in the near-infrared (NIR) than in the visible, which improves the visibility of the glint \citep{Zugger_2011}.
+NIR phase measurements also minimize the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+
+Phase variations of polarized light could also reveal other non-Lambertian cloud scattering phenomena.
+The degree of polarization of light scattered by spherical cloud droplets strongly depends on the scattering angle \citep[see][]{hansentravis1974}, and consequently, on the phase angle for an Earth-like exoplanet with liquid water clouds \citep[see][]{stam2008}.
+The sign of the degree of polarization (i.e.
+the direction of polarization) changes sign, and show peaks at the primary and secondary rainbow angles  \citep{bailey2007,2018MNRAS.480.1613B}.
+A measurement of those rainbows would be a strong indicator of spherically shaped particles in the exoplanet's atmosphere.
+The precise phase angle locations in the the degree of polarization of the rainbow peaks depend on wavelength \citep[see Fig.
+5 of][]{trees2022}, the particle size ...
+and on the type of liquid \citep{bailey2007}.
+A famous example...
+Venus ...
+%A famous example of using polarimetry to characetrize clouds within our Solar System is the characterization of the main cloud deck of Venus.
+\citep{hansenhovenier1974} derived that the main cloud deck of Venus consists of a 75\% sulfuric acid solution.
+%
+... glory ...\citep{2014A&A...566L...1G} ... and halos ... ice clouds ...
+
+ \textbf{Victor/Kim: please complete this paragraph with a few sentences describing what one learns from rainbow and glory, ideally backed with good references!}
 
@@
 
-\subsection{Polarized Phase Curves}
-\textcolor{blue}{Recap all of polarized phase curve work, including stuff not about ocean glint. Introduce the polarization signals of Rayleigh scattering, rainbow and glory.}
+%\subsection{Polarized Phase Curves}
+%\textcolor{blue}{Recap all of polarized phase curve work, including stuff not about ocean glint. Introduce the polarization signals of Rayleigh scattering, rainbow and glory.}
 %compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,
@@
 
- Phase curves of planets are more complex than simple Lambertian curves.
+ %Phase curves of planets are more complex than simple Lambertian curves.
  %
- The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
+ %The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
  %
- Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
+ %Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
  %
- In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
+ %In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
  %
- Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
+ %Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
 
- In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
+ %In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
 Notably, in the polarized phase curve the rainbow may be an especially strong feature and the maximum reflectance signal (from Rayleigh scattering) occurs
@@
 \begin{equation}
-    \label{eq:Delta_phi}
     \Delta \varphi 
     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
-    =  2(\varphi_\mathrm{max} - 90) 
-    =  2(\varphi_\mathrm{min} + 90)
+    =  2(\varphi_\mathrm{max} - 90^\circ) 
+    =  2(90^\circ-\varphi_\mathrm{min} ) \notag
 \end{equation}
+% \begin{equation}
+%     \label{eq:Delta_phi}
+%     \Delta \varphi 
+%     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
+%     =  2(\varphi_\mathrm{max} - 90) 
+%     =  2(\varphi_\mathrm{min} + 90)
+% \end{equation}
+
+% {\color{Green} Victor Trees: Do you mean:
+%INDEED, THERE MUST HAVE BEEN A TYPO, THX!
+% %
+% \begin{equation}
+%     \Delta \varphi 
+%     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
+%     =  2(\varphi_\mathrm{max} - 90^\circ) 
+%     =  2(90^\circ-\varphi_\mathrm{min} ) \notag
+% \end{equation}
+%
+% }
 
@@
 %
-The scattering phase angle $\varphi$ is this set up follows the equation,
+The scattering phase angle $\varphi$ in this set up follows the equation,
 \begin{equation}
@@
     \begin{cases}
-        2 \cdot i & \textrm{for} \cos(i) > \dfrac{\mathrm{IWA}\cdot d_* }{a}
+        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA}\; d_* }{a}
   \\ 
-        2 \cdot \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right)  & \textrm{for} \cos(i) < \dfrac{\mathrm{IWA}\cdot d_* }{a}
+        2  \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right)  & \textrm{for} \cos(i) < \frac{\mathrm{IWA}\; d_* }{a}
     \end{cases}\,.

### 2023-04-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+% Patch \hypertarget; for details, see 
+% https://tex.stackexchange.com/a/17138/97669
+\makeatletter
+\newcommand{\affiliationtarget}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\newcommand{\affiliationlink}[1]{\hyperlink{#1}{#1}}
+
 % Chasing rainbows with the Habitable Worlds Observatory
@@
 \author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan$^{1,}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
-    Kimberly Bott$^{2,28}$,
-    Sarah L. Casewell$^{3}$,
-    Nicolas B. Cowan$^{4}$,
-    David S. Doelman$^{5}$,\newauthor 
-    Timothy D. Gebhard$^{6}$,
-    Matthew Kenworthy$^{5}$,
-    Johan Mazoyer$^{7}$,
-    Maxwell A. Millar-Blanchaer$^{8}$,\newauthor 
+    Sophia R. Vaughan$^{\affiliationlink{1},}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
+    Kimberly Bott$^{\affiliationlink{2},\affiliationlink{28}}$,
+    Sarah L. Casewell$^{\affiliationlink{3}}$,
+    Nicolas B. Cowan$^{\affiliationlink{4}}$,
+    David S. Doelman$^{\affiliationlink{5,19}}$,
+    \newauthor 
+    Timothy D. Gebhard$^{\affiliationlink{6},\affiliationlink{29}}$,
+    Matthew Kenworthy$^{\affiliationlink{5}}$,
+    Johan Mazoyer$^{\affiliationlink{7}}$,
+    Maxwell A. Millar-Blanchaer$^{\affiliationlink{8}}$,
+    \newauthor 
     Olivier Absil$^{16}$,
@@
     Ruslan Belikov$^{24}$,
-    Alexis Bidot$^{9}$,\newauthor 
-    Markus Johannes Bonse$^{6}$,
+    Alexis Bidot$^{9}$,
+    Markus J. Bonse$^{29}$,
+    \newauthor 
     Bernhard Brandl$^{5}$,
@@
     Elodie Choquet$^{13}$,
-    Niyati Desai$^{17}$,\newauthor
+    Niyati Desai$^{17}$,
     Kevin Fogarty$^{24}$,
-    Jules Fowler$^{12}$
+    Jules Fowler$^{12}$,
+    \newauthor
     Yann Gutierrez$^{7,14,22}$,
     Olivier Guyon$^{15,25,26,27}$,
-    Sebastiaan Y. Haffert,$^{15}$, \newauthor
+    Sebastiaan Y. Haffert$^{15}$,
     Olivier Herscovici-Schiller$^{14}$, 
+    \newauthor
     Adrien Hours$^{9}$,
-    Roser Juanola-Parramon $^{20,21}$,
-    Elina Kleisioti $^{5}$, \newauthor
+    Roser Juanola-Parramon$^{20,21}$,
+    Elina Kleisioti $^{5}$,
     Lorenzo König$^{16}$,
     Mariya Krasteva$^{23}$, 
+    \newauthor
     Iva Laginja$^{7}$,
@@
     Lucie Leboulleux$^{9}$,
-    David Mouillet$^{9}$,\newauthor
+    David Mouillet$^{9}$,
     Mamadou N’Diaye$^{10}$,
     Emiel H. Por$^{18}$,
+    \newauthor
     Laurent Pueyo$^{18}$,
@@
     Dirk van Dam$^{5}$,
-    Kyle van Gorkom, $^{15}$,\newauthor
+    Kyle van Gorkom$^{15}$,
     Maaike van Kooten$^{11}$ 
-    \newauthor 
-    \\
-    $^{1}$ Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK\\
-    $^{2}$ Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\
-    $^{3}$ Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK\\
-    $^{4}$ Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada\\
-    $^{5}$ Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
-    $^{6}$ Max Planck Institute for Intelligent Systems, Max Planck Ring 4, 72076 Tübingen, Germany \& ETH Zurich, Switzerland \\
-    $^{7}$ LESIA, Observatoire de Paris, Universit\'e PSL, CNRS, Sorbonne Universit\'e, Universit\'e de Paris, F-92195 Meudon, France \\
-    $^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA \\
-$^{9}$ Univ. Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\
-$^{10}$ Universit\'e C\^ote d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange. France \\
-$^{11}$ National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C. Canada \\
-$^{12}$ University of California, Santa Cruz  \\
-$^{13}$ Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\
-$^{14}$ DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\
-$^{15}$ Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, Arizona \\
-$^{16}$ STAR Institute, Universit\'e de Li\`ege, All\'ee du Six Ao\^ut 19c, 4000 Li\`ege, Belgium \\
-$^{17}$ Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\
-$^{18}$ Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\
-$^{19}$ SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\
-$^{20}$ NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771,United States \\
-$^{21}$ University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, United States \\
-$^{22}$ DOTA, ONERA, 92322 Ch\^atillon, France \\
-$^{23}$ European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\
-$^{24}$ NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\
-$^{25}$ Subaru Telescope, NAOJ, United States \\
-$^{26}$ College of Optical Sciences, University of Arizona, Tucson, AZ 87521, United States \\
-$^{27}$ Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\
-$^{28}$ NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA}
+    \newauthor \\%
+    Affiliations are listed at the end of the paper
+    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
+    % https://academic.oup.com/mnras/article/498/2/2354/5900562
+    % I've also added some LaTeX magic to create links for the affiliations; if y'all agree that it's a good idea I can extend it to all authors?
+}
 \date{Draft version: \today}
@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory, a coronagraphic mission which would detect rocky planets in the habitable zone, establish their habitability, and search them for biosignatures.
+NASA recently announced the Habitable Worlds Observatory (HWO), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
@@
 Surface liquid water is central to the definition of planetary habitability.
-%
 Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
@@
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
-The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$20 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$55 systems, assuming a 60 mas inner working angle (\IWA).
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.
-
-\end{abstract}
+The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$20 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$55 systems, assuming a 60\,mas inner working angle (\IWA).
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.     \end{abstract}
 
 \begin{keywords}
-planets and satellites: terrestrial planets -- instrumentation: high angular resolution -- planets and satellites: atmospheres
+planets and satellites: terrestrial planets -- 
+instrumentation: high angular resolution -- 
+planets and satellites: atmospheres
 \end{keywords}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%--------------------------------------------------------------------------------
 \begin{figure*}%[t]
@@
    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-   \caption{BOTT PLOT this is a holder until Kim updates the figure}
+   \caption{BOTT PLOT this is a holder until Kim updates the figure {\color{Green}Victor Trees: The glory (in the backscattering direction) and diffraction peak (in the forward scattering direction) are distinct features in the total reflected flux, but not in the degree of polarization.
+In principle, the degree of polarization goes to zero in the forward and backward scattering directions.
+See e.g.
+Fig.
+1 of \citet{stam2008}, Fig.
+5 of \citet{trees2022}, Fig.
+3 of \citet{karalidi2011} and \citet{hansentravis1974}.}}
     \label{fig:bottplot}
 \end{figure*}
+%--------------------------------------------------------------------------------
 
@@
 
-Surface liquid water is closely related to planetary habitability: water is required for life as we know it, and life on exoplanets will most likely on be detectable if water is present at the surface.
+Liquid water is closely related to planetary habitability because it is essential for life as we know it. 
+Indeed, to be able to detect such life forms, this liquid water should be on the surface of an exoplanet instead of below it.
 To first order, the presence of surface liquid water can be predicted by the flux a planet receives from its star.
-The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which an Earth-like planet could maintain surface liquid water \citep{kasting93}.
-The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis.
-We would therefore like to establish the presence of liquid water on the surface of an exoplanet before trying to interpret its atmospheric biosignatures.
-%Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
+The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which a rocky planet would have the right temperature to maintain liquid surface water \citep{kasting93}.
+The HZ is a convenient concept because with it a planet can be deemed habitable based on solely its orbital separation, but it remains an untested hypothesis. 
+We therefore prefer to establish the presence of liquid surface water on an exoplanet before attempting to interpret biosignatures in the planet's atmosphere. 
+% Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
 
@@
 \subsection{Ocean Glint, Cloud Glory and Rainbows}
-\cite{2001Natur.412..885F} noted the potential for time-resolved reflected light photometry to identify water oceans on the surface of an exoplanet.
-\cite{2008Icar..195..927W} performed simulations of scattered light from directly imaged terrestrial exoplanets.
+%
+Starlight incident on liquid water on the surface of an Earth-like exoplanet is reflected back into space towards the observer.
+This Fresnel (or specular) reflection leaves a mirror image of the star on the water surface, called the glint, which broadens with increasing water surface roughness driven by the wind speed \citep{CoxMunk1954}.
+The strength of the Fresnel reflection increases with increasing reflection angle locally on the planet, resulting in a stronger glint.
+Consequently, in the total reflected flux by the exoplanet as a whole, the signal of the glint is most prominent at the largest planetary phase angles.
+%Within our Solar System, the glint has been used to detect methane lakes on Titan} \citep[][]{2010GeoRL..37.7104S}.
+
+\cite{2001Natur.412..885F} noted the potential for time-resolved reflected 
+light photometry to identify landmasses and oceans of liquid on the surface of 
+a rotating exoplanet.
+\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to clouds, with a relatively small contribution from ocean glint.
+In practice, clouds complicate the detection of ocean glint because they can obscure the underlying surface and can act as strong forward scatterers, mimicking the ocean glint in the total reflected flux at large phase angles.
+Scattering of light by the atmospheric gas, i.e.
+Rayleigh scattering, also blurs the glint.
+The Rayleigh optical thickness of a planet's atmosphere is smaller in the near-infrared (NIR) than in the visible, which improves the visibility of the glint \citep{Zugger_2011}.
+NIR phase measurements also minimize the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+
+ %Rotational mapping of an exoplanet at gibbous phases would reveal oceans as dark regions \citep{2009ApJ...700..915C} whose albedo would increase sharply at crescent phases \citep{lustig2019}.
+
+Except for only the total reflected flux at various wavelengths, the degree of polarization (the polarized flux divided by the total flux) and direction of polarization of the light could also be measured.
+The degree of polarization has practical advantages, because it is a relative measure: instrument errors could be cancelled out upon division of the polarized flux by the total flux, depending on the telescope design.
+As the starlight of a solar-type star is often considered to be unpolarized \citep{kemp1987}, and gets polarized during the reflection by the exoplanet, measuring polarized light increases the contrast between the exoplanet and its host star.
+The degree of polarization itself also provides information about the exoplanet's atmosphere and surface.
+In fact, most scattering phenomena in the atmosphere are more sensitive to the scattering angle in the degree of polarization than in the total (polarized plus unpolarized) flux \citep{hansentravis1974}.
+Also for detecting exo-oceans, measuring the degree of polarization could be an asset.
+The glint is a strong source of polarized reflected flux, and the degree of polarization of the light reflected by the (cloud-free) water surface peaks at the Brewster angle, which occurs at larger angular planet-star separations than the glint in the total flux \citep[see e.g.][]{Zugger_2010,treesandstam2019}.
+
 %to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
-They showed that oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
-%Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but
-Specular reflection, a.k.a.
-glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., Titan's methane lakes;][]{2010GeoRL..37.7104S}.
-%\citet{lustig2019} explored mapping exoplanets at crescent phases using ocean glint.
-Specular reflection appears at similar phase angles in both unpolarized and polarized light and its peak orbital phase can be extended by atmospheric winds, which produce waves and hence increase the size of the glint spot \citep{2011Icar..211..722B,kopparla2018}.
+%They showed that oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases (Figure~\ref{fig:bottplot}). 
+%Specular reflection appears at similar phase angles in both unpolarized and polarized light and its peak orbital phase can be extended by atmospheric winds, which fuel surface waves and hence broaden the size of the glint spot
 %\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
-\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean, proving that the method works in practice for an unresolved world.
-
-%\noindent {\color{Green}<Modifications by Victor Trees>}
-
 %\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
 They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
-
 % NEEDS Paragraph on mapping connection to tectonics
-
 % NEEDS KOPPARLA (prob refer to , BERDYUGINA, ROSSI?
@@
 blend cloud hinderance into argument for context to lead into rainbows, mention Lietzow and jeremy there., maybe some more mention of the Earthshine observations Karalidi, Gordon, etc
-
 %"Stam & Hovenier (2006) independently examined the observability of the polarized signatures of an Earth-like extrasolar planet, including Rayleigh scattering and sea-surface glint.
@@
 
-% Move to Clouds?
-In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and can act as strong forward scatterers, mimicking ocean glint in the total  reflected flux (polarized plus unpolarized).
-\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
+\cite{2008Icar..195..927W} performed simulations of reflected light from directly imaged exoplanets with oceans, and found strong peaks in the degree of polarization at the Brewster angles, but did not take into account Rayleigh scattering gas on top of the oceans.
+\cite{Zugger_2010} simulated photometric and polarized phase variations at visible wavelengths for ocean planets including the effects of Rayleigh scattering gas and a horizontally homogeneous cloud cover.
+They concluded that detection of an ocean glint signature in polarization is severely limited by the presence of clouds.
+\citet{kopparla2018} explored the ocean glint signatures in the signal from a multi-planetary (TRAPPIST-1) system, with TRAPPIST-1e being the ocean planet.
+They concluded that, although measuring in polarized light would increase the contrast between the ocean planet and the total signal of the multi-planetary system, the glint detectability is limited by Rayleigh scattering and clouds on the ocean planet and its neighbour planets.
 
-\cite{Zugger_2010} simulated photometric and polarized phase variations at visible wavelengths for Earth-like planets including the effects of %partial cloud coverage
-Rayleigh scattering and a horizontally homogeneous cloud cover.
-They concluded that detection of an ocean glint signature in polarization is severely limited by the presence of clouds.
 %Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they %again 
@@
 %{\color{Green} 
-\citet{treesstam2019} simulated the total and polarized fluxes and the degree of polarization of light reflected by ocean planets with patchy clouds.
-They found that the alternation of cloud-free and cloud-covered glint, as the planet rotates and its cloud cover changes in time, can result in strong fluctuations in the degree of polarization, hence resulting in noisier phase curves.
-Nonetheless, polarization phase curves of ocean planets at optical and near infrared wavelengths intersect \citep{treesstam2019}, providing a possible avenue to distinguish between an exo-ocean and forward scattering clouds.%}%}
+\citet{treesandstam2019} simulated the total and polarized fluxes and the degree of polarization of light reflected by ocean planets with a \textit{patchy} cloud cover.
+They showed that the alternation of cloud-free and cloud-covered glint, as the cloud pattern changes in time, can result in strong fluctuations in the degree of polarization phase curves at continuum wavelengths (i.e.
+at wavelengths at which the light is not absorbed by gases in the exoplanet's atmosphere), with generally high polarization
+when the glint is visible and low polarization when the glint is covered.
+Fluctuations of the polarization continuum phase curves are also expected when continents rotate over the glint \citep{groot2020}.
+If the telescope's integration time is too long to resolve those temporal variations, the fluctuations will results in noisier phase curves.
+Nonetheless, the polarization phase curves of partially cloudy planets with oceans at optical and near infrared wavelengths intersect \citep{treesandstam2019}, providing a possible avenue to distinguish between an exo-ocean and forward scattering clouds.
 
-%{\color{Green}
-The Rayleigh scattering optical thickness of a planet's atmosphere is smaller in the near-infrared (NIR) than in the visible, which improves the visibility of the glint \citep{Zugger_2011}.
-NIR phase measurements also minimize the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
+\textit{<Observing the Earth as an exoplanet>}
+Measurements of Earthshine (i.e.
+the reflected sunlight by the Earth incident on the dark side of the Moon) indeed reported higher degrees of polarization in the continuum at long wavelengths than at short wavelengths, which could be attributed to the ocean glint \citep{Emde2017,sterzik2019,takahashi2021}.
 
-Phase variations of polarized light could also reveal other non-Lambertian cloud scattering phenomena.
-The degree of polarization of light scattered by spherical cloud droplets strongly depends on the scattering angle \citep[see][]{hansentravis1974}, and consequently, on the phase angle for an Earth-like exoplanet with liquid water clouds \citep[see][]{stam2008}.
-The sign of the degree of polarization (i.e.
-the direction of polarization) changes sign, and show peaks at the primary and secondary rainbow angles  \citep{bailey2007,2018MNRAS.480.1613B}.
-A measurement of those rainbows would be a strong indicator of spherically shaped particles in the exoplanet's atmosphere.
-The precise phase angle locations in the the degree of polarization of the rainbow peaks depend on wavelength \citep[see Fig.
-5 of][]{trees2022}, the particle size ...
-and on the type of liquid \citep{bailey2007}.
-A famous example...
-Venus ...
-%A famous example of using polarimetry to characetrize clouds within our Solar System is the characterization of the main cloud deck of Venus.
-\citep{hansenhovenier1974} derived that the main cloud deck of Venus consists of a 75\% sulfuric acid solution.
-%
-... glory ...\citep{2014A&A...566L...1G} ... and halos ... ice clouds ...
+%\textit{\citet{sterzik2019} and \citet{Takahashi2021} used polarimetric Earthshine 
+%measurements to detect the presence of the ocean, proving that the method 
+%works in practice for a spatially unresolved world.}
 
- \textbf{Victor/Kim: please complete this paragraph with a few sentences describing what one learns from rainbow and glory, ideally backed with good references!}
+The optical properties of an Earth-like exoplanet orbiting its host star may vary along the phase curve, e.g., because of the planet's seasons and changing weather systems in time.
+Recent simulations by \citet{trees2022} showed that signatures of the glint could also be found in the degree of polarization spectra (i.e., as functions of wavelength with 1 nm resolution), measured at a single phase angle.
+Taking into account the absorption of light by gases in the exoplanet's atmosphere, they computed that the degree of polarization of ocean planets is lower across gaseous absorption bands with respect to the continuum, while for dry planets, it is higher \citep[see][]{stam2008}.
+Those dips, instead of peaks, can already be searched for at intermediate phase angles ($90^\circ$-$130^\circ$), thus at the largest angular planet-star separations, and also occur for high planetary cloud fractions as long as the glint is (partially) cloud-free.
+Comparing the degree of polarization in the continuum and at a wavelength (band) where gases partly absorb the light, could potentially allow for an ocean detection with only one measurement in time.
+In this paper, we neglect absorption by gases and do not focus on phase variations of the spectra, but on the phase variations at single wavelengths in the continuum.
+
+%Phase variations of polarized light could also reveal other non-Lambertian scattering phenomena (Figure~\ref{fig:bottplot}).
+
+The degree of polarization of light scattered by spherical cloud droplets strongly depends on the scattering angle \citep[see][]{hansentravis1974}, and consequently, on the phase angle for an Earth-like exoplanet with liquid water clouds \citep{stam2008}.
+The scattering angle dependent variations of the degree of polarization for different droplet sizes are more distinct than in the total reflected flux \citep{karalidi2011}, and the sign of the degree of polarization (i.e.
+the direction of polarization) changes at various angles \citep{stam2008,trees2022}.
+In particular, the degree of polarization shows peaks at the primary and secondary rainbow angles, caused by reflection of light inside the spherical droplets, and the shapes of the peaks depend on the wavelength, particle size (distribution) and type of liquid  \citep{hansentravis1974,bailey2007,bailey2018}.
+A measurement of rainbow features in polarized light would be a strong indicator of spherically shaped (i.e.
+liquid) particles in the exoplanet's atmosphere, as they are not expected to originate from reflection by the planet's surface or non-spherical (ice) cloud particles \citep{Karalidi2012ice}.
+In the total reflected flux, thus without the separation of the polarized light, the rainbow features are virtually unobservable \citep{karalidi2011}.
+The benefits of those rainbow features in polarized light for cloud characterization on other planets than the Earth have been proven by \citet{hansenhovenier1974}, who fitted model calculations through measurements of the degree of polarization phase curves of Venus.
+%Because Venus is an inner planet as seen from the Earth, it can be observed from the Earth at phase angles ranging from almost $0^\circ$ to $180^\circ$ .
+%They retrieved that the main cloud deck of Venus consists of a 75\% sulfuric acid solution, with a droplet effective radius of 1.05 $\mu$m and effective variance of 0.07.
+
+%... glory ...\citep{2014A&A...566L...1G} ... and halos ... ice clouds ...
 
@@
 
-%\subsection{Rainbows}
-%calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below
+\subsection{Rainbows {\color{Green} [Victor: The rainbows are already discussed in previous section.]}}
 
-%The rai
+A well-characterized rainbow feature in the phase curve of a planet provides a wealth of information about the atmosphere and condensates.
+The rainbow feature is visible in both polarized and unpolarized phase curves, but is particularly prominent in polarized light because it is strongly polarized and the fractional polarization from other sources like Rayleigh scattering is reduced at narrow scattering angles where many rainbows occur.
+The location of the rainbow peak in the phase curve shifts depending on the refractive index of the condensate and thus will shift depending on the species, and for most species the peak will also shift slightly with wavelength.
+The relative height of the rainbow compared to the Rayleigh peak with wavelength provides evidence of the depth of the cloud top height.
+The size distribution of droplets also effects the polarized rainbow and other optical effects.
+% Shape/phase
+% 
+% cite Bailey
 
-%cloud coverage paper in polarisation :
+As a characterization tool the rainbow and other cloud optical features are vital, and in studies of habitability they can provide evidence that water is the dominant condensate providing context to the glint feature.
+In this work we identify our ideal rainbow angle as %...
+
+In our own Solar System the additional information inherent in condensate polarization allowed the identification of the species and droplet size distribution of the clouds on Venus \citep{Hansen1974}, the discernment between a snowy surface and hazy thick atmosphere on Titan \citep{Zellner1973}, and may have eliminated some prospects for the unknown UV absorber on Venus \citep{Petrova2018}.
+
 %\cite{rossi17} computed flux and polarisation of reflected starlight for different types of cloud cover on Earth-like planets and determined different types of cloud cover can be distinguished from each other.
@@
 \subsection{Outline}
-In this paper we quantify the accessible scattering phase angles for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars.
-In \S 2.1 we describe the target list, in \S 2.2 we derive expressions for the range of scattering phase angles in a few different limits, in \S 2.3 we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4 we outline how the planet/star contrast changes as a function of orbital phase.
+
+In this paper, we quantify the phase angles that are accessible for direct
+observations for hypothetical terrestrial planets orbiting in the habitable 
+zones of HabWorlds target stars. 
+In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the 
+range of phase angles given a few different limits, in \S 2.3, we describe 
+Monte Carlo simulations to marginalize over the unknown orbital inclination and 
+eccentricity of planets in these systems, and in \S 2.4, we outline how the 
+planet/star contrast changes as a function of the planet's orbital phase.  
 We present and discuss our results in \S 3 and conclude in \S 4. 
@@
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\hwo\ is still in the early stages of development, and the exact telescope and instrument design is undecided. 
-%
-In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
-%
-In addition, the observatory will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
-%
-The \IWA\ of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
+\hwo\ is still in the early stages of development, and the exact 
+telescope and instrument design is as of yet undecided. 
+In this work, we assume a \SI{6}{\meter} primary mirror, which is the 
+currently favoured design for the observatory.
+In addition, the observatory will utilise a coronagraph to suppress the 
+stellar light and facilitate high-contrast observations of faint 
+companying exoplanet. 
+The \IWA\ of the coronagraph is currently undecided and will be driven 
+by both the science requirements and technological limitations. 
 %In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
@@
 For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-
-For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.
+For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations
+corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at representative visible and near infrared wavelengths of $\lambda = \SI{600}{\nano\meter}$ and $\SI{1}{\micro\meter}$.
 
@@
         Conversion between $\lambda / D$ and \si{\mas} for different wavelengths, assuming $D = \SI{6}{\meter}$. 
-        \todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
+        %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
     }
@@
     \toprule
-     & $\lambda/D$ & mas (at \SI{600}{\nano\meter}) & mas (at \SI{700}{\nano\meter}) \\
+     & $\lambda/D$ & mas (at \SI{600}{\nano\meter}) & mas (at \SI{1}{\micro\meter}) \\
     \midrule
     \midrule
-    IWA & 1 & 20.63 & 24.06 \\
-    IWA & 2 & 41.25 & 48.13 \\
-    IWA & 3 & 61.88 & 72.19 \\
-    IWA & 4 & 82.51 & 96.26 \\
+    IWA & 1 & 20.63 & 34.38 \\
+    IWA & 2 & 41.25 & 68.75 \\
+    IWA & 3 & 61.88 & 103.13 \\
+    IWA & 4 & 82.51 & 137.51 \\
     \midrule
-    OWA & 32 &  660.05 & 770.06 \\
-    OWA & 64 & 1320.09 & 1540.11 \\
+    OWA & 32 &  660.05 & 1100.08 \\
+    OWA & 64 & 1320.09 & 2200.16 \\
     \bottomrule
@@
     \caption{
-        The coronagraph can obscure the exoplanet at high and low scattering phase angles.
+        An illustration of an edge-on circular orbit showing that the coronagraph can obscure the exoplanet at high and low scattering phase angles.
 This can prevent observations of scattering phenomena which could be used to indicate the habitability and atmospheric composition of the exoplanet.
-\todo{Add why scattering is cool}
+        %\timmy{Maybe mention that this is a simplified illustration with a circular orbit, plus we are basically looking at an edge-on system?}
     }
@@
 \begin{equation}
+ \label{eq:Delta_phi}
     \Delta \varphi 
@@
     =  2(\varphi_\mathrm{max} - 90^\circ) 
-    =  2(90^\circ-\varphi_\mathrm{min} ) \notag
+    =  2(90^\circ-\varphi_\mathrm{min} ) 
 \end{equation}
@@
 
-\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the HabWorlds target list assuming an IWA of 62 mas ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
+\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo\ target list, assuming an IWA of \SI{62}{\mas} ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
+\timmy{60 (as in the abstract), or 62?}
 %
@@
     \caption{
-        \timmy{I believe those should be the results for circular, edge-on orbits, but let's double-check with Max.}
-        Scatter plot at 3 $\lambda/D$ for the target sample, showing stellar effective temperature and stellar distance. 
+        Scatter plot for the target sample, showing stellar effective temperature and stellar distance. 
         The size of the points represents the angular separation of the star and planet in milliarcseconds as presented in the target list. 
-        The colour of the points shows the atmospheric phenomenon that can be detected with darker colours, including all lighter (yellow) colour phenomenon. 
+        The colour of the points shows $\Delta \varphi$ assuming a circular edge-on orbit at a semi-major axis corresponding to an Earth-like instellation for an IWA of 62 mas.
+Additionally, the colour bar indicates the atmospheric phenomenon that can be detected where phenomena from the bottom of the colour bar up to the indicated colour are detectable for that system.
         Thus, dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
@@
 
-\begin{figure}[htb]
+\begin{figure}%[htb]
    \centering
@@
 
-For each star on the \hwo\ target list, we generate XX hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
+For each star on the \hwo\ target list, we generate $1000$ \todo{I think this is what the code says, check} hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
 %
@@
 We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
-The solid lines in Figure \ref{fig:betaallofit} show the cumulative distribution for the most extreme scattering phase angles accessible for the \hwo\ target list, depending on the choice of inner working angle.
-
-\todo{what does this plot show, how many systems could we see?}
+%
+The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible scattering phase angles normalised to the total number of systems in \hwo\ target list, depending on the choice of inner working angle.
 
@@
     \caption{
-        The big kahuna
-        \timmy{What are the dashed vs. solid lines?}
-        \timmy{How did this plot aggregate the simulations for each planet? mean? median? min/max?
-        }
+        \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
+        Cumulative distributions of the most extreme scattering phase angle accessible for different \IWA where the solid lines are for randomly inclined circular orbits and the dashed lines are for randomly orientated elliptical orbits.
+The top x axis indicates the minimum scattering phase angle and the bottom x axis indicates the maximum.
+These are symmetric about quadrature (90 degrees) \todo{is this true for elliptical orbits?}.
+The y axis indicates the number of systems divided by the number of Monte Carlo samples and is thus normalised to the number of systems in the target list.
+The secondary y axis indicates the number of systems assuming only 24 percent of them contain an Earth-like exoplanet in the habitable zone.
     }
@@
 Eccentricities were drawn from the beta distribution, and inclinations were uniform in $\cos i$.
-%
-We then filtered on \IWA's of 1, 2, 3 and 4 times $\lambda / D$, before identifying which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
-%
-\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's.
-The dashed lines in Figure \ref{fig:betaallofit} show the cumulative distribution for the most extreme scattering phase angles accessible, when accounting for orbital eccentricity.
-Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph IWA and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
-
-\todo{how was the angle coverage calculated in the monte carlo?}
-
+\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's
 %
 %Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
+%
+We then calculated the minimum and maximum accessible scattering phase angles for of \IWA's of 21, 42, 63 and 84 mas which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\SI{600}{\nano\meter}$.
+The dashed lines in \cref{fig:betaallofit} shows the normalised cumulative distribution of the accessible angles for each of the \IWA, when accounting for orbital eccentricity.
+%
+We also calculate which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+The cumulative distribution over \IWA\ of the number of systems, normalised to the total number of systems, where each scattering phenomena is visible is shown in \cref{fig:accessible_phase_angles}.
+%
 
 \subsection{Simulating contrast curves}
-\textbf{Kim and David}
+\todo{Kim and David: is this done?}
+
+A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, that is, the relative flux of the reflected light from the planet divided by the stellar flux.
 %
-A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, i.e.
-the relative flux of the reflected light from the planet divided by the stellar flux.
-%
-While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.https://www.overleaf.com/project/63f87efaa38311f6a4b2a551
+While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as a function of scattering phase angles.
 %
@@
 %
-We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of an Earth-like planet \citep{ treesstam2019,trees2022}.
+We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of an Earth-like planet \citep{ treesandstam2019,trees2022}.
 
@@
 %
-For their calculations they made the simplified assumptions of circular orbits, a geometric albedo of $p=0.2$, and that the scattering phase function, $\phi (\alpha)$, can be described by a simple Lambertian reflectance phase function, e.g.
-isotropic scattering.
+The assumptions that went into these calculations are that the orbits are circular, the geometric albedo is $p=0.2$, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, for example, isotropic scattering.
 %
@@
 
-We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
+We use the scattering phase functions from \cite{treesandstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
 %
-The $\SI{670}{\nano\meter}$ is close to the center of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
+The $\SI{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
 %
@@
 %
-Together they give the reflected light contrast as function of separation for a full orbit.
+Together, they give the reflected light contrast as a function of separation for a full orbit.
 %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\todo{Is there a plot for this? How many systems could we see? Why is it different to non-eccentric cases?}
-
-\timmy{No plot / results yet; see above.}
-
-\todo{Sophia: what is features\_plot\_peak.png showing?}
-
-\timmy{Good question --- maybe Matt or Max know how the data for this was generated?
+%\todo{Is there a plot for this? How many systems could we see? Why is it different to non-eccentric cases?}
+%\timmy{No plot / results yet; see above.}
+%\todo{Sophia: what is features\_plot\_peak.png showing?}
+%\timmy{Good question --- maybe Matt or Max know how the data for this was generated?
 Matt: yes, it's all in showyourwork right now.
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{The affect of eccentricity}
+\label{sec:result_eccentricity}
+As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable affect on the number of systems that can be observed at each scattering phase angle.
+This is because, Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA\ and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
+The eccentricity does not significantly effect the number of systems for smaller inner working angles as the systems in \hwo\ target list where chosen based on their on-sky separation quadrature and therefore the shape of the orbit has little affect for \IWA\ smaller than cut used in the selection.
+It is, however, important to account for orbital eccentricity when the \IWA\ of the coronoagraph is larger than the \IWA\ used to filter the target list otherwise the number of systems observable at extreme phase angles would be under predicted.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Observable scattering phenomenon}
+\label{sec:results_scattering_phenomena}
+
+\cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
+The \IWA\ of \hwo's coronograph will depend on the primary mirror diameter, observing wavelength and how well a coronograph of a given size can suppress the glare of the host star.
+The smaller the \IWA, the more features we will be observable and in more systems.
+
+The Rayleigh scattering feature appears near quadrature and is therefore observable in the most systems.
+This feature is caused by atmospheric gases and can therefore be used to determine if a planet has an atmosphere.
+The rainbow feature appears further from quadrature but will still be observable in a significant fraction of systems.
+This feature will only be present if the atmosphere contains spherical droplets, potentially indicating the presence of liquid water.
+The position and shape of the peak is sensitive to the properties of the droplets allowing their nature to be determined if the peak of the rainbow feature can be reached.
+The ocean glint feature occurs even further from quadrature and so is obserable in fewer systems.
+In most systems, only the first half of the feature is observable but this can still be used to identify a liquid ocean on the surface of the planet.
+A key indicator of habitabitily.
+Finally, it is possible to view the glory scattering feature in one system, Alpha Cen A.
+\todo{maybe we want a table of some numbers in this section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Contrast Curves}
+\label{sec:results_contrast}
+
+The accessible scattering phase angles is not the factor that needs to be considered when determining what scattering features are observable.
+The contrast of the planet is also an important factor to consider.
+\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo\ target list assuming egde-on orbits.
+The over all contrast varies significantly for each system and throughout the orbit.
+The observability of these planets will depend on the star light suppression of the coronograph which is influenced by its \IWA.
+It is not possible to predict how much the state of the art of coronaography will have advanced by the time \hwo\ is built.
+However, when its design is being finalised, it will be important to consider both the accessible scattering phase angles and contrast limits when deciding on the \IWA\ of the coronograph, if these features are to be observed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\subsection{The glory of Alpha Cen}%Halelujah
+%\label{sec:ealpha-cen}
+%\textbf{Glory science, what it tells us (Kim)}
+% Kim will write para on science of glory (what tells us, vaguely why appears here)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
         The orbits are scaled by the Earth-equivalent flux distance. 
-        The concentric circles marked by the dashed lines indicate inner working angles of 1, 2, 3, and 4\,$\lambda / D$, corresponding to 20.6, 41.3, 61.9, and \SI{82.5}{\mas}, respectively (assuming $\lambda = \SI{600}{\nano\meter}$ and $D = \SI{6}{\meter}$).
+        The concentric circles marked by the dashed lines indicate inner working angles of 20.6, 41.3, 61.9, and \SI{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$ respectively (assuming $\lambda = \SI{600}{\nano\meter}$ and $D = \SI{6}{\meter}$).
         The figure illustrates that the \IWA\ can significantly affect the range of scattering phases observable with each orbit.
@@
     \caption{
-        Number of systems vs inner working angle
+        The number of systems for which key scattering features (Rayleigh scattering, rainbows, ocean glint and glories) can be observed as a function of \IWA.
+Each subplot contains three lines where the dashed line shows the number of systems where the start of the scattering feature can be observed, the solid line shows the number of systems that reach the peak of the feature and the dot-dashed line the number where the scattering angles of the end of the feature are observable.
+The bottom x axis gives the \IWA\ of the coronograph in milliarcseconds which is converted to $\lambda/D$ for $\lambda=\SI{600}{\nano\meter}$ and $D=\SI{6}{\meter}$ for the top x axis.
+The secondary y axis indicates the number of systems that each feature would be observable in assuming only 24 percent of them contain an Earth-like exoplanet in the habitable zone.
+It is worth noting that the Glory feature is only observable in one system, Alpha Cen A.
     }
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{The glory of Alpha Cen}%Halelujah
-\label{sec:ealpha-cen}
-\textbf{Glory science, what it tells us (Kim)}
-% Kim will write para on science of glory (what tells us, vaguely why appears here)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040's \todo{check}. 
+The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040's. 
 %
@@
 %
-In this work, we investigated the possibility of using \hwo\ to detect different scattering phenomena which can be used to indicate the habitability of an exoplanet. \todo{more of why this is cool}
+In this work, we investigated the possibility of using \hwo\ to detect scattering phenomena on terrestrial exoplanets, including ocean glint, which could establish a planet's habitability.
 
-Coronagraphy is a rapidly advancing field and it is hard to predict what the capabilities of \hwo\ will be.
-%
 We primarily focused on the range of scattering phase angles and therefore which scattering phenomena could potentially be observed given the limitation of the coronagraphic \IWA.
 %
-Assuming a 60 mas \IWA, all of the systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, ocean glint would be accessible in $\sim$20 systems, and the rainbows of water clouds in $\sim$55 systems.
+Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \hwo\ will be.
 %
-The number of systems increases with decreasing \IWA\ and therefore the occurrence rate of Earth-like exoplanets will drive the choice of coronagraphic \IWA\ for this mission.
+Assuming a \SI{60}{\mas} \IWA, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$55 systems, and ocean glint in $\sim$20 systems.
+The habitability of Earth analogs in the remaining 140 systems would have to be established by less direct means like rotational mapping to identify continents and oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
+%
+
+The number of habitable zones that can be probed by \hwo\ increases with decreasing \IWA. 
+%
+The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic frequency of such planets, will drive the choice of coronagraphic \IWA\ for this mission.
+%
+We have shown that only $\sim$10\% of Earth-like exoplanets imaged with \hwo\ could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
+%
+The habitability of remaining targets would have to be established by less direct means.
 %
@@
 
+\section*{Author affiliations}
+\begingroup
+% \footnotesize
+\itshape
+\affiliationtarget{1}{$^{1}$} Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK\\
+\affiliationtarget{2}{$^{2}$} Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\
+\affiliationtarget{3}{$^{3}$} Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK\\
+\affiliationtarget{4}{$^{4}$} Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada\\
+\affiliationtarget{5}{$^{5}$} Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
+\affiliationtarget{6}{$^{6}$} Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 Tübingen, Germany \\
+$^{7}$ LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-92195 Meudon, France \\
+$^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA \\
+$^{9}$ Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\
+$^{10}$ Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\
+$^{11}$ National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C. Canada \\
+$^{12}$ University of California, Santa Cruz, USA  \\
+$^{13}$ Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\
+$^{14}$ DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\
+$^{15}$ Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, Arizona \\
+$^{16}$ STAR Institute, Université de Li\`ege, All\'ee du Six Ao\^ut 19c, 4000 Li\`ege, Belgium \\
+$^{17}$ Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\
+$^{18}$ Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\
+$^{19}$ SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\
+$^{20}$ NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\
+$^{21}$ University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\
+$^{22}$ DOTA, ONERA, 92322 Châtillon, France \\
+$^{23}$ European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\
+$^{24}$ NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\
+$^{25}$ Subaru Telescope, NAOJ, USA \\
+$^{26}$ College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\
+$^{27}$ Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\
+\affiliationtarget{28}{$^{28}$} NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\
+\affiliationtarget{29}{$^{29}$} ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland
+\endgroup
+
 \end{document}

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Frans Snik$^{5}$,
+    Daphne M. Stam$^{30}$,
+    Victor Trees$^{31,32}$,
     Dirk van Dam$^{5}$,
@@
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
-The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$20 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$55 systems, assuming a 60\,mas inner working angle (\IWA).
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.     \end{abstract}
+The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
+Increasing the \IWA to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.
+\end{abstract}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
@@
    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-   \caption{BOTT PLOT this is a holder until Kim updates the figure {\color{Green}Victor Trees: The glory (in the backscattering direction) and diffraction peak (in the forward scattering direction) are distinct features in the total reflected flux, but not in the degree of polarization.
+   \caption{BOTT PLOT this is a holder until Kim updates the figure}
+    \label{fig:bottplot}
+\end{figure*}
+%--------------------------------------------------------------------------------
+
+Signatures of habitability like ocean glint and optical effects from condensates--- like the rainbow---should be some of the most distinct signals in an exoplanet's phase curve, particularly in polarized light.
+However, depending on system architectures, some of these features can be obscured by the inner working angle of a coronagraph.
+
+Surface liquid water is closely related to planetary habitability because it is essential for life as we know it. 
+To first order, the presence of liquid surface water can be predicted by the flux a planet receives from its star.
+The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
+ However, there are many factors that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and geological activity. 
+Establishing the presence of liquid surface water on an exoplanet therefore provides important context for interpreting possible biosignatures.
+Providing additional context to the habitability of an exoplanet is the presence of a water-cycle as evidenced by the characterization of a rainbow and other cloud optical effects.
+
+%The phase curves showing these features are distinct depending on the characteristics of the planet in both polarized and unpolarized measurements.
+Polarized and unpolarized phase curves and spectra each have their own strengths in characterization outlined below.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Exoplanet Phase Curves}
+An exoplanet observed in an edge-on orbit will move through all phases and thus all scattering angles, while those viewed at higher inclinations will pass through a subset of these scattering angles.
+If the planet is rotating, this supplies additional regular variations that can be used to place constraints on the heterogeneity of the planet and potentially map it.
+
+The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scatter, rainbows, and glories.
+In polarized light an added potential benefit is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade \citep{kemp1987}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Ocean Glint}
+
+%As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
+Ocean glint is from the specular (Fresnel) reflection of light off of smooth liquid surfaces at crescent phases.
+In the total flux the signal of the glint will be most prominent at small planetary phase (large scattering) angles.
+The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53 degrees (a scattering angle of 127 degrees) \citep[see, e.g.
+][]{2008Icar..195..927W}.
+It therefore varies slightly by wavelength and by sea and atmosphere composition.
+In general on a planet the glint will be strongest in red wavelengths where the Rayleigh contribution to both the phase curve and attenuation are minimized \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also minimized \citep{2012ApJ...752L...3C}.
+The strengthened glint signature at longer wavelengths has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across scattering angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
+
+The glint provides evidence of surface liquid bodies, which taken in the context of the temperature and condensates may be interpreted as the presence of liquid water.
+The forward scattering from clouds can be confused with glint \citep{Robinson_2010} further motivating also probing the rainbow angle for context.
+The cloud coverage fraction on ocean worlds can also be constrained with observations spanning the ultraviolet to near-infrared by the point phase curves at all wavelengths intersect in polarized light \citep{treesandstam2019}.
+Additionally,
+ \citet{trees2022} showed that glint can also be detected in spectropolarimetry while the planet is at phase angles placing it at the wide separations more easily probed with coronagraphs.
+%how use to map, 
+%seen in Earthshine
+%strong signal for exoplanets
+% disentangled from cloud coupled with rainbow
+%disentangled from lattitude albedo coupled in polarization
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Rainbows and Glories}
+Rainbows, glories and other optical effects from condensates provide a wealth of information about an exoplanet \citep{karalidi2012rainbow, stam2008}.
+In our own Solar System the polarized back scatter from Titan provided evidence that the world's reflectance was due to haze rather than a surface \citep{zellner1973polarization}, on Venus the rainbow and phase curve behavior with wavelength showed the clouds were $\sim$75\% H$_2$SO$_4$ -- water mixture also constraining the droplet size distribution \citep{hansenhovenier1974}, and the glory was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
+
+In unpolarized light the forward scattering and backscatter from clouds can dramatically change the shape of a phase curve from Lambertian.
+While the glory and rainbow are technically visible in any phase curve, their strongly polarized nature coupled with the suppression of signal from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
+The rainbow angle depends on the refractive index and therefore varies with wavelength and depending on the species.
+The feature being far more narrow than the glint allows more exact determination of the species.
+The droplet size distribution also effects the shape of the rainbow and clouds of species in a solid phase can diverge from the Mie scattering that produces these features greatly, producing other optical effects particularly at narrow scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
+In this paper we define our optimal angle for rainbow detection as the peak of the water droplet rainbow in visible wavelengths.
+However, extending observations to angles closer to the star (smaller scattering angles, larger phase angles) would allow the full shape of the rainbow to be characterized and ensure the characterization of rainbows from other possible HZ condensates.
+The glory is produced when the droplet radii much smaller than the wavelength of light and can allow further characterization as their angle, intensity, and shape also vary with the complex refractive index but primarily occur within 20 degrees scattering angle very close to full phase.
+%The glory (in the backscattering direction) and diffraction peak (in the forward scattering direction) are distinct features in the total reflected flux, but not in the degree of polarization.
 In principle, the degree of polarization goes to zero in the forward and backward scattering directions.
@@
 3 of \citet{karalidi2011} and \citet{hansentravis1974}.}}
-    \label{fig:bottplot}
-\end{figure*}
-%--------------------------------------------------------------------------------
-
-%\textcolor{red}{introduce motivation, specs for habitable worlds observatory and the document we got the target list from.}
-
-Liquid water is closely related to planetary habitability because it is essential for life as we know it. 
-Indeed, to be able to detect such life forms, this liquid water should be on the surface of an exoplanet instead of below it.
-To first order, the presence of surface liquid water can be predicted by the flux a planet receives from its star.
-The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which a rocky planet would have the right temperature to maintain liquid surface water \citep{kasting93}.
-The HZ is a convenient concept because with it a planet can be deemed habitable based on solely its orbital separation, but it remains an untested hypothesis. 
-We therefore prefer to establish the presence of liquid surface water on an exoplanet before attempting to interpret biosignatures in the planet's atmosphere. 
-% Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.
-
-%\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
-
-%\textbf{Why glint [Kim]}
-% probe of surface conditions; habitability
-
-\subsection{Ocean Glint, Cloud Glory and Rainbows}
-%
-Starlight incident on liquid water on the surface of an Earth-like exoplanet is reflected back into space towards the observer.
-This Fresnel (or specular) reflection leaves a mirror image of the star on the water surface, called the glint, which broadens with increasing water surface roughness driven by the wind speed \citep{CoxMunk1954}.
-The strength of the Fresnel reflection increases with increasing reflection angle locally on the planet, resulting in a stronger glint.
-Consequently, in the total reflected flux by the exoplanet as a whole, the signal of the glint is most prominent at the largest planetary phase angles.
-%Within our Solar System, the glint has been used to detect methane lakes on Titan} \citep[][]{2010GeoRL..37.7104S}.
-
-\cite{2001Natur.412..885F} noted the potential for time-resolved reflected 
-light photometry to identify landmasses and oceans of liquid on the surface of 
-a rotating exoplanet.
-\cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to clouds, with a relatively small contribution from ocean glint.
-In practice, clouds complicate the detection of ocean glint because they can obscure the underlying surface and can act as strong forward scatterers, mimicking the ocean glint in the total reflected flux at large phase angles.
-Scattering of light by the atmospheric gas, i.e.
-Rayleigh scattering, also blurs the glint.
-The Rayleigh optical thickness of a planet's atmosphere is smaller in the near-infrared (NIR) than in the visible, which improves the visibility of the glint \citep{Zugger_2011}.
-NIR phase measurements also minimize the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
-
- %Rotational mapping of an exoplanet at gibbous phases would reveal oceans as dark regions \citep{2009ApJ...700..915C} whose albedo would increase sharply at crescent phases \citep{lustig2019}.
-
-Except for only the total reflected flux at various wavelengths, the degree of polarization (the polarized flux divided by the total flux) and direction of polarization of the light could also be measured.
-The degree of polarization has practical advantages, because it is a relative measure: instrument errors could be cancelled out upon division of the polarized flux by the total flux, depending on the telescope design.
-As the starlight of a solar-type star is often considered to be unpolarized \citep{kemp1987}, and gets polarized during the reflection by the exoplanet, measuring polarized light increases the contrast between the exoplanet and its host star.
-The degree of polarization itself also provides information about the exoplanet's atmosphere and surface.
-In fact, most scattering phenomena in the atmosphere are more sensitive to the scattering angle in the degree of polarization than in the total (polarized plus unpolarized) flux \citep{hansentravis1974}.
-Also for detecting exo-oceans, measuring the degree of polarization could be an asset.
-The glint is a strong source of polarized reflected flux, and the degree of polarization of the light reflected by the (cloud-free) water surface peaks at the Brewster angle, which occurs at larger angular planet-star separations than the glint in the total flux \citep[see e.g.][]{Zugger_2010,treesandstam2019}.
-
-%to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
-%They showed that oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases (Figure~\ref{fig:bottplot}). 
-%Specular reflection appears at similar phase angles in both unpolarized and polarized light and its peak orbital phase can be extended by atmospheric winds, which fuel surface waves and hence broaden the size of the glint spot
-%\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
-%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets.
-They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.
-% NEEDS Paragraph on mapping connection to tectonics
-% NEEDS KOPPARLA (prob refer to , BERDYUGINA, ROSSI?
-(cloud?), HU?, added...
-blend cloud hinderance into argument for context to lead into rainbows, mention Lietzow and jeremy there., maybe some more mention of the Earthshine observations Karalidi, Gordon, etc
-%"Stam & Hovenier (2006) independently examined the observability of the polarized signatures of an Earth-like extrasolar planet, including Rayleigh scattering and sea-surface glint.
-Schneider (2006) had similar ideas." -KB
-
-\cite{2008Icar..195..927W} performed simulations of reflected light from directly imaged exoplanets with oceans, and found strong peaks in the degree of polarization at the Brewster angles, but did not take into account Rayleigh scattering gas on top of the oceans.
-\cite{Zugger_2010} simulated photometric and polarized phase variations at visible wavelengths for ocean planets including the effects of Rayleigh scattering gas and a horizontally homogeneous cloud cover.
-They concluded that detection of an ocean glint signature in polarization is severely limited by the presence of clouds.
-\citet{kopparla2018} explored the ocean glint signatures in the signal from a multi-planetary (TRAPPIST-1) system, with TRAPPIST-1e being the ocean planet.
-They concluded that, although measuring in polarized light would increase the contrast between the ocean planet and the total signal of the multi-planetary system, the glint detectability is limited by Rayleigh scattering and clouds on the ocean planet and its neighbour planets.
-
-%Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they %again 
-%find that oceans are detectable and distinguishable from other sources of polarization. 
-%{\color{Green} 
-\citet{treesandstam2019} simulated the total and polarized fluxes and the degree of polarization of light reflected by ocean planets with a \textit{patchy} cloud cover.
-They showed that the alternation of cloud-free and cloud-covered glint, as the cloud pattern changes in time, can result in strong fluctuations in the degree of polarization phase curves at continuum wavelengths (i.e.
-at wavelengths at which the light is not absorbed by gases in the exoplanet's atmosphere), with generally high polarization
-when the glint is visible and low polarization when the glint is covered.
-Fluctuations of the polarization continuum phase curves are also expected when continents rotate over the glint \citep{groot2020}.
-If the telescope's integration time is too long to resolve those temporal variations, the fluctuations will results in noisier phase curves.
-Nonetheless, the polarization phase curves of partially cloudy planets with oceans at optical and near infrared wavelengths intersect \citep{treesandstam2019}, providing a possible avenue to distinguish between an exo-ocean and forward scattering clouds.
-
-\textit{<Observing the Earth as an exoplanet>}
-Measurements of Earthshine (i.e.
-the reflected sunlight by the Earth incident on the dark side of the Moon) indeed reported higher degrees of polarization in the continuum at long wavelengths than at short wavelengths, which could be attributed to the ocean glint \citep{Emde2017,sterzik2019,takahashi2021}.
-
-%\textit{\citet{sterzik2019} and \citet{Takahashi2021} used polarimetric Earthshine 
-%measurements to detect the presence of the ocean, proving that the method 
-%works in practice for a spatially unresolved world.}
-
-The optical properties of an Earth-like exoplanet orbiting its host star may vary along the phase curve, e.g., because of the planet's seasons and changing weather systems in time.
-Recent simulations by \citet{trees2022} showed that signatures of the glint could also be found in the degree of polarization spectra (i.e., as functions of wavelength with 1 nm resolution), measured at a single phase angle.
-Taking into account the absorption of light by gases in the exoplanet's atmosphere, they computed that the degree of polarization of ocean planets is lower across gaseous absorption bands with respect to the continuum, while for dry planets, it is higher \citep[see][]{stam2008}.
-Those dips, instead of peaks, can already be searched for at intermediate phase angles ($90^\circ$-$130^\circ$), thus at the largest angular planet-star separations, and also occur for high planetary cloud fractions as long as the glint is (partially) cloud-free.
-Comparing the degree of polarization in the continuum and at a wavelength (band) where gases partly absorb the light, could potentially allow for an ocean detection with only one measurement in time.
-In this paper, we neglect absorption by gases and do not focus on phase variations of the spectra, but on the phase variations at single wavelengths in the continuum.
-
-%Phase variations of polarized light could also reveal other non-Lambertian scattering phenomena (Figure~\ref{fig:bottplot}).
-
-The degree of polarization of light scattered by spherical cloud droplets strongly depends on the scattering angle \citep[see][]{hansentravis1974}, and consequently, on the phase angle for an Earth-like exoplanet with liquid water clouds \citep{stam2008}.
-The scattering angle dependent variations of the degree of polarization for different droplet sizes are more distinct than in the total reflected flux \citep{karalidi2011}, and the sign of the degree of polarization (i.e.
-the direction of polarization) changes at various angles \citep{stam2008,trees2022}.
-In particular, the degree of polarization shows peaks at the primary and secondary rainbow angles, caused by reflection of light inside the spherical droplets, and the shapes of the peaks depend on the wavelength, particle size (distribution) and type of liquid  \citep{hansentravis1974,bailey2007,bailey2018}.
-A measurement of rainbow features in polarized light would be a strong indicator of spherically shaped (i.e.
-liquid) particles in the exoplanet's atmosphere, as they are not expected to originate from reflection by the planet's surface or non-spherical (ice) cloud particles \citep{Karalidi2012ice}.
-In the total reflected flux, thus without the separation of the polarized light, the rainbow features are virtually unobservable \citep{karalidi2011}.
-The benefits of those rainbow features in polarized light for cloud characterization on other planets than the Earth have been proven by \citet{hansenhovenier1974}, who fitted model calculations through measurements of the degree of polarization phase curves of Venus.
-%Because Venus is an inner planet as seen from the Earth, it can be observed from the Earth at phase angles ranging from almost $0^\circ$ to $180^\circ$ .
-%They retrieved that the main cloud deck of Venus consists of a 75\% sulfuric acid solution, with a droplet effective radius of 1.05 $\mu$m and effective variance of 0.07.
-
-%... glory ...\citep{2014A&A...566L...1G} ... and halos ... ice clouds ...
-
-%%% KIM's BOOKMARK FOR HERLSELF, PLEASE DO NOT DELETE %%%
-
-%\subsection{Polarized Phase Curves}
-%\textcolor{blue}{Recap all of polarized phase curve work, including stuff not about ocean glint. Introduce the polarization signals of Rayleigh scattering, rainbow and glory.}
-%compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,
-
-% add latitude albedo degeneracy break here \citep{2012ApJ...752L...3C}
-
-%This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}.
-
- %\cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds.
-
- %Phase curves of planets are more complex than simple Lambertian curves.
- %
- %The phase curve of an edge on system in unpolarized light is dominated by the reflectance and backscatter of the surface, clouds, and atmosphere.
- %
- %Though this behaviour is not Lambertian, the planet will, in general, get brighter at fuller phases (small scattering angles) with a peak at full phase from cloud backscatter \citep[see, for example,][]{kopparla2018}.
- %
- %In cases where surface liquid (sea) is present, the specular reflection can create a notable addition to the phase curve at broad scattering angles near crescent phase \citep{Robinson_2010}.
- %
- %Exoplanet phase curves can also contain a rainbow signature at gibbous phases (moderately narrow scattering angles), though this is a small feature in comparison to the reflectance.
-
- %In polarized light, additional structure in the phase curve can enable phenomena like ocean glint, rainbows, and the Rayleigh scattering to be more easily distinguished.
-Notably, in the polarized phase curve the rainbow may be an especially strong feature and the maximum reflectance signal (from Rayleigh scattering) occurs
-
-%Polarimetry provides the possibility to detect molecules, hazes, clouds and different surfaces  as they all produce unique polarization signatures which cannot be disentagled by using broadband flux alone.
-Polarized measurements of terrestrial exoplanet atmospheres therefore have the potential to provide empirical constraints on atmospheric properties that are inaccessible through other observational techniques.
-Polarimetry has previously played an important role in the understanding of solar system planet atmospheres, for instance, polarimetric measurements of Venus’ atmosphere suggested that the upper atmosphere of Venus is dominated by a haze of liquid sulphuric acid droplets \citep{Venus74}, which has been confirmed by in situ observations.
-
-\subsection{Rainbows {\color{Green} [Victor: The rainbows are already discussed in previous section.]}}
-
-A well-characterized rainbow feature in the phase curve of a planet provides a wealth of information about the atmosphere and condensates.
-The rainbow feature is visible in both polarized and unpolarized phase curves, but is particularly prominent in polarized light because it is strongly polarized and the fractional polarization from other sources like Rayleigh scattering is reduced at narrow scattering angles where many rainbows occur.
-The location of the rainbow peak in the phase curve shifts depending on the refractive index of the condensate and thus will shift depending on the species, and for most species the peak will also shift slightly with wavelength.
-The relative height of the rainbow compared to the Rayleigh peak with wavelength provides evidence of the depth of the cloud top height.
-The size distribution of droplets also effects the polarized rainbow and other optical effects.
-% Shape/phase
-% 
-% cite Bailey
-
-As a characterization tool the rainbow and other cloud optical features are vital, and in studies of habitability they can provide evidence that water is the dominant condensate providing context to the glint feature.
-In this work we identify our ideal rainbow angle as %...
-
-In our own Solar System the additional information inherent in condensate polarization allowed the identification of the species and droplet size distribution of the clouds on Venus \citep{Hansen1974}, the discernment between a snowy surface and hazy thick atmosphere on Titan \citep{Zellner1973}, and may have eliminated some prospects for the unknown UV absorber on Venus \citep{Petrova2018}.
-
-%\cite{rossi17} computed flux and polarisation of reflected starlight for different types of cloud cover on Earth-like planets and determined different types of cloud cover can be distinguished from each other.
-%Alternative options or papers here....
-Kim can write this tomorrow , needs other Stam, Karalidi, Kopparla, maybe Gordon, need to reference rainbows in Bailey paper.
-Look at citations in Gordon paper (our bib)...
-he has a really thorough lit review --KB
 
 \subsection{Habitable Worlds Observatory}
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
+
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 habitable zone planets.
 %
-This assertion requires an instrument with a coronograph capable of high contrast imaging in the optical-near infrared.
+This assertion requires an instrument with a coronograph capable of high contrast imaging at optical to near infrared wavelengths.
 %
@@
 %
-The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
+The Inner Working angle (\IWA) of the resulting telescope will depend on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
 
-Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \SIrange{200}{2000}{\nano\meter} instrument for exoplanet characterisation, with a similar coronagraph operating between \num{450} and \SI{1800}{\nano\meter} planned for HabEx.
+The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \SIrange{200}{2000}{\nano\meter} instrument for exoplanet characterisation, with a similar coronagraph operating between \num{450} and \SI{1800}{\nano\meter} planned for HabEx.
 %
-POLLUX, a proposed UV instrument for the LUVOIR study could also allow the detection of polarised light from hot Jupiters at \SI{300}{\nano\meter} \citep{Bouret2018_pollux}.
+POLLUX, a proposed UV instrument for the LUVOIR study could also detect  polarised light from hot Jupiters at \SI{300}{\nano\meter} \citep{Bouret2018_pollux}.
 %
-However, the challenges associated with the design of a coronagraph in UV are enormous, and it is unlikely that \hwo will include a UV high-contrast instrument sensitive to Earth-like planets.
+However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \hwo will include a UV high-contrast instrument sensitive to Earth-like planets.
 
@@
 
-In this paper, we quantify the phase angles that are accessible for direct
-observations for hypothetical terrestrial planets orbiting in the habitable 
-zones of HabWorlds target stars. 
+In this paper, we quantify the planetary phase angles that would be 
+accessible for direct observations for hypothetical terrestrial planets 
+orbiting in the habitable zones of HabWorlds target stars. 
 In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the 
@@
 
+In this section we consider the number of stars around which different scattering phenomena could be observable for different \IWA.
+We first define the phase angles over which various these phenomena are accessible, in Table~\ref{tab:phase_ranges}.
+We note that these ranges should all be considered approximate, as the exact phase angles where these effects are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
+Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
+
 \cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
@@
 The smaller the \IWA, the more features we will be observable and in more systems.
+
+\begin{table}
+    \centering
+    \caption{
+        The ranges of phase angles for the different scattering phenomena considered here.
+    }
+    \label{tab:phase_ranges}
+    \begin{tabular}{ c c c c } 
+    \toprule
+     & $\varphi_{min}$ & $\varphi_{peak}$ & $\varphi_{max}$ \\
+    \midrule
+    \midrule
+    Rainbows & 22 & 42 & 63 \\
+    Rayleigh & 50 & 70 & 110 \\
+    Ocean Glint & 130 & 150 & 170 \\
+    Glory & 0 & 5 & 10 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+    % features = {r'Rainbow': (127,138,158),
+    %         r"Rayleigh": (90,110,130),
+    %         r"Ocean Glint": (50,30,10),
+    %         r"Glory": (10,5,0),
+    %         }
 
@@
 \todo{maybe we want a table of some numbers in this section}
+
+\begin{table}
+    \centering
+    \caption{
+        Number of stars around which the peak phase angle of each phenomena is detectable, expressed in $mas$ and $\lambda/D$ for $\lambda=600$nm. 
+    }
+    \label{tab:phase_ranges}
+    \begin{tabular}{ c c c c } 
+    \toprule
+      & \multicolumn{3}{c}{IWA}
+    \\
+     & 21$mas$ & 41$mas$ & 62$mas$
+    \\
+     & $2\lambda/D$ & $3\lambda/D$ & $4\lambda/D$ 
+    \\
+    \midrule
+    \midrule
+    Rainbows & 90 & 42 & 18 \\
+    Rayleigh & 152 & 115 & 65 \\
+    Ocean Glint & 43 & 14 & 5 \\
+    % Glory & 0 & 5 & 10 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 
@@
 \affiliationtarget{28}{$^{28}$} NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\
-\affiliationtarget{29}{$^{29}$} ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland
+\affiliationtarget{29}{$^{29}$} ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland \\
+\affiliationtarget{30}{$^{30}$} Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\
+\affiliationtarget{32}{$^{31}$} Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, the Netherlands \\
+\affiliationtarget{31}{$^{32}$} Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\
 \endgroup

### 2023-04-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Kyle van Gorkom$^{15}$,
+    \newauthor
     Maaike van Kooten$^{11}$ 
@@
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
-The glint signature at scattering phases of $\sim$50--70$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
+The glint signature at planetary phase angles of 
+%$\sim$50--70$^\circ$ 
+$\sim$110--130$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
 Increasing the \IWA to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
@@
 In the total flux the signal of the glint will be most prominent at small planetary phase (large scattering) angles.
-The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53 degrees (a scattering angle of 127 degrees) \citep[see, e.g.
+The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53 degrees  \citep[a scattering angle of 127 degrees; see e.g.
 ][]{2008Icar..195..927W}.
@@
 In this paper we define our optimal angle for rainbow detection as the peak of the water droplet rainbow in visible wavelengths.
-However, extending observations to angles closer to the star (smaller scattering angles, larger phase angles) would allow the full shape of the rainbow to be characterized and ensure the characterization of rainbows from other possible HZ condensates.
+However, extending observations to angles closer to the star (smaller scattering angles, i.e.\ larger phase angles) would allow the full shape of the rainbow to be characterized and ensure the characterization of rainbows from other possible HZ condensates.
 The glory is produced when the droplet radii much smaller than the wavelength of light and can allow further characterization as their angle, intensity, and shape also vary with the complex refractive index but primarily occur within 20 degrees scattering angle very close to full phase.

### 2023-04-16 src/tex/ms.tex

@@
 Surface liquid water is central to the definition of planetary habitability.
+%
 Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
+%
 Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
+%
 The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
+%
 Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
+%
 We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
+%
 We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
+%
 The glint signature at planetary phase angles of 
@@
 $\sim$110--130$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
+%
 Increasing the \IWA to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.
-\end{abstract}
+%
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.     \end{abstract}
 
@@
 
-Signatures of habitability like ocean glint and optical effects from condensates--- like the rainbow---should be some of the most distinct signals in an exoplanet's phase curve, particularly in polarized light.
+Signatures of habitability, from ocean glint to optical effects from condensates --- like the rainbow---should be some of the most distinct signals in an exoplanet's phase curve, particularly in polarized light.
+%
 However, depending on system architectures, some of these features can be obscured by the inner working angle of a coronagraph.
@@
 Surface liquid water is closely related to planetary habitability because it is essential for life as we know it.
+%
 To first order, the presence of liquid surface water can be predicted by the flux a planet receives from its star.
+%
 The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
+%
 However, there are many factors that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and geological activity. 
+%
 Establishing the presence of liquid surface water on an exoplanet therefore provides important context for interpreting possible biosignatures.
+%
 Providing additional context to the habitability of an exoplanet is the presence of a water-cycle as evidenced by the characterization of a rainbow and other cloud optical effects.
@@
 \subsection{Exoplanet Phase Curves}
+
 An exoplanet observed in an edge-on orbit will move through all phases and thus all scattering angles, while those viewed at higher inclinations will pass through a subset of these scattering angles.
+%
 If the planet is rotating, this supplies additional regular variations that can be used to place constraints on the heterogeneity of the planet and potentially map it.
@@
 The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scatter, rainbows, and glories.
+%
 In polarized light an added potential benefit is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade \citep{kemp1987}.
@@
 Ocean glint is from the specular (Fresnel) reflection of light off of smooth liquid surfaces at crescent phases.
+%
 In the total flux the signal of the glint will be most prominent at small planetary phase (large scattering) angles.
+%
 The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53 degrees  \citep[a scattering angle of 127 degrees; see e.g.
 ][]{2008Icar..195..927W}.
+%
 It therefore varies slightly by wavelength and by sea and atmosphere composition.
+%
 In general on a planet the glint will be strongest in red wavelengths where the Rayleigh contribution to both the phase curve and attenuation are minimized \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also minimized \citep{2012ApJ...752L...3C}.
+%
 The strengthened glint signature at longer wavelengths has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+%
 The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across scattering angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
@@
 The glint provides evidence of surface liquid bodies, which taken in the context of the temperature and condensates may be interpreted as the presence of liquid water.
+%
 The forward scattering from clouds can be confused with glint \citep{Robinson_2010} further motivating also probing the rainbow angle for context.
+%
 The cloud coverage fraction on ocean worlds can also be constrained with observations spanning the ultraviolet to near-infrared by the point phase curves at all wavelengths intersect in polarized light \citep{treesandstam2019}.
-Additionally,
- \citet{trees2022} showed that glint can also be detected in spectropolarimetry while the planet is at phase angles placing it at the wide separations more easily probed with coronagraphs.
+%
+Additionally, \citet{trees2022} showed that glint can also be detected in spectropolarimetry while the planet is at phase angles placing it at the wide separations more easily probed with coronagraphs.
 %how use to map, 
@@
 \subsection{Rainbows and Glories}
+
 Rainbows, glories and other optical effects from condensates provide a wealth of information about an exoplanet \citep{karalidi2012rainbow, stam2008}.
+%
 In our own Solar System the polarized back scatter from Titan provided evidence that the world's reflectance was due to haze rather than a surface \citep{zellner1973polarization}, on Venus the rainbow and phase curve behavior with wavelength showed the clouds were $\sim$75\% H$_2$SO$_4$ -- water mixture also constraining the droplet size distribution \citep{hansenhovenier1974}, and the glory was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
@@
 In unpolarized light the forward scattering and backscatter from clouds can dramatically change the shape of a phase curve from Lambertian.
+%
 While the glory and rainbow are technically visible in any phase curve, their strongly polarized nature coupled with the suppression of signal from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
+%
 The rainbow angle depends on the refractive index and therefore varies with wavelength and depending on the species.
+%
 The feature being far more narrow than the glint allows more exact determination of the species.
+%
 The droplet size distribution also effects the shape of the rainbow and clouds of species in a solid phase can diverge from the Mie scattering that produces these features greatly, producing other optical effects particularly at narrow scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
+%
 In this paper we define our optimal angle for rainbow detection as the peak of the water droplet rainbow in visible wavelengths.
+%
 However, extending observations to angles closer to the star (smaller scattering angles, i.e.\ larger phase angles) would allow the full shape of the rainbow to be characterized and ensure the characterization of rainbows from other possible HZ condensates.
+%
 The glory is produced when the droplet radii much smaller than the wavelength of light and can allow further characterization as their angle, intensity, and shape also vary with the complex refractive index but primarily occur within 20 degrees scattering angle very close to full phase.
+%
 %The glory (in the backscattering direction) and diffraction peak (in the forward scattering direction) are distinct features in the total reflected flux, but not in the degree of polarization.
@@
 %
-This assertion requires an instrument with a coronograph capable of high contrast imaging at optical to near infrared wavelengths.
+This assertion requires an instrument with a coronagraph capable of high contrast imaging at optical to near infrared wavelengths.
 %
@@
 
-In this paper, we quantify the planetary phase angles that would be 
-accessible for direct observations for hypothetical terrestrial planets 
-orbiting in the habitable zones of HabWorlds target stars. 
-In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the 
-range of phase angles given a few different limits, in \S 2.3, we describe 
-Monte Carlo simulations to marginalize over the unknown orbital inclination and 
-eccentricity of planets in these systems, and in \S 2.4, we outline how the 
-planet/star contrast changes as a function of the planet's orbital phase.  
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars.
+%
+In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the range of phase angles given a few different limits, in \S 2.3, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
+%
 We present and discuss our results in \S 3 and conclude in \S 4. 
@@
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\hwo\ is still in the early stages of development, and the exact 
-telescope and instrument design is as of yet undecided. 
-In this work, we assume a \SI{6}{\meter} primary mirror, which is the 
-currently favoured design for the observatory.
-In addition, the observatory will utilise a coronagraph to suppress the 
-stellar light and facilitate high-contrast observations of faint 
-companying exoplanet. 
-The \IWA\ of the coronagraph is currently undecided and will be driven 
-by both the science requirements and technological limitations. 
+\hwo\ is still in the early stages of development, and the exact telescope and instrument design is as of yet undecided. 
+%
+In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
+%
+The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint accompanying exoplanet. 
+%
+The \IWA{} of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
 %In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
@@
 For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations
-corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at representative visible and near infrared wavelengths of $\lambda = \SI{600}{\nano\meter}$ and $\SI{1}{\micro\meter}$.
+For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at representative visible and near infrared wavelengths of $\lambda = \SI{600}{\nano\meter}$ and $\SI{1}{\micro\meter}$.
 
@@
 
-\hwo\ is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
+\hwo\ is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.
+%
 \footnote{\url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
@@
 This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}. 
+%
 The authors of the report determined that these are the nearby stars (maximum distance \SI{25}{\parsec}) for which exo-Earths would be the most observable for an imaging survey of habitable zones with a \SI{6}{\meter} telescope.
@@
 In what follows, we consider hypothetical Earth-size planets orbiting at the right distance from each star to receive the same flux as the Earth receives from the Sun.
+%
 By construction, such planets should be observable at quadrature phase by HabWorlds.
@@
 %
-The range of scattering phase angles that can be observed will depend on the \IWA\ of the coronagraph as well as the shape and orientation of the exoplanet's orbit.
+The range of scattering phase angles that can be observed will depend on the \IWA{} of the coronagraph as well as the shape and orientation of the exoplanet's orbit.
 
@@
         An illustration of an edge-on circular orbit showing that the coronagraph can obscure the exoplanet at high and low scattering phase angles.
+        %
         This can prevent observations of scattering phenomena which could be used to indicate the habitability and atmospheric composition of the exoplanet. 
@@
         The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major $a$ at a distance $d_*$.
+        %
         The on-sky angular separation as viewed by the observer is $\delta$.
+        %
         The scattering phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.  
@@
         Scatter plot for the target sample, showing stellar effective temperature and stellar distance.
+        %
         The size of the points represents the angular separation of the star and planet in milliarcseconds as presented in the target list.
+        %
         The colour of the points shows $\Delta \varphi$ assuming a circular edge-on orbit at a semi-major axis corresponding to an Earth-like instellation for an IWA of 62 mas.
+        %
         Additionally, the colour bar indicates the atmospheric phenomenon that can be detected where phenomena from the bottom of the colour bar up to the indicated colour are detectable for that system.
+        %
         Thus, dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
@@
 
-(1)~If the inclination is relatively face-on, then the coronagraph mask does not obscure any part of the sky-projected orbit, and the maximum scattering phase angle coverage depends solely on the disk inclination ;
-
-(2)~When parts of the orbit  are hidden by the focal plane mask, the scattering phase angle coverage only depends on the \IWA\ and orbital semi-major axis.
+\begin{enumerate}
+    \item If the inclination is relatively face-on, then the coronagraph mask does not obscure any part of the sky-projected orbit, and the maximum scattering phase angle coverage depends solely on the disk inclination ;
+    \item When parts of the orbit  are hidden by the focal plane mask, the scattering phase angle coverage only depends on the \IWA{} and orbital semi-major axis. 
+\end{enumerate}
 
@@
         The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle.
+        %
         If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
+        %
         For orbital inclinations close to edge-on (bottom panels) the planet goes through all phases, but the scattering phase angles near 0$^\circ$ and 180$^\circ$ are not accessible due to obscuration by the coronagraph mask; for circular orbits, the accessible scattering phases are $\phi \in 90^\circ \pm \arcsin({\rm IWA}\cdot d_*/a)$.
@@
 
-For each star on the \hwo\ target list, we generate $1000$ \todo{I think this is what the code says, check} hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
+For each star on the \hwo\ target list, we generate $1000$  hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
 %
@@
         Cumulative distributions of the most extreme scattering phase angle accessible for different \IWA where the solid lines are for randomly inclined circular orbits and the dashed lines are for randomly orientated elliptical orbits.
+        %
         The top x axis indicates the minimum scattering phase angle and the bottom x axis indicates the maximum.
+        %
         These are symmetric about quadrature (90 degrees) \todo{is this true for elliptical orbits?}.
+        %
         The y axis indicates the number of systems divided by the number of Monte Carlo samples and is thus normalised to the number of systems in the target list.
+        %
         The secondary y axis indicates the number of systems assuming only 24 percent of them contain an Earth-like exoplanet in the habitable zone.
@@
 We then calculated the minimum and maximum accessible scattering phase angles for of \IWA's of 21, 42, 63 and 84 mas which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\SI{600}{\nano\meter}$.
+%
 The dashed lines in \cref{fig:betaallofit} shows the normalised cumulative distribution of the accessible angles for each of the \IWA, when accounting for orbital eccentricity.
@@
 We also calculate which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+%
 The cumulative distribution over \IWA\ of the number of systems, normalised to the total number of systems, where each scattering phenomena is visible is shown in \cref{fig:accessible_phase_angles}.
@@
 \end{equation}
-where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
+where $p$ is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
 %
@@
     The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of $90^\circ$ for the stars $\alpha$ Cen A, $\epsilon$ Eri and Lalande 21185.
+    %
     The solid line indicates the contrast in unpolarised light with the contrast at quadrature marked by a solid dot.
+    %
     The polarised component is indicated by the colored dots for which the color represents the scattering phase angle from quadrature.
+    %
     The light grey points show the quadrature contrasts of the other targets in the star list and the dashed lines indicate $1,2$ and $3$ times the \IWA\ for \hwo.
@@
 
-\subsection{The affect of eccentricity}
+\subsection{The effect of eccentricity}
 \label{sec:result_eccentricity}
 As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable affect on the number of systems that can be observed at each scattering phase angle.
-This is because, Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA\ and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
-The eccentricity does not significantly effect the number of systems for smaller inner working angles as the systems in \hwo\ target list where chosen based on their on-sky separation quadrature and therefore the shape of the orbit has little affect for \IWA\ smaller than cut used in the selection.
-It is, however, important to account for orbital eccentricity when the \IWA\ of the coronoagraph is larger than the \IWA\ used to filter the target list otherwise the number of systems observable at extreme phase angles would be under predicted.
+%
+This is because, Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA{} and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
+%
+The eccentricity does not significantly effect the number of systems for smaller inner working angles as the systems in \hwo{} target list where chosen based on their on-sky separation quadrature and therefore the shape of the orbit has little affect for \IWA{} smaller than cut used in the selection.
+It is, however, important to account for orbital eccentricity when the \IWA{} of the coronoagraph is larger than the \IWA{} used to filter the target list otherwise the number of systems observable at extreme phase angles would be under predicted.
 
@@
 
-In this section we consider the number of stars around which different scattering phenomena could be observable for different \IWA.
+In this section we consider the number of stars around which different scattering phenomena could be observable for different \IWA{}.
+%
 We first define the phase angles over which various these phenomena are accessible, in Table~\ref{tab:phase_ranges}.
+%
 We note that these ranges should all be considered approximate, as the exact phase angles where these effects are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
-Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
+%
+Nonetheless, this exercise can provide valuable insights on the effects of \IWA{} choice.
 
-\cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
-The \IWA\ of \hwo's coronograph will depend on the primary mirror diameter, observing wavelength and how well a coronograph of a given size can suppress the glare of the host star.
-The smaller the \IWA, the more features we will be observable and in more systems.
+\cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA{}.
+%
+The \IWA{} of \hwo's coronagraph will depend on the primary mirror diameter, observing wavelength and how well a coronagraph of a given size can suppress the glare of the host star.
+%
+The smaller the \IWA{}, the more features we will be observable and in more systems.
 
@@
 The Rayleigh scattering feature appears near quadrature and is therefore observable in the most systems.
+%
 This feature is caused by atmospheric gases and can therefore be used to determine if a planet has an atmosphere.
+%
 The rainbow feature appears further from quadrature but will still be observable in a significant fraction of systems.
+%
 This feature will only be present if the atmosphere contains spherical droplets, potentially indicating the presence of liquid water.
+%
 The position and shape of the peak is sensitive to the properties of the droplets allowing their nature to be determined if the peak of the rainbow feature can be reached.
-The ocean glint feature occurs even further from quadrature and so is obserable in fewer systems.
+%
+The ocean glint feature occurs even further from quadrature and so is observable in fewer systems.
 In most systems, only the first half of the feature is observable but this can still be used to identify a liquid ocean on the surface of the planet.
-A key indicator of habitabitily.
-Finally, it is possible to view the glory scattering feature in one system, Alpha Cen A.
+%
+A key indicator of habitability. Finally, it is possible to view the glory scattering feature in one system, Alpha Cen A.
+%
 \todo{maybe we want a table of some numbers in this section}
@@
     }
-    \label{tab:phase_ranges}
+    \label{tab:nstars_detect}
     \begin{tabular}{ c c c c } 
@@
 The accessible scattering phase angles is not the factor that needs to be considered when determining what scattering features are observable.
+%
 The contrast of the planet is also an important factor to consider.
-\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo\ target list assuming egde-on orbits.
+%
+\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo\ target list assuming edge-on orbits.
+%
 The over all contrast varies significantly for each system and throughout the orbit.
-The observability of these planets will depend on the star light suppression of the coronograph which is influenced by its \IWA.
-It is not possible to predict how much the state of the art of coronaography will have advanced by the time \hwo\ is built.
-However, when its design is being finalised, it will be important to consider both the accessible scattering phase angles and contrast limits when deciding on the \IWA\ of the coronograph, if these features are to be observed.
+%
+The observability of these planets will depend on the star light suppression of the coronagraph which is influenced by its \IWA{}.
+%
+It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \hwo{} is built.
+%
+However, when its design is being finalised, it will be important to consider both the accessible scattering phase angles and contrast limits when deciding on the \IWA{} of the coronagraph, if these features are to be observed.
 
@@
         Random examples of the eccentric orbits generated for the stellar sample.
+        %
         The orbits are scaled by the Earth-equivalent flux distance. 
+        %
         The concentric circles marked by the dashed lines indicate inner working angles of 20.6, 41.3, 61.9, and \SI{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$ respectively (assuming $\lambda = \SI{600}{\nano\meter}$ and $D = \SI{6}{\meter}$).
+        %
         The figure illustrates that the \IWA\ can significantly affect the range of scattering phases observable with each orbit.
@@
     \caption{
-        The number of systems for which key scattering features (Rayleigh scattering, rainbows, ocean glint and glories) can be observed as a function of \IWA.
+        The number of systems for which key scattering features (Rayleigh scattering, rainbows, ocean glint and glories) can be observed as a function of \IWA{}.
+        %
         Each subplot contains three lines where the dashed line shows the number of systems where the start of the scattering feature can be observed, the solid line shows the number of systems that reach the peak of the feature and the dot-dashed line the number where the scattering angles of the end of the feature are observable.
-The bottom x axis gives the \IWA\ of the coronograph in milliarcseconds which is converted to $\lambda/D$ for $\lambda=\SI{600}{\nano\meter}$ and $D=\SI{6}{\meter}$ for the top x axis.
-The secondary y axis indicates the number of systems that each feature would be observable in assuming only 24 percent of them contain an Earth-like exoplanet in the habitable zone.
+        %
+        The bottom $x$ axis gives the \IWA{} of the coronograph in milliarcseconds which is converted to $\lambda/D$ for $\lambda=\SI{600}{\nano\meter}$ and $D=\SI{6}{\meter}$ for the top $x$ axis.
+        %
+        The secondary $y$ axis indicates the number of systems that each feature would be observable in assuming only 24\% of them contain an Earth-like exoplanet in the habitable zone.
+        %
         It is worth noting that the Glory feature is only observable in one system, Alpha Cen A. 
@@
 %
-In this work, we investigated the possibility of using \hwo\ to detect scattering phenomena on terrestrial exoplanets, including ocean glint, which could establish a planet's habitability.
+In this work, we investigated the possibility of using \hwo{} to detect scattering phenomena on terrestrial exoplanets, including ocean glint, which could establish a planet's habitability.
 
-We primarily focused on the range of scattering phase angles and therefore which scattering phenomena could potentially be observed given the limitation of the coronagraphic \IWA.
+We primarily focused on the range of scattering phase angles and therefore which scattering phenomena could potentially be observed given the limitation of the coronagraphic \IWA{}.
 %
-Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \hwo\ will be.
+Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \hwo{} will be.
 %
-Assuming a \SI{60}{\mas} \IWA, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$55 systems, and ocean glint in $\sim$20 systems.
+Assuming a \SI{60}{\mas} \IWA{}, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$55 systems, and ocean glint in $\sim$20 systems.
+%
 The habitability of Earth analogs in the remaining 140 systems would have to be established by less direct means like rotational mapping to identify continents and oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
@@
 
-The number of habitable zones that can be probed by \hwo\ increases with decreasing \IWA. 
+The number of habitable zones that can be probed by \hwo{} increases with decreasing \IWA{}. 
 %
-The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic frequency of such planets, will drive the choice of coronagraphic \IWA\ for this mission.
+The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic frequency of such planets, will drive the choice of coronagraphic \IWA{} for this mission.
 %
-We have shown that only $\sim$10\% of Earth-like exoplanets imaged with \hwo\ could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
+We have shown that only $\sim$10\% of Earth-like exoplanets imaged with \hwo{} could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
 %
@@
 
-The NASA ExEP target list containing the data used for these simulations is available online.% 
+The NASA ExEP target list containing the data used for these simulations is available online.
+%
 \footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
@@
 \section*{Acknowledgements}
+
 The workshop on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, Netherlands.
+%
 This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).
-
+%
 This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
-
+%
 SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
+%
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
+%
 TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
+%
 Specific grant funding and any other specific requirements
+%
 KMB acknowledges support from NASA Habitable
+%
 Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.

### 2023-04-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The glint signature at planetary phase angles of 
-%$\sim$50--70$^\circ$ 
-$\sim$110--130$^\circ$ would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
+$\sim$50--70$^\circ$ 
+%\sim$110--130$^\circ$ 
+would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
 %
@@
 %
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA.     \end{abstract}
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.    \end{abstract}
 
@@
    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-   \caption{BOTT PLOT this is a holder until Kim updates the figure}
+   \caption{An illustration of phase curves for two homogeneous planets in an edge on system with oceans and an Earth-like atmosphere.
+The cloudy scenario includes an Earth-like water cloud deck at low altitudes.
+Below the plot, the features that the phase angles and scattering angles are most sensitive to are shown.
+In reality, as can be seen by the examples above, these features overlap.
+Throughout the paper we approximate feature sensitivity ranges as those that capture the rise an fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (scattering angles: glint $\sim$50--70$^\circ$; rainbow $\sim$140--160$^\circ$}; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest scattering angles).
     \label{fig:bottplot}
@@
 %
-However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \hwo will include a UV high-contrast instrument sensitive to Earth-like planets.
+However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \hwo{} will include a UV high-contrast instrument sensitive to Earth-like planets.
 
@@
 
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars.
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds (HWO) target stars.
 %
@@
 We present and discuss our results in \S 3 and conclude in \S 4.
+It is important to note that the phases for each feature throughout this paper are approximate, that features overlap, and that the maximum peak can vary with composition and wavelength.
 %the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
@@
 
-\subsection{\hwo\ and Stellar Sample}
+\subsection{\hwo{} and Stellar Sample}
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\hwo\ is still in the early stages of development, and the exact telescope and instrument design is as of yet undecided. 
+\hwo{} is still in the early stages of development, and the exact telescope and instrument design is as of yet undecided. 
 %
@@
 The \IWA{} of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
-%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
+%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA{} of the observatory. 
 %We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
@@
 
-\hwo\ is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.
+\hwo{} is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.
 %
@@
         This can prevent observations of scattering phenomena which could be used to indicate the habitability and atmospheric composition of the exoplanet.
+The phase angle, $\varphi$, is shown in red and is the complement to the scattering angle in edge on systems.
         %\timmy{Maybe mention that this is a simplified illustration with a circular orbit, plus we are basically looking at an edge-on system?}
@@
 
-\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo\ target list, assuming an IWA of \SI{62}{\mas} ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
+\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo{} target list, assuming an IWA of \SI{62}{\mas} ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
 \timmy{60 (as in the abstract), or 62?}
@@
 
-For each star on the \hwo\ target list, we generate $1000$  hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
+For each star on the \hwo{} target list, we generate $1000$  hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
 %
@@
 %
-The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible scattering phase angles normalised to the total number of systems in \hwo\ target list, depending on the choice of inner working angle.
+The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible scattering phase angles normalised to the total number of systems in \hwo{} target list, depending on the choice of inner working angle.
 
@@
 %
-The cumulative distribution over \IWA\ of the number of systems, normalised to the total number of systems, where each scattering phenomena is visible is shown in \cref{fig:accessible_phase_angles}.
+The cumulative distribution over \IWA{} of the number of systems, normalised to the total number of systems, where each scattering phenomena is visible is shown in \cref{fig:accessible_phase_angles}.
 %
@@
     %
-    The light grey points show the quadrature contrasts of the other targets in the star list and the dashed lines indicate $1,2$ and $3$ times the \IWA\ for \hwo.
+    The light grey points show the quadrature contrasts of the other targets in the star list and the dashed lines indicate $1,2$ and $3$ times the \IWA{} for \hwo.
     }
@@
 %
-\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo\ target list assuming edge-on orbits.
+\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo{} target list assuming edge-on orbits.
 %
@@
         %
-        The figure illustrates that the \IWA\ can significantly affect the range of scattering phases observable with each orbit.
+        The figure illustrates that the \IWA{} can significantly affect the range of scattering phases observable with each orbit.
     }

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\newcommand{\rainbows}{%
+\newcommand{\rainbows}{% MAK: KEEP this formatting for MNRAS submission. It deserves this!
     \textcolor{C0}{r}%
@@
 %
-The glint signature at planetary phase angles of 
-$\sim$50--70$^\circ$ 
+The glint signature at planetary phase angles of $\sim$50--70$^\circ$ 
 %\sim$110--130$^\circ$ 
-would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA; 3$\lambda/D$ for a 6-m telescope at 600nm).
+would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA{}; 3$\lambda/D$ for a 6-m telescope at 600nm).
 %
-Increasing the \IWA to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
+Increasing the \IWA{} to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
 %
@@
 %
-The Inner Working angle (\IWA) of the resulting telescope will depend on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
+The Inner Working angle (\IWA{}) of the resulting telescope will depend on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
 
@@
 \Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo{} target list, assuming an IWA of \SI{62}{\mas} ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
-\timmy{60 (as in the abstract), or 62?}
+%\timmy{60 (as in the abstract), or 62?}
 %
@@
         \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
-        Cumulative distributions of the most extreme scattering phase angle accessible for different \IWA where the solid lines are for randomly inclined circular orbits and the dashed lines are for randomly orientated elliptical orbits.
+        Cumulative distributions of the most extreme scattering phase angle accessible for different \IWA{} where the solid lines are for randomly inclined circular orbits and the dashed lines are for randomly orientated elliptical orbits.
         %
@@
 Eccentricities were drawn from the beta distribution, and inclinations were uniform in $\cos i$.
-\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA's
+\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA{}'s
 %
@@
 %
-We then calculated the minimum and maximum accessible scattering phase angles for of \IWA's of 21, 42, 63 and 84 mas which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\SI{600}{\nano\meter}$.
+We then calculated the minimum and maximum accessible scattering phase angles for of \IWA{}'s of 21, 42, 63 and 84 mas which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\SI{600}{\nano\meter}$.
 %
@@
     \\
-     & 21$mas$ & 41$mas$ & 62$mas$
+     & 21\,$mas$ & 41\,$mas$ & 62\,$mas$
     \\
@@
 Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
+%
+This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
 

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
+In this work, we assume a \SI{6}{\meter} primary mirror, as suggested in the The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
 %
@@
 \end{equation}
-% \begin{equation}
-%     \label{eq:Delta_phi}
-%     \Delta \varphi 
-%     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
-%     =  2(\varphi_\mathrm{max} - 90) 
-%     =  2(\varphi_\mathrm{min} + 90)
-% \end{equation}
-
-% {\color{Green} Victor Trees: Do you mean:
-%INDEED, THERE MUST HAVE BEEN A TYPO, THX!
-% %
-% \begin{equation}
-%     \Delta \varphi 
-%     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
-%     =  2(\varphi_\mathrm{max} - 90^\circ) 
-%     =  2(90^\circ-\varphi_\mathrm{min} ) \notag
-% \end{equation}
-%
-% }
 
@@
 
-\begin{figure*}[t]
+\begin{figure*}%[t]
     \centering
@@
         %
-        These are symmetric about quadrature (90 degrees) \todo{is this true for elliptical orbits?}.
+        These are symmetric about quadrature (90 degrees).
         %
@@
 \subsection{Simulating contrast curves}
-\todo{Kim and David: is this done?}
 
@@
 %
-The eccentricity does not significantly effect the number of systems for smaller inner working angles as the systems in \hwo{} target list where chosen based on their on-sky separation quadrature and therefore the shape of the orbit has little affect for \IWA{} smaller than cut used in the selection.
+The eccentricity does not significantly effect the number of systems for \IWA{} smaller than cut used in the selection of the systems in \hwo{} target list.
 It is, however, important to account for orbital eccentricity when the \IWA{} of the coronoagraph is larger than the \IWA{} used to filter the target list otherwise the number of systems observable at extreme phase angles would be under predicted.
@@
 
-The Rayleigh scattering feature appears near quadrature and is therefore observable in the most systems.
-%
-This feature is caused by atmospheric gases and can therefore be used to determine if a planet has an atmosphere.
+The Rayleigh scattering feature, which can used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
 %
 The rainbow feature appears further from quadrature but will still be observable in a significant fraction of systems.
-%
-This feature will only be present if the atmosphere contains spherical droplets, potentially indicating the presence of liquid water.
 %
@@
 The ocean glint feature occurs even further from quadrature and so is observable in fewer systems. 
+%
 In most systems, only the first half of the feature is observable but this can still be used to identify a liquid ocean on the surface of the planet.
@@
 A key indicator of habitability. Finally, it is possible to view the glory scattering feature in one system, Alpha Cen A.
-%
-\todo{maybe we want a table of some numbers in this section}
 
@@
     \caption{
-        Number of stars around which the peak phase angle of each phenomena is detectable, expressed in $mas$ and $\lambda/D$ for $\lambda=600$nm. 
+        Number of stars around which the peak phase angle of each phenomena is detectable, expressed in $mas$ and $\lambda/D$ for $\lambda=600$nm and  $D=6$m. 
     }
@@
 
-The accessible scattering phase angles is not the factor that needs to be considered when determining what scattering features are observable.
+The accessible scattering phase angles is not the only factor that needs to be considered when determining what scattering features are observable.
 %
@@
 However, when its design is being finalised, it will be important to consider both the accessible scattering phase angles and contrast limits when deciding on the \IWA{} of the coronagraph, if these features are to be observed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsection{The glory of Alpha Cen}%Halelujah
-%\label{sec:ealpha-cen}
-%\textbf{Glory science, what it tells us (Kim)}
-% Kim will write para on science of glory (what tells us, vaguely why appears here)
 
@@
 
-%\begin{figure*}
-%    \centering
-%    \includegraphics[width=0.6\textwidth]{figures/features_plot_peak.png}  
-%    \caption{
-%        Number of systems vs inner working angle
-%    }
-%    \label{fig:accessible_phase_angles}
-%\end{figure*}
-
 \begin{figure*}
@@
 
-%\subsection{Contrast Curves}
-%\label{sec:contrast}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-Assuming a \SI{60}{\mas} \IWA{}, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$55 systems, and ocean glint in $\sim$20 systems.
+Assuming a \SI{62}{\mas} \IWA{}, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$42 systems, and ocean glint in $\sim$14 systems.
 %
-The habitability of Earth analogs in the remaining 140 systems would have to be established by less direct means like rotational mapping to identify continents and oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
+The habitability of Earth analogs in the remaining 146 systems would have to be established by less direct means like rotational mapping to identify continents and oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
 %
-
 The number of habitable zones that can be probed by \hwo{} increases with decreasing \IWA{}. 
@@
 %
-We have shown that only $\sim$10\% of Earth-like exoplanets imaged with \hwo{} could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
+We have shown that with a $6$m primary mirror and \IWA $= 3\lambda/D$ at $600$nm coronograph, only $\sim$10\% of Earth-like exoplanets imaged with \hwo{} could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
 %
@@
 %
+This could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronograph.
+%
 While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
 %
-This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph.
+This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
 

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (HWO), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
+NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
@@
 
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds (HWO) target stars.
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
 %
@@
     \toprule
-     & $\varphi_{min}$ & $\varphi_{peak}$ & $\varphi_{max}$ \\
+     & $\varphi_\text{min}$ & $\varphi_\text{peak}$ & $\varphi_\text{max}$ \\
     \midrule

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Signatures of habitability, from ocean glint to optical effects from condensates --- like the rainbow---should be some of the most distinct signals in an exoplanet's phase curve, particularly in polarized light.
+Signatures of habitability, from ocean glint to optical effects from condensates --- like the rainbow--- should be some of the most distinct signals in an exoplanet's reflected light phase curve, particularly in polarization.
 %
@@
 
-Surface liquid water is closely related to planetary habitability because it is essential for life as we know it.
+Liquid surface water is closely related to planetary habitability because it is essential for life as we know it.
 %
@@
 %
-The habitable zone (HZ) describes the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
+The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
 %
-However, there are many factors that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and geological activity. 
+However, there are many factors besides the stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and e.g.\ geological activity.
 %
-Establishing the presence of liquid surface water on an exoplanet therefore provides important context for interpreting possible biosignatures.
+Establishing the presence of liquid surface water on an exoplanet would therefore provide important context for the interpretation of possible biosignatures.
 %
-Providing additional context to the habitability of an exoplanet is the presence of a water-cycle as evidenced by the characterization of a rainbow and other cloud optical effects.
+Providing additional context to the habitability of an exoplanet is the presence of a water-cycle as evidenced by the identification of a rainbow and other optical effects that can be attributed to water clouds.
 
@@
 %
-If the planet is rotating, this supplies additional regular variations that can be used to place constraints on the heterogeneity of the planet and potentially map it.
+If the planet is rotating, its reflected light signal is expected to show regular variations that can place constraints on the horizontal heterogeneity of the planet.
 
-The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scatter, rainbows, and glories.
+The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scatter, rainbows, and glories.
 %
-In polarized light an added potential benefit is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade \citep{kemp1987}.
+An added potential benefit of observing polarized light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
@@
 %As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
-Ocean glint is from the specular (Fresnel) reflection of light off of smooth liquid surfaces at crescent phases.
+Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces.
 %
-In the total flux the signal of the glint will be most prominent at small planetary phase (large scattering) angles.
+In the total flux the glint signal will be most prominent at small planetary phase (large scattering) angles.
 %
-The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53 degrees  \citep[a scattering angle of 127 degrees; see e.g.
+The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53$^\circ$ \citep[a scattering angle of 127 degrees; see e.g.
 ][]{2008Icar..195..927W}.
 %
-It therefore varies slightly by wavelength and by sea and atmosphere composition.
+This angle therefore varies slightly with wavelength and with sea and atmosphere composition.
 %
-In general on a planet the glint will be strongest in red wavelengths where the Rayleigh contribution to both the phase curve and attenuation are minimized \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also minimized \citep{2012ApJ...752L...3C}.
+In general, the glint on a planet will be strongest at red wavelengths where the Rayleigh scattering in the planet's atmosphere and attenuation through the atmosphere are smallest \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also small \citep{2012ApJ...752L...3C}.
 %
-The strengthened glint signature at longer wavelengths has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+The strengthened glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
 %
+{\color{red}{check the next sentence:}}
 The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across scattering angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
 
-The glint provides evidence of surface liquid bodies, which taken in the context of the temperature and condensates may be interpreted as the presence of liquid water.
+The glint provides evidence of liquid surface bodies, which taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
 %
-The forward scattering from clouds can be confused with glint \citep{Robinson_2010} further motivating also probing the rainbow angle for context.
+The forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
 %
-The cloud coverage fraction on ocean worlds can also be constrained with observations spanning the ultraviolet to near-infrared by the point phase curves at all wavelengths intersect in polarized light \citep{treesandstam2019}.
+The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
+angle at which the planet's polarization phase curves at various wavelengths intersect \citep{treesandstam2019}.
 %
-Additionally, \citet{trees2022} showed that glint can also be detected in spectropolarimetry while the planet is at phase angles placing it at the wide separations more easily probed with coronagraphs.
+While the glint feature can only be probed when a planet is at crescent phases, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry when the planet is at phase angles where it is at the wide separations more easily probed when using coronagraphs.
 %how use to map, 
@@
 
-Rainbows, glories and other optical effects from condensates provide a wealth of information about an exoplanet \citep{karalidi2012rainbow, stam2008}.
+Rainbows, glories and other optical effects from condensates provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,2014A&A...566L...1G}.
 %
-In our own Solar System the polarized back scatter from Titan provided evidence that the world's reflectance was due to haze rather than a surface \citep{zellner1973polarization}, on Venus the rainbow and phase curve behavior with wavelength showed the clouds were $\sim$75\% H$_2$SO$_4$ -- water mixture also constraining the droplet size distribution \citep{hansenhovenier1974}, and the glory was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
+In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to haze rather than a surface \citep{zellner1973polarization}, on Venus the rainbow and phase curve behavior with wavelength showed the clouds were $\sim$75\% H$_2$SO$_4$ -- water mixture also constraining the droplet size distribution \citep{hansenhovenier1974}, and the glory was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
 

### 2023-04-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     range-units = single,
-    range-phrase = {--},
-    separate-uncertainty = true,
-    multi-part-units = brackets,
-    product-units = brackets,
-    detect-weight = true,
-    per-mode = symbol,
-    mode = text,
+    range-phrase = \textup{--},
 }
@@
 \newcommand{\IWA}{\ensuremath{\mathrm{IWA}}}
+\newcommand{\OWA}{\ensuremath{\mathrm{OWA}}}
 \newcommand{\hwo}{HabWorlds}
@@
 %
-The glint signature at planetary phase angles of $\sim$50--70$^\circ$ 
+The glint signature at planetary phase angles of $\sim\qtyrange{50}{70}{\degree}$ 
 %\sim$110--130$^\circ$ 
-would be accessible to the Habitable Worlds Observatory in $\sim$14 systems, and the rainbow of water clouds at phases of 140--160$^\circ$ would be accessible in $\sim$42 systems, assuming a 62\,mas inner working angle (\IWA{}; 3$\lambda/D$ for a 6-m telescope at 600nm).
-%
-Increasing the \IWA{} to 2$\lambda/D$, increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
+would be accessible to the Habitable Worlds Observatory in $\sim\num{14}$ systems, and the rainbow of water clouds at phases of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} inner working angle (\IWA{}; $3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
+%
+Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
 %
@@
 %--------------------------------------------------------------------------------
-\begin{figure*}%[t]
+\begin{figure*}
     \centering
     \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-   \caption{An illustration of phase curves for two homogeneous planets in an edge on system with oceans and an Earth-like atmosphere.
+    \caption{
+        An illustration of phase curves for two homogeneous planets in an edge on system with oceans and an Earth-like atmosphere. 
         The cloudy scenario includes an Earth-like water cloud deck at low altitudes. 
-Below the plot, the features that the phase angles and scattering angles are most sensitive to are shown.
+        Below the plot, we are showing the features to which the phase angles and scattering angles are the most sensitive. 
         In reality, as can be seen by the examples above, these features overlap.  
-Throughout the paper we approximate feature sensitivity ranges as those that capture the rise an fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (scattering angles: glint $\sim$50--70$^\circ$; rainbow $\sim$140--160$^\circ$}; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest scattering angles).
+        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (scattering angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest scattering angles).
+    }
     \label{fig:bottplot}
@@
 
-Signatures of habitability, from ocean glint to optical effects from condensates --- like the rainbow--- should be some of the most distinct signals in an exoplanet's reflected light phase curve, particularly in polarization.
+Signatures of habitability, from ocean glint to optical effects from condensates---like the rainbow---should be some of the most distinct signals in an exoplanet's reflected light phase curve, particularly in polarization.
 %
@@
 %
-The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
-%
-However, there are many factors besides the stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and e.g.\ geological activity.
+The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
+%
+However, there are many factors besides the stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and, for example, geological activity.
 %
@@
 %
-Providing additional context to the habitability of an exoplanet is the presence of a water-cycle as evidenced by the identification of a rainbow and other optical effects that can be attributed to water clouds.
+Providing additional context to the habitability of an exoplanet is the presence of a water cycle, as evidenced by the identification of a rainbow and other optical effects that can be attributed to water clouds.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Exoplanet Phase Curves}
+\subsection{Exoplanet phase curves}
 
@@
 
-\subsection{Ocean Glint}
+\subsection{Ocean glint}
 
@@
 %
-In the total flux the glint signal will be most prominent at small planetary phase (large scattering) angles.
-%
-The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at 53$^\circ$ \citep[a scattering angle of 127 degrees; see e.g.
-][]{2008Icar..195..927W}.
+In the total flux, the glint signal will be most prominent at small planetary phase (large scattering) angles.
+%
+The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at \qty{53}{\degree} \citep[i.e., a scattering angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
 %
@@
 
-The glint provides evidence of liquid surface bodies, which taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
+The glint provides evidence of liquid surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
 %
@@
 %
-The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
-angle at which the planet's polarization phase curves at various wavelengths intersect \citep{treesandstam2019}.
+This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary phase angle depending on the cloud coverage fraction \citep{treesandstam2019}.
+%The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
+%angle at which the planet's polarization phase curves at various wavelengths intersect \citep{treesandstam2019}.
 %
@@
 
-\subsection{Rainbows and Glories}
+\subsection{Rainbows and glories}
 
@@
 %
-In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to haze rather than a surface \citep{zellner1973polarization}, on Venus the rainbow and phase curve behavior with wavelength showed the clouds were $\sim$75\% H$_2$SO$_4$ -- water mixture also constraining the droplet size distribution \citep{hansenhovenier1974}, and the glory was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
-
-In unpolarized light the forward scattering and backscatter from clouds can dramatically change the shape of a phase curve from Lambertian.
-%
-While the glory and rainbow are technically visible in any phase curve, their strongly polarized nature coupled with the suppression of signal from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
-%
-The rainbow angle depends on the refractive index and therefore varies with wavelength and depending on the species.
-%
-The feature being far more narrow than the glint allows more exact determination of the species.
-%
-The droplet size distribution also effects the shape of the rainbow and clouds of species in a solid phase can diverge from the Mie scattering that produces these features greatly, producing other optical effects particularly at narrow scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
-%
-In this paper we define our optimal angle for rainbow detection as the peak of the water droplet rainbow in visible wavelengths.
-%
-However, extending observations to angles closer to the star (smaller scattering angles, i.e.\ larger phase angles) would allow the full shape of the rainbow to be characterized and ensure the characterization of rainbows from other possible HZ condensates.
-%
-The glory is produced when the droplet radii much smaller than the wavelength of light and can allow further characterization as their angle, intensity, and shape also vary with the complex refractive index but primarily occur within 20 degrees scattering angle very close to full phase.
+In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to haze rather than a surface \citep{zellner1973polarization}, the phase curve behaviour of Venus showed not only that the clouds were $\sim$75\% H$_2$SO$_4$--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+The glory feature in Venus's phase curve was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
+
+In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
+Lambertian reflecting planet.
+%
+{\color{red}{check the sentence:}}
+While the glory and rainbow are technically visible in any phase curve
+in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
+%
+The rainbow angle depends on the refractive index of the scatterers and thus depends on the composition and the wavelength. 
+The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
+%
+Because the rainbow feature is more narrow than the glint, it would 
+allow for a more exact determination of the particle composition.
+%
+Note that the scattering by clouds that consist of solid particles, 
+such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
+%
+
+In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths.
+%
+Pushing observations to angles closer to the star (smaller scattering angles, i.e., larger phase angles) would allow the determination of the full shape of this rainbow, and enable the identification of rainbows due to the scattering of light by other possible HZ condensates.
+%
+The glory 
+%is produced when the droplet radii much smaller than the wavelength of light 
+can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within \qty{20}{\degree} scattering angle very close to full phase \citep{hansentravis1974}.
 %
@@
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 habitable zone planets.
-%
-This assertion requires an instrument with a coronagraph capable of high contrast imaging at optical to near infrared wavelengths.
-%
-Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
-%
-The precursor technology recommended by the Decadal Survey also lists ``direct imaging to probe polarized ocean glint on terrestrial planets'' as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
-%
-The performance and precise characteristics of this telescope (on-axis or off-axis, exact diameter, type of segmentation, type of coronagraph) are still to be determined.
-%
-The development will be heavily influenced by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
-%
-The Inner Working angle (\IWA{}) of the resulting telescope will depend on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
-
-The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \SIrange{200}{2000}{\nano\meter} instrument for exoplanet characterisation, with a similar coronagraph operating between \num{450} and \SI{1800}{\nano\meter} planned for HabEx.
-%
-POLLUX, a proposed UV instrument for the LUVOIR study could also detect  polarised light from hot Jupiters at \SI{300}{\nano\meter} \citep{Bouret2018_pollux}.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program, a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+%
+This requires an instrument with a coronagraph capable of high contrast imaging at optical to near infrared wavelengths.
+%
+Following the release of this survey, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
+%
+
+The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
+%
+The performance and precise characteristics of \hwo{} (on-axis or off-axis, diameter, type of segmentation, type of coronagraph) are still to be determined.
+%
+The development will be heavily informed by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
+%
+The Inner Working angle (\IWA{}) of the resulting telescope will depend on these characteristics, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
+
+The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
+%
+POLLUX, a proposed instrument for LUVOIR would detect polarized light from hot Jupiters at \qty{300}{\nano\meter} \citep{Bouret2018_pollux}.
 %
@@
 %
-In \S 2.1, we describe the target list, in \S 2.2 we derive expressions for the range of phase angles given a few different limits, in \S 2.3, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
-%
-We present and discuss our results in \S 3 and conclude in \S 4.
-It is important to note that the phases for each feature throughout this paper are approximate, that features overlap, and that the maximum peak can vary with composition and wavelength.
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
+%
+We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}. 
+It is important to note that the phases for the angular features throughout this paper are approximate, that features overlap, and that their maximum peak will usually vary with particle composition and wavelength.
 %the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
@@
 
-\subsection{\hwo{} and Stellar Sample}
+\subsection{\hwo{} and its stellar sample}
+\label{subsec:2.1}
+
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\hwo{} is still in the early stages of development, and the exact telescope and instrument design is as of yet undecided. 
-%
-In this work, we assume a \SI{6}{\meter} primary mirror, as suggested in the The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
-%
-The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint accompanying exoplanet. 
-%
-The \IWA{} of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
+\hwo{} is still in the early stages of development, and although the exact telescope and instrument design have not yet been determined, we assume a \qty{6}{\meter} primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
+%
+The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint accompanying exoplanets. 
+%
+The \IWA{} of the coronagraph will be driven by both the science requirements and technological limitations and is still unknown.
 %In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA{} of the observatory. 
@@
 For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at representative visible and near infrared wavelengths of $\lambda = \SI{600}{\nano\meter}$ and $\SI{1}{\micro\meter}$.
-
+For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths of $\lambda = \qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+
+%--------------------------------------------------------------------------
 \begin{table}
@@
     \caption{
-        Conversion between $\lambda / D$ and \si{\mas} for different wavelengths, assuming $D = \SI{6}{\meter}$. 
+        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., \IWA{} and \OWA{}, respectively).
         %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
@@
     \toprule
-     & $\lambda/D$ & mas (at \SI{600}{\nano\meter}) & mas (at \SI{1}{\micro\meter}) \\
+    & $\lambda/D$ & \unit{\mas} (at \qty{600}{\nano\meter}) & \unit{\mas} (at \qty{1}{\micro\meter}) \\
     \midrule
@@
 \end{table}
-
-\hwo{} is currently envisioned to observe the stars stated in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.
-%
-\footnote{\url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
-%
-This list comprises $\sim$160 stars, the majority of which are Sun-like dwarfs; 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
+%--------------------------------------------------------------------------
+
+\hwo{} is envisioned to observe the stars in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
+\footnote{\hwo{} target list: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
+%
+This list comprises $\sim\num{160}$ stars, the majority of which are Sun-like dwarfs: 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
 %
@@
 %with an occurrence rate of rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
-The corresponding adopted habitable zone limits are a semi-major axis of \SIrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \SIrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the bolometric luminosity normalized to the Sun.
+The corresponding adopted habitable zone limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \qtyrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the bolometric luminosity normalized to the Sun.
 %
@@
 %
-The authors of the report determined that these are the nearby stars (maximum distance \SI{25}{\parsec}) for which exo-Earths would be the most observable for an imaging survey of habitable zones with a \SI{6}{\meter} telescope.
-
-In what follows, we consider hypothetical Earth-size planets orbiting at the right distance from each star to receive the same flux as the Earth receives from the Sun.
-%
-By construction, such planets should be observable at quadrature phase by HabWorlds.
+The authors of the Decadal Survey determined that for an imaging survey of habitable zones with a \qty{6}{\meter} telescope, these are the nearby stars (maximum distance \qty{25}{\parsec}) around which exo-Earths would be the best observable.
+
+In what follows, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
+%
+By design, such planets should be observable by \hwo{} at least at their quadrature phase.
 
@@
 \subsection{Maximum scattering phase angle coverage}
-\label{sec:Delta_phi}
-
-The coronagraph can obscure part of the planet's orbit which, as demonstrated in \cref{fig:annotated-orbit}, can prevent the planet from being observed at high and low scattering angles. 
-%
-The range of scattering phase angles that can be observed will depend on the \IWA{} of the coronagraph as well as the shape and orientation of the exoplanet's orbit.
-
+\label{subsec:2.2}
+
+A coronagraph can obscure part of the planet's orbit which, as shown in \cref{fig:annotated-orbit}, can prevent the planet from being observable at high and low scattering angles. 
+%
+The range of observable scattering phase angles will depend on the \IWA{} of the coronagraph, as well as on the orientation and shape of the exoplanet's orbit.
+
+%--------------------------------------------------------------------------
+% Figure 2
+%--------------------------------------------------------------------------
 \begin{figure}
@@
     \caption{
-        An illustration of an edge-on circular orbit showing that the coronagraph can obscure the exoplanet at high and low scattering phase angles.
-        %
-        This can prevent observations of scattering phenomena which could be used to indicate the habitability and atmospheric composition of the exoplanet.
-The phase angle, $\varphi$, is shown in red and is the complement to the scattering angle in edge on systems.
-        %\timmy{Maybe mention that this is a simplified illustration with a circular orbit, plus we are basically looking at an edge-on system?}
+        An illustration of a planet in an edge-on circular orbit showing how a coronagraph can obscure the exoplanet at high and low scattering phase angles, and thus prevent observations of scattering features in an exoplanet's phase curve that could help to identify its atmospheric and/or surface composition and habitability.
+        The phase angle, $\varphi$, is shown in red and is the complement of the scattering angle in edge on systems.
     }
@@
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Circular edge-on orbits}
-
-In the non-eccentric case (i.e., with a circular orbit and where the coronagraphic mask centred on the star), minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) scattering phase angles accessible are symmetric about 90 degrees.
+%--------------------------------------------------------------------------
+
+\subsubsection{Circular, edge-on orbits}
+
+In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering phase angles are symmetric about 90 degrees.
 %
@@
     = \varphi_\mathrm{max} - \varphi_\mathrm{min}
-    =  2(\varphi_\mathrm{max} - 90^\circ) 
-    =  2(90^\circ-\varphi_\mathrm{min} ) 
+    =  2 \cdot (\varphi_\mathrm{max} - \qty{90}{\degree}) 
+    =  2 \cdot (\qty{90}{\degree} - \varphi_\mathrm{min} )  \,.
 \end{equation} 
-
-The on-sky projected planet distance, $r_\mathrm{proj}$ (in \si{\au}) is shown in \cref{fig:scattering-angle} assuming a non-eccentric circular orbit with a semi-major $a$ (in \si{\au}) at a distance $d_*$ (in \si{\parsec}) from the observer.
-%
-The scattering phase angle $\varphi$ in this set up follows the equation,
+The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a 
+semi-major orbital axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer. 
+%
+The scattering phase angle $\varphi$ is given by
 \begin{equation}
-    \sin(\varphi) = \frac{r_\mathrm{proj}}{a}
+    \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,.
 \end{equation}
@@
     \label{eq:scattering_angle}
-    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a}
+    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a} \,.
 \end{equation}
- %
+
+%--------------------------------------------------------------------------
+% Figure 3
+%--------------------------------------------------------------------------
 \begin{figure}
@@
     \caption{
-        The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major $a$ at a distance $d_*$.
-        %
+        The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major orbital axis $a$ at a distance $d_*$ from the observer.
         The on-sky angular separation as viewed by the observer is $\delta$.
-        %
         The scattering phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.  
@@
 \end{figure}
-
-\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo{} target list, assuming an IWA of \SI{62}{\mas} ($3 \lambda / D$ coronagraph at $\SI{600}{\nano\meter}$). 
-%\timmy{60 (as in the abstract), or 62?}
+%--------------------------------------------------------------------------
+
+\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo{} target list, assuming an IWA of \qty{62}{\mas} ($3\,\lambda / D$ coronagraph at $\qty{600}{\nano\meter}$, see \cref{tab:IWA_OWA}).
 %
@@
 %
-For edge-on circular orbits, it will be possible to observe Rayleigh scattering for all targets and rainbows will be accessible in some cases.
-
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature, while the rainbow angle will only be accessible for a subset of exoplanets.
+
+%--------------------------------------------------------------------------
+% Figure 4
+%--------------------------------------------------------------------------
 \begin{figure*}
@@
     \caption{
-        Scatter plot for the target sample, showing stellar effective temperature and stellar distance.
-        %
-        The size of the points represents the angular separation of the star and planet in milliarcseconds as presented in the target list.
-        %
-        The colour of the points shows $\Delta \varphi$ assuming a circular edge-on orbit at a semi-major axis corresponding to an Earth-like instellation for an IWA of 62 mas.
-        %
-        Additionally, the colour bar indicates the atmospheric phenomenon that can be detected where phenomena from the bottom of the colour bar up to the indicated colour are detectable for that system.
-        %
-        Thus, dark blue points are systems which have the most key features, as systems in which the angles required to see the rainbow are probed will also have the angles required to see the Rayleigh scattering probed.
+        Scatter plot for the target star sample, showing stellar effective 
+        temperatures and distances.
+        %
+        The circle sizes represent the angular separation between 
+        the star and its planet (in mas) as presented in the target list.
+        %
+        The colours show $\Delta \varphi$ assuming circular, edge-on orbits
+        at a semi-major axis $a$ corresponding to an Earth-like 
+        instellation and an IWA of \qty{62}{\mas}.
+        %
+        Additionally, the colour bar indicates the optical phenomena that 
+        can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
+        %
+        For example, dark blue circles indicate planets on which most phenomena would be observable because planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
     }
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Circular inclined orbits}
-
-Most systems will not be viewed as edge on from Earth.
-%
-For a given system with inclination $i$, we have identified 2 regimes, which are shown in \cref{fig:orb-grid}:
-
+%--------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Circular, inclined orbits}
+
+Most planetary systems will not be viewed as edge-on from Earth.
+%
+For a system with inclination angle $i$, we can identify the following 2 regimes (see also \cref{fig:orb-grid}):
 \begin{enumerate}
-    \item If the inclination is relatively face-on, then the coronagraph mask does not obscure any part of the sky-projected orbit, and the maximum scattering phase angle coverage depends solely on the disk inclination ;
-    \item When parts of the orbit  are hidden by the focal plane mask, the scattering phase angle coverage only depends on the \IWA{} and orbital semi-major axis. 
+    \item The coronagraph does not obscure any part of the sky-projected orbit: 
+    the maximum scattering phase angle coverage depends solely on $i$. This 
+    holds for orbits that are relatively face-on; 
+    \item The coronagraph does obscure parts of the sky-projected orbit: 
+    the scattering phase angle coverage depends only on the \IWA{} and 
+    the orbital semi-major axis $a$. 
 \end{enumerate}
 
-\begin{figure}%[htb]
+%--------------------------------------------------------------------------
+% Figure 5
+%--------------------------------------------------------------------------
+\begin{figure}
    \centering
@@
    \caption{
-        The scattering phase angles accessible to a direct imaging mission can be limited by orbital inclination or by the coronagraphic inner working angle.
-        %
-        If the orbital inclination is close to face-on (top panels) then the scattering phase angles probed throughout an orbit are limited by inclination; for circular orbits accessible scattering phases are $\phi \in 90^\circ \pm i$.
-        %
-        For orbital inclinations close to edge-on (bottom panels) the planet goes through all phases, but the scattering phase angles near 0$^\circ$ and 180$^\circ$ are not accessible due to obscuration by the coronagraph mask; for circular orbits, the accessible scattering phases are $\phi \in 90^\circ \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+        The scattering phase angles accessible to a direct imaging mission 
+        can be limited by either the orbital inclination angle $i$ or the 
+        coronagraphic \IWA{}.
+        %
+        If the orbit is close to face-on (top panels), the scattering phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
+        %
+        For orbits close to edge-on (bottom panels) the planet attains all 
+        phases, but the scattering phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
     }
@@
 \end{figure}
-
-Using \cref{eq:scattering_angle}, we deduce the equation for inclined circular orbits: 
+%--------------------------------------------------------------------------
+
+Using \cref{eq:scattering_angle}, we deduce the following for inclined 
+circular orbits: 
 \begin{equation}
@@
     \begin{cases}
-        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA}\; d_* }{a}
-  \\ 
+        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA}\; d_* }{a} \\ 
         2 \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA}\; d_* }{a}
-    \end{cases}\,.
+    \end{cases} \,.
 \end{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monte Carlo Simulations}
-
-\subsubsection{Circular randomly inclined orbits}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Monte Carlo simulations}
+\label{subsec:2.3}
+
+\subsubsection{Circular, randomly inclined orbits}
 \label{sec:circular}
 
-For each star on the \hwo{} target list, we generate $1000$  hypothetical habitable planets at the Earth-equivalent-flux semi-major axis. 
-%
-The orbital inclinations are randomly drawn from a distribution uniform in $\cos i$. 
-%
-We then use \cref{eq:Delta_phi_max} to determine the range of accessible scattering phase angles for each hypothetical planet.
-%
-The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible scattering phase angles normalised to the total number of systems in \hwo{} target list, depending on the choice of inner working angle.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+For each star on the \hwo{} target list, we generate \num{1000} random orbits of hypothetical habitable planets with Earth-equivalent-flux semi-major axes. 
+%
+The orbital inclination angles $i$ are randomly drawn from a distribution 
+uniform in $\cos i$. 
+%
+We then use \cref{eq:Delta_phi_max} to determine the range of 
+accessible scattering phase angles for each hypothetical planet.
+%
+The solid lines in \cref{fig:betaallofit} show the cumulative 
+distribution of the maximum and minimum accessible scattering phase 
+angles normalised to the total number of systems in the \hwo{} 
+target list for different values of the \IWA{}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%--------------------------------------------------------------------------
 \begin{figure*}%[t]
@@
         \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
-        Cumulative distributions of the most extreme scattering phase angle accessible for different \IWA{} where the solid lines are for randomly inclined circular orbits and the dashed lines are for randomly orientated elliptical orbits.
-        %
-        The top x axis indicates the minimum scattering phase angle and the bottom x axis indicates the maximum.
-        %
-        These are symmetric about quadrature (90 degrees).
-        %
-        The y axis indicates the number of systems divided by the number of Monte Carlo samples and is thus normalised to the number of systems in the target list.
-        %
-        The secondary y axis indicates the number of systems assuming only 24 percent of them contain an Earth-like exoplanet in the habitable zone.
+        The cumulative distributions of the most extreme scattering phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
+        (solid lines) and randomly inclined, elliptical orbits (dashed lines).
+        %
+        The top $x$-axis indicates the minimum and the 
+        bottom $x$-axis the maximum accessible scattering phase angle.
+        %
+        These angles are symmetric about quadrature (90 degrees).
+        %
+        The $y$-axis on the left indicates the number of planetary systems 
+        divided by the number of Monte Carlo samples and is thus normalised to 
+        the number of systems on the target list.
+        %
+        The $y$-axis on the right indicates the number of systems, assuming \qty{24}{\percent} of them have an Earth-like planet in their HZ.
     }
@@
 \end{figure*}
-
-\subsubsection{Eccentric randomly inclined orbits}
+%--------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Eccentric, randomly inclined orbits}
 \label{sec:eccentric}
-We repeat the Monte Carlo simulation from \cref{sec:circular} assuming the orbital eccentricity follows a beta distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
-%
-We randomly drew 1000 random orbits for each star and determined the sky-projected orbits following, e.g., \cite{2010exop.book...15M}.
-Eccentricities were drawn from the beta distribution, and inclinations were uniform in $\cos i$.
-\Cref{fig:ball-o-yarn} shows a sample of the generated orbits with the filtered \IWA{}'s
-%
-%Each plot has been scaled such that the habitable zone occupies the same radius, hence the \IWA's vary in apparent size compared to the orbit.
-%
-We then calculated the minimum and maximum accessible scattering phase angles for of \IWA{}'s of 21, 42, 63 and 84 mas which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\SI{600}{\nano\meter}$.
-%
-The dashed lines in \cref{fig:betaallofit} shows the normalised cumulative distribution of the accessible angles for each of the \IWA, when accounting for orbital eccentricity.
-%
-We also calculate which targets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
-%
-The cumulative distribution over \IWA{} of the number of systems, normalised to the total number of systems, where each scattering phenomena is visible is shown in \cref{fig:accessible_phase_angles}.
-%
-
+
+We repeat the Monte Carlo simulations from \cref{sec:circular} assuming 
+the orbital eccentricity follows a beta-distribution with shape parameters 
+$a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
+%
+We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
+Eccentricities are drawn from the beta distribution, and inclinations are uniform in $\cos i$. 
+\Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
+%
+
+We then calculate the minimum and maximum accessible scattering phase 
+angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
+%
+The dashed lines in \cref{fig:betaallofit} show the normalised cumulative 
+distribution of the accessible angles for each of \IWA for eccentric orbits. 
+%
+We also calculate which target planets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+%
+The cumulative distribution over the \IWA{} of the number of planetary systems where all optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulating contrast curves}
-
-A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, that is, the relative flux of the reflected light from the planet divided by the stellar flux.
-%
-While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as a function of scattering phase angles.
-%
-This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
-%
-Here, we aim to present some simplified examples of the impact of the scattering angle dependent contrast on the detectability of rainbow and glint features. 
-%
-We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of an Earth-like planet \citep{ treesandstam2019,trees2022}.
-
-The NASA Exoplanet Exploration Program’s Mission Star List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
-%
-The contrast is calculated as 
+\label{subsec:2.4}
+
+A second parameter that is crucial for detecting angular features like 
+rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the flux of the reflected light from the planet divided by the stellar flux. 
+%
+While this is not the main focus of this paper, it is important to consider that the total flux contrast and the polarized contrast generally 
+depend on the scattering phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
+%
+Here, we aim to present some simplified examples of the impact of the 
+scattering angle dependent contrast on the detectability of rainbow and 
+glint features. 
+%
+We note that these calculations are by no means complete or exact, given 
+the many variables that go into the modelling of reflected total and
+polarized fluxes of Earth-like exoplanets \citep{ treesandstam2019,trees2022}.
+%
+
+The NASA Exoplanet Exploration Program’s Mission Star List for \hwo{} 
+includes for every star an estimated contrast for an Earth-twin in 
+the habitable zone.
+%
+% The cited paper has all the equations that lead to this equation:
+The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
 \begin{equation}
-C = F_p/F_* = p \phi (\alpha) (R_p/a)^2,
+    C(\alpha) 
+    = F_p(\alpha) / F_* 
+    = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
 \end{equation}
-where $p$ is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
-%
-The assumptions that went into these calculations are that the orbits are circular, the geometric albedo is $p=0.2$, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, for example, isotropic scattering.
-%
-These calculations give a good order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
-%
-Therefore, in this work, we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
-%
-We note that we do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some of the work for the star list to be redone.
-
-We use the scattering phase functions from \cite{treesandstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
-%
-The $\SI{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
-%
-We normalize the reflected light curve at the value of 90 degrees (quadrature) and multiply it with the contrast.
-%
-In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given an inclination of $90$ degrees.
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at phase angle $\alpha$
+(normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.
+\todo{here, $\alpha$ is used for the phase angle ... and is the 
+'integral phase function' the same as the 'scattering phase function'?}
+%
+The assumptions that went into the contrast estimations for \hwo{}
+are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
+%
+These calculations give an order-of-magnitude calculation for the 
+contrast of Earth-like planets around nearby stars.
+%
+Therefore, in this work, we will assume that the calculated contrasts 
+in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
+%
+We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
+%
+
+We use the scattering phase functions from \citet{treesandstam2019} 
+for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
+%
+The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
+%
+We normalize the phase function at quadrature and multiply it with the 
+contrast.
+%
+In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given an inclination of \num{90} degrees.
 %
@@
 %
-We multiply this flux with the degree of polarisation and retrieve the polarised contrast.
-%
-The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri and the M~dwarf Lalande~21185.
-
+We multiply this flux with the degree of polarization and retrieve the polarized contrast.
+%
+The resulting contrast curves are shown in \cref{fig:contrasts} for 
+three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf 
+$\epsilon$~Eri, and the M~dwarf Lalande~21185.
+
+%--------------------------------------------------------------------------
 \begin{figure*}%[t]
@@
    \caption{
-    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of $90^\circ$ for the stars $\alpha$ Cen A, $\epsilon$ Eri and Lalande 21185.
-    %
-    The solid line indicates the contrast in unpolarised light with the contrast at quadrature marked by a solid dot.
-    %
-    The polarised component is indicated by the colored dots for which the color represents the scattering phase angle from quadrature.
-    %
-    The light grey points show the quadrature contrasts of the other targets in the star list and the dashed lines indicate $1,2$ and $3$ times the \IWA{} for \hwo.
+    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of \qty{90}{\degree} for the stars $\alpha$~Cen~A, $\epsilon$~Eri and Lalande~21185.
+    %
+    The solid line indicates the contrast in unpolarized light, with the contrast at quadrature marked by a solid dot.
+    %
+    The polarized component is indicated by the coloured dots, for which the colour represents the scattering phase angle from quadrature.
+    %
+    The light grey points show the quadrature contrasts of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
     }
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%--------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
+\label{sec:3}
 
@@
 \label{sec:result_eccentricity}
-As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable affect on the number of systems that can be observed at each scattering phase angle.
-%
-This is because, Kepler's second law ensures that eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA{} and extreme orbital phases (albeit with worse contrast due to the greater star--planet separation).
-%
-The eccentricity does not significantly effect the number of systems for \IWA{} smaller than cut used in the selection of the systems in \hwo{} target list.
-It is, however, important to account for orbital eccentricity when the \IWA{} of the coronoagraph is larger than the \IWA{} used to filter the target list otherwise the number of systems observable at extreme phase angles would be under predicted.
+
+As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable effect on the number of systems that can be observed at each scattering phase angle.
+%
+This is because---as Kepler's second law explains---eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA{} and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+%
+The eccentricity does not significantly affect the number of systems for \IWA{} smaller than cut used in the selection of the systems in the \hwo{} target list. 
+It is, however, important to account for orbital eccentricity when the \IWA{} of the coronagraph is larger than the \IWA{} used to filter the target list; otherwise, the number of systems observable at extreme phase angles would be under-predicted.
 
@@
 
-In this section we consider the number of stars around which different scattering phenomena could be observable for different \IWA{}.
-%
-We first define the phase angles over which various these phenomena are accessible, in Table~\ref{tab:phase_ranges}.
+In this section, we consider the number of stars around which different scattering phenomena could be observable for different \IWA{}.
+%
+We first define the phase angles over which various of these phenomena are accessible in \cref{tab:phase_ranges}.
 %
@@
 
-\cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA{}.
-%
-The \IWA{} of \hwo's coronagraph will depend on the primary mirror diameter, observing wavelength and how well a coronagraph of a given size can suppress the glare of the host star.
+\Cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA{}.
+%
+The \IWA{} of \hwo's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
 %
@@
 
+%--------------------------------------------------------------------------
+%--------------------------------------------------------------------------
 \begin{table}
@@
     \caption{
-        The ranges of phase angles for the different scattering phenomena considered here.
+        The ranges of phase angle for the different scattering phenomena considered here.
     }
     \label{tab:phase_ranges}
-    \begin{tabular}{ c c c c } 
+    \begin{tabular}{ l c c c } 
         \toprule
-     & $\varphi_\text{min}$ & $\varphi_\text{peak}$ & $\varphi_\text{max}$ \\
+        & $\varphi_\mathrm{min}$ & $\varphi_\mathrm{peak}$ & $\varphi_\mathrm{max}$ \\
         \midrule
@@
 \end{table}
-
+%--------------------------------------------------------------------------
     % features = {r'Rainbow': (127,138,158),
@@
 
-The Rayleigh scattering feature, which can used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-%
-The rainbow feature appears further from quadrature but will still be observable in a significant fraction of systems.
-%
-The position and shape of the peak is sensitive to the properties of the droplets allowing their nature to be determined if the peak of the rainbow feature can be reached.
+The Rayleigh scattering feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
+%
+The rainbow feature appears further from quadrature, but will still be observable in a significant fraction of systems.
+%
+The position and shape of the peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
 %
@@
 %
-In most systems, only the first half of the feature is observable but this can still be used to identify a liquid ocean on the surface of the planet.
-%
-A key indicator of habitability. Finally, it is possible to view the glory scattering feature in one system, Alpha Cen A.
-
+In most systems, only the first half of the feature is observable, but this can still be used to identify a liquid ocean on the surface of the planet.
+%
+A key indicator of habitability. 
+Finally, it is possible to view the glory scattering feature in one system, Alpha~Cen~A.
+
+%--------------------------------------------------------------------------
+%--------------------------------------------------------------------------
 \begin{table}
@@
     \caption{
-        Number of stars around which the peak phase angle of each phenomena is detectable, expressed in $mas$ and $\lambda/D$ for $\lambda=600$nm and  $D=6$m. 
+        Number of stars around which the peak phase angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
     }
     \label{tab:nstars_detect}
-    \begin{tabular}{ c c c c } 
+    \begin{tabular}{ l c c c } 
         \toprule
-      & \multicolumn{3}{c}{IWA}
-    \\
-     & 21\,$mas$ & 41\,$mas$ & 62\,$mas$
-    \\
-     & $2\lambda/D$ & $3\lambda/D$ & $4\lambda/D$ 
-    \\
+        & \multicolumn{3}{c}{IWA}\\
+        & \qty{21}{\mas} & \qty{41}{\mas} & \qty{62}{\mas} \\
+        & $2\,\lambda/D$ & $3\,\lambda/D$ & $4\,\lambda/D$\\
         \midrule
@@
 \end{table}
+%--------------------------------------------------------------------------
 
@@
 
-The accessible scattering phase angles is not the only factor that needs to be considered when determining what scattering features are observable.
+The accessible scattering phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
 %
@@
 %
-\cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo{} target list assuming edge-on orbits.
-%
-The over all contrast varies significantly for each system and throughout the orbit.
-%
-The observability of these planets will depend on the star light suppression of the coronagraph which is influenced by its \IWA{}.
+\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo{} target list assuming edge-on orbits.
+%
+The overall contrast varies significantly for each system and throughout the orbit.
+%
+The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA{}.
 %
@@
 
+%--------------------------------------------------------------------------
+% Figure 8:
+%--------------------------------------------------------------------------
 \begin{figure*}
@@
     \caption{
-        %Are we putting all of these somewhere?  Its prob about 7 pages worth in total?
-        Random examples of the eccentric orbits generated for the stellar sample.
-        %
-        The orbits are scaled by the Earth-equivalent flux distance. 
-        %
-        The concentric circles marked by the dashed lines indicate inner working angles of 20.6, 41.3, 61.9, and \SI{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$ respectively (assuming $\lambda = \SI{600}{\nano\meter}$ and $D = \SI{6}{\meter}$).
-        %
-        The figure illustrates that the \IWA{} can significantly affect the range of scattering phases observable with each orbit.
+        Examples of the eccentric orbits generated for the stellar sample.
+        %
+        The orbits are scaled by the Earth-equivalent incident flux distance. 
+        %
+        The concentric, dashed circles indicate \IWA{}s of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        %
+        Each system has been scaled such that its HZ has the same radius, 
+        hence the variation in the size of the \IWA{}s compared to the orbits.
+        It is clear that the \IWA{} can significantly affect the range of 
+        observable scattering phases for each orbit.
     }
@@
 \end{figure*}
-
+%--------------------------------------------------------------------------
+
+%--------------------------------------------------------------------------
+% Figure 9:
+%--------------------------------------------------------------------------
 \begin{figure*}
@@
     \caption{
-        The number of systems for which key scattering features (Rayleigh scattering, rainbows, ocean glint and glories) can be observed as a function of \IWA{}.
-        %
-        Each subplot contains three lines where the dashed line shows the number of systems where the start of the scattering feature can be observed, the solid line shows the number of systems that reach the peak of the feature and the dot-dashed line the number where the scattering angles of the end of the feature are observable.
-        %
-        The bottom $x$ axis gives the \IWA{} of the coronograph in milliarcseconds which is converted to $\lambda/D$ for $\lambda=\SI{600}{\nano\meter}$ and $D=\SI{6}{\meter}$ for the top $x$ axis.
-        %
-        The secondary $y$ axis indicates the number of systems that each feature would be observable in assuming only 24\% of them contain an Earth-like exoplanet in the habitable zone.
-        %
-        It is worth noting that the Glory feature is only observable in one system, Alpha Cen A. 
+        The number of planetary systems for which key optical phenomena 
+        (Rayleigh scattering, rainbow, ocean glint, glories) would be 
+        observable as functions of the \IWA{}.
+        %
+        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak of the phenomena is observable (solid lines), and where the end is observable (dot-dashed lines).
+        %
+        The $x$-axis at the bottom shows the \IWA{} of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA{} converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
+        %
+        The $y$-axis on the right indicates the number of systems where 
+        all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
+        %
+        It is worth noting that the Glory would be observable in only one system: Alpha~Cen~A. 
+        \todo{What is the reasoning with the colours?}
     }
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%--------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Conclusions}
+\label{sec:4}
+
 %What did we determine? 
@@
 
-The Habitable Worlds Observatory is a coronagraphic space mission envisioned to detect and characterise Earth-like exoplanets in the 2040's. 
-%
-The instrument design and capabilities are still under development and thus can be driven by the science goals of the mission.
-%
-In this work, we investigated the possibility of using \hwo{} to detect scattering phenomena on terrestrial exoplanets, including ocean glint, which could establish a planet's habitability.
-
-We primarily focused on the range of scattering phase angles and therefore which scattering phenomena could potentially be observed given the limitation of the coronagraphic \IWA{}.
+The Habitable Worlds Observatory is a coronagraphic space telescope 
+envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
+%
+The instrument design and capabilities are still under development and 
+should be driven by the mission's science goals.
+%
+In this work, we investigated the possibility of using \hwo{} to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory's, which could establish a planet's habitability.
+%
+We primarily focused on the range of accessible scattering phase angles given the limitations of the coronagraphic \IWA{} and aimed to investigate which phenomena could potentially be observed.
 %
@@
 %
-Assuming a \SI{62}{\mas} \IWA{}, all 160 systems in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory would be observable at scattering phases with Rayleigh features, but the rainbows of water clouds would only be accessible in $\sim$42 systems, and ocean glint in $\sim$14 systems.
-%
-The habitability of Earth analogs in the remaining 146 systems would have to be established by less direct means like rotational mapping to identify continents and oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
-%
-The number of habitable zones that can be probed by \hwo{} increases with decreasing \IWA{}. 
-%
-The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic frequency of such planets, will drive the choice of coronagraphic \IWA{} for this mission.
-%
-We have shown that with a $6$m primary mirror and \IWA $= 3\lambda/D$ at $600$nm coronograph, only $\sim$10\% of Earth-like exoplanets imaged with \hwo{} could be imaged at a sufficiently wide range of phase angles to establish the presence of surface liquid water via its specular reflection.
-%
-The habitability of remaining targets would have to be established by less direct means.
-%
-This could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronograph.
+The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets, will drive the choice of coronagraphic \IWA{} for this mission.
+%
+Assuming a \qty{62}{\mas} \IWA{}, all \num{160} planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \hwo{} would be observable at scattering phases where angular features due to Rayleigh scattering show up, but the rainbow of light that is scattered in droplets of water clouds would only be accessible in $\sim\num{42}$ systems, and ocean glint in $\sim\num{14}$ systems.
+%
+The habitability of Earth-like exoplanets in the remaining 146 systems would have to be established through less direct methods like rotational mapping to identify horizontally inhomogeneous surface coverage, for example due to presumably bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
+%
+The fraction of habitable zones around stars in the target list that can be probed by \hwo{} increases with decreasing \IWA{}. 
+%
+\todo{the paragraph below seems similar to the one above ... merge them?}
+We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of phase angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
+%
+This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
 %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Software/Data Availability}
 
-The NASA ExEP target list containing the data used for these simulations is available online.
-%
+\todo{check the naming and references/footnotes for this list throughout the paper}
+The NASA ExEP target list containing the data used for these simulations is  available online.%
 \footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowledgements}
@@
 %
-This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).
+This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No. 866001 - EXACT).
 %
@@
 
-% Add references
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% References
 \bibliographystyle{mnras}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Author affiliations}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2023-04-19 src/scripts/run_simulation.py

@@ -21,7 +21,7 @@ norbitsample = 1000
 op = snakemake.params['whichsim']
 #op = sys.argv[1]
 if op == 'iwa':
-	iwa = np.array([21, 42, 63, 84])
+	iwa = np.array([21, 41, 62, 83])
 	tag = ''
 elif op == 'iwa2':
 	iwa = np.linspace(20,120,5) # inner working angles (in mas) to compute betamax

### 2023-04-23 src/scripts/attic/alternate-simulate-elliptic-orbits.py
Move some unused script to the attic

@@ -1,5 +1,8 @@
 """
 Script to run Monte Carlo simulations of elliptic orbits.
+
+This is an alternative version of the data generation script that was
+NOT actually used in the end. It is kept here for reference.
 """
 
 # -----------------------------------------------------------------------------

### 2023-04-23 src/scripts/create-scatterplot.py
Move files with utility functions into `utils` folder

@@ -1,5 +1,6 @@
 """
-Create a color-coded scatter plot the the target list.
+Create Figure 4: Scatter plot of the target list color-coded by the
+degrees from quadrature when assuming edge-on, circular orbits.
 """
 
 # -----------------------------------------------------------------------------
@@ -8,7 +9,6 @@ Create a color-coded scatter plot the the target list.
 
 from typing import Any
 
-from matplotlib.colorbar import Colorbar
 from matplotlib.lines import Line2D
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
@@ -17,9 +17,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
-from utils import set_fontsize
-
-import paths
+from src.scripts.utils.plotting import set_fontsize
+from src.scripts.utils import paths
 
 
 # -----------------------------------------------------------------------------

### 2023-04-23 src/scripts/hwo.py

@@ -1,6 +1,0 @@
-d = 6.0 # diameter in m
-wavelength = 600e-9 # central wavelength in m
-
-if __name__ == "__main__":
-	print(f'Telescope diameter is {d:3.1f} m')	
-	print(f'Central wavelength is {wavelength*1e9:5.1f} nm')

### 2023-04-23 src/scripts/make_beta_max_3_lambda_over_d_csv.py

@@ -1,7 +1,7 @@
 import pandas as pd
 import numpy as np
-import paths
-from hwo import *
+from src.scripts.utils import paths
+from src.scripts.utils.constants import *
 
 # telescope d and wlen are imported from hwo file to make consistent for all the codes
 

### 2023-04-23 src/scripts/make_yarn_plot.py

@@ -17,11 +17,11 @@ import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
 
-from phases import kep3d
-from samplers import sample_e, sample_i
-from utils import CBF_COLORS
+from src.scripts.utils.phases import kep3d
+from src.scripts.utils.samplers import sample_e, sample_i
+from src.scripts.utils.plotting import CBF_COLORS
 
-import paths
+from src.scripts.utils import paths
 
 
 # -----------------------------------------------------------------------------

### 2023-04-23 src/scripts/plot-orb-grid.py

@@ -1,5 +1,5 @@
 import numpy as np
-import paths
+from src.scripts.utils import paths
 #from scipy.interpolate import interp1d
 #import astropy.constants as c
 import astropy.units as u
@@ -10,7 +10,7 @@ from matplotlib.patches import Arrow
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 from mpl_toolkits import mplot3d
 
-from orbfunc import xyz_position, phase
+from src.scripts.utils.orbfunc import xyz_position, phase
 
 ##########################################################################################
 

### 2023-04-23 src/scripts/plot_contrasts.py

@@ -10,7 +10,8 @@ import numpy as np
 import pandas as pd
 from scipy import interpolate
 
-import paths
+from src.scripts.utils import paths
+
 
 # -----------------------------------------------------------------------------
 # AUXILIARY FUNCTIONS

### 2023-04-23 src/scripts/plot_features_vs_iwa.py

@@ -2,8 +2,8 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 from scipy import interpolate as si
-import paths
-from hwo import *
+from src.scripts.utils import paths
+from src.scripts.utils.constants import *
 
 CBF_COLORS = [
     "#5790fc",  # blue

### 2023-04-23 src/scripts/plot_n_versus_phase.py

@@ -1,8 +1,7 @@
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
-import paths
-from hwo import *
+from src.scripts.utils import paths
 
 CBF_COLORS = [
     "#5790fc",  # blue

### 2023-04-23 src/scripts/utils/constants.py

@@ -1,0 +1,39 @@
+"""
+Define constants used in the scripts.
+"""
+
+import astropy.units as u
+
+d = 6.0  # diameter in m
+wavelength = 600e-9  # central wavelength in m
+
+# Assumed diameter of HWO
+DIAMETER = u.Quantity(6, u.m)
+
+# Assumed occurrence rate of rocky planets in the optimistic habitable zone
+# See: National Academy of Sciences Astronomy & Astrophysics 2020 Decadal
+# Survey, Section I.3.1. Value originally from Bryson et al. (2021).
+ETA_EARTH = 0.24
+
+# For each scattering feature, define the start, peak, and end scattering
+# angle (in degrees)
+SCATTERING_FEATURES = {
+    "Rainbow": (127, 138, 158),
+    "Rayleigh": (90, 110, 130),
+    "Ocean Glint": (50, 30, 10),
+    "Glory": (10, 5, 0),
+}
+
+# Assumed wavelength for the observations
+WAVELENGTH = u.Quantity(600, u.nm)
+
+
+# Compute lambda / D in mas
+with u.set_enabled_equivalencies(u.dimensionless_angles()):
+    LAMBDA_OVER_D_IN_MAS = (WAVELENGTH / DIAMETER).to(u.mas).value
+
+
+if __name__ == "__main__":
+    print(f'Telescope diameter is {d:3.1f} m')
+    print(f'Central wavelength is {wavelength * 1e9:5.1f} nm')
+    print("LAMBDA_OVER_D_IN_MAS:", LAMBDA_OVER_D_IN_MAS)

### 2023-04-23 src/scripts/utils/paths.py

@@ -1,11 +1,12 @@
 """
 Exposes common paths useful for manipulating datasets and generating figures.
-
 """
+
 from pathlib import Path
 
+
 # Absolute path to the top level of the repository
-root = Path(__file__).resolve().parents[2].absolute()
+root = Path(__file__).resolve().parents[3].absolute()
 
 # Absolute path to the `src` folder
 src = root / "src"

### 2023-04-23 src/scripts/create-scatterplot.py
Fix relative imports

@@ -17,8 +17,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
-from src.scripts.utils.plotting import set_fontsize
-from src.scripts.utils import paths
+from utils.plotting import set_fontsize
+from utils import paths
 
 
 # -----------------------------------------------------------------------------

### 2023-04-23 src/scripts/make_beta_max_3_lambda_over_d_csv.py

@@ -1,11 +1,11 @@
 import pandas as pd
 import numpy as np
-from src.scripts.utils import paths
-from src.scripts.utils.constants import *
+from utils import paths
+from utils.constants import *
 
 # telescope d and wlen are imported from hwo file to make consistent for all the codes
 
-# givea a nan for the first element because it tries to calculate 
+# givea a nan for the first element because it tries to calculate
 # arccos of 1.032, presumably this is alpha or proxima cen
 
 df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')

### 2023-04-23 src/scripts/make_yarn_plot.py

@@ -17,11 +17,10 @@ import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
 
-from src.scripts.utils.phases import kep3d
-from src.scripts.utils.samplers import sample_e, sample_i
-from src.scripts.utils.plotting import CBF_COLORS
-
-from src.scripts.utils import paths
+from utils.phases import kep3d
+from utils.samplers import sample_e, sample_i
+from utils.plotting import CBF_COLORS
+from utils import paths
 
 
 # -----------------------------------------------------------------------------

### 2023-04-23 src/scripts/plot-orb-grid.py

@@ -1,23 +1,21 @@
-import numpy as np
-from src.scripts.utils import paths
-#from scipy.interpolate import interp1d
-#import astropy.constants as c
-import astropy.units as u
 from astropy.time import Time
-
-import matplotlib.pyplot as plt
 from matplotlib.patches import Arrow
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-from mpl_toolkits import mplot3d
 
-from src.scripts.utils.orbfunc import xyz_position, phase
+import astropy.units as u
+import numpy as np
+import matplotlib.pyplot as plt
+
+from utils import paths
+from utils.orbfunc import xyz_position, phase
+
 
 ##########################################################################################
 
 
 
 params = {
-    'time periapsis passage':Time('2021-09-29T00:00:00', format='fits'), 
+    'time periapsis passage':Time('2021-09-29T00:00:00', format='fits'),
     'orbital period':1*u.day,
     'semi-major axis':1*u.au,
     'eccentricity':0.0,
@@ -36,7 +34,7 @@ def setup_plot(fig, i, oplane=False, losplane=False):
     ax = fig.add_subplot(4, 3, i, projection='3d')
     ax.set_proj_type('ortho')
     ax.margins(-0.499)
-    
+
     # plot orbit
     ax.plot3D(xyz[:,0], xyz[:,1], xyz[:,2], 'k-', lw=2)
 
@@ -44,7 +42,7 @@ def setup_plot(fig, i, oplane=False, losplane=False):
     if losplane==True:
         lossurfx, lossurfy = np.meshgrid(np.linspace(-0.65,0.65,10),np.linspace(-0.65,0.65,10))
         x1 = ax.plot_surface(lossurfx, lossurfy, np.zeros((10,10)), color='g', edgecolor=(0,0,0,0.15), alpha=0.1, label='Plane of Nodes')
-        
+
         ### Trying to catch a version dependent error in matplotlib, may not work for all versions ###
         try:
             x1._facecolors2d = x1._facecolor3d
@@ -54,10 +52,10 @@ def setup_plot(fig, i, oplane=False, losplane=False):
                 x1._facecolors2d = x1._facecolor3d
                 x1._edgecolors2d = x1._edgecolor3d
             except:
-            	pass 	
+            	pass
         ###############################################################################################
-        
-        
+
+
     # plane of orbit
     if oplane==True:
[…]

### 2023-04-23 src/scripts/plot_contrasts.py

@@ -10,7 +10,7 @@ import numpy as np
 import pandas as pd
 from scipy import interpolate
 
-from src.scripts.utils import paths
+from utils import paths
 
 
 # -----------------------------------------------------------------------------
@@ -42,7 +42,7 @@ df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')
 
 hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
 contrasts =np.array(df['Contrast'].values[1:],dtype=float)
-names = np.array(df['CommonID'].values[1:],dtype=str) 
+names = np.array(df['CommonID'].values[1:],dtype=str)
 
 data_trees = np.loadtxt(paths.data / 'Reflected_light_curves_Trees_Stam_2019.txt')
 
@@ -68,7 +68,7 @@ pad_inches = 0.025
 fig,axs = plt.subplots(nrows=1, ncols=3,
     figsize=(9 - 2 * pad_inches, 3.5 - 2 * pad_inches)
 )
-             
+
 xlims= np.array([1e3,2e2,1e2])
 
 for i,name_one in enumerate(names_array):
@@ -99,7 +99,7 @@ for i,name_one in enumerate(names_array):
     ax.set_yscale('log')
 
     if i>0:
-       ax.get_yaxis().set_visible(False)  
+       ax.get_yaxis().set_visible(False)
 
     ax.set_xlim(0,xlims[i])
     ax.set_ylim(8e-12,3e-8)
@@ -109,7 +109,7 @@ for i,name_one in enumerate(names_array):
 
     ax.set_xlabel('Separation [mas]')
     ax.set_ylabel('Contrast')
-    
+
     # Manually construct a color bar and add it to the plot
     if i==2:
         divider = make_axes_locatable(ax)

### 2023-04-23 src/scripts/plot_features_vs_iwa.py

@@ -2,8 +2,8 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 from scipy import interpolate as si
-from src.scripts.utils import paths
-from src.scripts.utils.constants import *
+from utils import paths
+from utils.constants import *
 
 CBF_COLORS = [
     "#5790fc",  # blue
@@ -21,8 +21,8 @@ def make_features_plot_multi(betas,hab_zones,features,iwa_list = [1,2,3,4],eta_e
                        save_fig=False,cmap_name='Spectral'):
     '''
     A function to make a multi-panneled plot of four different scattering features
-    Showing the numner of accessible systems for the 
-    low end, middle and upper end of the phase angle ranges of each feature, as a function of IWA. 
+    Showing the numner of accessible systems for the
+    low end, middle and upper end of the phase angle ranges of each feature, as a function of IWA.
 
     betas - the maximum scattering angle parameter, here defined as beta = phase_angle_max - 90
     iwa_list - list of IWAs you want to use, shape [n_iwas]
@@ -39,7 +39,7 @@ def make_features_plot_multi(betas,hab_zones,features,iwa_list = [1,2,3,4],eta_e
 
     # if ax is None:
     #     fig, axis = plt.subplots(1,1)
-    # else: 
+    # else:
     #     axis=ax
 
     fig,axes = plt.subplots(2,2,figsize=(12,5),sharex=True,sharey=True)
@@ -51,8 +51,8 @@ def make_features_plot_multi(betas,hab_zones,features,iwa_list = [1,2,3,4],eta_e
         highs = []
 
         for i in range(len(iwa_list)):
-        
-            #This helps identify which systems aren't detected at all. 
+
+            #This helps identify which systems aren't detected at all.
             # There are easier ways, but this is legacy and works
             beta_max = np.degrees(np.arccos(iwa_list[i]/(hab_zones)))
 
@@ -65,17 +65,17 @@ def make_features_plot_multi(betas,hab_zones,features,iwa_list = [1,2,3,4],eta_e
             cdf = np.cumsum(pdf)
 
             normalized_cdf = (1-cdf)*np.sum(np.isfinite(beta_max))
-            
+
             #Deal with angles below 90
             these_angles = np.array(features[feature])
             # print(these_angles)
-            for m,angle in enumerate(these_angles): 
+            for m,angle in enumerate(these_angles):
                 if angle < 90:
[…]

### 2023-04-23 src/scripts/plot_n_versus_phase.py

@@ -1,7 +1,7 @@
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
-from src.scripts.utils import paths
+from utils import paths
 
 CBF_COLORS = [
     "#5790fc",  # blue
@@ -18,7 +18,7 @@ def make_dual_cdf_plot(beta_maxes,betas,iwa_list,hab_zones,label_locations,rotat
                        n_bins=20,savefig=False,filename="accessible_phase_angles.png"):
     '''
     A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle
-    for a given inner working ange. 
+    for a given inner working ange.
 
     beta_maxes - An array of maximum scattering angles, generated by the dynamical simulations, shape [n_simulations,n_systems,n_iwas]
     betas - An array of beta values, generated for circular orbits, shape [n_simulations,n_systems,n_iwas]
@@ -33,18 +33,18 @@ def make_dual_cdf_plot(beta_maxes,betas,iwa_list,hab_zones,label_locations,rotat
     filename - the filename of the figure you want to save, string
     '''
 
-    #If an axis isn't provided, then make one. 
+    #If an axis isn't provided, then make one.
     if ax is None:
         # fig, axis = plt.subplots(1,1,figsize=(10,4))
         fig, axis = plt.subplots(1,1,figsize=(10,4))
-    else: 
+    else:
         axis=ax
 
-    #Cycle through each IWA to plot a line. 
+    #Cycle through each IWA to plot a line.
     for i,iwa in enumerate(iwa_list):
         #### First the dynamical betas.
         # (Although we're calling them betas, really they're the maximum scattering angles)
-        
+
         # Getting data of the histogram bins to build the CDF
         count, bins_count = np.histogram(beta_maxes[:,:,i], bins=np.linspace(90,180,15))
         # finding the PDF of the histogram using count values
@@ -59,25 +59,25 @@ def make_dual_cdf_plot(beta_maxes,betas,iwa_list,hab_zones,label_locations,rotat
         axis.plot(bins_count[1:], (1-cdf)*len(np.where(np.isfinite(beta_maxes[:,:,i]))[0])/beta_maxes[:,:,i].shape[0], color='C{:d}'.format(i),
                 linestyle='--',label=r"IWA = {:d} (mas)".format(iwa))
 
-        #### Now the circular case. 
+        #### Now the circular case.
         # (these ones are actually betas.)
 
-        #This is the analytical beta_max for a given iwa. 
[…]

### 2023-04-23 src/scripts/attic/run_simulation.py
Replace simulation of elliptical orbits with a parallelized version

@@ -1,13 +1,17 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.patches import Circle
+"""
+This is the original, non-parallelized version of the data generation
+script `simulate-elliptical-orbits.py`. It is kept here for reference.
 
-from phases import *
-from samplers import *
+[I double-checked the MD5 hash of the output files generated by this
+script against the output files generated by the parallelized version
+ensure that the results are identical.]
+"""
+
+from src.scripts.utils.phases import *
+from src.scripts.utils.samplers import *
 
 from tqdm import tqdm
-import time
-import paths
+from src.scripts.utils import paths
 
 np.random.seed(42) # set a seed so the plots and data output are reporducible for showyourwork!
 
@@ -37,7 +41,7 @@ print(f'option selected was {op}')
 from astropy.io import ascii
 
 fname = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
-t = ascii.read(paths.data / fname ,header_start = 1, data_start = 2)
+t = ascii.read(paths.data / fname, header_start = 1, data_start = 2)
 
 nstars = len(t)
 print(f'# read in {nstars} stars from {fname}')
@@ -58,8 +62,8 @@ for l in tqdm(t): # loop over all the stars in the table
 	# draw e and i from these distributions
 	esamp = sample_e(nplanets)
 	isamp = sample_i(nplanets) * u.deg
-	anodesamp = np.random.random_sample((nplanets,)) * 360*u.deg 
-	wsamp = np.random.random_sample((nplanets,)) * 360*u.deg 
+	anodesamp = np.random.random_sample((nplanets,)) * 360*u.deg
+	wsamp = np.random.random_sample((nplanets,)) * 360*u.deg
 
 #	for n in tqdm(np.arange(nplanets)):
 	for n in np.arange(nplanets):
@@ -111,4 +115,3 @@ print(f'saving the data with command line option {op}')
 np.savez_compressed(paths.data / f'iwa_all{tag}.npz', iwa=iwa, betamax=betamax, betamin=betamin)
 
 
-	

### 2023-04-23 src/scripts/simulate-elliptical-orbits.py

@@ -1,0 +1,176 @@
+"""
+Simulate 1,000 elliptical orbits for every planet in the target list.
+"""
+
+from warnings import catch_warnings, filterwarnings
+
+from astropy.io import ascii
+from joblib import Parallel, delayed
+from tqdm import tqdm
+
+import astropy.units as u
+import numpy as np
+
+from utils.phases import kep3d, xyztoscatter
+from utils.samplers import sample_e, sample_i
+from utils import paths
+
+
+# -----------------------------------------------------------------------------
+# DEFINITIONS
+# -----------------------------------------------------------------------------
+
+# noinspection PyUnresolvedReferences
+def simulate_orbit(
+    a: float,
+    e: float,
+    i: float,
+    w: float,
+    anode: float,
+) -> tuple[np.ndarray, np.ndarray]:
+    """
+    Simulate an elliptical orbit and return the minimum and maximum
+    scattering angles for each IWA.
+
+    Args:
+        a: The semi-major axis of the orbit (in mas).
+        e: The eccentricity of the orbit.
+        i: The inclination of the orbit (in degrees).
+        w: The argument of periapsis of the orbit (in degrees).
+        anode: The longitude of the ascending node of the orbit
+            (in degrees).
+
+    Returns:
+        scatmin: The minimum scattering angle for each IWA (in degrees).
+        scatmax: The maximum scattering angle for each IWA (in degrees).
+    """
+
+    P = 1.0 * u.year
+    tperi = 2050.0 * u.year
+
+    # Define the epochs to sample
+    epochs = np.linspace(tperi, tperi + P, n_epochs, endpoint=True)
+
+    # Calculate the orbit
+    _, _, Xsa, Ysa, Zsa, _, _, _ = kep3d(epochs, P, tperi, a, e, i, w, anode)
+
+    # Calculate current projected separation
+    rho = np.sqrt(Xsa * Xsa + Ysa * Ysa)
+
+    # Calculate scattering angle
+    scang = xyztoscatter(Xsa, Ysa, Zsa)
+
+    # For each IWA, set points on orbit that are behind the coronagraph to True
+    notvisible = rho[:, np.newaxis] < iwa
+
+    # Define dummy array in preparation for max and min
+    ttt = scang[:, np.newaxis] * np.ones(iwa.size)
+
+    # For each IWA column, set the points behind the mask as NaN
+    ttt[notvisible] = np.nan
+
[…]

### 2023-04-23 src/scripts/attic/plot_features_vs_iwa.py
Add new script for creating Figure 9

@@ -1,3 +1,8 @@
+"""
+This is the original script for generating Figure 9. It is not used in
+the paper, but is kept here for reference.
+"""
+
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt

### 2023-04-23 src/scripts/create-figure-9.py

@@ -1,0 +1,187 @@
+"""
+Create Figure 9: Number of planets accessible as a function of IWA,
+for different scattering features.
+"""
+
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+
+from warnings import catch_warnings, filterwarnings
+
+from matplotlib.lines import Line2D
+from scipy import interpolate as si
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from utils import paths
+from utils.plotting import set_fontsize, CBF_COLORS
+from utils.constants import (
+    ETA_EARTH,
+    LAMBDA_OVER_D_IN_MAS,
+    SCATTERING_FEATURES,
+)
+
+
+# -----------------------------------------------------------------------------
+# MAIN CODE
+# -----------------------------------------------------------------------------
+
+# Read in the target list
+df = pd.read_csv(paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv")
+
+# The angular separation of the Earth Equivalent Instellation (EEI) in mas
+hab_zones = np.array(df["EEIDmas"].values[1:], dtype=float)
+
+# Compute the minimum and maximum scattering angles (pre-computed)
+iwa = np.load(paths.data / "iwa_all3.npz")
+phase_max = iwa["betamax"]
+phase_min = iwa["betamin"]
+iwa_list = iwa["iwa"]
+
+# Create a new figure
+pad_inches = 0.025
+fig, axes = plt.subplots(
+    ncols=4,
+    figsize=(7 - 2 * pad_inches, 2.5 - 2 * pad_inches),
+)
+
+# Add twin axes for the top and right
+top_axes = [ax.twiny() for ax in axes]
+right_axes = [ax.twinx() for ax in axes]
+
+# Create a subplot for each scattering feature
+for k, (ax, top_ax, right_ax, feature_name) in enumerate(
+    zip(axes, top_axes, right_axes, SCATTERING_FEATURES.keys())
+):
+    # Set the fontsize for all axes
+    set_fontsize(ax=ax, fontsize=6)
+    set_fontsize(ax=top_ax, fontsize=6)
+    set_fontsize(ax=right_ax, fontsize=6)
+
+    # Add the title
+    ax.set_title(feature_name, fontsize=8, pad=10)
+
+    # Store values for plotting
+    lows = []
[…]

### 2023-04-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\usepackage{libertine}
-\usepackage[libertine]{newtxmath}
-\usepackage[scaled=0.76]{beramono}
+% Prettier fonts 
+% (We might have to go back to defaults for MNRAS submission)
+% \usepackage{libertine}
+% \usepackage[libertine]{newtxmath}
+% \usepackage[scaled=0.76]{beramono}
+
+% Those seem to be the "right" fonts for MNRAS
+\usepackage{newtxtext}
+\usepackage{newtxmath}
+
 \usepackage{microtype}
@@
     range-phrase = \textup{--},
+    per-mode = symbol,
 }
@@
 
-\usepackage{astro_bib_macro} %shortcuts of journals defined in this file
+% Load journal macros (e.g., \aa for A&A)
+% Source: https://ui.adsabs.harvard.edu/help/actions/journal-macros
+\usepackage{aas_macros}
+% \usepackage{astro_bib_macro}  % where is this one from?
 
@@
 \makeatletter
-\newcommand{\affiliationtarget}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\newcommand{\affdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}$^\mathrm{#1}$}
 \makeatother
-\newcommand{\affiliationlink}[1]{\hyperlink{#1}{#1}}
+\newcommand{\afflink}[1]{\hyperlink{#1}{#1}}
 
 % Chasing rainbows with the Habitable Worlds Observatory
-\title{Chasing \rainbows{} with the Habitable Worlds Observatory}
-%Scattering phase angle probed by directly imaged planets, or
+\title{Chasing \rainbows{} and ocean glint:\\  Inner working angle constraints for the Habitable Worlds Observatory}
+
+%phase angle probed by directly imaged planets, or
 %Don't ask us about contrast
@@
 \author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan$^{\affiliationlink{1},}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
-    Kimberly Bott$^{\affiliationlink{2},\affiliationlink{28}}$,
-    Sarah L. Casewell$^{\affiliationlink{3}}$,
-    Nicolas B. Cowan$^{\affiliationlink{4}}$,
-    David S. Doelman$^{\affiliationlink{5,19}}$,
+    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{\afflink{1}},
+    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
+    Sarah L. Casewell\textsuperscript{\afflink{5}},
+    Nicolas B. Cowan\textsuperscript{\afflink{6}},
+    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
     \newauthor
-    Timothy D. Gebhard$^{\affiliationlink{6},\affiliationlink{29}}$,
-    Matthew Kenworthy$^{\affiliationlink{5}}$,
-    Johan Mazoyer$^{\affiliationlink{7}}$,
-    Maxwell A. Millar-Blanchaer$^{\affiliationlink{8}}$,
+    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
+    Matthew Kenworthy\textsuperscript{\afflink{7}},
+    Johan Mazoyer\textsuperscript{\afflink{11}},
+    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
     \newauthor
-    Olivier Absil$^{16}$,
-    Lisa Altinier$^{13}$,
-    Pierre Baudoz$^{7}$,
-    Ruslan Belikov$^{24}$,
-    Alexis Bidot$^{9}$,
-    Markus J. Bonse$^{29}$,
+    Olivier Absil\textsuperscript{\afflink{13}},
+    Lisa Altinier\textsuperscript{\afflink{14}},
+    Pierre Baudoz\textsuperscript{\afflink{11}},
+    Ruslan Belikov\textsuperscript{\afflink{15}},
+    Alexis Bidot\textsuperscript{\afflink{16}},
+    Markus J. Bonse\textsuperscript{\afflink{10}},
     \newauthor
-    Bernhard Brandl$^{5}$,
-    Alexis Carlotti$^{9}$,
-    Elodie Choquet$^{13}$,
-    Niyati Desai$^{17}$,
-    Kevin Fogarty$^{24}$,
-    Jules Fowler$^{12}$,
+    Bernhard Brandl\textsuperscript{\afflink{7}},
+    Alexis Carlotti\textsuperscript{\afflink{16}},
+    Elodie Choquet\textsuperscript{\afflink{14}},
+    Niyati Desai\textsuperscript{\afflink{17}},
+    Kevin Fogarty\textsuperscript{\afflink{15}},
+    Jules Fowler\textsuperscript{\afflink{18}},
     \newauthor
-    Yann Gutierrez$^{7,14,22}$,
-    Olivier Guyon$^{15,25,26,27}$,
-    Sebastiaan Y. Haffert$^{15}$,
-    Olivier Herscovici-Schiller$^{14}$, 
+    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
+    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
+    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
+    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
     \newauthor
-    Adrien Hours$^{9}$,
-    Roser Juanola-Parramon$^{20,21}$,
-    Elina Kleisioti $^{5}$,
-    Lorenzo König$^{16}$,
-    Mariya Krasteva$^{23}$, 
+    Adrien Hours\textsuperscript{\afflink{16}},
+    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
+    Elina Kleisioti\textsuperscript{\afflink{7}},
+    Lorenzo König\textsuperscript{\afflink{13}},
+    Mariya Krasteva\textsuperscript{\afflink{27}},
     \newauthor
-    Iva Laginja$^{7}$,
-    Rico Landman$^{5}$,
-    Lucie Leboulleux$^{9}$,
-    David Mouillet$^{9}$,
-    Mamadou N’Diaye$^{10}$,
-    Emiel H. Por$^{18}$,
+    Iva Laginja\textsuperscript{\afflink{11}},
+    Rico Landman\textsuperscript{\afflink{7}},
+    Lucie Leboulleux\textsuperscript{\afflink{16}},
+    David Mouillet\textsuperscript{\afflink{16}},
+    Mamadou N’Diaye\textsuperscript{\afflink{28}},
     \newauthor
-    Laurent Pueyo$^{18}$,
-    Frans Snik$^{5}$,
-    Daphne M. Stam$^{30}$,
-    Victor Trees$^{31,32}$,
-    Dirk van Dam$^{5}$,
-    Kyle van Gorkom$^{15}$,
+    Emiel H. Por\textsuperscript{\afflink{29}},
+    Laurent Pueyo\textsuperscript{\afflink{29}},
+    Frans Snik\textsuperscript{\afflink{7}},
+    Daphne M. Stam\textsuperscript{\afflink{30}},
+    Victor Trees\textsuperscript{\afflink{31},\afflink{32}},
+    Dirk van Dam\textsuperscript{\afflink{7}},
     \newauthor
-    Maaike van Kooten$^{11}$ 
+    Kyle van Gorkom\textsuperscript{\afflink{21}},
+    Maaike van Kooten\textsuperscript{\afflink{33}}
     \newauthor \\%
@@
     % https://academic.oup.com/mnras/article/498/2/2354/5900562
-    % I've also added some LaTeX magic to create links for the affiliations; if y'all agree that it's a good idea I can extend it to all authors?
 }
@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
+NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
@@
 %
-Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
-%
-Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
-%
-The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
-%
-Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
-%
-We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
-%
-We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
-%
-The glint signature at planetary phase angles of $\sim\qtyrange{50}{70}{\degree}$ 
+Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to ocean glint, and to rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
+%
+Direct imaging missions are optimised for planets near quadrature, and \hwo's coronagraph may obscure the phase angles where such characteristic features are strongest. 
+%
+The range of accessible phase angles for a given exoplanet will depend on the planet's orbital inclination and/or the coronagraph's inner working angle (IWA). 
+%
+Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraphic obscuration. 
+%
+We use \hwo'~catalog of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering by gas. 
+%
+We find that any exo-Earths in systems in this catalog will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
+%
+The ocean glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
 %\sim$110--130$^\circ$ 
-would be accessible to the Habitable Worlds Observatory in $\sim\num{14}$ systems, and the rainbow of water clouds at phases of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} inner working angle (\IWA{}; $3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
-%
-Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
-%
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.  \end{abstract}
+would be accessible to \hwo~in $\sim\num{14}$ systems, and the rainbow due to water clouds at phase angles of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} \IWA{} ($3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
+%
+Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of roughly 2 and 3, respectively.
+%
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with the planet's orbital eccentricity and scales approximately inversely with the \IWA{}.
+\end{abstract}
 
@@
 %--------------------------------------------------------------------------------
+%\begin{figure*}
+    %\centering
+    %\includegraphics[width=\linewidth]%{figures/Fig1IWAPol.png}
+%    \caption{
+%        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit scattering angles 
+%        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
+%        The cloudy scenario includes a low-altitude, Earth-like water 
+%        cloud deck. 
+%        Below the plot, we are showing the scattering angle ranges that cover the
+%        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
+%        with each other).  
+%        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
+%    }
+%    \label{fig:bottplot}
+%\end{figure*}
+
 \begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-    \caption{
-        An illustration of phase curves for two homogeneous planets in an edge on system with oceans and an Earth-like atmosphere. 
-        The cloudy scenario includes an Earth-like water cloud deck at low altitudes. 
-        Below the plot, we are showing the features to which the phase angles and scattering angles are the most sensitive. 
-        In reality, as can be seen by the examples above, these features overlap.  
-        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (scattering angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest scattering angles).
+    \includegraphics[width=\linewidth]{figures/Phase_curves.pdf}
+    \caption{The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from 0$^\circ$ to 180$^\circ$ along its orbit).
+The wind speed over the ocean is 7~m/s.
+The clouds in the planet's atmosphere consist of spherical water droplets.
+The cloud coverage fraction is 0.5.
+The cloud pattern is patchy and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
+9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the 'noise' at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
+The
+    polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). Note that the reference plane for the
+    polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
+ %   and the latter
+ %   is equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
+        %The cloudy scenario includes a low-altitude, Earth-like water 
+        %cloud deck. 
+        Below the plot, we show the phase angle ranges that cover the
+        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
+        with each other).  
+        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
     }
@@
 \end{figure*}
+
 %--------------------------------------------------------------------------------
 
-Signatures of habitability, from ocean glint to optical effects from condensates---like the rainbow---should be some of the most distinct signals in an exoplanet's reflected light phase curve, particularly in polarization.
-%
-However, depending on system architectures, some of these features can be obscured by the inner working angle of a coronagraph.
+The field of exoplanet research has made great advances in the last few decades.
+The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve.
+The Habitable Worlds Observatory will perform a dedicated direct imaging survey of nearby stars searching for rocky planets in the habitable zone.
+This space-based coronographic telescope will detect and characterise exoplanets using their reflection spectrum.
+Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
+
+Variations of the flux and in particular the degree of polarization of
+light of the parent star that is reflected by an exoplanet as it moves
+along its orbit that are due to an ocean glint or optical phenomena from scattering by condensates --- like the rainbow --- could be some of the most distinct signatures of habitability.
+%
+However, depending on the architecture of a planetary system, some of these features can be obscured by the inner working angle (\IWA{}) of a telescope's coronagraph.
 
@@
 %
-The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
-%
-However, there are many factors besides the stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and, for example, geological activity.
+The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
+%
+However, there are many factors besides the incident stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and, for example, geological activity.
 %
@@
 %
-Providing additional context to the habitability of an exoplanet is the presence of a water cycle, as evidenced by the identification of a rainbow and other optical effects that can be attributed to water clouds.
+Providing additional context to the habitability of an exoplanet would be the presence of a water cycle, as evidenced by the identification of a glint, i.e.\ the specular reflection of incident light by a liquid surface, or by a rainbow and other optical phenomena that can be attributed to light that is scattered by water clouds, as detailed in the next sections.
 
@@
 
-An exoplanet observed in an edge-on orbit will move through all phases and thus all scattering angles, while those viewed at higher inclinations will pass through a subset of these scattering angles.
-%
-If the planet is rotating, its reflected light signal is expected to show regular variations that can place constraints on the horizontal heterogeneity of the planet.
-
-The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scatter, rainbows, and glories.
-%
-An added potential benefit of observing polarized light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as these stars themselves can be considered to be unpolarized \citep{kemp1987}.
+An exoplanet in an edge-on orbit moves through phase angles ranging from $0^\circ$ (full phase), through $90^\circ$ (quadrature), and to $180^\circ$ (the planet's nightside is towards the observer) (cf.\ Fig.~\ref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
+%
+If the planet is rotating, its reflected light signal in both total and polarized flux is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet \citep{2001Natur.412..885F,stam2008,2022A&A...664A..59M}.
+
+The strengths of detailed total and polarized flux phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
+%
+An added potential benefit of observing the state of polarization of the reflected light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
@@
 %As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
-Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces.
-%
-In the total flux, the glint signal will be most prominent at small planetary phase (large scattering) angles.
-%
-The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at \qty{53}{\degree} \citep[i.e., a scattering angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
-%
-This angle therefore varies slightly with wavelength and with sea and atmosphere composition.
-%
+Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces. While glint is a familiar phenomenon on Earth, there is only
+one other body in the Solar System where it has been observed: 
+the Cassini orbiter captured the glint of sunlight reflecting off the
+liquid hydrocarbon lakes near Titan's north pole as it was observing
+the moon from the nightside \citep{2010GeoRL..37.7104S}.
+%
+In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
+%
+The maximum polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
+%
+This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
+%
+
 In general, the glint on a planet will be strongest at red wavelengths where the Rayleigh scattering in the planet's atmosphere and attenuation through the atmosphere are smallest \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also small \citep{2012ApJ...752L...3C}.
 %
-The strengthened glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
-%
-{\color{red}{check the next sentence:}}
-The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across scattering angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
+The stronger glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+%
+The width across phase angles of the glint feature in total and polarized flux generally
+broadens with increasing wave height \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citep{CoxMunk1954}.
+%
 
@@
 %
-The forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
-%
-This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary phase angle depending on the cloud coverage fraction \citep{treesandstam2019}.
+At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
+%
+This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}, as shown in Fig.~\ref{fig:bottplot}.
 %The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
@@
 %
-While the glint feature can only be probed when a planet is at crescent phases, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry when the planet is at phase angles where it is at the wide separations more easily probed when using coronagraphs.
+While the glint feature can only be probed when a planet is at crescent, i.e.\ large phase angles, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
+This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
 %how use to map, 
@@
 
-\subsection{Rainbows and glories}
-
-Rainbows, glories and other optical effects from condensates provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,2014A&A...566L...1G}.
-%
-In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to haze rather than a surface \citep{zellner1973polarization}, the phase curve behaviour of Venus showed not only that the clouds were $\sim$75\% H$_2$SO$_4$--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
-The glory feature in Venus's phase curve was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
-
-In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
+\subsection{Rainbows, cloudbows and glories}
+
+Rainbows, cloudbows, glories and other optical phenomena from light that is 
+scattered by atmospheric condensates can dramatically influence the shape of a planet's total flux and polarized phase curves and provide a wealth of information 
+about exoplanets \citep{karalidi2012rainbow, stam2008,Bailey2007,2014A&A...566L...1G}.%
+\footnote{
+    Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light.
+Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white \citet[see][for further discussion]{Bailey2007}.
+}
+%
+In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the polarized phase curve of Venus at three wavelengths showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+The glory feature in Venus's total flux phase curve was later used to rule out potential solutions for the unknown UV absorber \citep{petrova2018glory}.
+
+In unpolarized light {\color{Green}{do you mean polarized light?}}, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
 Lambertian reflecting planet.
 %
-{\color{red}{check the sentence:}}
-While the glory and rainbow are technically visible in any phase curve
-in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
-%
-The rainbow angle depends on the refractive index of the scatterers and thus depends on the composition and the wavelength. 
+While the glory and rainbow are technically visible both in the total flux
+and the polarized flux phase curves
+in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized flux than in total flux \citep{karalidi2011, stam2008, treesandstam2019}.
+%
+The rainbow angle depends on the refractive index of the scatterers and thus depends on their composition and the wavelength. 
 The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
 %
-Because the rainbow feature is more narrow than the glint, it would 
-allow for a more exact determination of the particle composition.
+Because the rainbow feature is narrower than the glint due to the surface reflection, it would 
+allow for a more exact determination of the composition of the condensates.
 %
 Note that the scattering by clouds that consist of solid particles, 
-such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
-%
-
-In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths.
-%
-Pushing observations to angles closer to the star (smaller scattering angles, i.e., larger phase angles) would allow the determination of the full shape of this rainbow, and enable the identification of rainbows due to the scattering of light by other possible HZ condensates.
+such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
+%
+
+In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of 40$^\circ$ (see Fig.~\ref{fig:bottplot}).
+%
+Pushing observations to angles closer to the star (smaller phase angles, i.e., larger scattering angles) would allow the determination of the full shape of this rainbow, and enable the identification of rainbows due to the scattering of light by other possible HZ condensates.
 %
@@
 %is produced when the droplet radii much smaller than the wavelength of light 
-can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within \qty{20}{\degree} scattering angle very close to full phase \citep{hansentravis1974}.
+can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
 %
@@
 %
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes along a planet's orbits.
 %
 We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}. 
-It is important to note that the phases for the angular features throughout this paper are approximate, that features overlap, and that their maximum peak will usually vary with particle composition and wavelength.
-%the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
-% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.
+It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
+
+The conventions for angle names vary throughout the literature on exoplanet phase curves.
+In this paper we define the \textbf{scattering} angle as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
+A planet with a face-on orbit will thus have a phase angle of 90$^\circ$ everywhere along its orbit.
+In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
+We also note that we are considering here the (linearly) polarized flux
+which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information could be garnered from full Stokes characterization.
+
+%the   the way that inclination and coronagraph inner working angle can limit the accessible scattering angles  
+% We minimally explore the contrast ratio as we are primarily exploring whether these phase angles are reachable with inner working angle and outer working angle considered.
 
@@
 %\begin{enumerate}
-%    \item \st{Cartoon defining various phase angles: orbital phase, scattering phase, $\beta$ ? (Sophia)}
+%    \item \st{Cartoon defining various scattering angles: orbital phase, scattering phase, $\beta$ ? (Sophia)}
 %    \item Bott plot: annotated polarized and unpolarized phase curves for ocean world and cloudy worlds (Kim)
@@
 For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths of $\lambda = \qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
 
@@
     \caption{
-        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., \IWA{} and \OWA{}, respectively).
+        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., the \IWA{} and \OWA{}, respectively).
         %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
@@
 
-\hwo{} is envisioned to observe the stars in the NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory.%
-\footnote{\hwo{} target list: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
+\hwo{} is envisioned to observe the stars in the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
+\footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
 %
@@
 %
-By design, such planets should be observable by \hwo{} at least at their quadrature phase.
+By design, such planets should be observable by \hwo{} at least at their quadrature phase ($\alpha=90^\circ$).
 
@@
 
-\subsection{Maximum scattering phase angle coverage}
+\subsection{Maximum phase angle coverage}
 \label{subsec:2.2}
 
-A coronagraph can obscure part of the planet's orbit which, as shown in \cref{fig:annotated-orbit}, can prevent the planet from being observable at high and low scattering angles. 
-%
-The range of observable scattering phase angles will depend on the \IWA{} of the coronagraph, as well as on the orientation and shape of the exoplanet's orbit.
+A coronagraph can obscure part of a planet's orbit which, as illustrated in \cref{fig:annotated-orbit}, can prevent the planet from being observable at small and large phase angles, where it is close to its star.
+%
+The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the inclination and shape, i.e.\ eccentricity, of the exoplanet's orbit.
 
@@
     \caption{
-        An illustration of a planet in an edge-on circular orbit showing how a coronagraph can obscure the exoplanet at high and low scattering phase angles, and thus prevent observations of scattering features in an exoplanet's phase curve that could help to identify its atmospheric and/or surface composition and habitability.
-        The phase angle, $\varphi$, is shown in red and is the complement of the scattering angle in edge on systems.
+        An illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of scattering features in an exoplanet's phase curves that could help to characterize its atmospheric and/or surface composition and habitability.
+        The scattering angle, $\varphi$, is shown in red. It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
     }
@@
 
-In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering phase angles are symmetric about 90 degrees.
+For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about 90$^\circ$ (see Fig.~\ref{fig:accessible_phase_angles}).
 %
@@
 %
-The scattering phase angle $\varphi$ is given by
+Angle $\varphi$ is given by
 \begin{equation}
     \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,.
+    \label{eq:sinphi}
 \end{equation}
 %
-Using small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the angular separation on-sky.
-%
-We access the minimum scattering angle when the projected distance reaches the inner working angle separation $\delta = \mathrm{IWA}$. 
-%
-Using \cref{eq:Delta_phi}, we deduce:
+Assuming small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the on-sky angular separation.
+%
+We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
+%
+Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
 %
@@
     \label{eq:scattering_angle}
-    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA \cdot d_*}}{a} \,.
+    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA} 
+    \cdot d_*}{a} \,.
 \end{equation}
@@
         The on-sky angular separation as viewed by the observer is $\delta$.
-        The scattering phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.  
+%        Angle $\varphi$ is measured between the line of sight to the star and the line connecting the star and planet.  
     }
@@
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature, while the rainbow angle will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature ($\alpha=90^\circ$), while the rainbow angle (around $\alpha=40^\circ$) will only be accessible for a subset of exoplanets.
 
@@
     \centering
-    \includegraphics{figures/scatterplot.pdf}
+    \includegraphics{figures/figure-4-scatterplot.pdf}
+    \script{create-scatterplot.py}
     \caption{
@@
     \label{fig:scatterplot}
-    \script{create-scatterplot.py}
 \end{figure*}
@@
 
-Most planetary systems will not be viewed as edge-on from Earth.
-%
-For a system with inclination angle $i$, we can identify the following 2 regimes (see also \cref{fig:orb-grid}):
+Most planets are not in edge-on orbits as seen from Earth.
+%
+For an orbit with an inclination angle $i$, we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
 \begin{enumerate}
     \item The coronagraph does not obscure any part of the sky-projected orbit: 
-    the maximum scattering phase angle coverage depends solely on $i$. This 
+    the maximum phase angle coverage depends solely on $i$. This 
     holds for orbits that are relatively face-on; 
     \item The coronagraph does obscure parts of the sky-projected orbit: 
-    the scattering phase angle coverage depends only on the \IWA{} and 
+    the phase angle coverage depends only on the \IWA{} and 
     the orbital semi-major axis $a$. 
@@
 %--------------------------------------------------------------------------
-\begin{figure}
-   \centering
-   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
-   \caption{
-        The scattering phase angles accessible to a direct imaging mission 
-        can be limited by either the orbital inclination angle $i$ or the 
-        coronagraphic \IWA{}.
-        %
-        If the orbit is close to face-on (top panels), the scattering phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
-        %
-        For orbits close to edge-on (bottom panels) the planet attains all 
-        phases, but the scattering phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
-    }
-    \label{fig:orb-grid}
-    \script{plot-orb-grid.py}
-\end{figure}
+%\begin{figure}
+%   \centering
+%   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
+%   \script{create-figure-5-orb-grid.py}
+%   \caption{
+%        The phase angles accessible to a direct imaging mission 
+%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
+        %
+%        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
+        %
+%        For orbits close to edge-on (bottom panels) the planet attains all 
+%        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+%    }
+%    \label{fig:orb-grid}
+%\end{figure}
 %--------------------------------------------------------------------------
@@
 
-For each star on the \hwo{} target list, we generate \num{1000} random orbits of hypothetical habitable planets with Earth-equivalent-flux semi-major axes. 
+For each star on the \hwo{} target list, we generate \num{1000} random orbits of hypothetical habitable planets in orbits with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
 %
 The orbital inclination angles $i$ are randomly drawn from a distribution 
-uniform in $\cos i$. 
+that is uniform in $\cos i$. 
 %
 We then use \cref{eq:Delta_phi_max} to determine the range of 
-accessible scattering phase angles for each hypothetical planet.
+accessible phase angles for each hypothetical planet.
 %
 The solid lines in \cref{fig:betaallofit} show the cumulative 
-distribution of the maximum and minimum accessible scattering phase 
+distribution of the maximum and minimum accessible phase 
 angles normalised to the total number of systems in the \hwo{} 
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
+    \includegraphics[width=0.95\textwidth]{figures/figure-6-n-over-phase.pdf}
+    \script{create-figure-6-n-over-phase.py}
     \caption{
-        \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
-        The cumulative distributions of the most extreme scattering phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
+        The cumulative distributions of the most extreme phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
         (solid lines) and randomly inclined, elliptical orbits (dashed lines).
@@
         The top $x$-axis indicates the minimum and the 
-        bottom $x$-axis the maximum accessible scattering phase angle.
+        bottom $x$-axis the maximum accessible phase angle.
         %
@@
     \label{fig:betaallofit}
-    \script{plot_n_versus_phase.py}
 \end{figure*}
@@
 We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
-Eccentricities are drawn from the beta distribution, and inclinations are uniform in $\cos i$. 
+Orbital eccentricities are drawn from the beta distribution, and 
+inclination angles are uniform in $\cos i$. 
 \Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
@@
 
-We then calculate the minimum and maximum accessible scattering phase 
+We then calculate the minimum and maximum accessible phase 
 angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
@@
 The dashed lines in \cref{fig:betaallofit} show the normalised cumulative 
-distribution of the accessible angles for each of \IWA for eccentric orbits. 
-%
-We also calculate which target planets would be observable at the scattering phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+distribution of the accessible phase angles for each value of the 
+\IWA{} for eccentric orbits. 
+%
+We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
 %
@@
 A second parameter that is crucial for detecting angular features like 
-rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the flux of the reflected light from the planet divided by the stellar flux. 
+rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e.\ 
+the ratio of the flux of the starlight that is reflected by the planet 
+and the direct stellar flux. 
 %
 While this is not the main focus of this paper, it is important to consider that the total flux contrast and the polarized contrast generally 
-depend on the scattering phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
+depend on the phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
 %
 Here, we aim to present some simplified examples of the impact of the 
-scattering angle dependent contrast on the detectability of rainbow and 
+phase angle dependent contrast on the detectability of rainbow and 
 glint features. 
@@
     = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
+\label{eq:contrast}
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at phase angle $\alpha$
-(normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.
-\todo{here, $\alpha$ is used for the phase angle ... and is the 
-'integral phase function' the same as the 'scattering phase function'?}
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$
+(normalized to \num{1.0} at full phase, thus at $\alpha=0^\circ$), 
+and $a$ is the distance between the planet and its host star.
 %
 The assumptions that went into the contrast estimations for \hwo{}
-are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
+are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
 %
@@
 Therefore, in this work, we will assume that the calculated contrasts 
-in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
+in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
 %
@@
 
-We use the scattering phase functions from \citet{treesandstam2019} 
+We use the phase functions from \citet{treesandstam2019} 
 for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
@@
 %
-In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given an inclination of \num{90} degrees.
+In addition, we map the orbital phase for circular orbits to the phase angle and the on-sky separation given an inclination of \num{90} degrees.
 %
@@
    \centering
-   \includegraphics[width=0.99\textwidth]{figures/Contrast_vs_separation_inclination_angle_90.pdf}
+   \includegraphics[width=0.99\textwidth]{figures/figure-7-contrast-over-separation_i=90.pdf}
+    \script{create-figure-7-contrast-over-separation.py}
    \caption{
-    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of \qty{90}{\degree} for the stars $\alpha$~Cen~A, $\epsilon$~Eri and Lalande~21185.
-    %
-    The solid line indicates the contrast in unpolarized light, with the contrast at quadrature marked by a solid dot.
-    %
-    The polarized component is indicated by the coloured dots, for which the colour represents the scattering phase angle from quadrature.
-    %
-    The light grey points show the quadrature contrasts of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
+    The contrast (Eq.~\ref{eq:contrast}) and orbital separation 
+    of an Earth-like planet with an ocean surface and patchy clouds 
+    along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A, $\epsilon$~Eri, and Lalande~21185.
+    %
+    The solid, black line indicates the contrast in unpolarized light, 
+    with the contrast at quadrature marked by the black dot.
+    %
+    The polarized component is indicated by the coloured dots, with the colour representing the phase angle from quadrature.
+    %
+    The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
     }
     \label{fig:contrasts}
-    \script{plot_contrasts.py}
 \end{figure*}
@@
 I ran 1000 sims for each target with a random eccentric orbit, then pulled out the betamax and betamin angles for each of the 1000 realisations.
-Max's plots encode the completeness as a function of phase angle, showing that you get about 10 percent more phenomena.
+Max's plots encode the completeness as a function of scattering angle, showing that you get about 10 percent more phenomena.
 }
@@
 
-As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable effect on the number of systems that can be observed at each scattering phase angle.
+As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable effect on the number of systems that can be observed at each phase angle.
 %
@@
 The eccentricity does not significantly affect the number of systems for \IWA{} smaller than cut used in the selection of the systems in the \hwo{} target list. 
-It is, however, important to account for orbital eccentricity when the \IWA{} of the coronagraph is larger than the \IWA{} used to filter the target list; otherwise, the number of systems observable at extreme phase angles would be under-predicted.
+It is, however, important to account for orbital eccentricity when the \IWA{} of the coronagraph is larger than the \IWA{} used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
 
@@
 
-\subsection{Observable scattering phenomenon}
+\subsection{Observable optical phenomenon}
 \label{sec:results_scattering_phenomena}
 
-In this section, we consider the number of stars around which different scattering phenomena could be observable for different \IWA{}.
-%
-We first define the phase angles over which various of these phenomena are accessible in \cref{tab:phase_ranges}.
-%
-We note that these ranges should all be considered approximate, as the exact phase angles where these effects are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
+In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA{}.
+%
+We first define the scattering angles over which various of these phenomena are accessible in \cref{tab:phase_ranges}.
+%
+We note that these ranges should all be considered approximate, as the exact scattering angles where these phenomena are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
 %
@@
     \caption{
-        The ranges of phase angle for the different scattering phenomena considered here.
+        The approximate ranges of scattering angles for the different optical phenomena considered in this paper.
     }
@@
         \midrule
-        Rainbows    & 22    & 42    & 63 \\
-        Rayleigh    & 50    & 70    & 110 \\
-        Ocean Glint & 130   & 150   & 170 \\
-        Glory       & 0     & 5     & 10 \\
+        Rainbow    & 22$^\circ$    & 42$^\circ$    & 63$^\circ$ \\
+        Rayleigh    & 50$^\circ$    & 70$^\circ$    & 110$^\circ$ \\
+        Ocean Glint & 130$^\circ$   & 150$^\circ$   & 170$^\circ$ \\
+        Glory       & 0$^\circ$     & 5$^\circ$     & 10$^\circ$ \\
         \bottomrule
@@
 
-The Rayleigh scattering feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-%
-The rainbow feature appears further from quadrature, but will still be observable in a significant fraction of systems.
-%
-The position and shape of the peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
-%
-The ocean glint feature occurs even further from quadrature and so is observable in fewer systems. 
-%
-In most systems, only the first half of the feature is observable, but this can still be used to identify a liquid ocean on the surface of the planet.
+The Rayleigh feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
+%
+The rainbow feature appears further from quadrature (around $\alpha=40^\circ$), but will still be observable in a significant fraction of systems.
+%
+The position and shape of the rainbow peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
+%
+The ocean glint feature occurs even further from quadrature and so would be observable in fewer systems. 
+%
+In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet.
 %
@@
     \caption{
-        Number of stars around which the peak phase angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
+        Number of stars around which the peak scattering angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
     }
@@
 
-\subsection{Contrast Curves}
+\subsection{Contrast curves}
 \label{sec:results_contrast}
 
-The accessible scattering phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
+The accessible phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
 %
@@
 %
-However, when its design is being finalised, it will be important to consider both the accessible scattering phase angles and contrast limits when deciding on the \IWA{} of the coronagraph, if these features are to be observed.
+However, when its design is being finalised, it will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA{} of the coronagraph, if these features are to be observed.
 
@@
     \centering
-    \includegraphics{figures/ball_of_yarn.pdf}  
+    \includegraphics{figures/figure-9-yarnball.pdf}  
+    \script{create-figure-8-yarnball.py}
     \caption{
@@
     \label{fig:ball-o-yarn}
-    \script{make_yarn_plot.py} % most awesome command in this paper.  This knitter approves.
 \end{figure*}
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/nplanets_vs_iwa.png}  
+    \includegraphics{figures/figure-9-n-over-iwa.pdf}
+    \script{create-figure-9-n-over-iwa.py}
     \caption{
-        The number of planetary systems for which key optical phenomena 
-        (Rayleigh scattering, rainbow, ocean glint, glories) would be 
-        observable as functions of the \IWA{}.
-        %
-        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak of the phenomena is observable (solid lines), and where the end is observable (dot-dashed lines).
+        The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glories) would be observable as functions of the \IWA.
+        %
+        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
         %
@@
         It is worth noting that the Glory would be observable in only one system: Alpha~Cen~A.
-        \todo{What is the reasoning with the colours?}
     }
     \label{fig:accessible_phase_angles}
-    \script{plot_features_vs_iwa.py}
 \end{figure*}
@@
 %
-We primarily focused on the range of accessible scattering phase angles given the limitations of the coronagraphic \IWA{} and aimed to investigate which phenomena could potentially be observed.
+We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA{} and aimed to investigate which phenomena could potentially be observed.
 %
@@
 %
-\todo{the paragraph below seems similar to the one above ... merge them?}
-We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of phase angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
-%
-This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
+Additionally, decreasing the \IWA\ also increases the range of scattering angles observable in these systems. 
+%
+Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
 %
@@
 This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
+
+%\todo{the paragraph below seems similar to the one above ... merge them?}
+%We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA\ of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of scattering angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
+%
+%This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
+%
+%While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
+%
+%This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
 
@@
 
-The workshop on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, Netherlands.
-%
-This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No. 866001 - EXACT).
-%
+%The workshop on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, Netherlands.
+%
+%This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No. 866001 - EXACT).
+%
+%This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
+%
+%SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
+%
+%SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
+%
+%TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
+%
+%Specific grant funding and any other specific requirements
+%
+%KMB acknowledges support from NASA Habitable
+%
+%Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
+%
+%This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
+
+The Optimal Exoplanet Imaging workshop (2023) on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, The Netherlands. 
+%
+The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).
+
 This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France.
-%
+
 SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
-%
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
-%
 TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-%
-Specific grant funding and any other specific requirements
-%
-KMB acknowledges support from NASA Habitable
-%
-Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
-%
+KMB acknowledges support from NASA Habitable Worlds grant No.
+80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
+ILa acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
+PBa, ILa, and YGu were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
+EHP is supported by the NASA Hubble Fellowship grant \#HST-HF2-51467.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
+L.A. and E.C. acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement No. 101044152).
+O.A. and L.K. acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No 819155)
+SYH was funded by the generous support of the Heising-Simons Foundation.
+RJP is supported by NASA under award number 80GSFC21M0002.
+
 This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \clearpage
+
 % References
@@
 \begingroup
-% \footnotesize
     \itshape
-\affiliationtarget{1}{$^{1}$} Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK\\
-\affiliationtarget{2}{$^{2}$} Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\
-\affiliationtarget{3}{$^{3}$} Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK\\
-\affiliationtarget{4}{$^{4}$} Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada\\
-\affiliationtarget{5}{$^{5}$} Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
-\affiliationtarget{6}{$^{6}$} Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 Tübingen, Germany \\
-$^{7}$ LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, F-92195 Meudon, France \\
-$^{8}$ Department of Physics, University of California, Santa Barbara, CA 93106, USA \\
-$^{9}$ Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\
-$^{10}$ Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\
-$^{11}$ National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C. Canada \\
-$^{12}$ University of California, Santa Cruz, USA  \\
-$^{13}$ Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\
-$^{14}$ DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\
-$^{15}$ Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, Arizona \\
-$^{16}$ STAR Institute, Université de Li\`ege, All\'ee du Six Ao\^ut 19c, 4000 Li\`ege, Belgium \\
-$^{17}$ Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\
-$^{18}$ Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\
-$^{19}$ SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\
-$^{20}$ NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\
-$^{21}$ University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\
-$^{22}$ DOTA, ONERA, 92322 Châtillon, France \\
-$^{23}$ European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\
-$^{24}$ NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\
-$^{25}$ Subaru Telescope, NAOJ, USA \\
-$^{26}$ College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\
-$^{27}$ Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\
-\affiliationtarget{28}{$^{28}$} NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\
-\affiliationtarget{29}{$^{29}$} ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland \\
-\affiliationtarget{30}{$^{30}$} Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\
-\affiliationtarget{32}{$^{31}$} Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, the Netherlands \\
-\affiliationtarget{31}{$^{32}$} Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\
+    \raggedright
+    \affdest{1}Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK \\[0.5mm]
+    \affdest{2}Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\[0.5mm]
+    \affdest{3}NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\[0.5mm]
+    \affdest{4}NASA Nexus for Exoplanet System Science, Terrestrial Polarization Team, Orlando, FL 32826, USA \\[0.5mm]
+    \affdest{5}Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK \\[0.5mm]
+    \affdest{6}Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada \\[0.5mm]
+    \affdest{7}Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands \\[0.5mm]
+    \affdest{8}SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\[0.5mm]
+    \affdest{9}Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 Tübingen, Germany \\[0.5mm]
+    \affdest{10}ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland \\[0.5mm]
+    \affdest{11}LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, 92195 Meudon, France \\[0.5mm]
+    \affdest{12}Department of Physics, University of California, Santa Barbara, CA 93106, USA \\[0.5mm]
+    \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
+    \affdest{14}Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
+    \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
+    \affdest{16}Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\[0.5mm]
+    \affdest{17}Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\[0.5mm]
+    \affdest{18}University of California, Santa Cruz, USA \\[0.5mm]
+    \affdest{19}DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\[0.5mm]
+    \affdest{20}DOTA, ONERA, 92322 Châtillon, France \\[0.5mm]
+    \affdest{21}Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85719, USA \\[0.5mm]
+    \affdest{22}Subaru Telescope, NAOJ, USA \\[0.5mm]
+    \affdest{23}College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\[0.5mm]
+    \affdest{24}Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\[0.5mm]
+    \affdest{25}NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\[0.5mm]
+    \affdest{26}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
+    \affdest{27}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
+    \affdest{28}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
+    \affdest{29}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
+    \affdest{30}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
+    \affdest{31}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
+    \affdest{32}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
+    \affdest{33}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
 \endgroup

### 2023-04-24 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-04-19 src/tex/ms.tex
@@
 \makeatletter
-\newcommand{\affiliationtarget}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\newcommand{\affdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}$^\mathrm{#1}$}
 \makeatother
-\newcommand{\affiliationlink}[1]{\hyperlink{#1}{#1}}
+\newcommand{\afflink}[1]{\hyperlink{#1}{#1}}
 
@@
 \title{Chasing \rainbows{} with the Habitable Worlds Observatory}
+
 %Scattering phase angle probed by directly imaged planets, or
@@
 \author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan$^{\affiliationlink{1},}$\thanks{Correspondence:  \url{sophia.vaughan@physics.ox.ac.uk}}
-    Kimberly Bott$^{\affiliationlink{2},\affiliationlink{28}}$,
-    Sarah L. Casewell$^{\affiliationlink{3}}$,
-    Nicolas B. Cowan$^{\affiliationlink{4}}$,
-    David S. Doelman$^{\affiliationlink{5,19}}$,
+    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
+    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
+    Sarah L. Casewell\textsuperscript{\afflink{5}},
+    Nicolas B. Cowan\textsuperscript{\afflink{6}},
+    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
     \newauthor
-    Timothy D. Gebhard$^{\affiliationlink{6},\affiliationlink{29}}$,
-    Matthew Kenworthy$^{\affiliationlink{5}}$,
-    Johan Mazoyer$^{\affiliationlink{7}}$,
-    Maxwell A. Millar-Blanchaer$^{\affiliationlink{8}}$,
+    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
+    Matthew Kenworthy\textsuperscript{\afflink{7}},
+    Johan Mazoyer\textsuperscript{\afflink{11}},
+    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
     \newauthor
-    Olivier Absil$^{16}$,
-    Lisa Altinier$^{13}$,
-    Pierre Baudoz$^{7}$,
-    Ruslan Belikov$^{24}$,
-    Alexis Bidot$^{9}$,
-    Markus J. Bonse$^{29}$,
+    Olivier Absil\textsuperscript{\afflink{13}},
+    Lisa Altinier\textsuperscript{\afflink{14}},
+    Pierre Baudoz\textsuperscript{\afflink{11}},
+    Ruslan Belikov\textsuperscript{\afflink{15}},
+    Alexis Bidot\textsuperscript{\afflink{16}},
+    Markus J. Bonse\textsuperscript{\afflink{10}},
     \newauthor
-    Bernhard Brandl$^{5}$,
-    Alexis Carlotti$^{9}$,
-    Elodie Choquet$^{13}$,
-    Niyati Desai$^{17}$,
-    Kevin Fogarty$^{24}$,
-    Jules Fowler$^{12}$,
+    Bernhard Brandl\textsuperscript{\afflink{7}},
+    Alexis Carlotti\textsuperscript{\afflink{16}},
+    Elodie Choquet\textsuperscript{\afflink{14}},
+    Niyati Desai\textsuperscript{\afflink{17}},
+    Kevin Fogarty\textsuperscript{\afflink{15}},
+    Jules Fowler\textsuperscript{\afflink{18}},
     \newauthor
-    Yann Gutierrez$^{7,14,22}$,
-    Olivier Guyon$^{15,25,26,27}$,
-    Sebastiaan Y. Haffert$^{15}$,
-    Olivier Herscovici-Schiller$^{14}$, 
+    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
+    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
+    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
+    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
     \newauthor
-    Adrien Hours$^{9}$,
-    Roser Juanola-Parramon$^{20,21}$,
-    Elina Kleisioti $^{5}$,
-    Lorenzo König$^{16}$,
-    Mariya Krasteva$^{23}$, 
+    Adrien Hours\textsuperscript{\afflink{16}},
+    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
+    Elina Kleisioti\textsuperscript{\afflink{7}},
+    Lorenzo König\textsuperscript{\afflink{13}},
+    Mariya Krasteva\textsuperscript{\afflink{27}},
     \newauthor
-    Iva Laginja$^{7}$,
-    Rico Landman$^{5}$,
-    Lucie Leboulleux$^{9}$,
-    David Mouillet$^{9}$,
-    Mamadou N’Diaye$^{10}$,
-    Emiel H. Por$^{18}$,
+    Iva Laginja\textsuperscript{\afflink{11}},
+    Rico Landman\textsuperscript{\afflink{7}},
+    Lucie Leboulleux\textsuperscript{\afflink{16}},
+    David Mouillet\textsuperscript{\afflink{16}},
+    Mamadou N’Diaye\textsuperscript{\afflink{28}},
     \newauthor
-    Laurent Pueyo$^{18}$,
-    Frans Snik$^{5}$,
-    Daphne M. Stam$^{30}$,
-    Victor Trees$^{31,32}$,
-    Dirk van Dam$^{5}$,
-    Kyle van Gorkom$^{15}$,
+    Emiel H. Por\textsuperscript{\afflink{29}},
+    Laurent Pueyo\textsuperscript{\afflink{29}},
+    Frans Snik\textsuperscript{\afflink{7}},
+    Daphne M. Stam\textsuperscript{\afflink{30}},
+    Victor Trees\textsuperscript{\afflink{31},\afflink{32}},
+    Dirk van Dam\textsuperscript{\afflink{7}},
     \newauthor
-    Maaike van Kooten$^{11}$ 
+    Kyle van Gorkom\textsuperscript{\afflink{21}},
+    Maaike van Kooten\textsuperscript{\afflink{33}}
     \newauthor \\%
@@
 %
-Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering effects. 
+Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering phenomena. 
 %
@@
     \caption{
-        An illustration of phase curves for two homogeneous planets in an edge on system with oceans and an Earth-like atmosphere. 
-        The cloudy scenario includes an Earth-like water cloud deck at low altitudes. 
-        Below the plot, we are showing the features to which the phase angles and scattering angles are the most sensitive. 
-        In reality, as can be seen by the examples above, these features overlap.  
+        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit phase angles 
+        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
+        The cloudy scenario includes a low-altitude, Earth-like water 
+        cloud deck. 
+        Below the plot, we are showing the phase angle ranges that cover the
+        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
+        with each other).  
         Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (scattering angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest scattering angles).
@@
 
-Signatures of habitability, from ocean glint to optical effects from condensates---like the rainbow---should be some of the most distinct signals in an exoplanet's reflected light phase curve, particularly in polarization.
-%
-However, depending on system architectures, some of these features can be obscured by the inner working angle of a coronagraph.
+Variations of the flux and in particular the degree of polarization of
+light of the parent star that is reflected by an exoplanet as it moves
+along its orbit that are due to an ocean glint or optical phenomena from scattering by condensates --- like the rainbow --- could be some of the most distinct signatures of habitability.
+%
+However, depending on system architectures, some of these features can be obscured by the inner working angle (\IWA{}) of a telescope's coronagraph.
 
@@
 %
-Providing additional context to the habitability of an exoplanet is the presence of a water cycle, as evidenced by the identification of a rainbow and other optical effects that can be attributed to water clouds.
+Providing additional context to the habitability of an exoplanet would be the presence of a water cycle, as evidenced by the identification of a glint, i.e.\ specular reflection by a liquid surface, or by a rainbow and other optical phenomena that can be attributed to water clouds, as detailed in the next sections.
 
@@
 %As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
-Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces.
-%
-In the total flux, the glint signal will be most prominent at small planetary phase (large scattering) angles.
+Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces. While glint is a familiar phenomenon on Earth, there is only
+one other body in the Solar System where it has been observed: 
+the Cassini orbiter captured the glint of sunlight reflecting off the
+liquid hydrocarbon lakes near Titan's north pole as it was observing
+the moon from the nightside \citep{2010GeoRL..37.7104S}.
+%
+In the total flux, the glint signal is indeed most prominent at small planetary phase (large scattering) angles.
 %
@@
 
-Rainbows, glories and other optical effects from condensates provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,2014A&A...566L...1G}.
-%
-In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to haze rather than a surface \citep{zellner1973polarization}, the phase curve behaviour of Venus showed not only that the clouds were $\sim$75\% H$_2$SO$_4$--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+Rainbows, glories and other optical phenomena from light that is 
+scattered by atmospheric condensates provide a wealth of information 
+about exoplanets \citep{karalidi2012rainbow, stam2008,2014A&A...566L...1G}.%
+\footnote{
+    Note that these are not the familiar rainbows that we see when sunlight is scattered by rain droplets, which are huge particles compared to the wavelength of the light: with our eyes, we cannot see the rainbows that are due to sunlight that is scattered by cloud droplets and that are the focus of this paper.
+    However, the presence of these bows (which are also referred to as \enquote{cloudbows}) can be revealed by polarimetric measurements.
+}
+%
+In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the phase curve behaviour of Venus showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
 The glory feature in Venus's phase curve was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
 
-In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
+In unpolarized light {\color{Green}{do you mean polarized light?}}, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
 Lambertian reflecting planet.
@@
 Note that the scattering by clouds that consist of solid particles, 
-such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, Karalidi2012ice}.
+such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, karalidi2012rainbow}.
 %
@@
 
-In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering phase angles are symmetric about 90 degrees.
+In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering phase angles are symmetric about 90$^\circ$.
 %
@@
 %
-For a system with inclination angle $i$, we can identify the following 2 regimes (see also \cref{fig:orb-grid}):
+For a system with inclination angle $i$, we can identify the following two regimes (see also \cref{fig:orb-grid}):
 \begin{enumerate}
@@
         The scattering phase angles accessible to a direct imaging mission 
-        can be limited by either the orbital inclination angle $i$ or the 
-        coronagraphic \IWA{}.
+        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
         %
@@
     = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
+\label{eq:contrast}
 \end{equation}
@@
    \caption{
-    The reflected light contrast and orbital separation of an Earth-like planet with an ocean surface and patchy clouds over a planetary orbit assuming an orbital inclination of \qty{90}{\degree} for the stars $\alpha$~Cen~A, $\epsilon$~Eri and Lalande~21185.
-    %
-    The solid line indicates the contrast in unpolarized light, with the contrast at quadrature marked by a solid dot.
-    %
-    The polarized component is indicated by the coloured dots, for which the colour represents the scattering phase angle from quadrature.
-    %
-    The light grey points show the quadrature contrasts of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
+    The contrast (Eq.~\ref{eq:contrast}) and orbital separation 
+    of an Earth-like planet with an ocean surface and patchy clouds 
+    along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A, $\epsilon$~Eri, and Lalande~21185.
+    %
+    The solid, black line indicates the contrast in unpolarized light, 
+    with the contrast at quadrature marked by the black dot.
+    %
+    The polarized component is indicated by the coloured dots, with the colour representing the scattering phase angle from quadrature.
+    %
+    The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
     }
@@
 
-\subsection{Observable scattering phenomenon}
+\subsection{Observable optical phenomenon}
 \label{sec:results_scattering_phenomena}
 
-In this section, we consider the number of stars around which different scattering phenomena could be observable for different \IWA{}.
+In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA{}.
 %
@@
 %
-We note that these ranges should all be considered approximate, as the exact phase angles where these effects are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
+We note that these ranges should all be considered approximate, as the exact phase angles where these phenomena are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
 %
@@
     \caption{
-        The ranges of phase angle for the different scattering phenomena considered here.
+        The approximate ranges of phase angles for the different optical phenomena considered in this paper.
     }
@@
 
-\subsection{Contrast Curves}
+\subsection{Contrast curves}
 \label{sec:results_contrast}
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/nplanets_vs_iwa.png}  
-    \caption{
-        The number of planetary systems for which key optical phenomena 
-        (Rayleigh scattering, rainbow, ocean glint, glories) would be 
-        observable as functions of the \IWA{}.
-        %
-        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak of the phenomena is observable (solid lines), and where the end is observable (dot-dashed lines).
+    \includegraphics{figures/nplanets_vs_iwa.pdf}
+    \script{create-figure-9.py}
+    \caption{
+        The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glories) would be observable as functions of the \IWA.
+        %
+        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
         %
@@
 %
-\todo{the paragraph below seems similar to the one above ... merge them?}
-We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of phase angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
-%
-This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
+Additionally, decreasing the \IWA\ also increases the range of phase angles observable in these systems. 
+%
+Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
 %
@@
 This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
+
+%\todo{the paragraph below seems similar to the one above ... merge them?}
+%We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA\ of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of phase angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
+%
+%This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
+%
+%While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
+%
+%This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
 

### 2023-04-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
-    Matthew Kenworthy\textsuperscript{\afflink{7}},
+    Matthew A.  Kenworthy\textsuperscript{\afflink{7}},
     Johan Mazoyer\textsuperscript{\afflink{11}},
@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
+NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures. 
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
@@
 %
-Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
+Direct imaging missions are optimised to detect planets near quadrature, and the coronagraph may obscure the phases at which these features are strongest. 
 %
@@
 %
-Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
-%
-We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
-%
-We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
+Planets that orbit close to edge-on have accessible phase angles limited by coronagraph obscuration. 
+%
+We use the list of 160 target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
+%
+We find that any exo-Earths in systems listed in this \hwo\ catalog will have a detectable Rayleigh scattering peak. 
 %
@@
 
-In unpolarized light {\color{Green}{do you mean polarized light?}}, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
+In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
 Lambertian reflecting planet.
@@
     \caption{
-        \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
         The cumulative distributions of the most extreme scattering phase angles accessible for different \IWA{} and for randomly inclined, circular orbits

### 2023-05-01 src/scripts/create-figure-5-orb-grid.py
Rename plotting scripts, fix basic code style

@@ -1,3 +1,7 @@
+"""
+Create Figure 5: Grid of orbits with different inclinations and IWAs.
+"""
+
 from astropy.time import Time
 from matplotlib.patches import Arrow
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
@@ -197,4 +201,4 @@ ax.plot3D([0],[0],[0], 'ko', label='Mask', markersize=200, alpha=0.3)
 ax.view_init(elev=-40, azim=0)
 
 plt.subplots_adjust(wspace=0.0, hspace=-0.4)
-plt.savefig(paths.figures / 'orb-grid.pdf')
+plt.savefig(paths.figures / 'figure-5-orb-grid.pdf')

### 2023-05-01 src/scripts/create-figure-6-n-over-phase.py

@@ -1,0 +1,221 @@
+"""
+Create Figure 6: CDF of number of planets over accesible phase angle for
+different inner working angles.
+"""
+
+import matplotlib as mpl
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from utils import paths
+from utils.plotting import CBF_COLORS
+
+mpl.rcParams["axes.prop_cycle"] = mpl.cycler(color=CBF_COLORS)
+
+
+def make_dual_cdf_plot(
+    beta_maxes,
+    betas,
+    iwa_list,
+    hab_zones,
+    label_locations,
+    rotations,
+    ax=None,
+    eta_earth=0.24,
+    n_bins=20,
+    savefig=False,
+    filename="accessible_phase_angles.png",
+):
+    """
+    A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle
+    for a given inner working ange.
+
+    beta_maxes - An array of maximum scattering angles, generated by the dynamical simulations, shape [n_simulations,n_systems,n_iwas]
+    betas - An array of beta values, generated for circular orbits, shape [n_simulations,n_systems,n_iwas]
+    iwa_list - list of IWAs you want to use, shape [n_iwas]
+    hab_zones - list of habitable zone SMAs in mas, shape [n_systems]
+    label_locations - a list of locations for each iwa, shape [n_iwas,2]
+    rotations - a list of angles defining the label rotations for each iwa, shape [n_iwas]
+    ax - a matplotlib axes on which to plot
+    eta_earth - the assumption on the number of systems have interesting planets (basically a scale factor for the y-axis), scalar
+    n_bins - the number of bins to use when making hisotgram bins for the cdfs, scalar
+    savefig - whether to save the figure or not, boolean
+    filename - the filename of the figure you want to save, string
+    """
+
+    # If an axis isn't provided, then make one.
+    if ax is None:
+        # fig, axis = plt.subplots(1,1,figsize=(10,4))
+        fig, axis = plt.subplots(1, 1, figsize=(10, 4))
+    else:
+        axis = ax
+
+    # Cycle through each IWA to plot a line.
+    for i, iwa in enumerate(iwa_list):
+        #### First the dynamical betas.
+        # (Although we're calling them betas, really they're the maximum scattering angles)
+
+        # Getting data of the histogram bins to build the CDF
+        count, bins_count = np.histogram(
[…]

### 2023-05-01 src/scripts/create-figure-7-contrast-over-separation.py

@@ -1,0 +1,164 @@
+"""
+Create Figure 7: Contrast over separation for three example targets.
+"""
+
+from matplotlib.colorbar import Colorbar
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from scipy import interpolate
+
+from utils import paths
+from utils.plotting import set_fontsize
+
+
+df = pd.read_csv(paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv")
+
+hab_zones = np.array(df["EEIDmas"].values[1:], dtype=float)  # in mas
+contrasts = np.array(df["Contrast"].values[1:], dtype=float)
+names = np.array(df["CommonID"].values[1:], dtype=str)
+
+data_trees = np.loadtxt(
+    paths.data / "Reflected_light_curves_Trees_Stam_2019.txt"
+)
+
+
+P_670 = interpolate.interp1d(data_trees[0], data_trees[2])
+F_670 = interpolate.interp1d(data_trees[0], data_trees[1])
+
+orbital_phase = np.linspace(0, 2 * np.pi, 200)
+
+inclination = np.radians(90)
+
+alpha = np.arccos(np.cos(orbital_phase) * np.sin(inclination))
+
+# Define a mapping of quantity (e.g., beta value) to a color
+norm = mpl.colors.Normalize(vmin=0, vmax=90)
+cmap = mpl.cm.get_cmap("viridis_r")
+
+
+names_array = np.array(["Rigil Kentaurus", "Ran", "Lalande 21185"])
+common_names = np.array(
+    [r"$\alpha$ Cen A", r"$\epsilon$ Eri", "Lalande 21185"]
+)
+
+pad_inches = 0.025
+fig, axs = plt.subplots(
+    nrows=1, ncols=3, figsize=(9 - 2 * pad_inches, 3.5 - 2 * pad_inches)
+)
+
+xlims = np.array([1e3, 2e2, 1e2])
+
+for i, name_one in enumerate(names_array):
+    index2 = np.where(names == name_one)
+
+    dtel = 6
+    wl1 = 670e-9
+    iwa1 = wl1 / dtel * 180 / np.pi * 3600 * 1000
+
+    x2 = np.sin(orbital_phase)
+    y2 = np.cos(inclination) * np.cos(orbital_phase)
+
[…]

### 2023-05-01 src/scripts/create-figure-8-yarnball.py

@@ -1,5 +1,5 @@
 """
-Create the yarn ball plot with random eccentric orbits.
+Create Figure 8: The yarn ball plot with random eccentric orbits.
 """
 
 # -----------------------------------------------------------------------------
@@ -207,14 +207,11 @@ if __name__ == "__main__":
     # -------------------------------------------------------------------------
 
     # Define inner working angles (IWA) in mas for computing beta_max
-    iwas = np.array([convert_lod_to_mas(_) for _ in [1, 2, 3, 4]])
-
-    # Set up output arrays
-    beta_min = np.zeros((n_stars, n_planets, iwas.size))
-    beta_max = np.zeros((n_stars, n_planets, iwas.size))
+    iwas = np.array([convert_lod_to_mas(iwa) for iwa in [1, 2, 3, 4]])
 
     # Loop over all the stars in the table
     print("Simulating orbits:", flush=True)
+    # noinspection PyTypeChecker
     for n, target in tqdm(enumerate(targets), total=n_stars, ncols=80):
 
         # Define shortcuts
@@ -246,9 +243,6 @@ if __name__ == "__main__":
             w = wsamp[m]
             anode = anodesamp[m]
 
-            # TODO: This is where we need to compute beta_min and beta_max
-            #   for the given IWA values and store them in the output arrays
-
             # Plot some orbits (if desired)
             if create_plot and (n < len(axes)) and (m < n_orbits_to_plot):
                 epochs = np.linspace(tperi, tperi + P, 100, endpoint=True)
@@ -263,15 +257,12 @@ if __name__ == "__main__":
     # Save results
     # -------------------------------------------------------------------------
 
-    # TODO: This is where we need to save the output arrays for beta_min and
-    #   beta_max to disk. (Currently, we do not compute values for them.)
-
     if create_plot:
 
         print('Saving plot...', end=' ', flush=True)
 
         fig.tight_layout(pad=0)
-        file_path = paths.figures / "ball_of_yarn.pdf"
+        file_path = paths.figures / "figure-9-yarnball.pdf"
         plt.savefig(
             file_path,
             bbox_inches="tight",

### 2023-05-01 src/scripts/plot_contrasts.py

@@ -1,148 +1,0 @@
-from typing import Any
-
-from matplotlib.colorbar import Colorbar
-from matplotlib.lines import Line2D
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-from scipy import interpolate
-
-from utils import paths
-
-
-# -----------------------------------------------------------------------------
-# AUXILIARY FUNCTIONS
-# -----------------------------------------------------------------------------
-
-def set_fontsize(ax: plt.Axes, fontsize: float) -> None:
-    """
-    Auxiliary function to set the fontsize of all elements in a plot.
-    """
-
-    for item in (
-        [ax.title, ax.xaxis.label, ax.yaxis.label]
-        + ax.get_xticklabels()
-        + ax.get_yticklabels()
-    ):
-        item.set_fontsize(fontsize)
-
-def quantity_to_ms(q: float) -> float:
-    return 5 * (1 + q / 500)
-
-
-# Define a function to map a quantity (e.g., beta value) onto a color
-def quantity_to_color(q: float) -> Any:
-    return cmap(norm(q))
-
-
-df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')
-
-hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
-contrasts =np.array(df['Contrast'].values[1:],dtype=float)
-names = np.array(df['CommonID'].values[1:],dtype=str)
-
-data_trees = np.loadtxt(paths.data / 'Reflected_light_curves_Trees_Stam_2019.txt')
-
-
-P_670 = interpolate.interp1d(data_trees[0],data_trees[2])
-F_670 = interpolate.interp1d(data_trees[0],data_trees[1])
-
-orbital_phase = np.linspace(0,2*np.pi,200)
-
-inclination = np.radians(90)
-
-alpha = np.arccos(np.cos(orbital_phase)*np.sin(inclination))
-
-# Define a mapping of quantity (e.g., beta value) to a color
-norm = mpl.colors.Normalize(vmin=0, vmax=90)
-cmap = mpl.cm.get_cmap('viridis_r')
-
-
[…]

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-04-23 src/tex/ms.tex
@@
 \author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{\afflink{1}},
+    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
     Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
@@
 %
-Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraphic obscuration. 
+Planets that orbit close to edge-on have accessible scattering angles limited by coronagraph obscuration. 
 %
@@
 %
-The ocean glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
+The glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
 %\sim$110--130$^\circ$ 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/Phase_curves.pdf}
-    \caption{The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from 0$^\circ$ to 180$^\circ$ along its orbit).
-The wind speed over the ocean is 7~m/s.
-The clouds in the planet's atmosphere consist of spherical water droplets.
-The cloud coverage fraction is 0.5.
-The cloud pattern is patchy and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
-9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the 'noise' at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
-The
-    polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). Note that the reference plane for the
-    polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
- %   and the latter
- %   is equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
-        %The cloudy scenario includes a low-altitude, Earth-like water 
-        %cloud deck. 
-        Below the plot, we show the phase angle ranges that cover the
+    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
+    \caption{
+        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit scattering angles 
+        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
+        The cloudy scenario includes a low-altitude, Earth-like water 
+        cloud deck. 
+        Below the plot, we are showing the scattering angle ranges that cover the
         characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
         with each other).  
-        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
+        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
     }
@@
 
-An exoplanet in an edge-on orbit moves through phase angles ranging from $0^\circ$ (full phase), through $90^\circ$ (quadrature), and to $180^\circ$ (the planet's nightside is towards the observer) (cf.\ Fig.~\ref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
+An exoplanet observed in an edge-on orbit will move through all phases and thus all phase angles, while those viewed at higher inclinations will pass through a subset of these phase angles.
 %
@@
 %
-The maximum polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
+The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at \qty{53}{\degree} \citep[i.e., a phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
 %
@@
 %
+{\color{red}{check the next sentence:}}
+The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across phase angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
 
@@
 %
-This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}, as shown in Fig.~\ref{fig:bottplot}.
+This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}.
 %The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
@@
 %
-While the glint feature can only be probed when a planet is at crescent, i.e.\ large phase angles, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
-This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
+While the glint feature can only be probed when a planet is at crescent phases, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry when the planet is at scattering angles where it is at the wide separations more easily probed when using coronagraphs.
 %how use to map, 
@@
 Note that the scattering by clouds that consist of solid particles, 
-such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
+such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{bailey2007, hansentravis1974, karalidi2012rainbow}.
 %
@@
 %is produced when the droplet radii much smaller than the wavelength of light 
-can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
+can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within \qty{20}{\degree} phase angle very close to full phase \citep{hansentravis1974}.
 %
@@
 
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
-%
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes along a planet's orbits.
+In this paper, we quantify the planetary scattering angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
+%
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of scattering angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
 %
 We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}. 
-It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
+It is important to note that the phases angles selected for the features throughout this paper are approximate, that features overlap, and that their maximum peak will usually vary with chemical species and wavelength.
 
 The conventions for angle names vary throughout the literature on exoplanet phase curves.
-In this paper we define the \textbf{scattering} angle as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
-A planet with a face-on orbit will thus have a phase angle of 90$^\circ$ everywhere along its orbit.
-In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
-We also note that we are considering here the (linearly) polarized flux
-which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information could be garnered from full Stokes characterization.
+In this paper we define the \textbf{scattering} angle as the local divergence in the light direction (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
+A planet with a face-on orbit will thus have a phase angle value of 90 degrees for the duration of its orbit.
+In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge on orbits), which is helpful in considering the visibility of the planet with inner working angle.
+We also note that we are considering here the polarization flux which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information can be garnered from full Stokes characterization.
 
@@
 
-A coronagraph can obscure part of a planet's orbit which, as illustrated in \cref{fig:annotated-orbit}, can prevent the planet from being observable at small and large phase angles, where it is close to its star.
-%
-The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the inclination and shape, i.e.\ eccentricity, of the exoplanet's orbit.
+A coronagraph can obscure part of the planet's orbit which, as shown in \cref{fig:annotated-orbit}, can prevent the planet from being observable at high and low phase angles. 
+%
+The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the orientation and shape of the exoplanet's orbit.
 
@@
     \caption{
-        An illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of scattering features in an exoplanet's phase curves that could help to characterize its atmospheric and/or surface composition and habitability.
-        The scattering angle, $\varphi$, is shown in red. It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
+        An illustration of a planet in an edge-on circular orbit showing how a coronagraph can obscure the exoplanet at high and low phase angles, and thus prevent observations of scattering features in an exoplanet's phase curve that could help to identify its atmospheric and/or surface composition and habitability.
+        The scattering angle, $\varphi$, is shown in red and is the complement of the phase angle in edge on systems.
     }
@@
 
-For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about 90$^\circ$ (see Fig.~\ref{fig:accessible_phase_angles}).
+In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible phase angles are symmetric about 90$^\circ$.
 %
@@
 %
-Angle $\varphi$ is given by
+The phase angle $\varphi$ is given by
 \begin{equation}
@@
 %
-We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
+We access the minimum phase angle when the projected distance reaches the inner working angle separation $\delta = \mathrm{IWA}$. 
 %
@@
         The on-sky angular separation as viewed by the observer is $\delta$.
-%        Angle $\varphi$ is measured between the line of sight to the star and the line connecting the star and planet.  
+        The phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.  
     }
@@
 %--------------------------------------------------------------------------
-%\begin{figure}
-%   \centering
-%   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
-%   \script{create-figure-5-orb-grid.py}
-%   \caption{
-%        The phase angles accessible to a direct imaging mission 
-%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
-        %
-%        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
-        %
-%        For orbits close to edge-on (bottom panels) the planet attains all 
-%        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
-%    }
-%    \label{fig:orb-grid}
-%\end{figure}
+\begin{figure}
+   \centering
+   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
+   \caption{
+        The phase angles accessible to a direct imaging mission 
+        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
+        %
+        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
+        %
+        For orbits close to edge-on (bottom panels) the planet attains all 
+        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+    }
+    \label{fig:orb-grid}
+    \script{plot-orb-grid.py}
+\end{figure}
 %--------------------------------------------------------------------------
@@
 %
-We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
+We also calculate which target planets would be observable at the phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
 %
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$
-(normalized to \num{1.0} at full phase, thus at $\alpha=0^\circ$), 
-and $a$ is the distance between the planet and its host star.
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at scattering angle $\alpha$
+(normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.
+\todo{here, $\alpha$ is used for the scattering angle ... and is the 
+'integral phase function' the same as the 'scattering phase function'?}
 %

### 2023-05-01 src/scripts/create-figure-6-n-over-phase.py
Update some more file names

@@ -24,7 +24,7 @@ def make_dual_cdf_plot(
     eta_earth=0.24,
     n_bins=20,
     savefig=False,
-    filename="accessible_phase_angles.png",
+    filename="figure-6-n-over-phase.pdf",
 ):
     """
     A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle

### 2023-05-01 src/scripts/create-figure-7-contrast-over-separation.py

@@ -156,9 +156,9 @@ for i, name_one in enumerate(names_array):
             color="white",
         )
 
-plt.savefig(
-    paths.figures
-    / "Contrast_vs_separation_inclination_angle_{0}.pdf".format(
-        int(np.degrees(inclination))
-    )
+file_name = (
+    f"figure-7-contrast-over-separation_i={int(np.degrees(inclination))}.pdf"
 )
+file_path = paths.figures / file_name
+
+plt.savefig(file_path, bbox_inches="tight")

### 2023-05-01 src/scripts/create-figure-9-n-over-iwa.py

@@ -177,7 +177,7 @@ fig.tight_layout(pad=0)
 plt.subplots_adjust(wspace=0.1)
 
 # Save the figure
-file_name = "nplanets_vs_iwa.pdf"
+file_name = "figure-9-n-over-iwa.pdf"
 file_path = paths.figures / file_name
 fig.savefig(
     file_path,

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to ocean glint, and to rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
+Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to oceans, and are sensitive to rainbows and other phenomena that are due to scattering by cloud particles.
 %
@@
 %
-Planets that orbit close to edge-on have accessible scattering angles limited by coronagraph obscuration. 
+Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraph obscuration. 
 %
@@
     \includegraphics{figures/figure-4-scatterplot.pdf}
-    \script{create-scatterplot.py}
+    \script{create-figure-4-scatterplot.py}
     \caption{
@@
    \centering
-   \includegraphics[width=0.99\columnwidth]{figures/orb-grid.pdf}
+   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
+   \script{create-figure-5-orb-grid.py}
    \caption{
@@
     \label{fig:orb-grid}
-    \script{plot-orb-grid.py}
 \end{figure}

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to oceans, and are sensitive to rainbows and other phenomena that are due to scattering by cloud particles.
+Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to ocean glint, and to rainbows and other phenomena caused by scattering by cloud particles.
 %
@@
 %
-We use \hwo'~catalog of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering by gas. 
+We use the list of 160 target stars for \hwo~to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
 %

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-05-01 src/tex/ms.tex
@@
 %
-Planets that orbit close to edge-on have accessible scattering angles limited by coronagraph obscuration. 
+Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraphic obscuration. 
 %
@@
 %
-The glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
+The ocean glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
 %\sim$110--130$^\circ$ 
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-    \caption{
-        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit scattering angles 
-        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
-        The cloudy scenario includes a low-altitude, Earth-like water 
-        cloud deck. 
-        Below the plot, we are showing the scattering angle ranges that cover the
+    \includegraphics[width=\linewidth]{figures/Phase_curves.pdf}
+    \caption{The phase curves of the polarized flux of a planet with an ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from 0$^\circ$ to 180$^\circ$ along its orbit).
+The cloud fraction is 0.5, the cloud pattern is patchy, and the glint is kept cloud-free to maximize the glint signature \citep[see ][]{treesandstam2019}.
+        %The cloudy scenario includes a low-altitude, Earth-like water 
+        %cloud deck. 
+        Below the plot, we are showing the phase angle ranges that cover the
         characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
@@
     \label{fig:bottplot}
-\end{figure*}
+\end{figure}
 
 %--------------------------------------------------------------------------------
-
-The field of exoplanet research has made great advances in the last few decades.
-The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve.
-The Habitable Worlds Observatory will perform a dedicated direct imaging survey of nearby stars searching for rocky planets in the habitable zone.
-This space-based coronographic telescope will detect and characterise exoplanets using their reflection spectrum.
-Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
 
@@
 
-An exoplanet observed in an edge-on orbit will move through all phases and thus all phase angles, while those viewed at higher inclinations will pass through a subset of these phase angles.
-%
-If the planet is rotating, its reflected light signal in both total and polarized flux is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet \citep{2001Natur.412..885F,stam2008,2022A&A...664A..59M}.
-
-The strengths of detailed total and polarized flux phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
+An exoplanet in an edge-on orbit will move through phase angles ranging from $0^\circ$ (full phase), through $90^\circ$ (quadrature), and to $180^\circ$ (the planet's nightside is towards the observer), while planets higher inclination orbits will pass through a subset of phase angles.
+%
+If the planet is rotating, its reflected light signal is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet.
+
+The strengths of detailed phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
 %
@@
 %
-In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
+In the total flux, the glint signal is indeed most prominent at small planetary phase (large scattering) angles.
 %
@@
 %
-This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
-%
-
+This angle therefore varies slightly with wavelength and with sea and atmosphere composition.
+%
 In general, the glint on a planet will be strongest at red wavelengths where the Rayleigh scattering in the planet's atmosphere and attenuation through the atmosphere are smallest \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also small \citep{2012ApJ...752L...3C}.
 %
-The stronger glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
-%
-The width across phase angles of the glint feature in total and polarized flux generally
-broadens with increasing wave height \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citep{CoxMunk1954}.
+The strengthened glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
 %
@@
 %
-At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
+The forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
 %
@@
 Rainbows, cloudbows, glories and other optical phenomena from light that is 
-scattered by atmospheric condensates can dramatically influence the shape of a planet's total flux and polarized phase curves and provide a wealth of information 
-about exoplanets \citep{karalidi2012rainbow, stam2008,Bailey2007,2014A&A...566L...1G}.%
+scattered by atmospheric condensates provide a wealth of information 
+about exoplanets \citep{karalidi2012rainbow, stam2008,2014A&A...566L...1G}.%
 \footnote{
-    Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light.
-Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white \citet[see][for further discussion]{Bailey2007}.
-}
-%
-In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the polarized phase curve of Venus at three wavelengths showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
-The glory feature in Venus's total flux phase curve was later used to rule out potential solutions for the unknown UV absorber \citep{petrova2018glory}.
+    Rainbows occur when light is scattered by rain droplets, which are large compared to the wavelength of the light.
+Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white (\citet[see][for further discussion]{Bailey2007}).
+}
+%
+In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the phase curve behaviour of Venus showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+The glory feature in Venus's phase curve was later used to rule out some solutions for the unknown UV absorber \citep{petrova2018glory}.
 
@@
 %
-While the glory and rainbow are technically visible both in the total flux
-and the polarized flux phase curves
-in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized flux than in total flux \citep{karalidi2011, stam2008, treesandstam2019}.
-%
-The rainbow angle depends on the refractive index of the scatterers and thus depends on their composition and the wavelength. 
+{\color{red}{check the sentence:}}
+While the glory and rainbow are technically visible in any phase curve
+in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
+%
+The rainbow angle depends on the refractive index of the scatterers and thus depends on the composition and the wavelength. 
 The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
 %
-Because the rainbow feature is narrower than the glint due to the surface reflection, it would 
-allow for a more exact determination of the composition of the condensates.
+Because the rainbow feature is more narrow than the glint, it would 
+allow for a more exact determination of the particle composition.
 %
@@
 
-In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of 40$^\circ$ (see Fig.~\ref{fig:bottplot}).
+In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths.
 %
@@
 
-In this paper, we quantify the planetary scattering angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
-%
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of scattering angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
+%
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
 %
 We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}. 
-It is important to note that the phases angles selected for the features throughout this paper are approximate, that features overlap, and that their maximum peak will usually vary with chemical species and wavelength.
+It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will usually depend on the chemical species and wavelength.
 
 The conventions for angle names vary throughout the literature on exoplanet phase curves.
-In this paper we define the \textbf{scattering} angle as the local divergence in the light direction (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
+In this paper we define the \textbf{scattering} angle as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
 A planet with a face-on orbit will thus have a phase angle value of 90 degrees for the duration of its orbit.
@@
 For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths of $\lambda = \qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
 
@@
 %
-By design, such planets should be observable by \hwo{} at least at their quadrature phase ($\alpha=90^\circ$).
+By design, such planets should be observable by \hwo{} at least at their quadrature phase.
 
@@
     \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,.
-    \label{eq:sinphi}
 \end{equation}
 %
-Assuming small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the on-sky angular separation.
+Using small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the angular separation on-sky.
 %
@@
 %
-Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
+Using \cref{eq:Delta_phi}, we deduce:
 %
@@
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature ($\alpha=90^\circ$), while the rainbow angle (around $\alpha=40^\circ$) will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature, while the rainbow angle will only be accessible for a subset of exoplanets.
 
@@
 
-Most planets are not in edge-on orbits as seen from Earth.
-%
-For an orbit with an inclination angle $i$, we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
+Most planetary systems will not be viewed as edge-on from Earth.
+%
+For a system with inclination angle $i$, we can identify the following two regimes (see also \cref{fig:orb-grid}):
 \begin{enumerate}
@@
 The solid lines in \cref{fig:betaallofit} show the cumulative 
-distribution of the maximum and minimum accessible phase 
+distribution of the maximum and minimum accessible scattering phase 
 angles normalised to the total number of systems in the \hwo{} 
@@
 We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
-Orbital eccentricities are drawn from the beta distribution, and 
-inclination angles are uniform in $\cos i$. 
+Eccentricities are drawn from the beta distribution, and inclinations are uniform in $\cos i$. 
 \Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
@@
 
-We then calculate the minimum and maximum accessible phase 
+We then calculate the minimum and maximum accessible scattering phase 
 angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
@@
 The dashed lines in \cref{fig:betaallofit} show the normalised cumulative 
-distribution of the accessible phase angles for each value of the 
-\IWA{} for eccentric orbits. 
+distribution of the accessible angles for each of \IWA for eccentric orbits. 
 %
@@
 A second parameter that is crucial for detecting angular features like 
-rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e.\ 
-the ratio of the flux of the starlight that is reflected by the planet 
-and the direct stellar flux. 
+rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the flux of the reflected light from the planet divided by the stellar flux. 
 %
@@
 The assumptions that went into the contrast estimations for \hwo{}
-are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
+are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
 %
@@
 Therefore, in this work, we will assume that the calculated contrasts 
-in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
+in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
 %
@@
 
-We use the phase functions from \citet{treesandstam2019} 
+We use the scattering phase functions from \citet{treesandstam2019} 
 for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
@@
 
-The Rayleigh feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-%
-The rainbow feature appears further from quadrature (around $\alpha=40^\circ$), but will still be observable in a significant fraction of systems.
-%
-The position and shape of the rainbow peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
-%
-The ocean glint feature occurs even further from quadrature and so would be observable in fewer systems. 
-%
-In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet.
+The Rayleigh scattering feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
+%
+The rainbow feature appears further from quadrature, but will still be observable in a significant fraction of systems.
+%
+The position and shape of the peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
+%
+The ocean glint feature occurs even further from quadrature and so is observable in fewer systems. 
+%
+In most systems, only the first half of the feature is observable, but this can still be used to identify a liquid ocean on the surface of the planet.
 %

### 2023-05-02 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-05-01 src/tex/ms.tex
@@
 %
-Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to oceans, and are sensitive to rainbows and other phenomena that are due to scattering by cloud particles.
+Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to ocean glint, and to rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
 %
@@
 %
-Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraph obscuration. 
+Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraphic obscuration. 
 %
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/Fig1IWAPol.png}
-    \caption{
-        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit scattering angles 
-        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
-        The cloudy scenario includes a low-altitude, Earth-like water 
-        cloud deck. 
-        Below the plot, we are showing the scattering angle ranges that cover the
+    \includegraphics[width=\linewidth]{figures/Phase_curves.pdf}
+    \caption{The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from 0$^\circ$ to 180$^\circ$ along its orbit).
+The wind speed over the ocean is 7~m/s.
+The clouds in the planet's atmosphere consist of spherical water droplets.
+The cloud coverage fraction is 0.5.
+The cloud pattern is patchy and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
+9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the 'noise' at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
+The
+    polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). Note that the reference plane for the
+    polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
+ %   and the latter
+ %   is equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
+        %The cloudy scenario includes a low-altitude, Earth-like water 
+        %cloud deck. 
+        Below the plot, we show the phase angle ranges that cover the
         characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
         with each other).  
-        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
+        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
     }
@@
 
-An exoplanet observed in an edge-on orbit will move through all phases and thus all phase angles, while those viewed at higher inclinations will pass through a subset of these phase angles.
+An exoplanet in an edge-on orbit moves through phase angles ranging from $0^\circ$ (full phase), through $90^\circ$ (quadrature), and to $180^\circ$ (the planet's nightside is towards the observer) (cf.\ Fig.~\ref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
 %
@@
 %
-The maximum polarization occurs at the Brewster angle determined by the refractive indices at the interface; between air and liquid water this occurs at \qty{53}{\degree} \citep[i.e., a phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
+The maximum polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
 %
@@
 %
-{\color{red}{check the next sentence:}}
-The broadening of the glint feature in the presence of wind-driven waves is well-described by the Cox-Munk solution \citep{CoxMunk1954}, shifts in both polarized and unpolarized light, and broadens across phase angles in polarized light \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}.
 
@@
 %
-This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}.
+This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}, as shown in Fig.~\ref{fig:bottplot}.
 %The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
@@
 %
-While the glint feature can only be probed when a planet is at crescent phases, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry when the planet is at scattering angles where it is at the wide separations more easily probed when using coronagraphs.
+While the glint feature can only be probed when a planet is at crescent, i.e.\ large phase angles, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
+This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
 %how use to map, 
@@
 
-In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
-Lambertian reflecting planet.
+%In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a Lambertian reflecting planet.
 %
@@
 Note that the scattering by clouds that consist of solid particles, 
-such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{bailey2007, hansentravis1974, karalidi2012rainbow}.
+such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
 %
@@
 %is produced when the droplet radii much smaller than the wavelength of light 
-can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within \qty{20}{\degree} phase angle very close to full phase \citep{hansentravis1974}.
+can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
 %
@@
 
-In this paper, we quantify the planetary scattering angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
-%
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2} we derive expressions for the range of scattering angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes as a function of the planet's orbital phase.
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
+%
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes along a planet's orbits.
 %
 We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}. 
-It is important to note that the phases angles selected for the features throughout this paper are approximate, that features overlap, and that their maximum peak will usually vary with chemical species and wavelength.
+It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
 
 The conventions for angle names vary throughout the literature on exoplanet phase curves.
-In this paper we define the \textbf{scattering} angle as the local divergence in the light direction (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
-A planet with a face-on orbit will thus have a phase angle value of 90 degrees for the duration of its orbit.
-In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge on orbits), which is helpful in considering the visibility of the planet with inner working angle.
-We also note that we are considering here the polarization flux which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information can be garnered from full Stokes characterization.
+In this paper we define the \textbf{scattering} angle as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
+A planet with a face-on orbit will thus have a phase angle of 90$^\circ$ everywhere along its orbit.
+In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
+We also note that we are considering here the (linearly) polarized flux
+which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information could be garnered from full Stokes characterization.
 
@@
 
-A coronagraph can obscure part of the planet's orbit which, as shown in \cref{fig:annotated-orbit}, can prevent the planet from being observable at high and low phase angles. 
-%
-The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the orientation and shape of the exoplanet's orbit.
+A coronagraph can obscure part of a planet's orbit which, as illustrated in \cref{fig:annotated-orbit}, can prevent the planet from being observable at small and large phase angles, where it is close to its star.
+%
+The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the inclination and shape, i.e.\ eccentricity, of the exoplanet's orbit.
 
@@
     \caption{
-        An illustration of a planet in an edge-on circular orbit showing how a coronagraph can obscure the exoplanet at high and low phase angles, and thus prevent observations of scattering features in an exoplanet's phase curve that could help to identify its atmospheric and/or surface composition and habitability.
-        The scattering angle, $\varphi$, is shown in red and is the complement of the phase angle in edge on systems.
+        An illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of scattering features in an exoplanet's phase curves that could help to characterize its atmospheric and/or surface composition and habitability.
+        The scattering angle, $\varphi$, is shown in red. It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
     }
@@
 
-In the non-eccentric case (i.e., a circular orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible phase angles are symmetric about 90$^\circ$.
+For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about 90$^\circ$ (see Fig.~\ref{fig:accessible_phase_angles}).
 %
@@
 %
-The phase angle $\varphi$ is given by
+Angle $\varphi$ is given by
 \begin{equation}
@@
 %
-We access the minimum phase angle when the projected distance reaches the inner working angle separation $\delta = \mathrm{IWA}$. 
+We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
 %
@@
         The on-sky angular separation as viewed by the observer is $\delta$.
-        The phase angle, $\varphi$, is the angle between the line of sight to the star and the line connecting the star and planet.  
+%        Angle $\varphi$ is measured between the line of sight to the star and the line connecting the star and planet.  
     }
@@
 %--------------------------------------------------------------------------
-\begin{figure}
-   \centering
-   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
-   \script{create-figure-5-orb-grid.py}
-   \caption{
-        The phase angles accessible to a direct imaging mission 
-        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
-        %
-        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
-        %
-        For orbits close to edge-on (bottom panels) the planet attains all 
-        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
-    }
-    \label{fig:orb-grid}
-\end{figure}
+%\begin{figure}
+%   \centering
+%   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
+%   \script{create-figure-5-orb-grid.py}
+%   \caption{
+%        The phase angles accessible to a direct imaging mission 
+%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
+        %
+%        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
+        %
+%        For orbits close to edge-on (bottom panels) the planet attains all 
+%        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
+%    }
+%    \label{fig:orb-grid}
+%\end{figure}
 %--------------------------------------------------------------------------
@@
 %
-We also calculate which target planets would be observable at the phase angles where ocean glint, rainbows, or the Rayleigh peak occur.
+We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
 %

### 2023-05-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \centering
-    \includegraphics{figures/figure-4-scatterplot.pdf}
-    \script{create-figure-4-scatterplot.py}
+    \includegraphics{figures/figure-scatterplot.pdf}
+    \script{create-figure-scatterplot.py}
     \caption{
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at scattering angle $\alpha$
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at pangle $\alpha$
 (normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.

### 2023-05-02 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-05-02 src/tex/ms.tex
@@
     \includegraphics{figures/figure-4-scatterplot.pdf}
-    \script{create-figure-4-scatterplot.py}
+    \script{create-scatterplot.py}
     \caption{
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the integral phase function at scattering angle $\alpha$
-(normalized to \num{1.0} at full phase), and $a$ is the distance between the planet and its host star.
-\todo{here, $\alpha$ is used for the scattering angle ... and is the 
-'integral phase function' the same as the 'scattering phase function'?}
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$
+(normalized to \num{1.0} at full phase, thus at $\alpha=0^\circ$), 
+and $a$ is the distance between the planet and its host star.
 %
 The assumptions that went into the contrast estimations for \hwo{}
-are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the scattering phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
+are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
 %

### 2023-05-02 src/scripts/create-scatterplot.py
[showyourwork] overleaf sync

@@ -191,7 +191,7 @@ if __name__ == "__main__":
 
     # Save the figure
     fig.tight_layout(pad=0)
-    file_path = paths.figures / "scatterplot.pdf"
+    file_path = paths.figures / "figure-4-scatterplot.pdf"
     plt.savefig(
         file_path,
         dpi=600,

### 2023-05-02 src/scripts/make_beta_max_3_lambda_over_d_csv.py
Rename and refactor data generation for circular orbits

@@ -1,14 +1,0 @@
-import pandas as pd
-import numpy as np
-from utils import paths
-from utils.constants import *
-
-# telescope d and wlen are imported from hwo file to make consistent for all the codes
-
-# givea a nan for the first element because it tries to calculate
-# arccos of 1.032, presumably this is alpha or proxima cen
-
-df = pd.read_csv(paths.data / '2646_NASA_ExEP_Target_List_HWO_Table.csv')
-hab_zones = np.array(df['EEIDmas'].values[1:],dtype=float) #in mas
-beta_max_3ld =  np.degrees(np.arccos(3*wavelength/d*206265/(hab_zones/1000)))
-np.savetxt(paths.data / "beta_max_3_lambda_over_d.csv",beta_max_3ld,delimiter=",")

### 2023-05-02 src/scripts/attic/create-figure-5-orb-grid.py
Move script for orbgrid plot (old Figure 5) to attic

@@ -1,0 +1,204 @@
+"""
+Create Figure 5: Grid of orbits with different inclinations and IWAs.
+"""
+
+from astropy.time import Time
+from matplotlib.patches import Arrow
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+
+import astropy.units as u
+import numpy as np
+import matplotlib.pyplot as plt
+
+from utils import paths
+from utils.orbfunc import xyz_position, phase
+
+
+##########################################################################################
+
+
+
+params = {
+    'time periapsis passage':Time('2021-09-29T00:00:00', format='fits'),
+    'orbital period':1*u.day,
+    'semi-major axis':1*u.au,
+    'eccentricity':0.0,
+    'argument of periapsis':np.pi*0/180*u.rad,
+    'periapsis precession':0*u.rad/u.s,
+    'inclination':np.pi*75/180*u.rad,
+    'longditude of ascending node':np.pi*90/180*u.rad
+}
+
+time = params['time periapsis passage'] - 0.5*u.day + np.arange(0,1,0.001)*u.day
+
+##########################################################################################
+
+def setup_plot(fig, i, oplane=False, losplane=False):
+
+    ax = fig.add_subplot(4, 3, i, projection='3d')
+    ax.set_proj_type('ortho')
+    ax.margins(-0.499)
+
+    # plot orbit
+    ax.plot3D(xyz[:,0], xyz[:,1], xyz[:,2], 'k-', lw=2)
+
+    # plane of los
+    if losplane==True:
+        lossurfx, lossurfy = np.meshgrid(np.linspace(-0.65,0.65,10),np.linspace(-0.65,0.65,10))
+        x1 = ax.plot_surface(lossurfx, lossurfy, np.zeros((10,10)), color='g', edgecolor=(0,0,0,0.15), alpha=0.1, label='Plane of Nodes')
+
+        ### Trying to catch a version dependent error in matplotlib, may not work for all versions ###
+        try:
+            x1._facecolors2d = x1._facecolor3d
+            x1._edgecolors2d = x1._edgecolor3d
[…]

### 2023-05-02 src/scripts/create-figure-5-orb-grid.py

@@ -1,204 +1,0 @@
-"""
-Create Figure 5: Grid of orbits with different inclinations and IWAs.
-"""
-
-from astropy.time import Time
-from matplotlib.patches import Arrow
-from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-
-import astropy.units as u
-import numpy as np
-import matplotlib.pyplot as plt
-
-from utils import paths
-from utils.orbfunc import xyz_position, phase
-
-
-##########################################################################################
-
-
-
-params = {
-    'time periapsis passage':Time('2021-09-29T00:00:00', format='fits'),
-    'orbital period':1*u.day,
-    'semi-major axis':1*u.au,
-    'eccentricity':0.0,
-    'argument of periapsis':np.pi*0/180*u.rad,
-    'periapsis precession':0*u.rad/u.s,
-    'inclination':np.pi*75/180*u.rad,
-    'longditude of ascending node':np.pi*90/180*u.rad
-}
-
-time = params['time periapsis passage'] - 0.5*u.day + np.arange(0,1,0.001)*u.day
-
-##########################################################################################
-
-def setup_plot(fig, i, oplane=False, losplane=False):
-
-    ax = fig.add_subplot(4, 3, i, projection='3d')
-    ax.set_proj_type('ortho')
-    ax.margins(-0.499)
-
-    # plot orbit
-    ax.plot3D(xyz[:,0], xyz[:,1], xyz[:,2], 'k-', lw=2)
-
-    # plane of los
-    if losplane==True:
-        lossurfx, lossurfy = np.meshgrid(np.linspace(-0.65,0.65,10),np.linspace(-0.65,0.65,10))
-        x1 = ax.plot_surface(lossurfx, lossurfy, np.zeros((10,10)), color='g', edgecolor=(0,0,0,0.15), alpha=0.1, label='Plane of Nodes')
-
-        ### Trying to catch a version dependent error in matplotlib, may not work for all versions ###
-        try:
-            x1._facecolors2d = x1._facecolor3d
-            x1._edgecolors2d = x1._edgecolor3d
[…]

### 2023-05-02 src/scripts/create-figure-5-n-over-phase.py
Update file names

@@ -24,7 +24,7 @@ def make_dual_cdf_plot(
     eta_earth=0.24,
     n_bins=20,
     savefig=False,
-    filename="figure-6-n-over-phase.pdf",
+    filename="figure-5-n-over-phase.pdf",
 ):
     """
     A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle

### 2023-05-02 src/scripts/create-figure-6-contrast-over-separation.py

@@ -157,7 +157,7 @@ for i, name_one in enumerate(names_array):
         )
 
 file_name = (
-    f"figure-7-contrast-over-separation_i={int(np.degrees(inclination))}.pdf"
+    f"figure-6-contrast-over-separation_i={int(np.degrees(inclination))}.pdf"
 )
 file_path = paths.figures / file_name
 

### 2023-05-02 src/scripts/create-figure-7-yarnball.py

@@ -262,7 +262,7 @@ if __name__ == "__main__":
         print('Saving plot...', end=' ', flush=True)
 
         fig.tight_layout(pad=0)
-        file_path = paths.figures / "figure-9-yarnball.pdf"
+        file_path = paths.figures / "figure-7-yarnball.pdf"
         plt.savefig(
             file_path,
             bbox_inches="tight",

### 2023-05-02 src/scripts/create-figure-8-n-over-iwa.py

@@ -177,7 +177,7 @@ fig.tight_layout(pad=0)
 plt.subplots_adjust(wspace=0.1)
 
 # Save the figure
-file_name = "figure-9-n-over-iwa.pdf"
+file_name = "figure-8-n-over-iwa.pdf"
 file_path = paths.figures / file_name
 fig.savefig(
     file_path,

### 2023-05-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-The maximum polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{127}{\degree}; see, e.g.,][]{2008Icar..195..927W}.
+The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.
 %
@@
 The width across phase angles of the glint feature in total and polarized flux generally
-broadens with increasing wave height \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citep{CoxMunk1954}.
-%
-
-The glint provides evidence of liquid surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
+broadens with increasing wave height \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citet{CoxMunk1954}.
+%
+
+The glint provides evidence of specular reflecting surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
 %
@@
     Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light.
-Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white \citet[see][for further discussion]{Bailey2007}.
+Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white \citep[see][for further discussion]{Bailey2007}.
 }
@@
 
-%--------------------------------------------------------------------------
 % Figure 4
-%--------------------------------------------------------------------------
 \begin{figure*}
@@
 
-%--------------------------------------------------------------------------
-% Figure 5
-%--------------------------------------------------------------------------
+% Old Figure 5
 %\begin{figure}
@@
 %\end{figure}
-%--------------------------------------------------------------------------
-
-Using \cref{eq:scattering_angle}, we deduce the following for inclined 
-circular orbits: 
+
+Using \cref{eq:scattering_angle}, we deduce the following for inclined circular orbits: 
 \begin{equation}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%--------------------------------------------------------------------------
+% Figure 5
 \begin{figure*}%[t]
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure-6-n-over-phase.pdf}
-    \script{create-figure-6-n-over-phase.py}
+    \includegraphics[width=0.95\textwidth]{figures/figure-5-n-over-phase.pdf}
+    \script{create-figure-5-n-over-phase.py}
     \caption{
@@
 
-%--------------------------------------------------------------------------
-\begin{figure*}%[t]
+% Figure 6
+\begin{figure*}
    \centering
-   \includegraphics[width=0.99\textwidth]{figures/figure-7-contrast-over-separation_i=90.pdf}
-    \script{create-figure-7-contrast-over-separation.py}
+    \includegraphics[width=0.99\textwidth]{figures/figure-6-contrast-over-separation_i=90.pdf}
+    \script{create-figure-6-contrast-over-separation.py}
     \caption{
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%--------------------------------------------------------------------------
-% Figure 8:
-%--------------------------------------------------------------------------
+% Figure 7: Yarn ball
 \begin{figure*}
     \centering
-    \includegraphics{figures/figure-9-yarnball.pdf}  
-    \script{create-figure-8-yarnball.py}
+    \includegraphics{figures/figure-7-yarnball.pdf}  
+    \script{create-figure-7-yarnball.py}
     \caption{
@@
 \end{figure*}
-%--------------------------------------------------------------------------
-
-%--------------------------------------------------------------------------
-% Figure 9:
-%--------------------------------------------------------------------------
+
+% Figure 8
 \begin{figure*}
     \centering
-    \includegraphics{figures/figure-9-n-over-iwa.pdf}
-    \script{create-figure-9-n-over-iwa.py}
+    \includegraphics{figures/figure-8-n-over-iwa.pdf}
+    \script{create-figure-8-n-over-iwa.py}
     \caption{

### 2023-05-02 src/scripts/utils/plotting.py
Re-order colors in CBF_COLORS

@@ -19,8 +19,8 @@ CBF_COLORS = [
     "#3F90DA",  # blue
     "#FFA90E",  # orange
     "#BD1F01",  # red
-    "#94A4A2",  # gray
     "#832DB6",  # purple
+    "#94A4A2",  # gray
     "#A96B59",  # brown
     "#E76300",  # orange
     "#B9AC70",  # tan

### 2023-05-03 src/tex/ms.tex

@@
     \centering
-    \includegraphics[width=\linewidth]{figures/Phase_curves.pdf}
-    \caption{The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from 0$^\circ$ to 180$^\circ$ along its orbit).
-The wind speed over the ocean is 7~m/s.
+    \includegraphics{figures/figure-1-bott-plot.pdf}
+    \script{create-figure-1-bott-plot.py}
+    \caption{
+        The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from \qty{0}{\degree} to \qty{180}{\degree} along its orbit).
+        The wind speed over the ocean is \qty{7}{\meter\per\second}. 
         The clouds in the planet's atmosphere consist of spherical water droplets. 
         The cloud coverage fraction is 0.5. 
-The cloud pattern is patchy and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
-9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the 'noise' at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
-The
-    polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). Note that the reference plane for the
-    polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
+        The cloud pattern is patchy, and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
+9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the \enquote{noise} at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
+        The polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). 
+        Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
  %   and the latter

### 2023-05-03 src/tex/ms.tex
[showyourwork]

@@
 In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the polarized phase curve of Venus at three wavelengths showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+%
 The glory feature in Venus's total flux phase curve was later used to rule out potential solutions for the unknown UV absorber \citep{petrova2018glory}.
@@
 While the glory and rainbow are technically visible both in the total flux
-and the polarized flux phase curves
-in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized flux than in total flux \citep{karalidi2011, stam2008, treesandstam2019}.
+and the polarized flux phase curves in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized flux than in total flux \citep{karalidi2011, stam2008, treesandstam2019}.
 %
 The rainbow angle depends on the refractive index of the scatterers and thus depends on their composition and the wavelength. 
+%
 The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
 %
-Because the rainbow feature is narrower than the glint due to the surface reflection, it would 
-allow for a more exact determination of the composition of the condensates.
-%
-Note that the scattering by clouds that consist of solid particles, 
-such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
-%
+Because the rainbow feature is narrower than the glint due to the surface reflection, it would allow for a more exact determination of the composition of the condensates.
+%
+Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
 
@@
 %
-The glory 
-%is produced when the droplet radii much smaller than the wavelength of light 
-can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
+The glory can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
 %

### 2023-05-03 src/scripts/create-figure-4-scatterplot.py
[showyourwork] overleaf sync

@@ -40,7 +40,7 @@ if __name__ == "__main__":
     data = pd.read_csv(file_path, header=1)
 
     # Load the beta values
-    file_path = paths.data / "beta_max_3_lambda_over_d.csv"
+    file_path = paths.data / "phi_max_3_lambda_over_d.csv"
     beta_values = np.loadtxt(file_path)
 
     print("Done!")

### 2023-05-07 src/scripts/create-figure-4-scatterplot.py
Update plot size and legend

@@ -17,6 +17,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 
+from utils.constants import PHASE_ANGLES_OF_FEATURES
 from utils.plotting import set_fontsize
 from utils import paths
 
@@ -27,11 +28,11 @@ from utils import paths
 
 if __name__ == "__main__":
 
+    print("\nCREATE FIGURE 4: SCATTERPLOT\n")
+
     # -------------------------------------------------------------------------
     # Load the data
     # -------------------------------------------------------------------------
-
-    print("\nCREATE SCATTERPLOT\n")
 
     print("Loading data...", end=" ", flush=True)
 
@@ -65,17 +66,26 @@ if __name__ == "__main__":
         return cmap(norm(q))
 
     # Create a new figure
+    # These numbers where carefully hand-tuned to ensure that the output has
+    # exactly the right width (7 inches) for the paper when saved as a PDF.
     pad_inches = 0.025
-    fig, ax = plt.subplots(figsize=(7 - 2 * pad_inches, 3 - 2 * pad_inches))
+    fig, ax = plt.subplots(figsize=(7.02 - 2 * pad_inches, 2 - 2 * pad_inches))
 
     # Set up the font size
     set_fontsize(ax, 6)
 
-    # Add a a grid to the plot (many journals don't like this?)
-    # ax.grid(ls='--', c='k', alpha=0.2, zorder=0)
+    # Add a a grid to the plot
+    ax.grid(
+        ls='--',
+        alpha=0.3,
+        color='k',
+        lw=0.5,
+        dash_capstyle='round',
+        dashes=(0.05, 2.5),
+    )
 
     # Set labels and limits for the axes
-    ax.set_xlabel(r"Stellar distance (pc)")
+    ax.set_xlabel("Stellar distance (pc)")
     ax.set_ylabel("Stellar effective temperature (K)")
     ax.set_xlim(0, 25)
     ax.set_ylim(3500, 7500)
@@ -100,7 +110,7 @@ if __name__ == "__main__":
 
     # Manually construct a color bar and add it to the plot
     divider = make_axes_locatable(ax)
-    cbar_ax = divider.new_horizontal(size="4%", pad=0.05)
+    cbar_ax = divider.new_horizontal(size="6%", pad=0.05)
     cbar = mpl.colorbar.ColorbarBase(
         cbar_ax,
         cmap=cmap,
@@ -116,44 +126,24 @@ if __name__ == "__main__":
[…]

### 2023-05-07 src/scripts/create-figure-5-n-over-phase.py
Refactor plotting script

@@ -1,221 +1,268 @@
 """
-Create Figure 6: CDF of number of planets over accesible phase angle for
-different inner working angles.
+Create Figure 5: CDF of number of planets over accessible phase angle
+for different inner working angles (both circular and elliptical).
 """
 
-import matplotlib as mpl
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
+
+from warnings import catch_warnings, filterwarnings
+
+from matplotlib.lines import Line2D
+
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
+
 from utils import paths
-from utils.plotting import CBF_COLORS
+from utils.constants import ETA_EARTH, RANDOM_SEED
+from utils.plotting import set_fontsize, CBF_COLORS
+from utils.samplers import sample_i
 
-mpl.rcParams["axes.prop_cycle"] = mpl.cycler(color=CBF_COLORS)
 
-
-def make_dual_cdf_plot(
-    beta_maxes,
-    betas,
-    iwa_list,
-    hab_zones,
-    label_locations,
-    rotations,
-    ax=None,
-    eta_earth=0.24,
-    n_bins=20,
-    savefig=False,
-    filename="figure-5-n-over-phase.pdf",
-):
-    """
-    A function that plots the 1-cumulative distribution to show the number of planets that can access a given phase angle
-    for a given inner working ange.
-
-    beta_maxes - An array of maximum scattering angles, generated by the dynamical simulations, shape [n_simulations,n_systems,n_iwas]
-    betas - An array of beta values, generated for circular orbits, shape [n_simulations,n_systems,n_iwas]
-    iwa_list - list of IWAs you want to use, shape [n_iwas]
-    hab_zones - list of habitable zone SMAs in mas, shape [n_systems]
-    label_locations - a list of locations for each iwa, shape [n_iwas,2]
-    rotations - a list of angles defining the label rotations for each iwa, shape [n_iwas]
-    ax - a matplotlib axes on which to plot
-    eta_earth - the assumption on the number of systems have interesting planets (basically a scale factor for the y-axis), scalar
-    n_bins - the number of bins to use when making hisotgram bins for the cdfs, scalar
-    savefig - whether to save the figure or not, boolean
-    filename - the filename of the figure you want to save, string
-    """
-
-    # If an axis isn't provided, then make one.
-    if ax is None:
[…]

### 2023-05-07 src/scripts/create-figure-7-yarnball.py
Reduce the number of panels in the yarn ball plot

@@ -1,5 +1,5 @@
 """
-Create Figure 8: The yarn ball plot with random eccentric orbits.
+Create Figure 7: The yarn ball plot with random eccentric orbits.
 """
 
 # -----------------------------------------------------------------------------
@@ -10,17 +10,17 @@ import time
 
 from astropy.io import ascii
 from matplotlib.patches import Circle
-from tqdm.auto import tqdm
 
 import astropy.units as u
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
 
+from utils.constants import RANDOM_SEED
 from utils.phases import kep3d
 from utils.samplers import sample_e, sample_i
 from utils.plotting import CBF_COLORS
-from utils import paths
+from utils.paths import data as data_dir, figures as figures_dir
 
 
 # -----------------------------------------------------------------------------
@@ -129,7 +129,6 @@ def prepare_ax(
 # -----------------------------------------------------------------------------
 
 if __name__ == "__main__":
-
     # -------------------------------------------------------------------------
     # Preliminaries
     # -------------------------------------------------------------------------
@@ -140,7 +139,7 @@ if __name__ == "__main__":
     script_start = time.time()
 
     # Set random seed (for reproducibility)
-    np.random.seed(42)
+    np.random.seed(RANDOM_SEED)
 
     # Use colorblind-friendly colors as default color cycle
     mpl.rcParams["axes.prop_cycle"] = mpl.cycler(color=CBF_COLORS)
@@ -148,14 +147,6 @@ if __name__ == "__main__":
     # -------------------------------------------------------------------------
     # Setup configuration for simulations and plot
     # -------------------------------------------------------------------------
-
-    # Whether or not to create a plot
-    # If False, only the output data of the simulations will be saved
-    create_plot = True
-
-    # Define grid size for plot
-    # We will make a plot with (grid_size, grid_size) panels
-    grid_size = 4
 
     # Number of orbits to plot for each star
     n_orbits_to_plot = 10
@@ -174,54 +165,40 @@ if __name__ == "__main__":
     print("Reading in target list...", end=" ", flush=True)
 
     file_name = "2646_NASA_ExEP_Target_List_HWO_Table.csv"
-    targets = ascii.read(paths.data / file_name, header_start=1, data_start=2)
+    targets = ascii.read(data_dir / file_name, header_start=1, data_start=2)
     n_stars = len(targets)
 
[…]

### 2023-05-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\documentclass[
-    usenatbib,
-]{mnras}
+\documentclass[usenatbib]{mnras}
 
 \usepackage[T1]{fontenc}
-\usepackage[dvipsnames, x11names]{xcolor}
+\usepackage[dvipsnames, x11names, svgnames]{xcolor}
 
@@
 % Prettier fonts 
-% (We might have to go back to defaults for MNRAS submission)
-% \usepackage{libertine}
-% \usepackage[libertine]{newtxmath}
-% \usepackage[scaled=0.76]{beramono}
-
 % Those seem to be the "right" fonts for MNRAS
@@
 \usepackage[capitalise]{cleveref}
-%\usepackage{xspace}
+\usepackage{xspace}
 
@@
     shapes,
+    shapes.symbols,
 }
@@
 
-%REMOVE LATER
-\usepackage{soul}
-
 % Load journal macros (e.g., \aa for A&A)
@@
 % Define macros
-\newcommand{\IWA}{\ensuremath{\mathrm{IWA}}}
-\newcommand{\OWA}{\ensuremath{\mathrm{OWA}}}
-\newcommand{\hwo}{HabWorlds}
+\newcommand{\IWA}{\ensuremath{\mathrm{IWA}}\xspace}
+\newcommand{\OWA}{\ensuremath{\mathrm{OWA}}\xspace}
+\newcommand{\HWO}{HabWorlds\xspace}
 
@@
     \textcolor{C7}{s}%
-    % \xspace
-}
+    \xspace
+}
+
+\newcommand{\istar}{%
+  \hspace{-0.125em}%
+  \tikz[baseline=(char.base)]{
+    \node [inner sep=0] (char) {i};
+    \node[starburst, starburst points=5, starburst point height=2.5mm, fill=White, draw=MidnightBlue, line width=0.1pt, minimum width=0.3mm, minimum height=0.1mm, inner sep=0.9pt, rotate=15] at (0.0015, 0.15) {};
+  }%
+  \hspace{-0.175em}%
+}
+\newcommand{\oceanglint}{\textcolor{MidnightBlue}{ocean gl\istar{}nt}\xspace}
 
@@
 
-% Chasing rainbows with the Habitable Worlds Observatory
-\title{Chasing \rainbows{} and ocean glint:\\  Inner working angle constraints for the Habitable Worlds Observatory}
-
-%phase angle probed by directly imaged planets, or
-%Don't ask us about contrast
+% Full title of the paper (MNRAS has no running title?)
+\title{Chasing \rainbows and \oceanglint:\\ Inner working angle constraints for the Habitable Worlds Observatory}
 
@@
     Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
-    Matthew A.  Kenworthy\textsuperscript{\afflink{7}},
+    Matthew Kenworthy\textsuperscript{\afflink{7}},
     Johan Mazoyer\textsuperscript{\afflink{11}},
@@
     Alexis Bidot\textsuperscript{\afflink{16}},
+    Jayne L. Birkby\textsuperscript{\afflink{1}},
+    \newauthor
     Markus J. Bonse\textsuperscript{\afflink{10}},
-    \newauthor
     Bernhard Brandl\textsuperscript{\afflink{7}},
@@
     Niyati Desai\textsuperscript{\afflink{17}},
+    \newauthor
     Kevin Fogarty\textsuperscript{\afflink{15}},
     Jules Fowler\textsuperscript{\afflink{18}},
-    \newauthor
     Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
@@
     Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
+    \newauthor
     Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
-    \newauthor
     Adrien Hours\textsuperscript{\afflink{16}},
     Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
-    Elina Kleisioti\textsuperscript{\afflink{7}},
+    Evangelia Kleisioti\textsuperscript{\afflink{7}},
+    \newauthor
     Lorenzo König\textsuperscript{\afflink{13}},
     Mariya Krasteva\textsuperscript{\afflink{27}},
-    \newauthor
     Iva Laginja\textsuperscript{\afflink{11}},
@@
     Lucie Leboulleux\textsuperscript{\afflink{16}},
+    \newauthor
     David Mouillet\textsuperscript{\afflink{16}},
     Mamadou N’Diaye\textsuperscript{\afflink{28}},
-    \newauthor
     Emiel H. Por\textsuperscript{\afflink{29}},
@@
     Frans Snik\textsuperscript{\afflink{7}},
+    \newauthor
     Daphne M. Stam\textsuperscript{\afflink{30}},
@@
     Dirk van Dam\textsuperscript{\afflink{7}},
-    \newauthor
     Kyle van Gorkom\textsuperscript{\afflink{21}},
@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
+NASA recently announced the Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
 %Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
@@
 %
-Photometric and polarimetric phase curves of starlight that is reflected by an exoplanet can reveal angular features that are due to ocean glint, and to rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
-%
-Direct imaging missions are optimised for planets near quadrature, and \hwo's coronagraph may obscure the phase angles where such characteristic features are strongest. 
+Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows, and other phenomena caused by scattering by clouds or atmospheric gas.
+%
+Direct imaging missions are optimised for planets near quadrature, and \HWO's coronagraph may obscure the phase angles where such scattering features are strongest. 
 %
@@
 %
-Planets in close-to-edge-on orbits have accessible phase angles limited by coronagraphic obscuration. 
-%
-We use \hwo'~catalog of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering by gas. 
-%
-We find that any exo-Earths in systems in this catalog will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
-%
-The ocean glint signature at planetary scattering angles of $\sim\qtyrange{50}{70}{\degree}$ 
-%\sim$110--130$^\circ$ 
-would be accessible to \hwo~in $\sim\num{14}$ systems, and the rainbow due to water clouds at phase angles of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} \IWA{} ($3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
-%
-Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of roughly 2 and 3, respectively.
-%
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with the planet's orbital eccentricity and scales approximately inversely with the \IWA{}.
+For planets in nearly edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
+%
+We use \HWO'~catalogue of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
+%
+We find that any exo-Earths in systems in this catalogue will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
+%
+The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO\ in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
+%
+Improving the \IWA to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of roughly 2 and 3, respectively.
+%
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.  %and scales approximately inversely with the \IWA.  
 \end{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Introduction}
@@
 
-%--------------------------------------------------------------------------------
 %\begin{figure*}
@@
         The polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). 
-        Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
+%        Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
  %   and the latter
@@
         Below the plot, we show the phase angle ranges that cover the
-        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
-        with each other).  
+        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles adopted in this paper.
+Note that the features in the actual curves partly overlap
+        with each other.  
         %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
@@
 
-The field of exoplanet research has made great advances in the last few decades.
-The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve.
-The Habitable Worlds Observatory will perform a dedicated direct imaging survey of nearby stars searching for rocky planets in the habitable zone.
-This space-based coronographic telescope will detect and characterise exoplanets using their reflection spectrum.
+%The field of exoplanet research has made great advances in the last few decades. 
+%The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve. 
+The Habitable Worlds Observatory (\HWO) will perform a dedicated direct imaging survey of nearby stars, searching for rocky planets in the habitable zone. 
+This space-based coronagraphic telescope will detect and characterise exoplanets using their reflection spectrum. 
 Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
 
-Variations of the flux and in particular the degree of polarization of
-light of the parent star that is reflected by an exoplanet as it moves
-along its orbit that are due to an ocean glint or optical phenomena from scattering by condensates --- like the rainbow --- could be some of the most distinct signatures of habitability.
-%
-However, depending on the architecture of a planetary system, some of these features can be obscured by the inner working angle (\IWA{}) of a telescope's coronagraph.
-
 Liquid surface water is closely related to planetary habitability because it is essential for life as we know it.
@@
 %
-The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star for which a rocky planet (or moon) may have the right temperature to maintain liquid surface water \citep{kasting93}.
-%
-However, there are many factors besides the incident stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and, for example, geological activity.
+The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet may have the right temperature to maintain liquid surface water \citep{kasting93}.
+%
+However, there are many factors besides the incident stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and geological activity.
 %
 Establishing the presence of liquid surface water on an exoplanet would therefore provide important context for the interpretation of possible biosignatures.
-%
-Providing additional context to the habitability of an exoplanet would be the presence of a water cycle, as evidenced by the identification of a glint, i.e.\ the specular reflection of incident light by a liquid surface, or by a rainbow and other optical phenomena that can be attributed to light that is scattered by water clouds, as detailed in the next sections.
+
+Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal the most distinct signatures of habitability: 
+specular reflection (a.k.a. glint) from water oceans, or optical phenomena from scattering by condensates---like the rainbow.
+%
+However, depending on the architecture of a planetary system, some of these features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph.
+%
+%Providing additional context to the habitability of an exoplanet would be the presence of a water cycle, as evidenced by the identification of a  as detailed in the next sections.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Exoplanet phase curves}
-
-An exoplanet in an edge-on orbit moves through phase angles ranging from $0^\circ$ (full phase), through $90^\circ$ (quadrature), and to $180^\circ$ (the planet's nightside is towards the observer) (cf.\ Fig.~\ref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
-%
-If the planet is rotating, its reflected light signal in both total and polarized flux is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet \citep{2001Natur.412..885F,stam2008,2022A&A...664A..59M}.
-
-The strengths of detailed total and polarized flux phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
-%
-An added potential benefit of observing the state of polarization of the reflected light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
+%\subsection{Exoplanet phase curves}
+
+%Exoplanets exhibit phase variations as they orbit their host star because
+
+%
+%If the planet is rotating, its reflected light signal in both total and polarized flux is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet \citep{2001Natur.412..885F,stam2008,2022A&A...664A..59M}.
+
+%The strengths of detailed total and polarized flux phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
+%
+%An added potential benefit of observing the state of polarization of the reflected light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
@@
 %As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
-Ocean glint is the specular (Fresnel) reflection of light off smooth liquid surfaces. While glint is a familiar phenomenon on Earth, there is only
-one other body in the Solar System where it has been observed: 
-the Cassini orbiter captured the glint of sunlight reflecting off the
-liquid hydrocarbon lakes near Titan's north pole as it was observing
-the moon from the nightside \citep{2010GeoRL..37.7104S}.
-%
+Ocean glint is the specular (Fresnel) reflection of light off a smooth liquid surface. 
+While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing the moon at crescent phases \citep{2010GeoRL..37.7104S}.
 In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
-%
-The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at an angle of reflection of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.
-%
+As shown in Figure \ref{fig:bottplot}, the maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a scattering angle of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]{2008Icar..195..927W,Zugger_2010,treesandstam2019}.
 This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
-%
-
-In general, the glint on a planet will be strongest at red wavelengths where the Rayleigh scattering in the planet's atmosphere and attenuation through the atmosphere are smallest \citep{Zugger_2011} and, conveniently, where the latitude-albedo effect (a degeneracy between polar brightness and glint) is also small \citep{2012ApJ...752L...3C}.
-%
+\looseness=-1
+
+In general, the glint on a planet will be strongest at red wavelengths because atmospheric Rayleigh scattering and attenuation are smallest \citep{Zugger_2011}.
 The stronger glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
-%
-The width across phase angles of the glint feature in total and polarized flux generally
-broadens with increasing wave height \citep{kopparla2018, Zugger_2010, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citet{CoxMunk1954}.
-%
-
-The glint provides evidence of specular reflecting surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
-%
-At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which further motivated probing the rainbow angle for context.
-%
-This confusion can be solved by measuring the intersection of phase curves at various wavelengths in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction \citep{treesandstam2019}, as shown in Fig.~\ref{fig:bottplot}.
+%A degeneracy between polar brightness and glint is also smaller at redder wavelengths \citep{2012ApJ...752L...3C}.
+%
+The width across phase angles of the glint feature in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citet{CoxMunk1954}.
+%
+
+Glint provides evidence of specular reflecting surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
+%
+At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which motivates probing the rainbow angle for context, as described below.
+%
+This confusion can be resolved by measuring polarized phase curves at multiple wavelengths %in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction
+\citep{treesandstam2019}, as shown in \cref{fig:bottplot}.
 %The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
@@
 %
-While the glint feature can only be probed when a planet is at crescent, i.e.\ large phase angles, \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
-This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
+%While the glint feature can only be probed when a planet is at crescent (i.e., large phase angles), \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
+%This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
 %how use to map, 
@@
 
-Rainbows, cloudbows, glories and other optical phenomena from light that is 
-scattered by atmospheric condensates can dramatically influence the shape of a planet's total flux and polarized phase curves and provide a wealth of information 
-about exoplanets \citep{karalidi2012rainbow, stam2008,Bailey2007,2014A&A...566L...1G}.%
+Rainbows, cloudbows, glories and other optical phenomena from light that is scattered by atmospheric condensates can dramatically influence the shape of a planet's total flux and polarized phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,Bailey2007,2014A&A...566L...1G}.%
 \footnote{
     Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light. 
-Cloudbows refer to features occurring at the same phase angles from smaller cloud particles with the spread in colors invisible to the human eye making them appear white \citep[see][for further discussion]{Bailey2007}.
+    Cloudbows refer to features occurring at the same phase angles from smaller cloud particles, with the spread in colours invisible to the human eye making them appear white \citep[see][for further discussion]{Bailey2007}.
 }
@@
 %
-The rainbow angle depends on the refractive index of the scatterers and thus depends on their composition and the wavelength. 
-%
-The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
+The rainbow angle depends on the refractive index of the scattering media, and thus depends on their composition and the wavelength. 
+%
+The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}.
 %
@@
 %
-Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halo's, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow} that also depend on the composition and shape of the particles \citep[see e.g.][for simulations of halo's due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
-
-In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of 40$^\circ$ (see Fig.~\ref{fig:bottplot}).
+Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halos, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow}, that also depend on the composition and shape of the particles \citep[see, e.g.,][for simulations of halos due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
+
+In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of \qty{40}{\degree} (see \cref{fig:bottplot}).
 %
@@
 %
-The glory can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase and close to the star \citep{hansentravis1974}, as can also be seen in Fig.~\ref{fig:bottplot}.
+The glory can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase when the on sky separation is small \citep{hansentravis1974}, as can also be seen in \cref{fig:bottplot}.
+
 %
@@
 The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program, a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
-%
 This requires an instrument with a coronagraph capable of high contrast imaging at optical to near infrared wavelengths.
-%
-Following the release of this survey, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
-%
+Following the release of this survey, NASA recently announced the start of the development of the \HWO.
 
@@
 %
-The performance and precise characteristics of \hwo{} (on-axis or off-axis, diameter, type of segmentation, type of coronagraph) are still to be determined.
+The performance and precise characteristics of \HWO (on-axis or off-axis, diameter, type of segmentation, type of coronagraph) are still to be determined.
 %
@@
 %
-The Inner Working angle (\IWA{}) of the resulting telescope will depend on these characteristics, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
-
-The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronograph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
+The Inner Working Angle (\IWA) of the resulting telescope will depend on these characteristics, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
+
+The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronagraph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
 %
@@
 %
-However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \hwo{} will include a UV high-contrast instrument sensitive to Earth-like planets.
+However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
 
@@
 
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations for hypothetical terrestrial planets orbiting in the habitable zones of \hwo{} target stars.
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
 %
@@
 
+\subsection{Angle conventions}
 The conventions for angle names vary throughout the literature on exoplanet phase curves. 
-In this paper we define the \textbf{scattering} angle as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \textbf{phase} angle $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
-A planet with a face-on orbit will thus have a phase angle of 90$^\circ$ everywhere along its orbit.
-In some plots we ``fold'' the half orbit so that we measure \textbf{degrees from quadrature} (so that both new and full phase are at maxima of 90$^\circ$ for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
-We also note that we are considering here the (linearly) polarized flux
-which takes into account the intensity of the polarization and reflectance (illumination) of the planet but additional information could be garnered from full Stokes characterization.
+In this paper, we define the \emph{scattering angle} as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \emph{phase angle} $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
+A planet with a face-on orbit will thus have a phase angle of \qty{90}{\degree} everywhere along its orbit. 
+In some plots, we \enquote{fold} the half orbit so that we measure \emph{degrees from quadrature} (so that both new and full phase are at maxima of \qty{90}{\degree} for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
+We also note that we are considering here the (linearly) polarized flux, which considers the intensity of the polarization and reflectance (illumination) of the planet. 
+Additional information could be garnered from full Stokes characterization.
+
+An exoplanet in an edge-on orbit moves through phase angles ranging from \qty{0}{\degree} (full phase), through \qty{90}{\degree} (quadrature), and to \qty{180}{\degree} (the planet's nightside is towards the observer) (cf.\ \cref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
 
@@
 
-\subsection{\hwo{} and its stellar sample}
+\subsection{\HWO and its stellar sample}
 \label{subsec:2.1}
@@
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\hwo{} is still in the early stages of development, and although the exact telescope and instrument design have not yet been determined, we assume a \qty{6}{\meter} primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
+\HWO is still in the early stages of development, and although the exact telescope and instrument design have not yet been determined, we assume a \qty{6}{\meter} primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
 %
@@
 %
-The \IWA{} of the coronagraph will be driven by both the science requirements and technological limitations and is still unknown.
-%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA{} of the observatory. 
+The \IWA of the coronagraph will be driven by both the science requirements and technological limitations and is still unknown.
+%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA of the observatory. 
 %We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
@@
     \caption{
-        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., the \IWA{} and \OWA{}, respectively).
+        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., the \IWA and \OWA{}, respectively).
         %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
@@
 
-\hwo{} is envisioned to observe the stars in the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
+\HWO is envisioned to observe the stars in the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
 \footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
@@
 
-In what follows, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
-%
-By design, such planets should be observable by \hwo{} at least at their quadrature phase ($\alpha=90^\circ$).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
+By design, such planets should be observable by \HWO at least at their quadrature phase (i.e., $\alpha=\qty{90}{\degree}$).
 
@@
 A coronagraph can obscure part of a planet's orbit which, as illustrated in \cref{fig:annotated-orbit}, can prevent the planet from being observable at small and large phase angles, where it is close to its star.
-%
-The range of observable phase angles will depend on the \IWA{} of the coronagraph, as well as on the inclination and shape, i.e.\ eccentricity, of the exoplanet's orbit.
-
-%--------------------------------------------------------------------------
+The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and shape (i.e., the eccentricity) of the exoplanet's orbit.
+
 % Figure 2
-%--------------------------------------------------------------------------
 \begin{figure}
@@
         An illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of scattering features in an exoplanet's phase curves that could help to characterize its atmospheric and/or surface composition and habitability.
-        The scattering angle, $\varphi$, is shown in red. It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
+        The scattering angle, $\varphi$, is shown in red. 
+        It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
     }
@@
 
-For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about 90$^\circ$ (see Fig.~\ref{fig:accessible_phase_angles}).
-%
+For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about \qty{90}{\degree} (see \cref{fig:accessible_phase_angles}).
 We define the angle $\Delta \varphi$ such that: 
@@
 semi-major orbital axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer. 
-%
-Angle $\varphi$ is given by
+The angle $\varphi$ is given by
 \begin{equation}
-    \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,.
     \label{eq:sinphi}
+    \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,,
 \end{equation}
-%
-Assuming small angles, $r_\mathrm{proj} = \delta d_*$ where $\delta$ is the on-sky angular separation.
-%
+assuming small angles and $r_\mathrm{proj} = \delta d_*$, where $\delta$ is the on-sky angular separation.
 We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
-%
 Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
-%
 \begin{equation}
@@
 
-%--------------------------------------------------------------------------
 % Figure 3
-%--------------------------------------------------------------------------
 \begin{figure}
@@
 \end{figure}
-%--------------------------------------------------------------------------
-
-\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \hwo{} target list, assuming an IWA of \qty{62}{\mas} ($3\,\lambda / D$ coronagraph at $\qty{600}{\nano\meter}$, see \cref{tab:IWA_OWA}).
+
+\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas} ($3\,\lambda / D$ coronagraph at $\qty{600}{\nano\meter}$, see \cref{tab:IWA_OWA}).
 %
@@
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature ($\alpha=90^\circ$), while the rainbow angle (around $\alpha=40^\circ$) will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle (around $\alpha=\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
 
@@
     \caption{
-        Scatter plot for the target star sample, showing stellar effective 
-        temperatures and distances.
-        %
-        The circle sizes represent the angular separation between 
-        the star and its planet (in mas) as presented in the target list.
-        %
-        The colours show $\Delta \varphi$ assuming circular, edge-on orbits
-        at a semi-major axis $a$ corresponding to an Earth-like 
-        instellation and an IWA of \qty{62}{\mas}.
-        %
-        Additionally, the colour bar indicates the optical phenomena that 
-        can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
-        %
+        Scatter plot for the target star sample, showing stellar effective temperatures and distances.
+        The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
+        The colours show $\Delta \varphi$ assuming circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
+        Additionally, the colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
         For example, dark blue circles indicate planets on which most phenomena would be observable because planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
@@
 \end{figure*}
-%--------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Most planets are not in edge-on orbits as seen from Earth.
-%
 For an orbit with an inclination angle $i$, we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
@@
     \item The coronagraph does not obscure any part of the sky-projected orbit: 
-    the maximum phase angle coverage depends solely on $i$. This 
-    holds for orbits that are relatively face-on; 
-    \item The coronagraph does obscure parts of the sky-projected orbit: 
-    the phase angle coverage depends only on the \IWA{} and 
-    the orbital semi-major axis $a$. 
+    the maximum phase angle coverage depends solely on $i$. 
+        This holds for orbits that are relatively face-on. 
+    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$. 
 \end{enumerate}
@@
 %        The phase angles accessible to a direct imaging mission 
-%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA{}.
+%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA.
         %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsubsection{Eccentric orbits}
-
-%The inclination $i$ is sampled such that $\cos(i)$ follows a uniform distribution, $\cos(i) \sim \mathcal{U}(0, 1)$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Monte Carlo simulations}
@@
 
-For each star on the \hwo{} target list, we generate \num{1000} random orbits of hypothetical habitable planets in orbits with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
-%
-The orbital inclination angles $i$ are randomly drawn from a distribution 
-that is uniform in $\cos i$. 
-%
-We then use \cref{eq:Delta_phi_max} to determine the range of 
-accessible phase angles for each hypothetical planet.
-%
-The solid lines in \cref{fig:betaallofit} show the cumulative 
-distribution of the maximum and minimum accessible phase 
-angles normalised to the total number of systems in the \hwo{} 
-target list for different values of the \IWA{}.
+For each star on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets in orbits with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
+The orbital inclination angles $i$ are randomly drawn from a distribution that is uniform in $\cos i$. 
+We then use \cref{eq:Delta_phi_max} to determine the range of accessible phase angles for each hypothetical planet.
+The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible phase angles normalised to the total number of systems in the \HWO target list for different values of the \IWA.
 
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{figures/figure-5-n-over-phase.pdf}
+    \includegraphics{figures/figure-5-n-over-phase.pdf}
     \script{create-figure-5-n-over-phase.py}
     \caption{
-        The cumulative distributions of the most extreme phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
-        (solid lines) and randomly inclined, elliptical orbits (dashed lines).
-        %
-        The top $x$-axis indicates the minimum and the 
-        bottom $x$-axis the maximum accessible phase angle.
-        %
+        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, elliptical orbits (dashed lines).
+        The top $x$-axis indicates the minimum and the bottom $x$-axis the maximum accessible phase angle.
         These angles are symmetric about quadrature (90 degrees).
-        %
-        The $y$-axis on the left indicates the number of planetary systems 
-        divided by the number of Monte Carlo samples and is thus normalised to 
-        the number of systems on the target list.
-        %
-        The $y$-axis on the right indicates the number of systems, assuming \qty{24}{\percent} of them have an Earth-like planet in their HZ.
+        The $y$-axis on the left indicates the number of planetary systems divided by the number of Monte Carlo samples, and is thus normalised to the number of systems on the target list.
+        The $y$-axis on the right indicates the number of systems, assuming $\eta_\oplus = \qty{24}{\percent}$ of them have an Earth-like planet in their HZ.
     }
@@
 \end{figure*}
-%--------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsubsection{Eccentric, randomly inclined orbits}
@@
 
-We repeat the Monte Carlo simulations from \cref{sec:circular} assuming 
-the orbital eccentricity follows a beta-distribution with shape parameters 
-$a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
-%
+We repeat the Monte Carlo simulations from \cref{sec:circular}, assuming the orbital eccentricity follows a beta-distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
 We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
-Orbital eccentricities are drawn from the beta distribution, and 
-inclination angles are uniform in $\cos i$. 
+Orbital eccentricities are drawn from the beta distribution, and inclination angles are uniform in $\cos i$. 
 \Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
-%
-
-We then calculate the minimum and maximum accessible phase 
-angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
-%
-The dashed lines in \cref{fig:betaallofit} show the normalised cumulative 
-distribution of the accessible phase angles for each value of the 
-\IWA{} for eccentric orbits. 
-%
+
+We then calculate the minimum and maximum accessible phase angles for inner working angles of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
+The dashed lines in \cref{fig:betaallofit} show the normalised cumulative distribution of the accessible phase angles for each value of the \IWA for eccentric orbits. 
 We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
-%
-The cumulative distribution over the \IWA{} of the number of planetary systems where all optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The cumulative distribution over the \IWA of the number of planetary systems where all optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
+
 \subsection{Simulating contrast curves}
@@
 
-A second parameter that is crucial for detecting angular features like 
-rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e.\ 
-the ratio of the flux of the starlight that is reflected by the planet 
-and the direct stellar flux. 
-%
-While this is not the main focus of this paper, it is important to consider that the total flux contrast and the polarized contrast generally 
-depend on the phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
-%
-Here, we aim to present some simplified examples of the impact of the 
-phase angle dependent contrast on the detectability of rainbow and 
-glint features. 
-%
-We note that these calculations are by no means complete or exact, given 
-the many variables that go into the modelling of reflected total and
-polarized fluxes of Earth-like exoplanets \citep{ treesandstam2019,trees2022}.
-%
-
-The NASA Exoplanet Exploration Program’s Mission Star List for \hwo{} 
-includes for every star an estimated contrast for an Earth-twin in 
-the habitable zone.
-%
+A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the ratio of the flux of the starlight that is reflected by the planet and the direct stellar flux.
+While this is not the main focus of this paper, it is important to consider that the total flux contrast and the polarized contrast generally depend on the phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
+Here, we aim to present some simplified examples of the impact of the phase angle dependent contrast on the detectability of rainbow and glint features. 
+We note that these calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
+
+The NASA Exoplanet Exploration Program’s Mission Star List for \HWO includes for every star an estimated contrast for an Earth-twin in the habitable zone.
 % The cited paper has all the equations that lead to this equation:
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$
-(normalized to \num{1.0} at full phase, thus at $\alpha=0^\circ$), 
-and $a$ is the distance between the planet and its host star.
-%
-The assumptions that went into the contrast estimations for \hwo{}
-are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
-%
-These calculations give an order-of-magnitude calculation for the 
-contrast of Earth-like planets around nearby stars.
-%
-Therefore, in this work, we will assume that the calculated contrasts 
-in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
-%
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
+The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
+These calculations give an order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
+Therefore, in this work, we will assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
 We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
-%
-
-We use the phase functions from \citet{treesandstam2019} 
-for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
-%
+
+We use the phase functions from \citet{treesandstam2019} for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
 The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
-%
-We normalize the phase function at quadrature and multiply it with the 
-contrast.
-%
+We normalize the phase function at quadrature and multiply it with the contrast.
 In addition, we map the orbital phase for circular orbits to the phase angle and the on-sky separation given an inclination of \num{90} degrees.
-%
 Together, they give the reflected light contrast as a function of separation for a full orbit.
-%
 We multiply this flux with the degree of polarization and retrieve the polarized contrast.
-%
-The resulting contrast curves are shown in \cref{fig:contrasts} for 
-three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf 
-$\epsilon$~Eri, and the M~dwarf Lalande~21185.
+The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri, and the M~dwarf Lalande~21185.
 
@@
     \centering
-    \includegraphics[width=0.99\textwidth]{figures/figure-6-contrast-over-separation_i=90.pdf}
+    \includegraphics[width=\textwidth]{figures/figure-6-contrast-over-separation_i=90.pdf}
     \script{create-figure-6-contrast-over-separation.py}
     \caption{
-        The contrast (Eq.~\ref{eq:contrast}) and orbital separation 
-        of an Earth-like planet with an ocean surface and patchy clouds 
-        along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A, $\epsilon$~Eri, and Lalande~21185.
-        %
-        The solid, black line indicates the contrast in unpolarized light, 
-        with the contrast at quadrature marked by the black dot.
-        %
+        The contrast (\cref{eq:contrast}) and orbital separation of an Earth-like planet with an ocean surface and patchy clouds along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A, $\epsilon$~Eri, and Lalande~21185.
+        The solid, black line indicates the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
         The polarized component is indicated by the coloured dots, with the colour representing the phase angle from quadrature.
-        %
-        The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA{} for \hwo.
+        The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA for \HWO.
     }
@@
 \end{figure*}
-%--------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Results}
@@
 
-%\todo{Is there a plot for this? How many systems could we see? Why is it different to non-eccentric cases?}
-%\timmy{No plot / results yet; see above.}
-%\todo{Sophia: what is features\_plot\_peak.png showing?}
-%\timmy{Good question --- maybe Matt or Max know how the data for this was generated?
-Matt: yes, it's all in showyourwork right now.
-I ran 1000 sims for each target with a random eccentric orbit, then pulled out the betamax and betamin angles for each of the 1000 realisations.
-Max's plots encode the completeness as a function of scattering angle, showing that you get about 10 percent more phenomena.
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{The effect of eccentricity}
@@
 As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable effect on the number of systems that can be observed at each phase angle.
-%
-This is because---as Kepler's second law explains---eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA{} and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
-%
-The eccentricity does not significantly affect the number of systems for \IWA{} smaller than cut used in the selection of the systems in the \hwo{} target list. 
-It is, however, important to account for orbital eccentricity when the \IWA{} of the coronagraph is larger than the \IWA{} used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This is because---as Kepler's second law explains---eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+The eccentricity does not significantly affect the number of systems for \IWA smaller than cut used in the selection of the systems in the \HWO target list. 
+It is, however, important to account for orbital eccentricity when the \IWA of the coronagraph is larger than the \IWA used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
 
@@
 
-In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA{}.
-%
+In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA.
 We first define the scattering angles over which various of these phenomena are accessible in \cref{tab:phase_ranges}.
-%
 We note that these ranges should all be considered approximate, as the exact scattering angles where these phenomena are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
-%
-Nonetheless, this exercise can provide valuable insights on the effects of \IWA{} choice.
-
-\Cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA{}.
-%
-The \IWA{} of \hwo's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
-%
-The smaller the \IWA{}, the more features we will be observable and in more systems.
-
-%--------------------------------------------------------------------------
-%--------------------------------------------------------------------------
+Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
+
+\Cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
+The \IWA of \HWO's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
+The smaller the \IWA, the more features we will be observable and in more systems.
+
 \begin{table}
@@
     \caption{
-        The approximate ranges of scattering angles for the different optical phenomena considered in this paper.
+        The approximate ranges of the phase angle $\varphi$ for the different optical phenomena considered in this paper.
     }
@@
         \midrule
-        Rainbow    & 22$^\circ$    & 42$^\circ$    & 63$^\circ$ \\
-        Rayleigh    & 50$^\circ$    & 70$^\circ$    & 110$^\circ$ \\
-        Ocean Glint & 130$^\circ$   & 150$^\circ$   & 170$^\circ$ \\
-        Glory       & 0$^\circ$     & 5$^\circ$     & 10$^\circ$ \\
+        Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
+        Rainbow    & \qty{22}{\degree}    & \qty{42}{\degree}    & \qty{63}{\degree} \\
+        Rayleigh    & \qty{50}{\degree}    & \qty{70}{\degree}    & \qty{110}{\degree} \\
+        Ocean Glint & \qty{130}{\degree}   & \qty{150}{\degree}   & \qty{170}{\degree} \\
         \bottomrule
@@
 \end{table}
-%--------------------------------------------------------------------------
-    % features = {r'Rainbow': (127,138,158),
-    %         r"Rayleigh": (90,110,130),
-    %         r"Ocean Glint": (50,30,10),
-    %         r"Glory": (10,5,0),
-    %         }
 
 The Rayleigh feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-%
-The rainbow feature appears further from quadrature (around $\alpha=40^\circ$), but will still be observable in a significant fraction of systems.
-%
+The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be observable in a significant fraction of systems.
 The position and shape of the rainbow peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
-%
-The ocean glint feature occurs even further from quadrature and so would be observable in fewer systems. 
-%
+The ocean glint feature occurs even further from quadrature, and so would be observable in fewer systems. 
 In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet.
-%
 A key indicator of habitability. 
@@
 
-%--------------------------------------------------------------------------
-%--------------------------------------------------------------------------
 \begin{table}
@@
     \caption{
-        Number of stars around which the peak scattering angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
+        The number of stars around which the peak scattering angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
     }
@@
 \end{table}
-%--------------------------------------------------------------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The accessible phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
-%
 The contrast of the planet is also an important factor to consider.
-%
-\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \hwo{} target list assuming edge-on orbits.
-%
+\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming edge-on orbits.
 The overall contrast varies significantly for each system and throughout the orbit.
-%
-The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA{}.
-%
-It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \hwo{} is built.
-%
-However, when its design is being finalised, it will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA{} of the coronagraph, if these features are to be observed.
+The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
+It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \HWO is built.
+However, when its design is being finalised, it will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the coronagraph, if these features are to be observed.
 
@@
         Examples of the eccentric orbits generated for the stellar sample.
-        %
         The orbits are scaled by the Earth-equivalent incident flux distance. 
-        %
-        The concentric, dashed circles indicate \IWA{}s of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
-        %
-        Each system has been scaled such that its HZ has the same radius, 
-        hence the variation in the size of the \IWA{}s compared to the orbits.
-        It is clear that the \IWA{} can significantly affect the range of 
-        observable scattering phases for each orbit.
+        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        Each system has been scaled such that its HZ has the same radius, hence the variation in the size of the \IWA{}s compared to the orbits.
+        It is clear that the \IWA can significantly affect the range of observable scattering phases for each orbit.
     }
@@
         The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glories) would be observable as functions of the \IWA.
-        %
         Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
-        %
-        The $x$-axis at the bottom shows the \IWA{} of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA{} converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        %
+        The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
         The $y$-axis on the right indicates the number of systems where 
         all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
-        %
         It is worth noting that the Glory would be observable in only one system: Alpha~Cen~A.
@@
 
-The Habitable Worlds Observatory is a coronagraphic space telescope 
-envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
-%
-The instrument design and capabilities are still under development and 
-should be driven by the mission's science goals.
-%
-In this work, we investigated the possibility of using \hwo{} to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory's, which could establish a planet's habitability.
-%
-We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA{} and aimed to investigate which phenomena could potentially be observed.
-%
-Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \hwo{} will be.
-%
-The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets, will drive the choice of coronagraphic \IWA{} for this mission.
-%
-Assuming a \qty{62}{\mas} \IWA{}, all \num{160} planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \hwo{} would be observable at scattering phases where angular features due to Rayleigh scattering show up, but the rainbow of light that is scattered in droplets of water clouds would only be accessible in $\sim\num{42}$ systems, and ocean glint in $\sim\num{14}$ systems.
-%
-The habitability of Earth-like exoplanets in the remaining 146 systems would have to be established through less direct methods like rotational mapping to identify horizontally inhomogeneous surface coverage, for example due to presumably bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
-%
-The fraction of habitable zones around stars in the target list that can be probed by \hwo{} increases with decreasing \IWA{}. 
-%
-Additionally, decreasing the \IWA\ also increases the range of scattering angles observable in these systems. 
-%
+The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
+The instrument design and capabilities are still under development and should be driven by the mission's science goals.
+In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory's, which could establish a planet's habitability.
+We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
+Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \HWO will be.
+The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets, will drive the choice of coronagraphic \IWA for this mission.
+Assuming a \qty{62}{\mas} \IWA, all \num{160} planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at scattering phases where angular features due to Rayleigh scattering show up, but the rainbow of light that is scattered in droplets of water clouds would only be accessible in $\sim\num{42}$ systems, and ocean glint in $\sim\num{14}$ systems.
+The habitability of Earth-like exoplanets in the remaining 146 systems would have to be established through less direct methods near quadrature: spectropolarimetry to deduce ocean reflection \citep{trees2022} or rotational mapping to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
+The fraction of habitable zones around stars in the target list that can be probed by \HWO increases with decreasing \IWA. 
+Additionally, decreasing the \IWA also increases the range of scattering angles observable in these systems. 
 Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
-%
 While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
-%
 This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
@@
 %\todo{the paragraph below seems similar to the one above ... merge them?}
-%We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA\ of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \hwo{} would be observable at a range of scattering angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
+%We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA\ of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \HWO would be observable at a range of scattering angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
 %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This paper has been formatted in the \texttt{showyourwork!} framework \citep{Luger2021}, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible.%
+\footnote{\url{https://github.com/mkenworthy/HWObows/}}
+
 \section*{Acknowledgements}
@@
 
-SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
+SRV and JB acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
@@
 
-This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \clearpage
-
 % References
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Author affiliations
 \section*{Author affiliations}
@@
     \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
-    \affdest{14}Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
+    \affdest{14}Aix Marseille Université, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
     \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2023-05-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows, and other phenomena caused by scattering by clouds or atmospheric gas.
+Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
 %
-Direct imaging missions are optimised for planets near quadrature, and \HWO's coronagraph may obscure the phase angles where such scattering features are strongest. 
+Direct imaging missions are optimised for planets near quadrature, and \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
 %
@@
 %
-For planets in nearly edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
+For planets in near edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
 %
@@
 %
-We find that any exo-Earths in systems in this catalogue will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
+We find that any exo-Earths in planetary systems in this catalogue will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
 %
-The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO\ in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
+The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO\ in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-1-bott-plot.pdf}
-    \script{create-figure-1-bott-plot.py}
-    \caption{
-        The phase curves of the normalized polarized flux of a planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (the planet thus exhibits phase angles ranging from \qty{0}{\degree} to \qty{180}{\degree} along its orbit).
-        The wind speed over the ocean is \qty{7}{\meter\per\second}. 
-        The clouds in the planet's atmosphere consist of spherical water droplets. 
-        The cloud coverage fraction is 0.5. 
-        The cloud pattern is patchy, and the glint is kept cloud-free to maximize the glint signature \citep[see also Fig.
-9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the \enquote{noise} at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
-        The polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). 
-%        Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
- %   and the latter
- %   is equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
-        %The cloudy scenario includes a low-altitude, Earth-like water 
-        %cloud deck. 
-        Below the plot, we show the phase angle ranges that cover the
-        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles adopted in this paper.
-Note that the features in the actual curves partly overlap
-        with each other.  
-        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
-    }
-    \label{fig:bottplot}
-\end{figure*}
-
 %--------------------------------------------------------------------------------
@@
 %The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve. 
-The Habitable Worlds Observatory (\HWO) will perform a dedicated direct imaging survey of nearby stars, searching for rocky planets in the habitable zone. 
-This space-based coronagraphic telescope will detect and characterise exoplanets using their reflection spectrum. 
-Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
+%The Habitable Worlds Observatory (\HWO) will perform a dedicated direct imaging survey of nearby stars, searching for rocky planets in the habitable zone. 
+%This space-based coronagraphic telescope will detect and characterise exoplanets using their reflection spectrum. 
+%Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
 
@@
 %
-To first order, the presence of liquid surface water can be predicted by the flux a planet receives from its star.
+To first order, its presence can be predicted by the flux a planet receives from its star.
 %
@@
 %
-However, there are many factors besides the incident stellar flux that influence a planet's surface temperature and ability to hold liquid surface water, such as the atmospheric surface pressure and geological activity.
+However, there are many factors besides the incident stellar flux that influence a planet's ability to hold liquid surface water, such as the atmospheric pressure and geological activity. 
 %
-Establishing the presence of liquid surface water on an exoplanet would therefore provide important context for the interpretation of possible biosignatures.
-
-Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal the most distinct signatures of habitability: 
-specular reflection (a.k.a. glint) from water oceans, or optical phenomena from scattering by condensates---like the rainbow.
+Detecting liquid surface water on an exoplanet would thus provide important context for the interpretation of possible biosignatures.
 %
-However, depending on the architecture of a planetary system, some of these features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph.
+Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection, i.e.
+glint, from water oceans \citep[][]{2008Icar..195..927W}.
+%
+However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in Fig.~\ref{fig:annotated-orbit}.
 %
@@
 
+\subsection{Angle conventions}
+%The conventions for angle names vary throughout the literature on exoplanet phase curves. 
+As shown in Fig.~\ref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle 
+%local divergence in the propagation direction of the light 
+(large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (zero at full phase and 180$^\circ$ at new phase). 
+A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = 90^\circ$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles.
+%a phase angle of \qty{90}{\degree} everywhere along its orbit. 
+In some plots, we report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.
+%, which is helpful when considering the observability of the planet given a certain inner working angle.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure 1
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+    \includestandalone{tikz/annotated-orbit}
+    \caption{
+        Illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of characteristic scattering features in an exoplanet's
+        phase curve. 
+        The scattering angle, $\varphi$, is shown in red and the supplementary phase angle, $\alpha$, is shown in \todo{blah}. Also see Figure~\ref{fig:scattering-angle}
+    }
+    \label{fig:annotated-orbit}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Ocean glint}
@@
 Ocean glint is the specular (Fresnel) reflection of light off a smooth liquid surface. 
-While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing the moon at crescent phases \citep{2010GeoRL..37.7104S}.
+While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing %the moon
+Titan at crescent phases \citep{2010GeoRL..37.7104S}.
 In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
-As shown in Figure \ref{fig:bottplot}, the maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a scattering angle of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]{2008Icar..195..927W,Zugger_2010,treesandstam2019}.
+%As shown in Figure \ref{fig:bottplot}, 
+The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a scattering angle of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]
+%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
+{Zugger_2010,treesandstam2019}.
 This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
+The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
 \looseness=-1
 
-In general, the glint on a planet will be strongest at red wavelengths because atmospheric Rayleigh scattering and attenuation are smallest \citep{Zugger_2011}.
-The stronger glint signature at longer wavelengths has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+In general, the glint on a planet will be strongest at red wavelengths because there atmospheric Rayleigh scattering and attenuation are smallest \citep{Zugger_2011}.
+This has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
 %A degeneracy between polar brightness and glint is also smaller at redder wavelengths \citep{2012ApJ...752L...3C}.
 %
-The width across phase angles of the glint feature in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}, which depends strongly on the wind speed over the surface, as described by \citet{CoxMunk1954}.
+%The width across phase angles of the glint feature in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}, which depends on the surface wind speed, as described by \citet{CoxMunk1954}.
 %
 
-Glint provides evidence of specular reflecting surface bodies, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for the presence of liquid water.
+Glint provides evidence of specular reflecting surfaces, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for liquid water.
 %
-At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}, which motivates probing the rainbow angle for context, as described below.
+At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
 %
-This confusion can be resolved by measuring polarized phase curves at multiple wavelengths %in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction
-\citep{treesandstam2019}, as shown in \cref{fig:bottplot}.
+This confusion can be resolved by measuring polarized phase curves at multiple wavelengths
+\citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
+the rainbow, see below.
+%in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction
+%\citep{treesandstam2019}, as shown in \cref{fig:bottplot}.
 %The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure 2
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+    \centering
+    \includegraphics{figures/figure-1-bott-plot.pdf}
+    \script{create-figure-1-bott-plot.py}
+    \caption{
+        The phase curves of the normalized linearly polarized flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit; adapted from \citet{treesandstam2019}.
+        %(the planet goes through all possible phase angles throughout its orbit). 
+        The wind speed over the ocean, which determines
+        the wave height \citep[][]{CoxMunk1954}, is \qty{7}{\meter\per\second}. 
+        The clouds consist of spherical water droplets. 
+        The cloud coverage fraction is 0.5 and while the  
+        cloud pattern is patchy, the glint is always cloud-free. 
+        %to maximize the glint signature \citep[see also Fig.
+9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the \enquote{noise} at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
+        %The polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). 
+        % Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
+        % and the latter
+        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
+        % The cloudy scenario includes a low-altitude, Earth-like water 
+        % cloud deck. 
+        Below the plot, we show the scattering angles of glories, rainbows, Rayleigh scattering, and ocean glint. Note that the features in the actual curves partly overlap with each other.
+        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
+    }
+    \label{fig:bottplot}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Rainbows, cloudbows and glories}
+\subsection{Rainbows, cloudbows, and glories}
 
-Rainbows, cloudbows, glories and other optical phenomena from light that is scattered by atmospheric condensates can dramatically influence the shape of a planet's total flux and polarized phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,Bailey2007,2014A&A...566L...1G}.%
+Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
 \footnote{
     Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light. 
-    Cloudbows refer to features occurring at the same phase angles from smaller cloud particles, with the spread in colours invisible to the human eye making them appear white \citep[see][for further discussion]{Bailey2007}.
+    Cloudbows refer to features occurring at the same phase angles from smaller cloud particles, with the spread in colours invisible to the human eye making them appear white \citep[for further discussion, see][]{bailey2007}.
 }
 %
-In our own Solar System, the polarized back scatter from Titan provided evidence that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, the polarized phase curve of Venus at three wavelengths showed not only that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture but also constrained the droplet size distribution \citep{hansenhovenier1974}.
+In the Solar System, the polarized back-scatter from Titan indicated that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, Venus's polarized phase curve at three wavelengths showed that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture and constrained the droplet size distribution \citep{hansenhovenier1974}.
 %
-The glory feature in Venus's total flux phase curve was later used to rule out potential solutions for the unknown UV absorber \citep{petrova2018glory}.
-
+The glory feature in Venus's flux phase curve was later used to rule out various species as the unknown UV absorber \citep{petrova2018glory}.
 %In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a Lambertian reflecting planet.
 %
-While the glory and rainbow are technically visible both in the total flux
-and the polarized flux phase curves in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized flux than in total flux \citep{karalidi2011, stam2008, treesandstam2019}.
+Glories and rainbows are technically visible both in total and polarized flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarized nature coupled with the suppression of other polarized features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
 %
-The rainbow angle depends on the refractive index of the scattering media, and thus depends on their composition and the wavelength. 
-%
-The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}.
-%
-Because the rainbow feature is narrower than the glint due to the surface reflection, it would allow for a more exact determination of the composition of the condensates.
-%
-Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halos, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow}, that also depend on the composition and shape of the particles \citep[see, e.g.,][for simulations of halos due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
+
+% The rainbow angle depends on the refractive index of the scattering media, and thus depends on their composition and the wavelength. 
+% %
+% The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}.
+% %
+% Because the rainbow feature is narrower than the glint due to the surface reflection, it would allow for a more exact determination of the composition of the condensates.
+% %
+% Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halos, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow}, that also depend on the composition and shape of the particles \citep[see, e.g.,][for simulations of halos due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
 
@@
 %
-Pushing observations to angles closer to the star (smaller phase angles, i.e., larger scattering angles) would allow the determination of the full shape of this rainbow, and enable the identification of rainbows due to the scattering of light by other possible HZ condensates.
+Pushing observations to angles closer to the star (smaller phase angles, i.e., larger scattering angles) would allow the determination of the full shape of the rainbow, and enable the identification of rainbows due to the scattering of light by other condensates.
 %
-The glory can allow further characterization of condensates as its angle and shape also depend on the complex refractive index, but this feature occurs within a phase angle of \qty{20}{\degree}, thus very close to full phase when the on sky separation is small \citep{hansentravis1974}, as can also be seen in \cref{fig:bottplot}.
+The glory can allow further characterization of condensates as its angle and shape also depend on the refractive index, but this glory occurs within a phase angle of \qty{20}{\degree}, when the on-sky separation is small, as can also be seen in \cref{fig:bottplot}.
+We note that we are considering here only linearly polarized fluxes, not circularly polarized fluxes.
+%which considers the intensity of the polarization and reflectance (illumination) of the planet. 
+%Additional information could be garnered from a full Stokes characterization.
 
@@
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program, a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
-This requires an instrument with a coronagraph capable of high contrast imaging at optical to near infrared wavelengths.
-Following the release of this survey, NASA recently announced the start of the development of the \HWO.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+This requires an instrument with a coronagraph capable of high-contrast imaging and low-resolution spectroscopy at optical to near-infrared wavelengths. 
+%It will measure the planet's reflected light at low spectral resolution to identify its atmospheric constituents.
+The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
 
-The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
 %
-The performance and precise characteristics of \HWO (on-axis or off-axis, diameter, type of segmentation, type of coronagraph) are still to be determined.
+Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter ``\HWO'').
 %
-The development will be heavily informed by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
+The performance and characteristics (on- or off-axis, diameter, type of segmentation and coronagraph, \IWA) of \HWO will significantly 
+impact the expected exo-Earth yield 
+\citep{Stark2019_exoplanetyield}
+and are still to be determined.
 %
-The Inner Working Angle (\IWA) of the resulting telescope will depend on these characteristics, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.
+They will likely be heavily informed by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
+%
+%\HWO' \IWA and thus the expected exo-Earth yield and characterization capabilities depend on these characteristics \citep{Stark2019_exoplanetyield}.
 
-The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronagraph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
-%
-POLLUX, a proposed instrument for LUVOIR would detect polarized light from hot Jupiters at \qty{300}{\nano\meter} \citep{Bouret2018_pollux}.
-%
-However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
+% The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronagraph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
+% %
+% POLLUX, a proposed instrument for LUVOIR would detect polarized light from hot Jupiters at \qty{300}{\nano\meter} \citep{Bouret2018_pollux}.
+% %
+% However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
 
@@
 %
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \cref{subsec:2.4}, we outline how the planet/star contrast changes along a planet's orbits.
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
 %
 We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
-It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
 
-\subsection{Angle conventions}
-The conventions for angle names vary throughout the literature on exoplanet phase curves. 
-In this paper, we define the \emph{scattering angle} as the local divergence in the propagation direction of the light (large angles for back-scattering, small angles for forward-scattering); the \emph{phase angle} $\varphi$ as the viewer-star-planet angle, which is zero at full phase and 180 at new phase, and which is the complement of the scattering angle.
-A planet with a face-on orbit will thus have a phase angle of \qty{90}{\degree} everywhere along its orbit. 
-In some plots, we \enquote{fold} the half orbit so that we measure \emph{degrees from quadrature} (so that both new and full phase are at maxima of \qty{90}{\degree} for edge-on orbits), which is helpful when considering the observability of the planet given a certain inner working angle.
-We also note that we are considering here the (linearly) polarized flux, which considers the intensity of the polarization and reflectance (illumination) of the planet. 
-Additional information could be garnered from full Stokes characterization.
+%It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
 
-An exoplanet in an edge-on orbit moves through phase angles ranging from \qty{0}{\degree} (full phase), through \qty{90}{\degree} (quadrature), and to \qty{180}{\degree} (the planet's nightside is towards the observer) (cf.\ \cref{fig:bottplot}), while planets in higher inclination orbits will pass through a subset of these phase angles.
+%An exoplanet in an edge-on orbit moves through phase angles ranging from \qty{0}{\degree} (full phase) through \qty{90}{\degree} (quadrature) to \qty{180}{\degree} (the planet's nightside is towards the observer; cf.\ \cref{fig:bottplot}), while planets in more inclined orbits will pass through a subset of these phase angles.
 
@@
 %\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\HWO is still in the early stages of development, and although the exact telescope and instrument design have not yet been determined, we assume a \qty{6}{\meter} primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
+\HWO is still in the early stages of development, so we assume a $D=\qty{6}{\meter}$ primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
 %
-The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint accompanying exoplanets. 
+The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint exoplanets. 
 %
-The \IWA of the coronagraph will be driven by both the science requirements and technological limitations and is still unknown.
+The \IWA of the coronagraph will be driven by science requirements and technological limitations. 
 %In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA of the observatory. 
@@
 
-%--------------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Table 1
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}
@@
     \caption{
-        Conversion between $\lambda / D$ and \unit{\mas} for wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming the telescope's diameter $D$ is $\qty{6}{\meter}$ for the Inner and Outer Working Angles (i.e., the \IWA and \OWA{}, respectively).
+        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milli-arcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
         %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
@@
 \end{table}
-%--------------------------------------------------------------------------
-
-\HWO is envisioned to observe the stars in the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \HWO is envisioned to observe the stars in the 
+A preliminary target list of nearby stars around which \HWO may search for habitable rocky planets has been compiled: the
+\emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
 \footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
 %
-This list comprises $\sim\num{160}$ stars, the majority of which are Sun-like dwarfs: 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
+This list consists of $\num{164}$ stars, the majority of which are Sun-like dwarfs: 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
 %
-The target list is constructed assuming a maximum planet magnitude of $R_c = 31$ and a contrast floor of \num{2.5e-11}. 
+The target list is constructed assuming a maximum planet magnitude in the Cousins Rc band of $R_c = 31$ and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
 %with an occurrence rate of rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
-The corresponding adopted habitable zone limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin, planet sizes between \qtyrange{0.8}{1.4}{} Earth radii, and for non-solar stars, scale as square root of the bolometric luminosity normalized to the Sun.
+The adopted HZ limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin, the planet sizes between \qtyrange{0.8}{1.4}{} Earth radii, and for non-solar stars, the semi-major axis scales as square root of the bolometric luminosity normalized to the Sun.
 %
@@
 %
-The authors of the Decadal Survey determined that for an imaging survey of habitable zones with a \qty{6}{\meter} telescope, these are the nearby stars (maximum distance \qty{25}{\parsec}) around which exo-Earths would be the best observable.
+The authors of the Decadal Survey determined that for an imaging survey of habitable zones with a \qty{6}{\meter} telescope, these are the nearby stars (maximum distance \qty{25}{\parsec}) around which exo-Earths would be observable.
 
 In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
-By design, such planets should be observable by \HWO at least at their quadrature phase (i.e., $\alpha=\qty{90}{\degree}$).
+By design, such planets should be observable by \HWO at quadrature phase. %(i.e., $\alpha=\qty{90}{\degree}$).
 
-\subsection{Maximum phase angle coverage}
+%\subsection{Relevant scattering angles}
+%In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA.
+
+\subsection{Accessible scattering angles}
 \label{subsec:2.2}
 
-A coronagraph can obscure part of a planet's orbit which, as illustrated in \cref{fig:annotated-orbit}, can prevent the planet from being observable at small and large phase angles, where it is close to its star.
-The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and shape (i.e., the eccentricity) of the exoplanet's orbit.
+We first define the phase angle ranges over which the various phenomena --- glint, Rayleigh peak, rainbow and glory --- manifest themselves in \cref{tab:phase_ranges}, also shown in Fig.~\ref{fig:bottplot}.
+These ranges are all approximate, as the exact angles where these phenomena dominate depend on the details of the atmospheric and surface properties, as well as on the wavelengths being observed.
+%Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
 
-% Figure 2
-\begin{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Table 2
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
     \centering
-    \includestandalone{tikz/annotated-orbit}
     \caption{
-        An illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of scattering features in an exoplanet's phase curves that could help to characterize its atmospheric and/or surface composition and habitability.
-        The scattering angle, $\varphi$, is shown in red. 
-        It is the complement of the phase angle $\alpha$, and, in an edge-on system, also the complement of the planet's orbital phase.
+        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. Fig.~\ref{fig:bottplot}).
     }
-    \label{fig:annotated-orbit}
-\end{figure}
-%--------------------------------------------------------------------------
+    \label{tab:phase_ranges}
+    \begin{tabular}{ l c c c } 
+        \toprule
+        & $\alpha_\mathrm{min}$ & $\alpha_\mathrm{peak}$ & $\alpha_\mathrm{max}$ \\
+        \midrule
+        \midrule
+        Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
+        Rainbow    & \qty{22}{\degree}    & \qty{42}{\degree}    & \qty{63}{\degree} \\
+        Rayleigh peak    & \qty{50}{\degree}    & \qty{70}{\degree}    & \qty{110}{\degree} \\
+        Ocean Glint & \qty{130}{\degree}   & \qty{150}{\degree}   & \qty{170}{\degree} \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A coronagraphic mask can obscure part of a planet's orbit.
+As illustrated in \cref{fig:annotated-orbit} obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
+The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and eccentricity of the exoplanet's orbit.
 
@@
 
-For the non-eccentric case (i.e., a circular planetary orbit with the coronagraphic mask centred on the star), the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible angles are symmetric about \qty{90}{\degree} (see \cref{fig:accessible_phase_angles}).
+For the non-eccentric case, 
+%(i.e., a circular planetary orbit with the coronagraphic mask centred on the star), 
+the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering angles are symmetric about \qty{90}{\degree}. 
 We define the angle $\Delta \varphi$ such that: 
@@
 The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a 
-semi-major orbital axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer. 
+semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer. 
 The angle $\varphi$ is given by
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Figure 3
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
     \caption{
-        The on-sky projected planet distance, $r_\mathrm{proj}$, of a planet with a semi-major orbital axis $a$ at a distance $d_*$ from the observer.
-        The on-sky angular separation as viewed by the observer is $\delta$.
-        % Angle $\varphi$ is measured between the line of sight to the star and the line connecting the star and planet.  
+        The angular separation of a planet from its star as seen by the observer, $\delta$, is related to the scattering angle, $\varphi$.
+The star is at a distance $d_*$ from the observer, the planet orbits a distance $a$ from the star, and  the projected distance between the star and planet is  $r_\mathrm{proj}$.
+Not to scale.
     }
@@
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the angular feature that is representative for Rayleigh scattering, as that occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle (around $\alpha=\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle (around $\alpha=\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Figure 4
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}
@@
     \caption{
-        Scatter plot for the target star sample, showing stellar effective temperatures and distances.
+        Scatter plot for the \HWO target stars, showing stellar effective temperatures and distances.
         The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
-        The colours show $\Delta \varphi$ assuming circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
-        Additionally, the colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
-        For example, dark blue circles indicate planets on which most phenomena would be observable because planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
+        The colours show $\Delta \varphi$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
+        The colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
+        For example, dark blue circles indicate planets for which most phenomena would be observable: planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
     }
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Most planets are not in edge-on orbits as seen from Earth.
-For an orbit with an inclination angle $i$, we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
+For an orbit with an inclination angle $i$ ($i=0^\circ$ is face-on), we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
 \begin{enumerate}
@@
     the maximum phase angle coverage depends solely on $i$. 
-        This holds for orbits that are relatively face-on. 
-    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$. 
+        This holds for orbits that are relatively face-on (small $i$). 
+    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$, following Eq.~\ref{eq:scattering_angle}.
 \end{enumerate}
@@
 
-Using \cref{eq:scattering_angle}, we deduce the following for inclined circular orbits: 
+%Using \cref{eq:scattering_angle}, we deduce the following 
+The range of accessible scattering angles is then 
 \begin{equation}
@@
         2 \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA}\; d_* }{a}
-    \end{cases} \,.
+    \end{cases} \,
 \end{equation}
 
-\subsection{Monte Carlo simulations}
+\subsection{Monte Carlo simulations of accessible orbits}
 \label{subsec:2.3}
@@
 
-For each star on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets in orbits with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
+For each star on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
 The orbital inclination angles $i$ are randomly drawn from a distribution that is uniform in $\cos i$. 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Figure 5
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}%[t]
@@
     \caption{
-        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, elliptical orbits (dashed lines).
-        The top $x$-axis indicates the minimum and the bottom $x$-axis the maximum accessible phase angle.
-        These angles are symmetric about quadrature (90 degrees).
+        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
+        The top and bottom $x$-axes indicate the minimum and maximum accessible phase angles, respectively.
+        These angles are symmetric about quadrature (90$^\circ$).
         The $y$-axis on the left indicates the number of planetary systems divided by the number of Monte Carlo samples, and is thus normalised to the number of systems on the target list.
-        The $y$-axis on the right indicates the number of systems, assuming $\eta_\oplus = \qty{24}{\percent}$ of them have an Earth-like planet in their HZ.
+        The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
+        {\color{red}{the plot says 'elliptical orbits'}}
     }
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure 7: Yarn ball
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+    \centering
+    \includegraphics{figures/figure-7-yarnball.pdf}  
+    \script{create-figure-7-yarnball.py}
+    \caption{
+        Examples of the eccentric orbits generated for the stellar sample. Each plot shows 10 examples of the orbits drawn by the Monte-Carlo for each star.
+        The orbits are scaled to ensure Earth-like incident flux. 
+        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
+        It is clear that the \IWA can significantly affect the range of observable scattering phases for each orbit.
+    }
+    \label{fig:ball-o-yarn}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-We then calculate the minimum and maximum accessible phase angles for inner working angles of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
+We then calculate the minimum and maximum accessible phase angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
 The dashed lines in \cref{fig:betaallofit} show the normalised cumulative distribution of the accessible phase angles for each value of the \IWA for eccentric orbits. 
 We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
-The cumulative distribution over the \IWA of the number of planetary systems where all optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
+%The cumulative distribution over the \IWA of the number of planetary systems where these optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
 
@@
 
-A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the ratio of the flux of the starlight that is reflected by the planet and the direct stellar flux.
-While this is not the main focus of this paper, it is important to consider that the total flux contrast and the polarized contrast generally depend on the phase angles, due to a combination of the changing illumination fraction of the observable planetary disk and the non-isotropic reflection by the planet.
-Here, we aim to present some simplified examples of the impact of the phase angle dependent contrast on the detectability of rainbow and glint features. 
-We note that these calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
+A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e., the planet/star flux ratio.
+%that is, the ratio of the flux of the starlight that is reflected by the planet and the direct stellar flux.
+While not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
+We present some simplified examples in Fig.~\ref{fig:contrasts}. %of the impact of the phase angle dependent contrast on the detectability of rainbow and glint features. 
+These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
 
@@
 where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
-The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are thus circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function, $\phi(\alpha)$, can be described by a simple Lambertian reflectance phase function, which assumes isotropic reflection by the planetary surface.
-These calculations give an order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
-Therefore, in this work, we will assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
-We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
+The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet.
+%These calculations give an order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
+We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
+%We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
 
-We use the phase functions from \citet{treesandstam2019} for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$ at $\qty{670}{\nano\meter}$ and covered by patchy clouds.
-The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
-We normalize the phase function at quadrature and multiply it with the contrast.
-In addition, we map the orbital phase for circular orbits to the phase angle and the on-sky separation given an inclination of \num{90} degrees.
+We use the phase functions at $\qty{670}{\nano\meter}$ from \citet{treesandstam2019} for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$  and covered by patchy clouds.
+%The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
+%We normalize the phase function at quadrature and multiply it with the contrast.
+We map the orbital phase for circular orbits to the phase angle and on-sky separation given an inclination $i$ of 90$^\circ$. 
 Together, they give the reflected light contrast as a function of separation for a full orbit.
-We multiply this flux with the degree of polarization and retrieve the polarized contrast.
-The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri, and the M~dwarf Lalande~21185.
+We multiply this flux with the degree of polarization to obtain the polarized contrast.
+\cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri, and the M~dwarf Lalande~21185.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Figure 6
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}
@@
     \caption{
-        The contrast (\cref{eq:contrast}) and orbital separation of an Earth-like planet with an ocean surface and patchy clouds along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A, $\epsilon$~Eri, and Lalande~21185.
+        The planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean surface and patchy clouds along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A (G2V), $\epsilon$~Eri (K2V), and Lalande~21185 (M2V).
         The solid, black line indicates the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
         The polarized component is indicated by the coloured dots, with the colour representing the phase angle from quadrature.
-        The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times the \IWA for \HWO.
+        The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times $\lambda / D$ (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
     }
@@
 
-As shown in \cref{fig:betaallofit}, allowing the planet's orbit to be eccentric can have a noticeable effect on the number of systems that can be observed at each phase angle.
-This is because---as Kepler's second law explains---eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peak out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
-The eccentricity does not significantly affect the number of systems for \IWA smaller than cut used in the selection of the systems in the \HWO target list. 
-It is, however, important to account for orbital eccentricity when the \IWA of the coronagraph is larger than the \IWA used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
+As shown in \cref{fig:betaallofit}, allowing for eccentric orbits increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
+This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+The eccentricity does not significantly affect the number of systems for \IWA smaller than that used in the selection of the \HWO target list (yellow and blue lines). 
+%It is, however, important to account for orbital eccentricity when the \IWA of the coronagraph is larger than the \IWA used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
 
-\subsection{Observable optical phenomenon}
-\label{sec:results_scattering_phenomena}
+\subsection{Accessible scattering phenomena}
+%\label{sec:results_scattering_phenomena}
 
-In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA.
-We first define the scattering angles over which various of these phenomena are accessible in \cref{tab:phase_ranges}.
-We note that these ranges should all be considered approximate, as the exact scattering angles where these phenomena are dominant depend on the details of the atmospheric and surface properties, as well as the wavelengths being observed.
-Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
-
-\Cref{fig:accessible_phase_angles} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
-The \IWA of \HWO's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
-The smaller the \IWA, the more features we will be observable and in more systems.
-
-\begin{table}
-    \centering
-    \caption{
-        The approximate ranges of the phase angle $\varphi$ for the different optical phenomena considered in this paper.
-    }
-    \label{tab:phase_ranges}
-    \begin{tabular}{ l c c c } 
-        \toprule
-        & $\varphi_\mathrm{min}$ & $\varphi_\mathrm{peak}$ & $\varphi_\mathrm{max}$ \\
-        \midrule
-        \midrule
-        Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
-        Rainbow    & \qty{22}{\degree}    & \qty{42}{\degree}    & \qty{63}{\degree} \\
-        Rayleigh    & \qty{50}{\degree}    & \qty{70}{\degree}    & \qty{110}{\degree} \\
-        Ocean Glint & \qty{130}{\degree}   & \qty{150}{\degree}   & \qty{170}{\degree} \\
-        \bottomrule
-    \end{tabular}
-\end{table}
-
+\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
+%The \IWA of \HWO's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
+%The smaller the \IWA, the more features we will be observable and in more systems. 
 The Rayleigh feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
@@
 The ocean glint feature occurs even further from quadrature, and so would be observable in fewer systems. 
-In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet.
-A key indicator of habitability. 
-Finally, it is possible to view the glory scattering feature in one system, Alpha~Cen~A.
+In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability. 
+%Finally, it is possible to view the glory scattering feature in one system, Alpha~Cen~A.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}
@@
     \caption{
-        The number of stars around which the peak scattering angle of each phenomenon is detectable, expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
+        The number of stars around which the peak scattering angle of each phenomenon is detectable, as a function of inner working angle (IWA).
+%expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
+        %\todo{Do these numbers refer to circular or elliptical orbits?}
+        %\todo{Many of these values (e.g., glint) don't really match Fig. 5?}
     }
     \label{tab:nstars_detect}
-    \begin{tabular}{ l c c c } 
+    % \begin{tabular}{ l c c c } 
+    %     \toprule
+    %     & \multicolumn{3}{c}{Inner Working Angle}\\
+    %     & \qty{21}{\mas} & \qty{41}{\mas} & \qty{62}{\mas} \\
+    %     & $2\,\lambda/D$ & $3\,\lambda/D$ & $4\,\lambda/D$\\
+    %     \midrule
+    %     \midrule
+    %     Rainbows    & 90  & 42    & 18 \\
+    %     Rayleigh    & 152 & 115   & 65 \\
+    %     Ocean Glint & 43  & 14    & 5 \\
+    %     % Glory       & 0   & 5     & 10 \\
+    %     \bottomrule
+    % \end{tabular}\\[5mm]
+    %\todo{Just for comparison, here are some values from Fig. 5 (elliptical orbits) that I pulled straight out of the plotting script:}\\[5mm]
+    \begin{tabular}{lcccc}
     \toprule
-        & \multicolumn{3}{c}{IWA}\\
-        & \qty{21}{\mas} & \qty{41}{\mas} & \qty{62}{\mas} \\
-        & $2\,\lambda/D$ & $3\,\lambda/D$ & $4\,\lambda/D$\\
+    & \multicolumn{4}{c}{Inner Working Angle}\\
+    \cmidrule(lr){2-5}
+     Feature     &   21 mas &   41 mas &   62 mas &   83 mas \\
     \midrule
-        \midrule
-        Rainbows    & 90  & 42    & 18 \\
-        Rayleigh    & 152 & 115   & 65 \\
-        Ocean Glint & 43  & 14    & 5 \\
-        % Glory       & 0   & 5     & 10 \\
+     Glory       &        0 &        0 &        0 &        0 \\
+     Ocean Glint &       79 &       43 &       16 &        6 \\
+     Rainbow     &      109 &       91 &       47 &       23 \\
+     Rayleigh    &      154 &      152 &      128 &       82 \\
     \bottomrule
-    \end{tabular}
+    \end{tabular}\\[5mm]
+    %\todo{There is also a mismatch with Fig. 8, though. Where does this discrepancy come frome? Something about phase angle vs. scattering angle?}\\
+    %\todo{Observation: The values in the original table seem like a good match for Fig.~8 if we assume the colums where "shifted by one", i.e., 62 mas is actually 83 mas, 41 mas is 62 mas, ...}
 \end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The accessible phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
-The contrast of the planet is also an important factor to consider.
-\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming edge-on orbits.
-The overall contrast varies significantly for each system and throughout the orbit.
+%The accessible phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
+The contrast of the planet is also an important factor in assessing the detectability of these scattering phenomena. 
+\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming an Earth-like planet on an edge-on orbit.
+The overall contrast varies significantly for each system and throughout the orbit. In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
 The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
-It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \HWO is built.
-However, when its design is being finalised, it will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the coronagraph, if these features are to be observed.
+%It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \HWO is built.
+It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph, if these scattering features are to be observed.
 
-% Figure 7: Yarn ball
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-7-yarnball.pdf}  
-    \script{create-figure-7-yarnball.py}
-    \caption{
-        Examples of the eccentric orbits generated for the stellar sample.
-        The orbits are scaled by the Earth-equivalent incident flux distance. 
-        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
-        Each system has been scaled such that its HZ has the same radius, hence the variation in the size of the \IWA{}s compared to the orbits.
-        It is clear that the \IWA can significantly affect the range of observable scattering phases for each orbit.
-    }
-    \label{fig:ball-o-yarn}
-\end{figure*}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Figure 8
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}
@@
     \caption{
-        The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glories) would be observable as functions of the \IWA.
-        Each subplot contains three lines showing the number of systems where the start of the phenomena is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
+        The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
+        Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
         The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
         The $y$-axis on the right indicates the number of systems where 
-        all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
-        It is worth noting that the Glory would be observable in only one system: Alpha~Cen~A.
+        the phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
+        The glory peak would be observable in only one system: Alpha~Cen~A.
     }
@@
 
-The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
+The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
 The instrument design and capabilities are still under development and should be driven by the mission's science goals.
-In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory's, which could establish a planet's habitability.
+In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
 We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
-Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \HWO will be.
-The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets, will drive the choice of coronagraphic \IWA for this mission.
+%Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \HWO will be.
+
+The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets around different stellar spectral types, will drive the choice of coronagraphic \IWA for \HWO.
 Assuming a \qty{62}{\mas} \IWA, all \num{160} planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at scattering phases where angular features due to Rayleigh scattering show up, but the rainbow of light that is scattered in droplets of water clouds would only be accessible in $\sim\num{42}$ systems, and ocean glint in $\sim\num{14}$ systems.
 The habitability of Earth-like exoplanets in the remaining 146 systems would have to be established through less direct methods near quadrature: spectropolarimetry to deduce ocean reflection \citep{trees2022} or rotational mapping to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
-The fraction of habitable zones around stars in the target list that can be probed by \HWO increases with decreasing \IWA. 
-Additionally, decreasing the \IWA also increases the range of scattering angles observable in these systems. 
-Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
-While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
-This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
+The number of habitable zones that can be probed by \HWO increases with decreasing \IWA, as does the range of scattering angles observable in these systems. 
+A larger-diameter telescope with a more aggressive coronagraph
+%Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
+%While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
+%This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
 
@@
 
-The Optimal Exoplanet Imaging workshop (2023) on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, The Netherlands. 
+The 2023 \emph{Optimal Exoplanet Imagers} workshop, which sparked the work presented in this manuscript, was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, The Netherlands.
 %
@@
 
-SRV and JB acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
+SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement n°805445.
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
 TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-KMB acknowledges support from NASA Habitable Worlds grant No.
-80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
-ILa acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
-PBa, ILa, and YGu were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
+KMB acknowledges support from NASA Habitable Worlds grant n°80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement n°NNA13AA93A.
+IL acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
+PB, IL, and YG were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
 EHP is supported by the NASA Hubble Fellowship grant \#HST-HF2-51467.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
-L.A. and E.C. acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement No. 101044152).
-O.A. and L.K. acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No 819155)
+LA and EC acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement n°101044152).
+OA and LK acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement n°819155)
 SYH was funded by the generous support of the Heising-Simons Foundation.
 RJP is supported by NASA under award number 80GSFC21M0002.
+OHS acknowledges funding from the direction scientifique générale de l'ONERA (ARE Alioth).
 

### 2023-05-12 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO\ in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
@@
 In general, the glint on a planet will be strongest at red wavelengths because there atmospheric Rayleigh scattering and attenuation are smallest \citep{Zugger_2011}.
-This has indeed been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
+This has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}.
 %A degeneracy between polar brightness and glint is also smaller at redder wavelengths \citep{2012ApJ...752L...3C}.
@@
         % and the latter
-        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$.
+        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$, scattering angle $\varphi = 180^\circ$.
         % The cloudy scenario includes a low-altitude, Earth-like water 
@@
 
-Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
+Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.% Detailed characterization of the rainbow in phase and wavelength space can constrain the condensate species, providing context to the habitability of the planet.
 \footnote{
@@
 Glories and rainbows are technically visible both in total and polarized flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarized nature coupled with the suppression of other polarized features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
+\looseness=-1
 %
@@
 
-\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} shows the number of systems in which key scattering phenomena will be observable for different \IWA.
+\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be observable for different \IWA.
 %The \IWA of \HWO's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
@@
 TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-KMB acknowledges support from NASA Habitable Worlds grant n°80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement n°NNA13AA93A.
+KB acknowledges support from NASA Habitable Worlds grant n°80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement n°NNA13AA93A.
 IL acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.

### 2023-05-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 }
-\newcommand{\oceanglint}{\textcolor{MidnightBlue}{ocean gl\istar{}nt}\xspace}
+\newcommand{\oceanglints}{\textcolor{MidnightBlue}{ocean gl\istar{}nts}\xspace}
 
@@
 % Full title of the paper (MNRAS has no running title?)
-\title{Chasing \rainbows and \oceanglint:\\ Inner working angle constraints for the Habitable Worlds Observatory}
+\title{Chasing \rainbows and \oceanglints:\\ Inner working angle constraints for the Habitable Worlds Observatory}
 
 % Authors are in alphabetical order:
+% DO NOT EDIT THIS MANUALLY
+% ALL ENTRIES ARE GENERATED BY A SCRIPT
+% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
 \author[Sophia R. Vaughan et al.]{%
@@
     Alexis Bidot\textsuperscript{\afflink{16}},
-    Jayne L. Birkby\textsuperscript{\afflink{1}},
+    Jayne Birkby\textsuperscript{\afflink{1}},
     \newauthor
@@
     Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
-    Evangelia Kleisioti\textsuperscript{\afflink{7}},
+    Elina Kleisioti\textsuperscript{\afflink{7},\afflink{27}},
     \newauthor
     Lorenzo König\textsuperscript{\afflink{13}},
-    Mariya Krasteva\textsuperscript{\afflink{27}},
+    Mariya Krasteva\textsuperscript{\afflink{28}},
     Iva Laginja\textsuperscript{\afflink{11}},
@@
     David Mouillet\textsuperscript{\afflink{16}},
-    Mamadou N’Diaye\textsuperscript{\afflink{28}},
-    Emiel H. Por\textsuperscript{\afflink{29}},
-    Laurent Pueyo\textsuperscript{\afflink{29}},
+    Mamadou N’Diaye\textsuperscript{\afflink{29}},
+    Emiel H. Por\textsuperscript{\afflink{30}},
+    Laurent Pueyo\textsuperscript{\afflink{30}},
     Frans Snik\textsuperscript{\afflink{7}},
     \newauthor
-    Daphne M. Stam\textsuperscript{\afflink{30}},
-    Victor Trees\textsuperscript{\afflink{31},\afflink{32}},
+    Daphne M. Stam\textsuperscript{\afflink{31}},
+    Victor Trees\textsuperscript{\afflink{32},\afflink{33}},
     Dirk van Dam\textsuperscript{\afflink{7}},
     Kyle van Gorkom\textsuperscript{\afflink{21}},
-    Maaike van Kooten\textsuperscript{\afflink{33}}
+    Maaike van Kooten\textsuperscript{\afflink{34}}
     \newauthor \\%
@@
 NASA recently announced the Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
-%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal.
-After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties.
+%
 Surface liquid water is central to the definition of planetary habitability.
@@
 %
-The rainbow due to water clouds at scattering angles of \qtyrange{140}{160}{\degree} would be accessible to \HWO in $\sim\num{42}$ systems, while ocean glint signature at scattering angles of $\qtyrange{50}{70}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible to \HWO in $\sim\num{42}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-%\begin{figure*}
-    %\centering
-    %\includegraphics[width=\linewidth]%{figures/Fig1IWAPol.png}
-%    \caption{
-%        The phase curves of two homogeneous planets with oceans and Earth-like atmospheres in edge-on orbits (the planets thus exhibit scattering angles 
-%        ranging from 0$^\circ$ to 180$^\circ$ along their orbits). 
-%        The cloudy scenario includes a low-altitude, Earth-like water 
-%        cloud deck. 
-%        Below the plot, we are showing the scattering angle ranges that cover the
-%        characteristic features of glories, rainbows, Rayleigh scattering, glint, and forward scattering by cloud particles (note that features partly overlap
-%        with each other).  
-%        Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
-%    }
-%    \label{fig:bottplot}
-%\end{figure*}
-
-%--------------------------------------------------------------------------------
-
-%The field of exoplanet research has made great advances in the last few decades. 
-%The characterisation of Earth-like planets in the habitable zones of their stars is now at the frontier of what we can achieve. 
-%The Habitable Worlds Observatory (\HWO) will perform a dedicated direct imaging survey of nearby stars, searching for rocky planets in the habitable zone. 
-%This space-based coronagraphic telescope will detect and characterise exoplanets using their reflection spectrum. 
-%Complementary information on the habitability of these worlds can be obtained by observing the change in the planet's brightness at different planetary phases.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in Fig.~\ref{fig:annotated-orbit}.
-%
-%Providing additional context to the habitability of an exoplanet would be the presence of a water cycle, as evidenced by the identification of a  as detailed in the next sections.
-
-%The phase curves showing these features are distinct depending on the characteristics of the planet in both polarized and unpolarized measurements.
-Polarized and unpolarized phase curves and spectra each have their own strengths in characterization outlined below.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Exoplanet phase curves}
-
-%Exoplanets exhibit phase variations as they orbit their host star because
-
-%
-%If the planet is rotating, its reflected light signal in both total and polarized flux is expected to show temporal variations that can place constraints on the horizontal heterogeneity of the planet \citep{2001Natur.412..885F,stam2008,2022A&A...664A..59M}.
-
-%The strengths of detailed total and polarized flux phase curve analysis include mapping \citep{2001Natur.412..885F, berdyugina2019surface, 2022A&A...664A..59M}, the detection of dark areas or glint indicative of seas \citep{groot2020, cowan2008inverting, lustig2019}, and the characterization of clouds through forward- and back-scattering, rainbows, and glories.
-%
-%An added potential benefit of observing the state of polarization of the reflected light is the suppression of starlight from quiescent stars without explicitly requiring a coronagraph or starshade as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
@@
 \subsection{Angle conventions}
-%The conventions for angle names vary throughout the literature on exoplanet phase curves. 
-As shown in Fig.~\ref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle 
-%local divergence in the propagation direction of the light 
-(large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (zero at full phase and 180$^\circ$ at new phase). 
+
+As shown in Fig.~\ref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle (large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (zero at full phase and 180$^\circ$ at new phase).
+%
 A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = 90^\circ$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles.
-%a phase angle of \qty{90}{\degree} everywhere along its orbit. 
+%
 In some plots, we report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.
-%, which is helpful when considering the observability of the planet given a certain inner working angle.
 
@@
     \caption{
-        Illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of characteristic scattering features in an exoplanet's
-        phase curve. 
-        The scattering angle, $\varphi$, is shown in red and the supplementary phase angle, $\alpha$, is shown in \todo{blah}. Also see Figure~\ref{fig:scattering-angle}
+        Illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of characteristic scattering features in an exoplanet's phase curve.
+        The scattering angle (the planet--star--observer angle), $\varphi$, is shown in red, and the supplementary phase angle, $\alpha$, is shown in blue. 
+        See also \cref{fig:scattering-angle} and \cref{tab:phase_ranges}.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Ocean glint}
 
-%As ocean glint is one of the few sources of specular reflection and is strongly polarized, it provides one of the best means to detect a signature of habitability in exoplanet phase curves. 
 Ocean glint is the specular (Fresnel) reflection of light off a smooth liquid surface. 
-While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing %the moon
-Titan at crescent phases \citep{2010GeoRL..37.7104S}.
+%
+While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing Titan at crescent phases \citep{2010GeoRL..37.7104S}.
+%
 In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
-%As shown in Figure \ref{fig:bottplot}, 
-The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a scattering angle of \qty{53}{\degree} \citep[i.e., at a planetary phase angle of \qty{106}{\degree}; see, e.g.,][]
-%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
-{Zugger_2010,treesandstam2019}.
+%
+The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
+{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
+%
 This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
-The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
-\looseness=-1
-
-In general, the glint on a planet will be strongest at red wavelengths because there atmospheric Rayleigh scattering and attenuation are smallest \citep{Zugger_2011}.
+%
+The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.\looseness=-1 \space In general, the glint on a planet will be strongest at red wavelengths because there Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}.
+%
 This has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}. 
-%A degeneracy between polar brightness and glint is also smaller at redder wavelengths \citep{2012ApJ...752L...3C}.
 %
-%The width across phase angles of the glint feature in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}, which depends on the surface wind speed, as described by \citet{CoxMunk1954}.
+At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
 %
+This confusion can be resolved by measuring polarized phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
+the rainbow, see below.
 
@@
 %
-At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
+The presence of liquid water on the surface of a planet implies the presence of a water cycle which provides key context to its habitability and the interpretation of atmospheric biosignatures.
 %
-This confusion can be resolved by measuring polarized phase curves at multiple wavelengths
-\citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
-the rainbow, see below.
-%in polarized light (the polarized colour changes from blue, through white, to red), which only occurs when the glint contributes to the signal and at a planetary scattering angle depending on the cloud coverage fraction
-%\citep{treesandstam2019}, as shown in \cref{fig:bottplot}.
-%The cloud coverage fraction on ocean worlds could be constrained with observations spanning the ultraviolet to the near-infrared by finding the 
-%angle at which the planet's polarization phase curves at various wavelengths intersect \citep{treesandstam2019}.
-%
-%While the glint feature can only be probed when a planet is at crescent (i.e., large phase angles), \citet{trees2022} showed that reflection by a liquid surface body can also be detected with spectropolarimetry, as the glint changes the difference in polarization between continuum wavelengths and absorption bands.
-%This could be applied when the planet is near quadrature, and thus more easily probed when using coronagraphs.
-%how use to map, 
-%seen in Earthshine
-%strong signal for exoplanets
-% disentangled from cloud coupled with rainbow
-%disentangled from lattitude albedo coupled in polarization
+Aside from detecting ocean glint, the presence of surface water on Earth-like exoplanets can be established through less direct methods near quadrature such as spectropolarimetry, to infer ocean reflection \citep{trees2022}, or through rotational mapping, to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
 
@@
         The phase curves of the normalized linearly polarized flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit; adapted from \citet{treesandstam2019}.
-        %(the planet goes through all possible phase angles throughout its orbit). 
-        The wind speed over the ocean, which determines
-        the wave height \citep[][]{CoxMunk1954}, is \qty{7}{\meter\per\second}. 
+        %
+        The wind speed over the ocean, which determines the wave height \citep[][]{CoxMunk1954}, is \qty{7}{\meter\per\second}.
+        %
         The clouds consist of spherical water droplets.
-        The cloud coverage fraction is 0.5 and while the  
-        cloud pattern is patchy, the glint is always cloud-free. 
-        %to maximize the glint signature \citep[see also Fig.
-9 of][from which we took the mean signal of 300 planets with randomly generated patchy cloud patterns + 2 standard deviations, representative for cloud-free glint scenarios; the \enquote{noise} at the larger phase angles and longer wavelengths is due to the variation in the location of cloud patches]{treesandstam2019}.
-        %The polarized flux is normalized to the total (unpolarized + polarized) flux (it is thus in fact the degree of polarization). 
-        % Note that the reference plane for the polarized flux is chosen such that only the polarized components perpendicular or parallel to the plane differ significantly from zero.
-        % and the latter
-        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$, scattering angle $\varphi = 180^\circ$.
-        % The cloudy scenario includes a low-altitude, Earth-like water 
-        % cloud deck. 
+        %
+        The cloud coverage fraction is 0.5 and while the cloud pattern is patchy, the glint is always cloud-free. 
+        %
+        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$ (scattering angle $\varphi = 180^\circ$).
+        %
         Below the plot, we show the scattering angles of glories, rainbows, Rayleigh scattering, and ocean glint. Note that the features in the actual curves partly overlap with each other.
-        %Throughout the paper we approximate feature sensitivity ranges as those that capture the rise and fall of peaks in polarized light for most habitable zone species in the UV, Optical, and IR (phase angles: glint $\sim\qtyrange{50}{70}{\degree}$; rainbow $\sim\SIrange{140}{160}{\degree}$; Rayleigh narrow of quadrature, and other forward and backscattering effects at the widest and narrowest phase angles).
     }
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Rainbows, cloudbows, and glories}
 
-Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.% Detailed characterization of the rainbow in phase and wavelength space can constrain the condensate species, providing context to the habitability of the planet.
+Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
 \footnote{
@@
 The glory feature in Venus's flux phase curve was later used to rule out various species as the unknown UV absorber \citep{petrova2018glory}.
-%In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a Lambertian reflecting planet.
 %
@@
 
-% The rainbow angle depends on the refractive index of the scattering media, and thus depends on their composition and the wavelength. 
-% %
-% The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}.
-% %
-% Because the rainbow feature is narrower than the glint due to the surface reflection, it would allow for a more exact determination of the composition of the condensates.
-% %
-% Note that the scattering by clouds that consist of solid particles, such as ice crystals, will produce other optical features, such as halos, particularly at small phase angles \citep{2022JQSRT.29008313M,Bailey2007,hansentravis1974, karalidi2012rainbow}, that also depend on the composition and shape of the particles \citep[see, e.g.,][for simulations of halos due to carbon dioxide ice crystals]{2023GeoRL..5003457D}.
-
+The phase angles and wavelength dependence of these features are sensitive to the chemical composition and droplet size distribution of condensates in the atmosphere. 
+%
+As with ocean glint, the detection of liquid condensates on an exoplanet can provide key context to their habitability, potential water cycle, and interpretation of atmospheric biosignatures.
+%
 In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of \qty{40}{\degree} (see \cref{fig:bottplot}).
 %
-Pushing observations to angles closer to the star (smaller phase angles, i.e., larger scattering angles) would allow the determination of the full shape of the rainbow, and enable the identification of rainbows due to the scattering of light by other condensates.
+Pushing observations to angles closer to the star (smaller and larger phase angles) would allow the determination of the full shape of the rainbow, and enable the identification of rainbows due to the scattering of light by other condensates.
 %
 The glory can allow further characterization of condensates as its angle and shape also depend on the refractive index, but this glory occurs within a phase angle of \qty{20}{\degree}, when the on-sky separation is small, as can also be seen in \cref{fig:bottplot}.
-We note that we are considering here only linearly polarized fluxes, not circularly polarized fluxes.
-%which considers the intensity of the polarization and reflectance (illumination) of the planet. 
-%Additional information could be garnered from a full Stokes characterization.
 
-%
-%The glory (in the backscattering direction) and diffraction peak (in the forward scattering direction) are distinct features in the total reflected flux, but not in the degree of polarization.
-In principle, the degree of polarization goes to zero in the forward and backward scattering directions.
-See e.g.
-Fig.
-1 of \citet{stam2008}, Fig.
-5 of \citet{trees2022}, Fig.
-3 of \citet{karalidi2011} and \citet{hansentravis1974}.}}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Habitable Worlds Observatory}
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey (Astro2020) \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+%
 This requires an instrument with a coronagraph capable of high-contrast imaging and low-resolution spectroscopy at optical to near-infrared wavelengths. 
-%It will measure the planet's reflected light at low spectral resolution to identify its atmospheric constituents.
+%
 The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
-
 %
@@
 %
-The performance and characteristics (on- or off-axis, diameter, type of segmentation and coronagraph, \IWA) of \HWO will significantly 
-impact the expected exo-Earth yield 
-\citep{Stark2019_exoplanetyield}
-and are still to be determined.
+The performance and characteristics (on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
 %
@@
 %
-%\HWO' \IWA and thus the expected exo-Earth yield and characterization capabilities depend on these characteristics \citep{Stark2019_exoplanetyield}.
-
-% The Extreme Coronagraph for Living Planetary Systems (ECLIPS) coronagraph for LUVOIR was a proposed \qtyrange{200}{2000}{\nano\meter} instrument for exoplanet characterization, with a similar coronagraph, albeit operating between \num{450} and \qty{1800}{\nano\meter}, planned for HabEx.
-% %
-% POLLUX, a proposed instrument for LUVOIR would detect polarized light from hot Jupiters at \qty{300}{\nano\meter} \citep{Bouret2018_pollux}.
-% %
-% However, because of the challenges associated with the design of a coronagraph in the UV, it appears to be unlikely that \HWO will include a UV high-contrast instrument sensitive to Earth-like planets.
-
-%We ignore contrast in this study
-%OWA -> should go in discussions
-
-%\textbf{In this paper we...}
-\subsection{Outline}
-
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
-%
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
-%
-We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
-
-%It is important to note that the phases angles selected for the optical phenomena throughout this paper are approximate, that phenomena overlap, and that their maximum peak will depend on the chemical species and wavelength.
-
-%An exoplanet in an edge-on orbit moves through phase angles ranging from \qty{0}{\degree} (full phase) through \qty{90}{\degree} (quadrature) to \qty{180}{\degree} (the planet's nightside is towards the observer; cf.\ \cref{fig:bottplot}), while planets in more inclined orbits will pass through a subset of these phase angles.
-
-%the   the way that inclination and coronagraph inner working angle can limit the accessible scattering angles  
-% We minimally explore the contrast ratio as we are primarily exploring whether these phase angles are reachable with inner working angle and outer working angle considered.
-
-% We do not consider the particulars of specific planet scenarios as that work is completed by foward models
-
-%We do not separate Stokes parameters as we are interested in where the net signal is achievable with IWA and OWA.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section{Plots to Include}
-%\label{sec:plots}
-%\begin{enumerate}
-%    \item \st{Cartoon defining various scattering angles: orbital phase, scattering phase, $\beta$ ? (Sophia)}
-%    \item Bott plot: annotated polarized and unpolarized phase curves for ocean world and cloudy worlds (Kim)
-%    \item \st{3x3 cartoon showing effect of inclination and IWA compared to orbit (Sophia)}
-%    \item Cumulative distribution functions, both simple (circular, edge-on), bit more realistic (circular, but range of inclinations), and full-on (range of inclinations and eccentricities).
-Max \& Matthew
-%    \item Scatter plot of stellar effective temperature vs system distance (Timmy)
-%\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section{Observing scattering phenomena}
-\section{Methods}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{\HWO and its stellar sample}
-\label{subsec:2.1}
-
-%\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
-\HWO is still in the early stages of development, so we assume a $D=\qty{6}{\meter}$ primary mirror, as suggested in The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey.
-%
-The observatory will utilise a coronagraph to suppress the stellar light and facilitate high-contrast observations of faint exoplanets. 
-%
-The \IWA of the coronagraph will be driven by science requirements and technological limitations. 
-%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA of the observatory. 
-%We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
-%From the HabWorlds report.
-Inner Working Angle: The IWA defines the region near the star that cannot be accessed for direct imaging due to a coronagraphic mask, starshade obscuration, or an interferometric null.
-It depends on the architecture of the telescope, the starlight suppression system used, and the observation wavelength.
-While a nominal value could be defined by making assumptions about the design for Habitable Worlds Observatory, we choose instead to derive the IWA from the star list itself and Astro2020’s requirement that ~100 cumulative habitable zones be surveyed by direct imaging.
-As shown in the main table, an IWA near ~70 mas will be needed to access ~100 cumulative habitable zones, and would need to be achieved at all wavelengths of interest for spectral characterization.
-For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.
-For ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a \qty{6}{\meter} telescope at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+For generality, in this work, we report angular separations in milliarcseconds however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
 
@@
         Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milli-arcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
-        %\todo{We should use consistent wavelengths for this table, for the Bott Plot, and for the M\&M cumulative distribution function. Also update to use the only wavelength we choose.}
     }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \HWO is envisioned to observe the stars in the 
-A preliminary target list of nearby stars around which \HWO may search for habitable rocky planets has been compiled: the
-\emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Outline}
+
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
+%
+In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
+%
+We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\HWO and its stellar sample}
+\label{subsec:2.1}
+
+A preliminary target list of nearby stars (maximum distance \qty{25}{\parsec}) around which \HWO may search for habitable rocky planets, assuming a \qty{6}{\meter} telescope, has been compiled: the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
 \footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
@@
 The target list is constructed assuming a maximum planet magnitude in the Cousins Rc band of $R_c = 31$ and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
-%with an occurrence rate of rocky planets in the optimistic habitable zone to be $\eta_{\oplus}$ = 0.24 \citep{decadal}. 
-The adopted HZ limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin, the planet sizes between \qtyrange{0.8}{1.4}{} Earth radii, and for non-solar stars, the semi-major axis scales as square root of the bolometric luminosity normalized to the Sun.
+% 
+The adopted HZ limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin and planet sizes between \qtyrange{0.8}{1.4}{} Earth radii. 
+%
+For non-solar stars, the semi-major axis scales as square root of the bolometric luminosity normalized to the Sun. 
 %
@@
 %
-The authors of the Decadal Survey determined that for an imaging survey of habitable zones with a \qty{6}{\meter} telescope, these are the nearby stars (maximum distance \qty{25}{\parsec}) around which exo-Earths would be observable.
+In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
+%
+By design, such planets should be observable by \HWO at quadrature phase.
 
-In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
-By design, such planets should be observable by \HWO at quadrature phase. %(i.e., $\alpha=\qty{90}{\degree}$).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\subsection{Relevant scattering angles}
-%In this section, we consider the number of stars around which different optical phenomena could be observable for different \IWA.
-
-\subsection{Accessible scattering angles}
+\subsection{Accessible phase angles}
 \label{subsec:2.2}
@@
 We first define the phase angle ranges over which the various phenomena --- glint, Rayleigh peak, rainbow and glory --- manifest themselves in \cref{tab:phase_ranges}, also shown in Fig.~\ref{fig:bottplot}.
+%
 These ranges are all approximate, as the exact angles where these phenomena dominate depend on the details of the atmospheric and surface properties, as well as on the wavelengths being observed.
-%Nonetheless, this exercise can provide valuable insights on the effects of \IWA choice.
 
@@
         \midrule
-        \midrule
         Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Circular, edge-on orbits}
 
-For the non-eccentric case, 
-%(i.e., a circular planetary orbit with the coronagraphic mask centred on the star), 
-the minimum ($\varphi_\mathrm{min}$) and maximum ($\varphi_\mathrm{max}$) accessible scattering angles are symmetric about \qty{90}{\degree}. 
-We define the angle $\Delta \varphi$ such that: 
+For the non-eccentric case, the minimum ($\alpha_\mathrm{min}$) and maximum ($\alpha_\mathrm{max}$) accessible phase angles are symmetric about \qty{90}{\degree}. 
+We define the angle $\Delta \alpha$ such that: 
 \begin{equation}
     \label{eq:Delta_phi}
-    \Delta \varphi 
-    = \varphi_\mathrm{max} - \varphi_\mathrm{min}
-    =  2 \cdot (\varphi_\mathrm{max} - \qty{90}{\degree}) 
-    =  2 \cdot (\qty{90}{\degree} - \varphi_\mathrm{min} )  \,.
+    \Delta \alpha 
+    = \alpha_\mathrm{max} - \alpha_\mathrm{min}
+    =  2 \cdot (\alpha_\mathrm{max} - \qty{90}{\degree}) 
+    =  2 \cdot (\qty{90}{\degree} - \alpha_\mathrm{min} )  \,.
 \end{equation} 
-The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a 
-semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer. 
-The angle $\varphi$ is given by
+The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer.
+%
+The scattering angle, $\varphi$, is given by
 \begin{equation}
     \label{eq:sinphi}
-    \sin(\varphi) = \frac{r_\mathrm{proj}}{a} \,,
+    \sin(\varphi) = - \sin(\alpha) = \frac{r_\mathrm{proj}}{a} \,,
 \end{equation}
 assuming small angles and $r_\mathrm{proj} = \delta d_*$, where $\delta$ is the on-sky angular separation.
+%
 We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
+%
 Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
@@
     \label{eq:scattering_angle}
-    \cos\left(\dfrac{\Delta \varphi}{2}\right) = \frac{\mathrm{IWA} 
+    \cos\left(\dfrac{\Delta \alpha}{2}\right) = \frac{\mathrm{IWA} 
     \cdot d_*}{a} \,.
@@
 
-\Cref{fig:scatterplot} shows $\Delta \varphi$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas} ($3\,\lambda / D$ coronagraph at $\qty{600}{\nano\meter}$, see \cref{tab:IWA_OWA}).
-%
-%This assumes each planet has a semi-major axis such that it receives an Earth-equivalent instellation, which is provided in the input star list.
+\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas}.
 %
@@
         Scatter plot for the \HWO target stars, showing stellar effective temperatures and distances.
+        %
         The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
-        The colours show $\Delta \varphi$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
+        %
+        The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
+        %
         The colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
+        %
         For example, dark blue circles indicate planets for which most phenomena would be observable: planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Circular, inclined orbits}
@@
 Most planets are not in edge-on orbits as seen from Earth.
-For an orbit with an inclination angle $i$ ($i=0^\circ$ is face-on), we can identify the following two regimes:% (see also \cref{fig:orb-grid}):
+For an orbit with an inclination angle $i$ ($i=0^\circ$ is face-on), we can identify the following two regimes:
 \begin{enumerate}
@@
 
-% Old Figure 5
-%\begin{figure}
-%   \centering
-%   \includegraphics[width=0.99\columnwidth]{figures/figure-5-orb-grid.pdf}
-%   \script{create-figure-5-orb-grid.py}
-%   \caption{
-%        The phase angles accessible to a direct imaging mission 
-%        can be limited by either the orbital inclination angle $i$ or the coronagraphic \IWA.
-        %
-%        If the orbit is close to face-on (top panels), the phase angles probed along the orbit are limited by the inclination angle $i$; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm i$.
-        %
-%        For orbits close to edge-on (bottom panels) the planet attains all 
-%        phases, but the phase angles near \qty{0}{\degree} and \qty{180}{\degree} are not observable due to obscuration by the coronagraph; for circular orbits, the accessible scattering phases are $\phi \in \qty{90}{\degree} \pm \arcsin({\rm IWA}\cdot d_*/a)$.
-%    }
-%    \label{fig:orb-grid}
-%\end{figure}
-
-%Using \cref{eq:scattering_angle}, we deduce the following 
-The range of accessible scattering angles is then 
+The range of accessible phase angles is then 
 \begin{equation}
 \label{eq:Delta_phi_max}
-    \Delta \varphi = 
+    \Delta \alpha = 
     \begin{cases}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Monte Carlo simulations of accessible orbits}
@@
 
-For each star on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
+For each of the 164 stars on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
 The orbital inclination angles $i$ are randomly drawn from a distribution that is uniform in $\cos i$. 
@@
         Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
+        %
         The top and bottom $x$-axes indicate the minimum and maximum accessible phase angles, respectively.
+        %
         These angles are symmetric about quadrature (90$^\circ$).
+        %
         The $y$-axis on the left indicates the number of planetary systems divided by the number of Monte Carlo samples, and is thus normalised to the number of systems on the target list.
+        %
         The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
-        {\color{red}{the plot says 'elliptical orbits'}}
+        {\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
     }
@@
     \caption{
-        Examples of the eccentric orbits generated for the stellar sample. Each plot shows 10 examples of the orbits drawn by the Monte-Carlo for each star.
-        The orbits are scaled to ensure Earth-like incident flux. 
-        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        Examples of the eccentric orbits generated for the stellar sample. Each plot shows 10 examples of orbits randomly drawn for each star.
+        The orbits are scaled to ensure an Earth-like incident flux. 
+        %
+        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
+        %, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        %
         Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
-        It is clear that the \IWA can significantly affect the range of observable scattering phases for each orbit.
+        %
+        It is clear that the \IWA can significantly affect the range of observable phases angles for each orbit.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Eccentric, randomly inclined orbits}
@@
 We repeat the Monte Carlo simulations from \cref{sec:circular}, assuming the orbital eccentricity follows a beta-distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
+%
 We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
+%
 Orbital eccentricities are drawn from the beta distribution, and inclination angles are uniform in $\cos i$. 
+%
 \Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
 
-We then calculate the minimum and maximum accessible phase angles for \IWA{}s of \qtylist{21; 42; 63; 84}{\mas} which correspond to 1, 2, 3 and 4 times $\lambda / D$ at $\qty{600}{\nano\meter}$, respectively (see \cref{tab:IWA_OWA} and the dashed lines in \Cref{fig:ball-o-yarn}).
+We then calculate the minimum and maximum accessible phase angles for \IWA{}s of $\qtylist{21; 42; 63; 84}{\mas}$.
+%
 The dashed lines in \cref{fig:betaallofit} show the normalised cumulative distribution of the accessible phase angles for each value of the \IWA for eccentric orbits. 
+%
 We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
-%The cumulative distribution over the \IWA of the number of planetary systems where these optical phenomena would be observable, normalised to the total number of systems, is shown in \cref{fig:accessible_phase_angles}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulating contrast curves}
@@
 A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e., the planet/star flux ratio. 
-%that is, the ratio of the flux of the starlight that is reflected by the planet and the direct stellar flux.
+%
 While not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
-We present some simplified examples in Fig.~\ref{fig:contrasts}. %of the impact of the phase angle dependent contrast on the detectability of rainbow and glint features. 
+%
+We present some simplified examples in Fig.~\ref{fig:contrasts}. 
+%
 These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
@@
 The NASA Exoplanet Exploration Program’s Mission Star List for \HWO includes for every star an estimated contrast for an Earth-twin in the habitable zone.
-% The cited paper has all the equations that lead to this equation:
+%
 The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
@@
 where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
+%
 The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet.
-%These calculations give an order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
-We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and normalize the phase functions to this point.
-%We do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some work for the star list to be redone.
+%
+We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and use Eq.
+\ref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$, i.e.
+$C(90) = 6.37\%$.
 
-We use the phase functions at $\qty{670}{\nano\meter}$ from \citet{treesandstam2019} for an Earth-like planet with an ocean surface with a wind-speed of $\qty{7}{\meter\per\second}$  and covered by patchy clouds.
-%The $\qty{670}{\nano\meter}$ is close to the centre of the Rc band for Vega ($\qty{642}{\nano\meter}$). 
-%We normalize the phase function at quadrature and multiply it with the contrast.
+We use the phase functions at $\qty{670}{\nano\meter}$ from \citet{treesandstam2019} for an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds.
+%
+We correct the Star List contrast by dividing by the $6.37\%$ reflected light fraction of the Lambertian phase function and multiply by the reflected light fraction of the \citet{treesandstam2019} model of $5.64\%$.
+%
 We map the orbital phase for circular orbits to the phase angle and on-sky separation given an inclination $i$ of 90$^\circ$. 
+%
 Together, they give the reflected light contrast as a function of separation for a full orbit.
+%
 We multiply this flux with the degree of polarization to obtain the polarized contrast.
-\cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri, and the M~dwarf Lalande~21185.
+%
+\cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A
 
@@
     \caption{
-        The planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean surface and patchy clouds along its orbit with an assumed inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A (G2V), $\epsilon$~Eri (K2V), and Lalande~21185 (M2V).
-        The solid, black line indicates the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
-        The polarized component is indicated by the coloured dots, with the colour representing the phase angle from quadrature.
-        The light grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate $1$, $2$ and $3$ times $\lambda / D$ (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
+        The planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit with an inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A (G2V), $\epsilon$~Eri (K2V), and Lalande~21185 (M2V).
+        %
+        The solid, black lines indicate the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
+        %
+        The polarized contrast (at $\lambda = \qty{670}{\nano\meter}$) is indicated by the coloured dots, with the colour representing the degrees from quadrature.
+        %
+        The light grey dots indicate the contrasts at quadrature of the other targets in the star list, and the dashed lines \IWA's of $\qtylist{21; 42; 63}{\mas}$.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
 \label{sec:3}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 As shown in \cref{fig:betaallofit}, allowing for eccentric orbits increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
+%
 This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+%
 The eccentricity does not significantly affect the number of systems for \IWA smaller than that used in the selection of the \HWO target list (yellow and blue lines).
-%It is, however, important to account for orbital eccentricity when the \IWA of the coronagraph is larger than the \IWA used to filter the target list; otherwise, the number of systems observable at extreme scattering angles would be under-predicted.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Accessible scattering phenomena}
@@
 \Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be observable for different \IWA.
-%The \IWA of \HWO's coronagraph will depend on the primary mirror diameter, observing wavelength, and how well a coronagraph of a given size can suppress the glare of the host star.
-%The smaller the \IWA, the more features we will be observable and in more systems. 
-The Rayleigh feature, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
+%
+The Rayleigh peak, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
 The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be observable in a significant fraction of systems.
-The position and shape of the rainbow peak is sensitive to the properties of the droplets, allowing their nature to be determined if the peak of the rainbow feature can be reached.
-The ocean glint feature occurs even further from quadrature, and so would be observable in fewer systems. 
+%
+The position and shape of the rainbow peak is sensitive to the properties of the cloud droplets, allowing their characterization if the peak of the rainbow feature can be reached.
+%
 In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability.
-%Finally, it is possible to view the glory scattering feature in one system, Alpha~Cen~A.
 
@@
     \caption{
-        The number of stars around which the peak scattering angle of each phenomenon is detectable, as a function of inner working angle (IWA).
+        The number of stars around which the peak phase angle of each phenomenon would be detectable, as a function of inner working angle (IWA).
 %expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
@@
         %\todo{Many of these values (e.g., glint) don't really match Fig. 5?}
+        {\color{red}{mention 164 total?}}
     }
@@
     \toprule
-    & \multicolumn{4}{c}{Inner Working Angle}\\
+    & \multicolumn{4}{c}{Inner Working Angle (IWA)}\\
     \cmidrule(lr){2-5}
@@
      Rainbow     &      109 &       91 &       47 &       23 \\
-     Rayleigh    &      154 &      152 &      128 &       82 \\
+     Rayleigh Peak   &      154 &      152 &      128 &       82 \\
     \bottomrule
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Contrast curves}
@@
 
-%The accessible phase angle is not the only factor that needs to be considered when determining what scattering features are observable.
 The contrast of the planet is also an important factor in assessing the detectability of these scattering phenomena. 
+%
 \Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming an Earth-like planet on an edge-on orbit.
+%
 The overall contrast varies significantly for each system and throughout the orbit. In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
+%
 The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
-%It is not possible to predict how much the state of the art of coronagraphy will have advanced by the time \HWO is built.
+%
 It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph, if these scattering features are to be observed.
@@
     \caption{
-        The number of systems for which key optical phenomena (Rayleigh scattering, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
+        The number of systems for which key optical phenomena (Rayleigh peak, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
+        %
         Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
+        %
         The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        The $y$-axis on the right indicates the number of systems where 
-        the phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
-        The glory peak would be observable in only one system: Alpha~Cen~A.
+        %
+        The $y$-axis on the right indicates the number of systems where each phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
 \label{sec:4}
-
-%What did we determine? 
-%Can we actually detect this stuff for anything in this sample? 
-%Do we need to not something about things like A stars which are not included in the sample - they were rejected by the list creators - do we need to comment on this?  - we decided nope.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
+%
 The instrument design and capabilities are still under development and should be driven by the mission's science goals.
-In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
+%
+In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized and unpolarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
+%
 We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
-%Coronagraphy is a rapidly advancing field, and it is hard to predict what the capabilities of \HWO will be.
+%
+While we do not consider the detect-ability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
+%
+A benefit of observing the polarized reflected light is that a coronograph is not explicitly required for quiescent stars as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
 The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets around different stellar spectral types, will drive the choice of coronagraphic \IWA for \HWO.
-Assuming a \qty{62}{\mas} \IWA, all \num{160} planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at scattering phases where angular features due to Rayleigh scattering show up, but the rainbow of light that is scattered in droplets of water clouds would only be accessible in $\sim\num{42}$ systems, and ocean glint in $\sim\num{14}$ systems.
-The habitability of Earth-like exoplanets in the remaining 146 systems would have to be established through less direct methods near quadrature: spectropolarimetry to deduce ocean reflection \citep{trees2022} or rotational mapping to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
-The number of habitable zones that can be probed by \HWO increases with decreasing \IWA, as does the range of scattering angles observable in these systems. 
-A larger-diameter telescope with a more aggressive coronagraph
-%Therefore, to increase the number of systems for which key features such as ocean glint can be observed, the primary mirror diameter could be increased and/or the \IWA of the coronagraph could be decreased.
-%While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
-%This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
-
-%\todo{the paragraph below seems similar to the one above ... merge them?}
-%We have shown that with a \qty{6}{\meter} primary mirror and a coronagraph with an \IWA\ of $3\,\lambda/D$ at \qty{600}{\nano\meter}, only $\sim\qty{10}{\percent}$ of the Earth-like exoplanets targeted by \HWO would be observable at a range of scattering angles that is wide enough to establish the presence of liquid surface water via the polarization signature of the glint (i.e., specular reflection).
 %
-%This number could be improved by increasing the primary mirror diameter or decreasing the \IWA of the coronagraph.
+Assuming a \qty{62}{\mas} \IWA, {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles where angular features due to Rayleigh scattering show up.
 %
-%While we briefly considered the contrast ratio of the scattering features, we did not investigate the feasibility of detecting them given their contrast. 
+The rainbow of light that is scattered in droplets of water clouds and ocean glint, both of which indicate the presence of a water cycle which provides additional context to the habitability of an exoplanet, would be accessible in {\color{red}{!!}} $\sim\qty{-1}{\percent}$ and {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of systems respectively.
 %
-%This will be an important task once the instrument design and capabilities are better defined and could drive the design of the coronagraph and the primary mirror diameter.
+The presence of a water cycle on Earth-like exoplanets in the remaining systems would have to be established through less direct methods: spectropolarimetry or rotational mapping.
+%
+To observe more scattering feature in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronograph and potentially a larger primary mirror.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowledgements}
-
-%The workshop on which this manuscript is based was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, Netherlands.
-%
-%This workshop was supported by NOVA  and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No. 866001 - EXACT).
-%
-%This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
-%
-%SRV acknowledges funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement No 805445.
-%
-%SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
-%
-%TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-%
-%Specific grant funding and any other specific requirements
-%
-%KMB acknowledges support from NASA Habitable
-%
-%Worlds grant No. 80NSSC20K152, and previous support for related work supported by NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement Number NNA13AA93A.
-%
-%This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
 
@@
 
+This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % References
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Author affiliations
+% DO NOT EDIT THIS MANUALLY
+% ALL ENTRIES ARE GENERATED BY A SCRIPT
+% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
 \section*{Author affiliations}
@@
     \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
-    \affdest{14}Aix Marseille Université, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
+    \affdest{14}Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
     \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
@@
     \affdest{26}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
-    \affdest{27}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
-    \affdest{28}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
-    \affdest{29}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
-    \affdest{30}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
-    \affdest{31}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
-    \affdest{32}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
-    \affdest{33}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
+    \affdest{27}Faculty of Aerospace Engineering, TU Delft, Building 62, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
+    \affdest{28}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
+    \affdest{29}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
+    \affdest{30}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
+    \affdest{31}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
+    \affdest{32}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
+    \affdest{33}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
+    \affdest{34}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
 \endgroup

### 2023-05-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in Fig.~\ref{fig:annotated-orbit}.
+However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in \cref{fig:annotated-orbit}.
 
@@
 
-As shown in Fig.~\ref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle (large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (zero at full phase and 180$^\circ$ at new phase).
+As shown in \cref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle (large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (\qty{0}{\degree} at full phase and \qty{180}{\degree} at new phase).
 %
-A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = 90^\circ$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles.
+A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = \qty{90}{\degree}$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles.
 %
@@
 %
-The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.\looseness=-1 \space In general, the glint on a planet will be strongest at red wavelengths because there Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}.
+The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
+In general, the glint on a planet will be strongest at red wavelengths because there, Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}. 
 %
@@
 %
-The presence of liquid water on the surface of a planet implies the presence of a water cycle which provides key context to its habitability and the interpretation of atmospheric biosignatures.
+The presence of liquid water on the surface of a planet implies the presence of a water cycle, which provides key context to its habitability and the interpretation of atmospheric biosignatures.
 %
@@
         %
-        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$ (scattering angle $\varphi = 180^\circ$).
+        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$ (scattering angle $\varphi = \qty{180}{\degree}$).
         %
@@
 %
-Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter ``\HWO'').
+Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter \enquote{\HWO}).
 %
@@
 %
-For generality, in this work, we report angular separations in milliarcseconds however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+For generality, in this work, we report angular separations in milliarcseconds; however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
 
@@
     \caption{
-        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milli-arcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
+        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
     }
@@
 %
-For non-solar stars, the semi-major axis scales as square root of the bolometric luminosity normalized to the Sun. 
+For non-solar stars, the semi-major axis scales as the square root of the bolometric luminosity normalized to the Sun. 
 %
@@
 
-We first define the phase angle ranges over which the various phenomena --- glint, Rayleigh peak, rainbow and glory --- manifest themselves in \cref{tab:phase_ranges}, also shown in Fig.~\ref{fig:bottplot}.
+We first define the phase angle ranges over which the various phenomena---glint, Rayleigh peak, rainbow, and glory---manifest themselves in \cref{tab:phase_ranges}, also shown in \cref{fig:bottplot}. 
 %
@@
     \caption{
-        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. Fig.~\ref{fig:bottplot}).
+        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. \cref{fig:bottplot}).
     }
@@
 A coronagraphic mask can obscure part of a planet's orbit. 
-As illustrated in \cref{fig:annotated-orbit} obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
+As illustrated in \cref{fig:annotated-orbit}, obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
 The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and eccentricity of the exoplanet's orbit.
@@
     \caption{
-        The angular separation of a planet from its star as seen by the observer, $\delta$, is related to the scattering angle, $\varphi$.
+        The angular separation of a planet from its star as seen by the observer, $\delta$, is related to the scattering angle $\varphi$. 
         The star is at a distance $d_*$ from the observer, the planet orbits a distance $a$ from the star, and  the projected distance between the star and planet is  $r_\mathrm{proj}$. 
@@
 Most planets are not in edge-on orbits as seen from Earth.
-For an orbit with an inclination angle $i$ ($i=0^\circ$ is face-on), we can identify the following two regimes:
+For an orbit with an inclination angle $i$ ($i=\qty{0}{\degree}$ is face-on), we can identify the following two regimes:
 \begin{enumerate}
@@
         This holds for orbits that are relatively face-on (small $i$). 
-    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$, following Eq.~\ref{eq:scattering_angle}.
+    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$, following \cref{eq:scattering_angle}. 
 \end{enumerate}
@@
         %
-        These angles are symmetric about quadrature (90$^\circ$).
+        These angles are symmetric about quadrature (\qty{90}{\degree}).
         %
@@
     \caption{
-        Examples of the eccentric orbits generated for the stellar sample. Each plot shows 10 examples of orbits randomly drawn for each star.
+        Examples of the eccentric orbits generated for the stellar sample. 
+        Each plot shows 10 examples of orbits randomly drawn for each star.
         The orbits are scaled to ensure an Earth-like incident flux. 
-        %
         The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
@@
         Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
-        %
         It is clear that the \IWA can significantly affect the range of observable phases angles for each orbit.
@@
 
-A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e., the planet/star flux ratio. 
+A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the planet/star flux ratio. 
 %
@@
 %
-We present some simplified examples in Fig.~\ref{fig:contrasts}. 
+We present some simplified examples in \cref{fig:contrasts}. 
 %
@@
 %
-The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet.
+The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_\mathrm{G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet.
 %
-We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and use Eq.
-\ref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$, i.e.
-$C(90) = 6.37\%$.
+We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
 
@@
 %
-We correct the Star List contrast by dividing by the $6.37\%$ reflected light fraction of the Lambertian phase function and multiply by the reflected light fraction of the \citet{treesandstam2019} model of $5.64\%$.
+We correct the Star List contrast by dividing by the $6.37\%$ reflected light fraction of the Lambertian phase function and multiplying by the reflected light fraction of the \citet{treesandstam2019} model of $5.64\%$.
 %
-We map the orbital phase for circular orbits to the phase angle and on-sky separation given an inclination $i$ of 90$^\circ$. 
+We map the orbital phase for circular orbits to the phase angle and on-sky separation, given an inclination $i$ of \qty{90}{\degree}. 
 %
@@
 %
-\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming an Earth-like planet on an edge-on orbit.
+\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list, assuming an Earth-like planet on an edge-on orbit.
 %
@@
 %
-It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph, if these scattering features are to be observed.
+It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph if these scattering features are to be observed.
 
@@
         %
-        The $y$-axis on the right indicates the number of systems where each phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
+        The $y$-axis on the right indicates the number of systems where all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
     }
@@
 
-The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
+The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
 %
@@
 %
-A benefit of observing the polarized reflected light is that a coronograph is not explicitly required for quiescent stars as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
+A benefit of observing the polarized reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
 
@@
 %
-To observe more scattering feature in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronograph and potentially a larger primary mirror.
+To observe more scattering features in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronagraph and potentially a larger primary mirror.
 
@@
 \footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
+%
 This work has made use of \textsf{numpy}
@@
 %
-The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).
-
+The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~866001 - EXACT).
+%
 This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
-
-SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement n°805445.
+%
+SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program under grant agreement \textnumero~805445.
+%
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
+%
 TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-KB acknowledges support from NASA Habitable Worlds grant n°80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement n°NNA13AA93A.
+%
+KB acknowledges support from NASA Habitable Worlds grant \textnumero 80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement \textnumero~NNA13AA93A.
+%
 IL acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
+%
 PB, IL, and YG were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
+%
 EHP is supported by the NASA Hubble Fellowship grant \#HST-HF2-51467.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
-LA and EC acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement n°101044152).
-OA and LK acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement n°819155)
+%
+LA and EC acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement \textnumero~101044152).
+%
+OA and LK acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~819155).
+%
 SYH was funded by the generous support of the Heising-Simons Foundation.
+%
 RJP is supported by NASA under award number 80GSFC21M0002.
-OHS acknowledges funding from the direction scientifique générale de l'ONERA (ARE Alioth).
-
-This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
+%
+OHS acknowledges funding from the Direction Scientifique Générale de l'ONERA (ARE Alioth).
+%
+%This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
 

### 2023-05-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \caption{
-        The phase curves of the normalized linearly polarized flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit; adapted from \citet{treesandstam2019}.
-        %
-        The wind speed over the ocean, which determines the wave height \citep[][]{CoxMunk1954}, is \qty{7}{\meter\per\second}.
-        %
-        The clouds consist of spherical water droplets.
-        %
-        The cloud coverage fraction is 0.5 and while the cloud pattern is patchy, the glint is always cloud-free. 
-        %
-        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$ (scattering angle $\varphi = \qty{180}{\degree}$).
-        %
-        Below the plot, we show the scattering angles of glories, rainbows, Rayleigh scattering, and ocean glint. Note that the features in the actual curves partly overlap with each other.
+        Phase curves of the normalised total and the normalised linearly polarised flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (adapted from \citealt{treesandstam2019}).
+        The wind speed over the ocean---which determines the wave height \citep{CoxMunk1954}---is \qty{7}{\meter\per\second}, and the clouds consist of spherical water droplets.
+        The cloud coverage fraction is 0.5, and while the cloud pattern is patchy, the glint is always cloud-free. 
+        The fluxes are normalised such that the total (unpolarised + polarised) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$.
+        In the bottom panel, we illustrate the scattering angle ranges of glories, rainbows, Rayleigh scattering, and ocean glint. 
+        Note that the features in the actual curves partly overlap with each other.
+        The glory illustration is taken from \citet{Zakovryashin_2020}; all other illustrations are from \href{https://unsplash.com}{unsplash.com} [\href{https://unsplash.com/photos/YWL4dqqiRdk}{1}, \href{https://unsplash.com/photos/BgFOMcg4HNs}{2}, \href{https://unsplash.com/photos/9YdePM6BrO4}{3}].
+        This figure is best viewed in colour.
+        \looseness=-1
     }
@@
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey (Astro2020) \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[Astro2020;][]{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
 %
@@
 %
-Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter \enquote{\HWO}).
+Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter \enquote{\HWO{}}).
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\HWO and its stellar sample}
+\subsection{Target list}
 \label{subsec:2.1}
@@
 Most planets are not in edge-on orbits as seen from Earth.
-For an orbit with an inclination angle $i$ ($i=\qty{0}{\degree}$ is face-on), we can identify the following two regimes:
+For an orbit with an inclination angle $i$ (where $i=\qty{0}{\degree}$ is face-on), we can identify the following two regimes:
 \begin{enumerate}
@@
     \begin{cases}
-        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA}\; d_* }{a} \\ 
-        2 \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA}\; d_* }{a}
+        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA} \cdot d_* }{a} \\ 
+        2 \cos^{-1}\left(\dfrac{\mathrm{IWA} \cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA} \cdot d_* }{a}
     \end{cases} \,
@@
         The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
-        %, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
-        %
         Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
@@
 We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
+\looseness=-1
 
@@
 %
-While not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
+While contrast is not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
 %
-We present some simplified examples in \cref{fig:contrasts}. 
+For the contrast calculations we start with the NASA Exoplanet Exploration Program’s Mission Star List for \HWO, which includes for every star an estimated contrast for an Earth-twin in the habitable zone.
 %
-These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
-
-The NASA Exoplanet Exploration Program’s Mission Star List for \HWO includes for every star an estimated contrast for an Earth-twin in the habitable zone.
-%
-The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
+The contrast is given by \citep[e.g.,][]{2023A&A...671A.165M}
 \begin{equation}
@@
 %
-The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_\mathrm{G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet.
+The contrast estimations for \HWO assumed circular orbits and a Lambertian phase function  \citep{Russell1916} with $A_\mathrm{G}=0.2$. 
 %
-We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
+For a more accurate representation of the contrast as function of separation we can update the phase function with the models shown in Fig. \Cref{fig:bottplot} from \citet{treesandstam2019}. 
+%
+To update the phase function we will first correct for the reflected flux at quadrature and then map the normalise total flux and polarized flux as function of separation.
 
-We use the phase functions at $\qty{670}{\nano\meter}$ from \citet{treesandstam2019} for an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds.
 %
-We correct the Star List contrast by dividing by the $6.37\%$ reflected light fraction of the Lambertian phase function and multiplying by the reflected light fraction of the \citet{treesandstam2019} model of $5.64\%$.
+We assume that the calculated contrasts in the Star List for \HWO correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
+%
+In addition, we calculate the reflected light fraction at quadrature with the updated model of \citet{treesandstam2019}, which assumes an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\qty{670}{\nano\meter}$.
+%
+We correct the star list contrast by dividing by the \qty{6.37}{\percent} reflected light fraction of the Lambertian phase function and multiplying by the reflected light fraction of the updated model.
 %
@@
 %
-\cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A
+\Cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A
+%
+These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
 
@@
 
-As shown in \cref{fig:betaallofit}, allowing for eccentric orbits increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
+As shown in \cref{fig:betaallofit}, a realistic eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
 %
@@
 %
-Assuming a \qty{62}{\mas} \IWA, {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles where angular features due to Rayleigh scattering show up.
+Assuming a \qty{62}{\mas} \IWA, {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarization from Rayleigh scattering.
 %
@@
 
-\todo{check the naming and references/footnotes for this list throughout the paper}
 The NASA ExEP target list containing the data used for these simulations is  available online.%
@@
 %
-This work has made use of \textsf{numpy}
- \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy},\footnote{\url{https://www.astropy.org}} a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
+This work has made use of \textsf{numpy} \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy}\footnote{\url{https://www.astropy.org}}, a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
+The colourblind-friendly colour schemes used in this paper were taken from \citet{Petroff_2021}.
 
@@
 OHS acknowledges funding from the Direction Scientifique Générale de l'ONERA (ARE Alioth).
-%
-%This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.
 

### 2023-05-22 src/scripts/create-figure-6-contrast-over-separation.py
Update plotting script for Figure 6

@@ -1,164 +1,281 @@
 """
-Create Figure 7: Contrast over separation for three example targets.
+Create Figure 6: Contrast over separation for three example targets.
 """
 
-from matplotlib.colorbar import Colorbar
-from mpl_toolkits.axes_grid1 import make_axes_locatable
+# -----------------------------------------------------------------------------
+# IMPORTS
+# -----------------------------------------------------------------------------
 
+import astropy.units as u
+import h5py
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
+
+from matplotlib.colorbar import Colorbar
 from scipy import interpolate
 
 from utils import paths
+from utils.constants import SCATTERING_ANGLES_OF_FEATURES, DIAMETER
 from utils.plotting import set_fontsize
 
 
-df = pd.read_csv(paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv")
+# -----------------------------------------------------------------------------
+# MAIN CODE
+# -----------------------------------------------------------------------------
 
-hab_zones = np.array(df["EEIDmas"].values[1:], dtype=float)  # in mas
-contrasts = np.array(df["Contrast"].values[1:], dtype=float)
-names = np.array(df["CommonID"].values[1:], dtype=str)
+if __name__ == "__main__":
 
-data_trees = np.loadtxt(
-    paths.data / "Reflected_light_curves_Trees_Stam_2019.txt"
-)
+    print("\nCREATE FIGURE 6: CONTRAST OVER SEPARATION\n")
 
+    # -------------------------------------------------------------------------
+    # Load data
+    # -------------------------------------------------------------------------
 
-P_670 = interpolate.interp1d(data_trees[0], data_trees[2])
-F_670 = interpolate.interp1d(data_trees[0], data_trees[1])
+    print("Loading data...", end=" ", flush=True)
 
-orbital_phase = np.linspace(0, 2 * np.pi, 200)
+    # Read in target list
+    file_path = paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv"
+    targets = pd.read_csv(file_path, header=[0, 1])
 
-inclination = np.radians(90)
+    # Select habitable zone values and target names
+    hz = targets["EEIDmas"].values.astype(float)  # in mas
[…]

### 2023-05-22 src/scripts/create-figure-6-yarnball.py
Fix names after swapping figures in the paper

@@ -231,7 +231,7 @@ if __name__ == "__main__":
     print("Saving plot...", end=" ", flush=True)
 
     fig.tight_layout(pad=0)
-    file_path = figures_dir / "figure-7-yarnball.pdf"
+    file_path = figures_dir / "figure-6-yarnball.pdf"
     plt.savefig(
         file_path,
         bbox_inches="tight",

### 2023-05-22 src/scripts/create-figure-7-contrast-over-separation.py

@@ -1,5 +1,5 @@
 """
-Create Figure 6: Contrast over separation for three example targets.
+Create Figure 7: Contrast over separation for three example targets.
 """
 
 # -----------------------------------------------------------------------------
@@ -270,7 +270,7 @@ if __name__ == "__main__":
     fig.tight_layout(pad=0, w_pad=0.4)
 
     print(f"Saving plot to PDF...", end=" ", flush=True)
-    file_name = "figure-6-contrast-over-separation.pdf"
+    file_name = "figure-7-contrast-over-separation.pdf"
     file_path = paths.figures / file_name
     fig.savefig(
         file_path,

### 2023-05-22 src/scripts/create-figure-7-contrast-over-separation.py
Harmonize file paths between scripts

@@ -16,8 +16,11 @@ import pandas as pd
 from matplotlib.colorbar import Colorbar
 from scipy import interpolate
 
-from utils import paths
 from utils.constants import SCATTERING_ANGLES_OF_FEATURES, DIAMETER
+from utils.paths import (
+    figures as figures_dir,
+    data as data_dir,
+)
 from utils.plotting import set_fontsize
 
 
@@ -36,7 +39,7 @@ if __name__ == "__main__":
     print("Loading data...", end=" ", flush=True)
 
     # Read in target list
-    file_path = paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv"
+    file_path = data_dir / "2646_NASA_ExEP_Target_List_HWO_Table.csv"
     targets = pd.read_csv(file_path, header=[0, 1])
 
     # Select habitable zone values and target names
@@ -46,7 +49,8 @@ if __name__ == "__main__":
     # Read in data from Trees and Stam (2019) and extract I, Q, P, and phases.
     # Note: We add 2*std to make the values consistent with Figure 2.
     # Note: The index [3] corresponds to the 670 nm band.
-    with h5py.File(paths.data / "v7_fc50.h5", "r") as hdf_file:
+    file_path = data_dir / "v7_fc50.h5"
+    with h5py.File(file_path, "r") as hdf_file:
         I_data = np.array(hdf_file["I"])[3] + 2 * np.array(hdf_file["Istd"])[3]
         Q_data = (
             np.abs(np.array(hdf_file["Q"]))[3]
@@ -271,7 +275,7 @@ if __name__ == "__main__":
 
     print(f"Saving plot to PDF...", end=" ", flush=True)
     file_name = "figure-7-contrast-over-separation.pdf"
-    file_path = paths.figures / file_name
+    file_path = figures_dir / file_name
     fig.savefig(
         file_path,
         dpi=600,

### 2023-05-22 src/scripts/utils/paths.py
Add `tables` path

@@ -26,5 +26,8 @@ tex = src / "tex"
 # Absolute path to the `src/tex/figures` folder (contains figure output)
 figures = tex / "figures"
 
+# Absolute path to the `src/tex/tables` folder (contains tables output)
+tables = tex / "tables"
+
 # Absolute path to the `src/tex/output` folder (contains other user-defined output)
 output = tex / "output"

### 2023-05-22 src/scripts/create-data-for-eccentric-orbits.py
Simplify script: Generate only one output file with all IWAs

@@ -1,5 +1,5 @@
 """
-Simulate 1,000 elliptical orbits for every planet in the target list.
+Simulate 1,000 eccentric orbits for every planet in the target list.
 """
 
 from warnings import catch_warnings, filterwarnings
@@ -30,7 +30,7 @@ def simulate_orbit(
     anode: float,
 ) -> tuple[np.ndarray, np.ndarray]:
     """
-    Simulate an elliptical orbit and return the minimum and maximum
+    Simulate an eccentric orbit and return the minimum and maximum
     scattering angles for each IWA.
 
     Args:
@@ -86,7 +86,7 @@ def simulate_orbit(
 
 if __name__ == "__main__":
 
-    print("\nSIMULATE ELLIPTICAL ORBITS (IN PARALLEL)\n")
+    print("\nSIMULATE ECCENTRIC ORBITS (IN PARALLEL)\n")
 
     # Fix the random seed to ensure reproducibility
     np.random.seed(RANDOM_SEED)
@@ -100,31 +100,11 @@ if __name__ == "__main__":
     print(f"Number of epochs to sample:   {n_epochs:,}\n")
 
     # Select values for the inner working angle (IWA) to simulate (in mas)
-    try:
-        op = snakemake.params["whichsim"]
-    except NameError:
-        print(80 * "-")
-        print(
-            "WARNING: This script is intended to be run via snakemake! \n"
-            "Since no value was provided for 'whichsim', the default 'iwa' "
-            "will be used."
-        )
-        print(80 * "-" + "\n")
-        op = "iwa"
-
-    if op == "iwa":
-        iwa = np.array([21, 41, 62, 83])
-        tag = ""
-    elif op == "iwa2":
-        iwa = np.linspace(20, 120, 5)
-        tag = "2"
-    elif op == "iwa3":
-        iwa = np.linspace(20, 120, 20)
-        tag = "3"
-    else:
-        raise ValueError(f"Invalid value for 'whichsim': {op}")
-    print(f'IWA option: "{op}"')
-    print(f"IWA values: {iwa} mas\n")
+    # We simulate IWAs of 20 to 120 mas in steps of 1 mas (101 values) so that
+    # we can use them both for Figure 5 and Figure 8.
+    iwa = np.arange(20, 121)
+    with np.printoptions(threshold=1):
+        print(f"IWA values (in mas): {iwa}\n")
 
     # Read in the stellar target list
[…]

### 2023-05-22 src/scripts/create-figure-5-n-over-phase.py
Complete refactoring

@@ -1,6 +1,6 @@
 """
-Create Figure 5: CDF of number of planets over accessible phase angle
-for different inner working angles (both circular and elliptical).
+Create Figure 5: Expected number of accessible planets over phase angle
+for different inner working angles (both circular and eccentric).
 """
 
 # -----------------------------------------------------------------------------
@@ -10,13 +10,15 @@ for different inner working angles (both circular and elliptical).
 from warnings import catch_warnings, filterwarnings
 
 from matplotlib.lines import Line2D
+from scipy.interpolate import interp1d
+from tabulate import tabulate
 
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 
 from utils import paths
-from utils.constants import ETA_EARTH, RANDOM_SEED
+from utils.constants import ETA_EARTH, RANDOM_SEED, PHASE_ANGLES_OF_FEATURES
 from utils.plotting import set_fontsize, CBF_COLORS
 from utils.samplers import sample_i
 
@@ -41,20 +43,20 @@ if __name__ == "__main__":
 
     # Read in the maximum and minimum scattering angles as calculated by the
     # dynamical simulations. This set of files only has 4 IWAs.
-    iwa = np.load(paths.data / "iwa_all.npz")
-    phase_max = iwa["betamax"]
-    phase_min = iwa["betamin"]
+    iwa = np.load(paths.data / "eccentric-orbits.npz")
+    phase_max = iwa["betamax"]  # in degrees
+    phase_min = iwa["betamin"]  # in degrees
     iwa_list = iwa["iwa"]  # in mas
 
     # Note: We only use phase_max for plotting.
-    # For a given system, phase_max and phase_min might differ (for elliptical
+    # For a given system, phase_max and phase_min might differ (for eccentric
     # orbits), but the distributions mirror each other (about 90 degrees) when
     # you average over enough orbits.
 
     print("Done!", flush=True)
 
     # -------------------------------------------------------------------------
-    # Run simulations for different inclinations
+    # Run simulations for different inclinations (circular orbits)
     # -------------------------------------------------------------------------
 
     print("Simulating different inclinations...", end=" ", flush=True)
@@ -71,7 +73,7 @@ if __name__ == "__main__":
     beta = np.zeros(shape=(n_sims, len(hz), len(iwa_list)))
 
     # Loop over all targets to run the simulations
-    for i in range(len(hz)):
+    for i, target_hz in enumerate(hz):
 
[…]

### 2023-05-22 src/scripts/create-figure-8-n-over-iwa.py

@@ -1,27 +1,30 @@
 """
-Create Figure 9: Number of planets accessible as a function of IWA,
-for different scattering features.
+Create Figure 8: Number of planets accessible as a function of IWA for
+different scattering features. This script also creates Table 3, which
+contains the number of systems accessible as a function of IWA.
 """
 
 # -----------------------------------------------------------------------------
 # IMPORTS
 # -----------------------------------------------------------------------------
 
-from warnings import catch_warnings, filterwarnings
-
 from matplotlib.lines import Line2D
-from scipy import interpolate as si
+from tabulate import tabulate
 
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 
-from utils import paths
+from utils.paths import (
+    data as data_dir,
+    figures as figures_dir,
+    tables as tables_dir,
+)
 from utils.plotting import set_fontsize, CBF_COLORS
 from utils.constants import (
     ETA_EARTH,
     LAMBDA_OVER_D_IN_MAS,
-    # PHASE_ANGLES_OF_FEATURES as SCATTERING_FEATURES,  # breaks Rayleigh
+    PHASE_ANGLES_OF_FEATURES as SCATTERING_FEATURES
 )
 
 
@@ -29,168 +32,216 @@ from utils.constants import (
 # MAIN CODE
 # -----------------------------------------------------------------------------
 
-# TODO: This should be imported from utils.constants, but it seems to break
-#   the plot for Rayleigh scattering. I'm not sure why.
-SCATTERING_FEATURES = {
-    "Rainbow": (127, 138, 158),
-    "Rayleigh": (90, 110, 130),
-    "Ocean Glint": (50, 30, 10),
-    "Glory": (10, 5, 0),
-}
+if __name__ == "__main__":
 
-# Read in the target list
-df = pd.read_csv(paths.data / "2646_NASA_ExEP_Target_List_HWO_Table.csv")
+    print("\nCREATE FIGURE 8: N_SYSTEMS OVER IWA\n")
 
-# The angular separation of the Earth Equivalent Instellation (EEI) in mas
-hab_zones = np.array(df["EEIDmas"].values[1:], dtype=float)
+    # -------------------------------------------------------------------------
+    # Read in the data
+    # -------------------------------------------------------------------------
 
-# Compute the minimum and maximum scattering angles (pre-computed)
-iwa = np.load(paths.data / "iwa_all3.npz")
[…]

### 2023-05-24 src/scripts/create-figure-5-n-over-phase.py
Improve output to terminal

@@ -9,6 +9,8 @@ for different inner working angles (both circular and eccentric).
 
 from warnings import catch_warnings, filterwarnings
 
+import time
+
 from matplotlib.lines import Line2D
 from scipy.interpolate import interp1d
 from tabulate import tabulate
@@ -29,7 +31,15 @@ from utils.samplers import sample_i
 
 if __name__ == "__main__":
 
+    # -------------------------------------------------------------------------
+    # Preliminaries
+    # -------------------------------------------------------------------------
+
+    print("\n" + 80 * "-")
     print("\nCREATE FIGURE 5: N_SYSTEMS OVER PHASE ANGLE\n")
+
+    # Start timer
+    script_start = time.time()
 
     # -------------------------------------------------------------------------
     # Read in the data
@@ -323,3 +333,10 @@ if __name__ == "__main__":
     )
 
     print("Done!", flush=True)
+
+    # -------------------------------------------------------------------------
+    # Postliminaries
+    # -------------------------------------------------------------------------
+
+    print(f"\nThis took {time.time() - script_start:.1f} seconds!")
+    print("\n" + 80 * "-" + "\n")

### 2023-05-24 src/scripts/create-figure-6-yarnball.py

@@ -134,11 +134,13 @@ def prepare_ax(
 # -----------------------------------------------------------------------------
 
 if __name__ == "__main__":
+
     # -------------------------------------------------------------------------
     # Preliminaries
     # -------------------------------------------------------------------------
 
-    print("\nCREATE YARN PLOT\n")
+    print("\n" + 80 * "-")
+    print("\nCREATE FIGURE 6: YARN BALL PLOT\n")
 
     # Start timer
     script_start = time.time()
@@ -249,4 +251,5 @@ if __name__ == "__main__":
     # Postliminaries
     # -------------------------------------------------------------------------
 
-    print(f"\nThis took {time.time() - script_start:.1f} seconds!\n")
+    print(f"\nThis took {time.time() - script_start:.1f} seconds!")
+    print("\n" + 80 * "-" + "\n")

### 2023-05-24 src/scripts/create-figure-8-n-over-iwa.py

@@ -8,6 +8,8 @@ contains the number of systems accessible as a function of IWA.
 # IMPORTS
 # -----------------------------------------------------------------------------
 
+import time
+
 from matplotlib.lines import Line2D
 from tabulate import tabulate
 
@@ -34,7 +36,15 @@ from utils.constants import (
 
 if __name__ == "__main__":
 
+    # -------------------------------------------------------------------------
+    # Preliminaries
+    # -------------------------------------------------------------------------
+
+    print("\n" + 80 * "-")
     print("\nCREATE FIGURE 8: N_SYSTEMS OVER IWA\n")
+
+    # Start timer
+    script_start = time.time()
 
     # -------------------------------------------------------------------------
     # Read in the data
@@ -245,3 +255,10 @@ if __name__ == "__main__":
     )
 
     print("Done!", flush=True)
+
+    # -------------------------------------------------------------------------
+    # Postliminaries
+    # -------------------------------------------------------------------------
+
+    print(f"\nThis took {time.time() - script_start:.1f} seconds!")
+    print("\n" + 80 * "-" + "\n")

### 2023-05-24 src/scripts/utils/constants.py
Add `lambda_over_d_in_mas()`

@@ -5,6 +5,9 @@ Define constants used in the scripts.
 import astropy.units as u
 
 
+# Random seed
+RANDOM_SEED = 42
+
 # Assumed diameter of HWO
 DIAMETER = u.Quantity(6, u.m)
 
@@ -28,42 +31,42 @@ SCATTERING_ANGLES_OF_FEATURES = {
     for key, (start, peak, end) in PHASE_ANGLES_OF_FEATURES.items()
 }
 
-# Assumed wavelength for the observations
+# Assumed default wavelength for most plots / calculations
 WAVELENGTH = u.Quantity(600, u.nm)
 
-# Compute lambda / D in mas
-with u.set_enabled_equivalencies(u.dimensionless_angles()):
-    LAMBDA_OVER_D_IN_MAS = (WAVELENGTH / DIAMETER).to(u.mas).value
 
-# Random seed
-RANDOM_SEED = 42
+# Compute lambda / D in mas
+# Make `wavelength` an optional argument here so that we can use this function
+# also to compute Table 2.
+def lambda_over_d_in_mas(
+    wavelength: u.Quantity = WAVELENGTH,
+) -> u.Quantity:
+    with u.set_enabled_equivalencies(u.dimensionless_angles()):
+        return (wavelength / DIAMETER).to(u.mas)
 
 
 if __name__ == "__main__":
 
     print("\nCONSTANTS USED IN THE SCRIPTS:\n")
 
-    print('Occurrence rate of rocky planets in the optimistic habitable zone:')
-    print(f'  ETA_EARTH = {ETA_EARTH:.2f}\n')
+    print("Occurrence rate of rocky planets in the optimistic habitable zone:")
+    print(f"  ETA_EARTH = {ETA_EARTH:.2f}\n")
 
-    print('Assumed wavelength:')
-    print(f'  WAVELENGTH = {WAVELENGTH.to(u.nm):.2f}\n')
+    print("Assumed wavelength:")
+    print(f"  WAVELENGTH = {WAVELENGTH.to(u.nm):.2f}\n")
 
-    print('Assumed telescope diameter:')
-    print(f'  DIAMETER = {DIAMETER.to(u.m):.2f}\n')
+    print("Assumed telescope diameter:")
+    print(f"  DIAMETER = {DIAMETER.to(u.m):.2f}\n")
 
-    print('lambda / D:')
[…]

### 2023-05-24 src/scripts/create-figure-5-n-over-phase.py
Homogenize IWA values (1, 2, 3, 4 lambda / D in mas) across scripts

@@ -20,7 +20,12 @@ import pandas as pd
 import matplotlib.pyplot as plt
 
 from utils import paths
-from utils.constants import ETA_EARTH, RANDOM_SEED, PHASE_ANGLES_OF_FEATURES
+from utils.constants import (
+    lambda_over_d_in_mas,
+    ETA_EARTH,
+    RANDOM_SEED,
+    PHASE_ANGLES_OF_FEATURES,
+)
 from utils.plotting import set_fontsize, CBF_COLORS
 from utils.samplers import sample_i
 
@@ -40,6 +45,14 @@ if __name__ == "__main__":
 
     # Start timer
     script_start = time.time()
+
+    # Get IWA values (in mas) for 1, 2, 3, 4 lambda / D at 600 nm and round
+    # to the nearest integer (because `create-data-for-eccentric-orbits.py`
+    # only computes the beta values for integer values of the IWA in mas).
+    # This should be [21, 41, 62, 83] mas.
+    n_lambda_over_d_in_mas = np.around(
+        lambda_over_d_in_mas().value * np.array([1, 2, 3, 4]), 0
+    )
 
     # -------------------------------------------------------------------------
     # Read in the data
@@ -187,7 +200,7 @@ if __name__ == "__main__":
     phase_angles = np.linspace(90, 180, 360)
 
     # Loop over IWA to plot a line for each one
-    for i, iwa in enumerate([21, 41, 62, 83]):
+    for i, iwa in enumerate(n_lambda_over_d_in_mas):
 
         # Get the index that belongs to the current IWA (since we are pulling
         # the data from a file that has all IWA values: 20, 21, ..., 120)
@@ -309,7 +322,7 @@ if __name__ == "__main__":
                 .unstack(level=0),
                 headers=(
                     ["Feature"]
-                    + [f"{iwa:.0f} mas" for iwa in (21, 41, 62, 83)]
+                    + [f"{iwa:.0f} mas" for iwa in n_lambda_over_d_in_mas]
                 ),
                 tablefmt="simple",  # latex_booktabs
             )

### 2023-05-24 src/scripts/create-figure-6-yarnball.py

@@ -16,7 +16,7 @@ import matplotlib as mpl
 import matplotlib.pyplot as plt
 import numpy as np
 
-from utils.constants import RANDOM_SEED
+from utils.constants import RANDOM_SEED, lambda_over_d_in_mas
 from utils.phases import kep3d
 from utils.samplers import sample_e, sample_i
 from utils.plotting import CBF_COLORS
@@ -27,12 +27,27 @@ from utils.paths import data as data_dir, figures as figures_dir
 # DEFINITIONS
 # -----------------------------------------------------------------------------
 
-def draw_iwa(ax: plt.Axes, iwas: np.ndarray) -> None:
+def prepare_ax(
+    ax: plt.Axes,
+    star_name: str,
+    distance: float,
+    hz: float,
+) -> None:
     """
-    Auxiliary function to draw inner working angles on a given axis.
+    Auxiliary function to prepare an `ax` for plotting.
     """
 
-    for iwa in iwas:
+    # Remove ticks and setup aspect ratio
+    ax.set_xticks([])
+    ax.set_yticks([])
+    ax.set_aspect("equal")
+
+    # Set up limits and draw inner working angles
+    ax.set_xlim(-2 * hz, 2 * hz)
+    ax.set_ylim(-2 * hz, 2 * hz)
+
+    # Draw inner working angles (1 to 4 lambda / D at 600 nm)
+    for iwa in np.arange(1, 5) * lambda_over_d_in_mas().value:
         for i, (ec, ls, alpha) in enumerate([
             ("white", "-", 0.5),
             ("black", "--", 1.0),
@@ -49,46 +64,6 @@ def draw_iwa(ax: plt.Axes, iwas: np.ndarray) -> None:
                     zorder=100 + i,
                 )
             )
-
-
-def convert_lod_to_mas(
-    lod: float,
-    wavelength: u.Quantity = 600 * u.nm,
-    diameter: u.Quantity = 6 * u.m,
-) -> float:
-    """
-    Convert a angular separation from units of lambda / D to mas.
-    """
-
-    return (
-        (lod * wavelength / diameter)
-        .to(
-            "mas",
-            equivalencies=u.dimensionless_angles(),
-        )
-        .value
-    )
-
-
-def prepare_ax(
-    ax: plt.Axes,
-    star_name: str,
-    distance: float,
-    hz: float,
-) -> None:
-    """
-    Auxiliary function to prepare an axis for plotting.
-    """
-
[…]

### 2023-05-24 src/scripts/create-figure-7-contrast-over-separation.py

@@ -18,7 +18,7 @@ import pandas as pd
 from matplotlib.colorbar import Colorbar
 from scipy import interpolate
 
-from utils.constants import SCATTERING_ANGLES_OF_FEATURES, DIAMETER
+from utils.constants import lambda_over_d_in_mas, SCATTERING_ANGLES_OF_FEATURES
 from utils.paths import (
     figures as figures_dir,
     data as data_dir,
@@ -41,6 +41,9 @@ if __name__ == "__main__":
 
     # Start timer
     script_start = time.time()
+
+    # Define wavelength for this figure (670 nm)
+    wavelength = 670 * u.nm
 
     # -------------------------------------------------------------------------
     # Load data
@@ -126,11 +129,6 @@ if __name__ == "__main__":
     inclination = np.deg2rad(90)
     alpha = np.rad2deg(np.arccos(np.cos(orbital_phase) * np.sin(inclination)))
 
-    # Compute the IWA at 670 nm
-    WAVELENGTH = 670 * u.nm
-    with u.set_enabled_equivalencies(u.dimensionless_angles()):
-        IWA = (WAVELENGTH / DIAMETER).to(u.mas)
-
     # -------------------------------------------------------------------------
     # Prepare the plot
     # -------------------------------------------------------------------------
@@ -170,8 +168,8 @@ if __name__ == "__main__":
         ax.set_ylim(1e-12, 1e-7)
         ax.get_yaxis().set_visible(i == 0)
 
-        # Plot vertical lines at 1, 2, and 3 IWA
-        for x in np.arange(1, 4) * IWA:
+        # Plot vertical lines at 1, 2, and 3 IWA (in mas)
+        for x in np.arange(1, 4) * lambda_over_d_in_mas(wavelength):
             ax.axvline(x=x.value, ls="--", c="k", alpha=0.5, lw=0.5)
 
         # Get index of current target in target list
@@ -215,7 +213,9 @@ if __name__ == "__main__":
     # Add labels for the IWA lines
     # -------------------------------------------------------------------------
 
-    for x in np.arange(1, 4) * IWA:
+    # Reminder: We get lines at 23, 46, and 69 mas here, because our
+    # wavelength is 670 nm (not 600 nm as in most of the other plots).
+    for x in np.arange(1, 4) * lambda_over_d_in_mas(wavelength):
         axes[0].text(
             x=x.value,
             y=2e-12,
@@ -224,7 +224,7 @@ if __name__ == "__main__":
             alpha=0.5,
[…]

### 2023-05-24 src/scripts/create-figure-8-n-over-iwa.py

@@ -24,8 +24,8 @@ from utils.paths import (
 )
 from utils.plotting import set_fontsize, CBF_COLORS
 from utils.constants import (
+    lambda_over_d_in_mas,
     ETA_EARTH,
-    LAMBDA_OVER_D_IN_MAS,
     PHASE_ANGLES_OF_FEATURES as SCATTERING_FEATURES
 )
 
@@ -45,6 +45,14 @@ if __name__ == "__main__":
 
     # Start timer
     script_start = time.time()
+
+    # Get IWA values (in mas) for 1, 2, 3, 4 lambda / D at 600 nm and round
+    # to the nearest integer (because `create-data-for-eccentric-orbits.py`
+    # only computes the beta values for integer values of the IWA in mas).
+    # This should be [21, 41, 62, 83] mas.
+    n_lambda_over_d_in_mas = np.around(
+        lambda_over_d_in_mas().value * np.array([1, 2, 3, 4]), 0
+    )
 
     # -------------------------------------------------------------------------
     # Read in the data
@@ -114,7 +122,7 @@ if __name__ == "__main__":
             ax.plot(iwa_list, y, ls=ls, lw=1, color=CBF_COLORS[k])
 
             if i == 1:
-                for iwa in [21, 41, 62, 83]:
+                for iwa in n_lambda_over_d_in_mas:
                     idx = np.where(iwa_list == iwa)[0][0]
                     expected_numbers.append(
                         {
@@ -133,7 +141,7 @@ if __name__ == "__main__":
         ax.set_xlabel(r"IWA (in mas)")
 
         # Set the options for the top x-axis
-        top_ax.set_xticks(np.arange(1, 7) * LAMBDA_OVER_D_IN_MAS)
+        top_ax.set_xticks(np.arange(1, 7) * lambda_over_d_in_mas().value)
         top_ax.set_xticklabels(np.arange(1, 7))
         top_ax.set_xlim(ax.get_xlim())
         top_ax.set_xlabel(r"IWA (in $\lambda / D$)")
@@ -202,7 +210,9 @@ if __name__ == "__main__":
     )
 
     # Define the column names
-    headers = ["Feature"] + [f"{iwa:.0f} mas" for iwa in (21, 41, 62, 83)]
+    headers = (
+        ["Feature"] + [f"{iwa:.0f} mas" for iwa in n_lambda_over_d_in_mas]
+    )
 
     # Print the table to the terminal
     table = tabulate(df, headers=headers, tablefmt="simple")

### 2023-05-24 src/scripts/create-data-for-eccentric-orbits.py
Update script to print total runtime upon completion

@@ -4,6 +4,8 @@ Simulate 1,000 eccentric orbits for every planet in the target list.
 
 from warnings import catch_warnings, filterwarnings
 
+import time
+
 from astropy.io import ascii
 from joblib import Parallel, delayed
 from tqdm import tqdm
@@ -86,10 +88,22 @@ def simulate_orbit(
 
 if __name__ == "__main__":
 
+    # -------------------------------------------------------------------------
+    # Preliminaries
+    # -------------------------------------------------------------------------
+
+    print("\n" + 80 * "-")
     print("\nSIMULATE ECCENTRIC ORBITS (IN PARALLEL)\n")
+
+    # Start timer
+    script_start = time.time()
 
     # Fix the random seed to ensure reproducibility
     np.random.seed(RANDOM_SEED)
+
+    # -------------------------------------------------------------------------
+    # Define parameters; set up output arrays
+    # -------------------------------------------------------------------------
 
     # Define number of planet orbits to simulate for each star and number of
     # epochs to sample along each orbit ("resolution of the orbit")
@@ -116,6 +130,10 @@ if __name__ == "__main__":
     # Set up output arrays
     betamin = np.zeros((n_orbits, n_stars, iwa.size))
     betamax = np.zeros((n_orbits, n_stars, iwa.size))
+
+    # -------------------------------------------------------------------------
+    # Run simulations in parallel
+    # -------------------------------------------------------------------------
 
     # Loop over all the stars in the target list
     print("Simulating orbits for stars in target list:")
@@ -146,6 +164,10 @@ if __name__ == "__main__":
             betamin[idx_orbit, idx_star, :] = scatmin
             betamax[idx_orbit, idx_star, :] = scatmax
 
+    # -------------------------------------------------------------------------
+    # Save results to *.npz file
+    # -------------------------------------------------------------------------
+
     file_name = "eccentric-orbits.npz"
     print(f"\nSaving data to file {file_name}...", end=" ")
     np.savez_compressed(
@@ -155,3 +177,10 @@ if __name__ == "__main__":
         betamin=betamin,
     )
     print("Done!\n")
+
+    # -------------------------------------------------------------------------
+    # Postliminaries
+    # -------------------------------------------------------------------------
+
+    print(f"\nThis took {time.time() - script_start:.1f} seconds!")
+    print("\n" + 80 * "-" + "\n")

### 2023-05-25 src/tex/ms.tex

@@
 %
-Direct imaging missions are optimised for planets near quadrature, and \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
+Direct imaging missions are optimised for planets near quadrature, but \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
 %
@@
 %
-We use \HWO'~catalogue of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
+We use \HWO'~catalogue of 164 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
 %
-We find that any exo-Earths in planetary systems in this catalogue will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
+We find that most exo-Earths planetary systems in this catalogue will be observable at phase angles at which the Rayleigh scattering peak occurs in their polarimetric phase curve. 
 %
-The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible to \HWO in $\sim\num{42}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be observable with \HWO in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be observable in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
-Improving the \IWA to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of roughly 2 and 3, respectively.
+Improving the \IWA to $2\,\lambda/D$ increases the number of systems with observable rainbows and glints by factors of approximately 2 and 3, respectively.
 %
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.  %and scales approximately inversely with the \IWA.  
+The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.
+%
+By observing these scattering features, \HWO could detect a surface ocean and water cycle, key indicators of habitability.
+\looseness=-1
 \end{abstract}
@@
 %
-Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection, i.e.
-glint, from water oceans \citep[][]{2008Icar..195..927W}.
+Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[][]{2008Icar..195..927W}.
 %
 However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in \cref{fig:annotated-orbit}.
+%
+In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
+\looseness=-1
 
@@
     \caption{
-        Phase curves of the normalised total and the normalised linearly polarised flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (adapted from \citealt{treesandstam2019}).
+        Phase curves of the normalised total and normalised linearly polarised flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (adapted from \citealt{treesandstam2019}).
         The wind speed over the ocean---which determines the wave height \citep{CoxMunk1954}---is \qty{7}{\meter\per\second}, and the clouds consist of spherical water droplets.
         The cloud coverage fraction is 0.5, and while the cloud pattern is patchy, the glint is always cloud-free. 
-        The fluxes are normalised such that the total (unpolarised + polarised) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$.
+        The fluxes are normalised such that the total (i.e., unpolarised + polarised) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$.
         In the bottom panel, we illustrate the scattering angle ranges of glories, rainbows, Rayleigh scattering, and ocean glint. 
-        Note that the features in the actual curves partly overlap with each other.
+        Note that the features in the actual curves partly overlap.
         The glory illustration is taken from \citet{Zakovryashin_2020}; all other illustrations are from \href{https://unsplash.com}{unsplash.com} [\href{https://unsplash.com/photos/YWL4dqqiRdk}{1}, \href{https://unsplash.com/photos/BgFOMcg4HNs}{2}, \href{https://unsplash.com/photos/9YdePM6BrO4}{3}].
         This figure is best viewed in colour.
-        \looseness=-1
     }
@@
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[Astro2020;][]{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[Astro2020;][]{decadal} recommended a telescope with the capability to detect signatures of habitability on approximately 25 HZ planets in a new \enquote{Great Observatories} program.
 %
@@
 %
-Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter \enquote{\HWO{}}).
+Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory.
 %
-The performance and characteristics (on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
+The performance and characteristics (e.g.
+on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
 %
@@
     \caption{
-        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
+        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter}, \qty{670}{\nano\meter}, and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of \qty{6}{\meter}.
     }
     \label{tab:IWA_OWA}
-    \begin{tabular}{ c c c c } 
-    \toprule
-    & $\lambda/D$ & \unit{\mas} (at \qty{600}{\nano\meter}) & \unit{\mas} (at \qty{1}{\micro\meter}) \\
-    \midrule
-    \midrule
-    IWA & 1 & 20.63 & 34.38 \\
-    IWA & 2 & 41.25 & 68.75 \\
-    IWA & 3 & 61.88 & 103.13 \\
-    IWA & 4 & 82.51 & 137.51 \\
-    \midrule
-    OWA & 32 &  660.05 & 1100.08 \\
-    OWA & 64 & 1320.09 & 2200.16 \\
-    \bottomrule
-    \end{tabular}
+    \script{create-table-1-iwa-values.py}
+    \input{tables/table-1-iwa-owa}
 \end{table}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Outline}
+% \subsection{Outline}
 
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
+% In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
 %
-In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
+% In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
 %
-We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
+% We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
 
@@
 %
-The target list is constructed assuming a maximum planet magnitude in the Cousins Rc band of $R_c = 31$ and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
+The target list is constructed assuming a maximum planet magnitude in the Cousins $R_c$ band of $31$ mag and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
 % 
@@
 %
-By design, such planets should be observable by \HWO at quadrature phase.
+By design, all such planets should be observable by \HWO at quadrature phase. 
+\looseness=-1
 
@@
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle (around $\alpha=\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle ($\alpha\sim\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
 
@@
         %
-        The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
+        %The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
         %
@@
         %
-        The colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
+        The colour bar indicates the optical phenomena that can in principle be detected: all the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
         %
@@
         The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
-        {\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
+        %{\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
     }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 7: Yarn ball
+% Figure 6: Yarn ball
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \centering
-    \includegraphics{figures/figure-7-yarnball.pdf}  
-    \script{create-figure-7-yarnball.py}
+    \includegraphics{figures/figure-6-yarnball.pdf}  
+    \script{create-figure-6-yarnball.py}
     \caption{
@@
 
-We then calculate the minimum and maximum accessible phase angles for \IWA{}s of $\qtylist{21; 42; 63; 84}{\mas}$.
+We then calculate the minimum and maximum accessible phase angles for \IWA{}s of $\qtylist{21; 41; 62; 83}{\mas}$.
 %
@@
 %
-For the contrast calculations we start with the NASA Exoplanet Exploration Program’s Mission Star List for \HWO, which includes for every star an estimated contrast for an Earth-twin in the habitable zone.
+For the contrast calculations, we start with the NASA Exoplanet Exploration Program’s Mission Star List for \HWO, which includes for every star an estimated contrast for an Earth-twin in the habitable zone.
 %
@@
 %
-For a more accurate representation of the contrast as function of separation we can update the phase function with the models shown in Fig. \Cref{fig:bottplot} from \citet{treesandstam2019}. 
+For a more accurate representation of the contrast as a function of separation, we can update the phase function with the models shown in Fig. \Cref{fig:bottplot} from \citet{treesandstam2019}. 
 %
-To update the phase function we will first correct for the reflected flux at quadrature and then map the normalise total flux and polarized flux as function of separation.
+To update the phase function, we will first correct for the reflected flux at quadrature and then map the normalised total flux and polarized flux as a function of separation.
 
@@
 %
-\Cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A
+\Cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 6
+% Figure 7
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/figure-6-contrast-over-separation_i=90.pdf}
-    \script{create-figure-6-contrast-over-separation.py}
+    \includegraphics{figures/figure-7-contrast-over-separation.pdf}
+    \script{create-figure-7-contrast-over-separation.py}
     \caption{
-        The planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit with an inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A (G2V), $\epsilon$~Eri (K2V), and Lalande~21185 (M2V).
+        Planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit  around Lalande~21185 (M2V), $\epsilon$~Eri (K2V), and $\alpha$~Cen~A (G2V) when assuming an inclination of $i = \qty{90}{\degree}$.
         %
-        The solid, black lines indicate the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
+        The solid black lines indicate the contrast in unpolarised light, with the contrast at quadrature marked by a black dot.
         %
-        The polarized contrast (at $\lambda = \qty{670}{\nano\meter}$) is indicated by the coloured dots, with the colour representing the degrees from quadrature.
+        The contrast for linearly polarised light at $\lambda = \qty{670}{\nano\meter}$ is indicated by the coloured lines, with the colour encoding the scattering angle.
         %
-        The light grey dots indicate the contrasts at quadrature of the other targets in the star list, and the dashed lines \IWA's of $\qtylist{21; 42; 63}{\mas}$.
+        The small grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate \IWA's of $\qtylist{23; 46; 69}{\mas}$.
     }
@@
     \caption{
-        The number of stars around which the peak phase angle of each phenomenon would be detectable, as a function of inner working angle (IWA).
-%expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        %\todo{Do these numbers refer to circular or elliptical orbits?}
-        %\todo{Many of these values (e.g., glint) don't really match Fig. 5?}
-        {\color{red}{mention 164 total?}}
+        The expected number of targets for which the peak phase angle of each phenomenon would be detectable when assuming randomly inclined, eccentric orbits (see \cref{sec:eccentric}).
+        For reference, the \HWO target list consists of 164 targets in total.
     }
     \label{tab:nstars_detect}
-    % \begin{tabular}{ l c c c } 
-    %     \toprule
-    %     & \multicolumn{3}{c}{Inner Working Angle}\\
-    %     & \qty{21}{\mas} & \qty{41}{\mas} & \qty{62}{\mas} \\
-    %     & $2\,\lambda/D$ & $3\,\lambda/D$ & $4\,\lambda/D$\\
-    %     \midrule
-    %     \midrule
-    %     Rainbows    & 90  & 42    & 18 \\
-    %     Rayleigh    & 152 & 115   & 65 \\
-    %     Ocean Glint & 43  & 14    & 5 \\
-    %     % Glory       & 0   & 5     & 10 \\
-    %     \bottomrule
-    % \end{tabular}\\[5mm]
-    %\todo{Just for comparison, here are some values from Fig. 5 (elliptical orbits) that I pulled straight out of the plotting script:}\\[5mm]
-    \begin{tabular}{lcccc}
-    \toprule
-    & \multicolumn{4}{c}{Inner Working Angle (IWA)}\\
-    \cmidrule(lr){2-5}
-     Feature     &   21 mas &   41 mas &   62 mas &   83 mas \\
-    \midrule
-     Glory       &        0 &        0 &        0 &        0 \\
-     Ocean Glint &       79 &       43 &       16 &        6 \\
-     Rainbow     &      109 &       91 &       47 &       23 \\
-     Rayleigh Peak   &      154 &      152 &      128 &       82 \\
-    \bottomrule
-    \end{tabular}\\[5mm]
-    %\todo{There is also a mismatch with Fig. 8, though. Where does this discrepancy come frome? Something about phase angle vs. scattering angle?}\\
-    %\todo{Observation: The values in the original table seem like a good match for Fig.~8 if we assume the colums where "shifted by one", i.e., 62 mas is actually 83 mas, 41 mas is 62 mas, ...}
+    \input{tables/table-3-expected-number}
+    \script{create-figure-8-n-over-iwa.py}
 \end{table}
@@
 %
-Assuming a \qty{62}{\mas} \IWA, {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarization from Rayleigh scattering.
+Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarization from Rayleigh scattering.
 %
-The rainbow of light that is scattered in droplets of water clouds and ocean glint, both of which indicate the presence of a water cycle which provides additional context to the habitability of an exoplanet, would be accessible in {\color{red}{!!}} $\sim\qty{-1}{\percent}$ and {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of systems respectively.
+The rainbow of light that is scattered in droplets of water clouds and ocean glint, both of which indicate the presence of a water cycle which provides additional context to the habitability of an exoplanet, would be accessible in $\sim\qty{28}{\percent}$ and $\sim\qty{10}{\percent}$ of systems respectively.
 %

### 2023-05-26 src/tex/affiliations.tex
[showyourwork] overleaf sync

--- 2023-05-25 src/tex/ms.tex
@@
-\RequirePackage{cmap}
-
-\documentclass[usenatbib]{mnras}
-
-\usepackage[T1]{fontenc}
-\usepackage[dvipsnames, x11names, svgnames]{xcolor}
-
-\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
-\usepackage{showyourwork}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{mathrsfs}
-
-\usepackage{xurl}
-\urlstyle{tt}
-
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage[
-    % colorlinks=true, 
-    % allcolors=blue,
-]{hyperref}
-
-% Prettier fonts 
-% Those seem to be the "right" fonts for MNRAS
-\usepackage{newtxtext}
-\usepackage{newtxmath}
-
-\usepackage{microtype}
-\usepackage{csquotes}
-\usepackage[capitalise]{cleveref}
-\usepackage{xspace}
-
-\usepackage{standalone}
-\usepackage{tikz}
-\usetikzlibrary{
-    calc,
-    positioning,
-    shapes,
-    shapes.symbols,
-}
-
-\usepackage{siunitx}
-\sisetup{
-    range-units = single,
-    range-phrase = \textup{--},
-    per-mode = symbol,
-}
-\DeclareSIUnit\parsec{pc}
-\DeclareSIUnit\au{au}
-\DeclareSIUnit\mas{mas}
-
-% Load journal macros (e.g., \aa for A&A)
-% Source: https://ui.adsabs.harvard.edu/help/actions/journal-macros
-\usepackage{aas_macros}
-% \usepackage{astro_bib_macro}  % where is this one from?
-
-\newcommand{\todo}[1]{\textcolor{red}{[#1]}}
-\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}} % amazing
-
-% Define macros
-\newcommand{\IWA}{\ensuremath{\mathrm{IWA}}\xspace}
-\newcommand{\OWA}{\ensuremath{\mathrm{OWA}}\xspace}
-\newcommand{\HWO}{HabWorlds\xspace}
-
-% Define the colors of the \rainbows
-\definecolor{C0}{HTML}{ff0000}
-\definecolor{C1}{HTML}{ff6d38}
-\definecolor{C2}{HTML}{ecc86f}
-\definecolor{C3}{HTML}{a4f89f}
-\definecolor{C4}{HTML}{5af8c8}
-\definecolor{C5}{HTML}{12c8e6}
-\definecolor{C6}{HTML}{386df9}
-\definecolor{C7}{HTML}{8000ff}
-
-\newcommand{\rainbows}{% MAK: KEEP this formatting for MNRAS submission. It deserves this!
-    \textcolor{C0}{r}%
-    \textcolor{C1}{a}%
-    \textcolor{C2}{i}%
-    \textcolor{C3}{n}%
-    \textcolor{C4}{b}%
-    \textcolor{C5}{o}%
-    \textcolor{C6}{w}%
-    \textcolor{C7}{s}%
-    \xspace
-}
-
-\newcommand{\istar}{%
-  \hspace{-0.125em}%
-  \tikz[baseline=(char.base)]{
-    \node [inner sep=0] (char) {i};
-    \node[starburst, starburst points=5, starburst point height=2.5mm, fill=White, draw=MidnightBlue, line width=0.1pt, minimum width=0.3mm, minimum height=0.1mm, inner sep=0.9pt, rotate=15] at (0.0015, 0.15) {};
-  }%
-  \hspace{-0.175em}%
-}
-\newcommand{\oceanglints}{\textcolor{MidnightBlue}{ocean gl\istar{}nts}\xspace}
-
-% Patch \hypertarget; for details, see 
-% https://tex.stackexchange.com/a/17138/97669
-\makeatletter
-\newcommand{\affdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}$^\mathrm{#1}$}
-\makeatother
-\newcommand{\afflink}[1]{\hyperlink{#1}{#1}}
-
-% Full title of the paper (MNRAS has no running title?)
-\title{Chasing \rainbows and \oceanglints:\\ Inner working angle constraints for the Habitable Worlds Observatory}
-
-% Authors are in alphabetical order:
 % DO NOT EDIT THIS MANUALLY
@@
 % IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
-\author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
-    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
-    Sarah L. Casewell\textsuperscript{\afflink{5}},
-    Nicolas B. Cowan\textsuperscript{\afflink{6}},
-    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
-    \newauthor
-    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
-    Matthew Kenworthy\textsuperscript{\afflink{7}},
-    Johan Mazoyer\textsuperscript{\afflink{11}},
-    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
-    \newauthor
-    Olivier Absil\textsuperscript{\afflink{13}},
-    Lisa Altinier\textsuperscript{\afflink{14}},
-    Pierre Baudoz\textsuperscript{\afflink{11}},
-    Ruslan Belikov\textsuperscript{\afflink{15}},
-    Alexis Bidot\textsuperscript{\afflink{16}},
-    Jayne Birkby\textsuperscript{\afflink{1}},
-    \newauthor
-    Markus J. Bonse\textsuperscript{\afflink{10}},
-    Bernhard Brandl\textsuperscript{\afflink{7}},
-    Alexis Carlotti\textsuperscript{\afflink{16}},
-    Elodie Choquet\textsuperscript{\afflink{14}},
-    Niyati Desai\textsuperscript{\afflink{17}},
-    \newauthor
-    Kevin Fogarty\textsuperscript{\afflink{15}},
-    Jules Fowler\textsuperscript{\afflink{18}},
-    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
-    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
-    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
-    \newauthor
-    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
-    Adrien Hours\textsuperscript{\afflink{16}},
-    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
-    Elina Kleisioti\textsuperscript{\afflink{7},\afflink{27}},
-    \newauthor
-    Lorenzo König\textsuperscript{\afflink{13}},
-    Mariya Krasteva\textsuperscript{\afflink{28}},
-    Iva Laginja\textsuperscript{\afflink{11}},
-    Rico Landman\textsuperscript{\afflink{7}},
-    Lucie Leboulleux\textsuperscript{\afflink{16}},
-    \newauthor
-    David Mouillet\textsuperscript{\afflink{16}},
-    Mamadou N’Diaye\textsuperscript{\afflink{29}},
-    Emiel H. Por\textsuperscript{\afflink{30}},
-    Laurent Pueyo\textsuperscript{\afflink{30}},
-    Frans Snik\textsuperscript{\afflink{7}},
-    \newauthor
-    Daphne M. Stam\textsuperscript{\afflink{31}},
-    Victor Trees\textsuperscript{\afflink{32},\afflink{33}},
-    Dirk van Dam\textsuperscript{\afflink{7}},
-    Kyle van Gorkom\textsuperscript{\afflink{21}},
-    Maaike van Kooten\textsuperscript{\afflink{34}}
-    \newauthor \\%
-    Affiliations are listed at the end of the paper
-    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
-    % https://academic.oup.com/mnras/article/498/2/2354/5900562
-}
-\date{Draft version: \today}
 
-\pagestyle{plain}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
-%
-Surface liquid water is central to the definition of planetary habitability.
-%
-Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
-%
-Direct imaging missions are optimised for planets near quadrature, but \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
-%
-The range of accessible phase angles for a given exoplanet will depend on the planet's orbital inclination and/or the coronagraph's inner working angle (IWA). 
-%
-For planets in near edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
-%
-We use \HWO'~catalogue of 164 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
-%
-We find that most exo-Earths planetary systems in this catalogue will be observable at phase angles at which the Rayleigh scattering peak occurs in their polarimetric phase curve. 
-%
-The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be observable with \HWO in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be observable in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
-%
-Improving the \IWA to $2\,\lambda/D$ increases the number of systems with observable rainbows and glints by factors of approximately 2 and 3, respectively.
-%
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.
-%
-By observing these scattering features, \HWO could detect a surface ocean and water cycle, key indicators of habitability.
-\looseness=-1
-\end{abstract}
-
-\begin{keywords}
-    planets and satellites: terrestrial planets -- 
-    instrumentation: high angular resolution -- 
-    planets and satellites: atmospheres
-\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:intro}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Liquid surface water is closely related to planetary habitability because it is essential for life as we know it.
-%
-To first order, its presence can be predicted by the flux a planet receives from its star.
-%
-The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet may have the right temperature to maintain liquid surface water \citep{kasting93}.
-%
-However, there are many factors besides the incident stellar flux that influence a planet's ability to hold liquid surface water, such as the atmospheric pressure and geological activity. 
-%
-Detecting liquid surface water on an exoplanet would thus provide important context for the interpretation of possible biosignatures.
-%
-Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[][]{2008Icar..195..927W}.
-%
-However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in \cref{fig:annotated-orbit}.
-%
-In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
-\looseness=-1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Angle conventions}
-
-As shown in \cref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle (large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (\qty{0}{\degree} at full phase and \qty{180}{\degree} at new phase).
-%
-A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = \qty{90}{\degree}$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles.
-%
-In some plots, we report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 1
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-    \centering
-    \includestandalone{tikz/annotated-orbit}
-    \caption{
-        Illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of characteristic scattering features in an exoplanet's phase curve.
-        The scattering angle (the planet--star--observer angle), $\varphi$, is shown in red, and the supplementary phase angle, $\alpha$, is shown in blue. 
-        See also \cref{fig:scattering-angle} and \cref{tab:phase_ranges}.
-    }
-    \label{fig:annotated-orbit}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Ocean glint}
-
-Ocean glint is the specular (Fresnel) reflection of light off a smooth liquid surface. 
-%
-While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing Titan at crescent phases \citep{2010GeoRL..37.7104S}.
-%
-In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
-%
-The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
-{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
-%
-This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
-%
-The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
-In general, the glint on a planet will be strongest at red wavelengths because there, Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}. 
-%
-This has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}. 
-%
-At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
-%
-This confusion can be resolved by measuring polarized phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
-the rainbow, see below.
-
-Glint provides evidence of specular reflecting surfaces, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for liquid water.
-%
-The presence of liquid water on the surface of a planet implies the presence of a water cycle, which provides key context to its habitability and the interpretation of atmospheric biosignatures.
-%
-Aside from detecting ocean glint, the presence of surface water on Earth-like exoplanets can be established through less direct methods near quadrature such as spectropolarimetry, to infer ocean reflection \citep{trees2022}, or through rotational mapping, to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 2
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-1-bott-plot.pdf}
-    \script{create-figure-1-bott-plot.py}
-    \caption{
-        Phase curves of the normalised total and normalised linearly polarised flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit (adapted from \citealt{treesandstam2019}).
-        The wind speed over the ocean---which determines the wave height \citep{CoxMunk1954}---is \qty{7}{\meter\per\second}, and the clouds consist of spherical water droplets.
-        The cloud coverage fraction is 0.5, and while the cloud pattern is patchy, the glint is always cloud-free. 
-        The fluxes are normalised such that the total (i.e., unpolarised + polarised) flux equals the planet's geometric albedo at phase angle $\alpha = \qty{0}{\degree}$.
-        In the bottom panel, we illustrate the scattering angle ranges of glories, rainbows, Rayleigh scattering, and ocean glint. 
-        Note that the features in the actual curves partly overlap.
-        The glory illustration is taken from \citet{Zakovryashin_2020}; all other illustrations are from \href{https://unsplash.com}{unsplash.com} [\href{https://unsplash.com/photos/YWL4dqqiRdk}{1}, \href{https://unsplash.com/photos/BgFOMcg4HNs}{2}, \href{https://unsplash.com/photos/9YdePM6BrO4}{3}].
-        This figure is best viewed in colour.
-    }
-    \label{fig:bottplot}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Rainbows, cloudbows, and glories}
-
-Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
-\footnote{
-    Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light. 
-    Cloudbows refer to features occurring at the same phase angles from smaller cloud particles, with the spread in colours invisible to the human eye making them appear white \citep[for further discussion, see][]{bailey2007}.
-}
-%
-In the Solar System, the polarized back-scatter from Titan indicated that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, Venus's polarized phase curve at three wavelengths showed that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture and constrained the droplet size distribution \citep{hansenhovenier1974}.
-%
-The glory feature in Venus's flux phase curve was later used to rule out various species as the unknown UV absorber \citep{petrova2018glory}.
-%
-Glories and rainbows are technically visible both in total and polarized flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarized nature coupled with the suppression of other polarized features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
-\looseness=-1
-%
-
-The phase angles and wavelength dependence of these features are sensitive to the chemical composition and droplet size distribution of condensates in the atmosphere. 
-%
-As with ocean glint, the detection of liquid condensates on an exoplanet can provide key context to their habitability, potential water cycle, and interpretation of atmospheric biosignatures.
-%
-In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of \qty{40}{\degree} (see \cref{fig:bottplot}).
-%
-Pushing observations to angles closer to the star (smaller and larger phase angles) would allow the determination of the full shape of the rainbow, and enable the identification of rainbows due to the scattering of light by other condensates.
-%
-The glory can allow further characterization of condensates as its angle and shape also depend on the refractive index, but this glory occurs within a phase angle of \qty{20}{\degree}, when the on-sky separation is small, as can also be seen in \cref{fig:bottplot}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Habitable Worlds Observatory}
-
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[Astro2020;][]{decadal} recommended a telescope with the capability to detect signatures of habitability on approximately 25 HZ planets in a new \enquote{Great Observatories} program.
-%
-This requires an instrument with a coronagraph capable of high-contrast imaging and low-resolution spectroscopy at optical to near-infrared wavelengths. 
-%
-The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
-%
-Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory.
-%
-The performance and characteristics (e.g.
-on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
-%
-They will likely be heavily informed by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
-%
-For generality, in this work, we report angular separations in milliarcseconds; however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table 1
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-    \centering
-    \caption{
-        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter}, \qty{670}{\nano\meter}, and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of \qty{6}{\meter}.
-    }
-    \label{tab:IWA_OWA}
-    \script{create-table-1-iwa-values.py}
-    \input{tables/table-1-iwa-owa}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{Outline}
-
-% In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
-%
-% In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
-%
-% We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Target list}
-\label{subsec:2.1}
-
-A preliminary target list of nearby stars (maximum distance \qty{25}{\parsec}) around which \HWO may search for habitable rocky planets, assuming a \qty{6}{\meter} telescope, has been compiled: the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
-\footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
-%
-This list consists of $\num{164}$ stars, the majority of which are Sun-like dwarfs: 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
-%
-The target list is constructed assuming a maximum planet magnitude in the Cousins $R_c$ band of $31$ mag and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
-% 
-The adopted HZ limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin and planet sizes between \qtyrange{0.8}{1.4}{} Earth radii. 
-%
-For non-solar stars, the semi-major axis scales as the square root of the bolometric luminosity normalized to the Sun. 
-%
-This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}. 
-%
-In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
-%
-By design, all such planets should be observable by \HWO at quadrature phase. 
-\looseness=-1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Accessible phase angles}
-\label{subsec:2.2}
-
-We first define the phase angle ranges over which the various phenomena---glint, Rayleigh peak, rainbow, and glory---manifest themselves in \cref{tab:phase_ranges}, also shown in \cref{fig:bottplot}. 
-%
-These ranges are all approximate, as the exact angles where these phenomena dominate depend on the details of the atmospheric and surface properties, as well as on the wavelengths being observed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Table 2
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-    \centering
-    \caption{
-        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. \cref{fig:bottplot}).
-    }
-    \label{tab:phase_ranges}
-    \begin{tabular}{ l c c c } 
-        \toprule
-        & $\alpha_\mathrm{min}$ & $\alpha_\mathrm{peak}$ & $\alpha_\mathrm{max}$ \\
-        \midrule
-        Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
-        Rainbow    & \qty{22}{\degree}    & \qty{42}{\degree}    & \qty{63}{\degree} \\
-        Rayleigh peak    & \qty{50}{\degree}    & \qty{70}{\degree}    & \qty{110}{\degree} \\
-        Ocean Glint & \qty{130}{\degree}   & \qty{150}{\degree}   & \qty{170}{\degree} \\
-        \bottomrule
-    \end{tabular}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A coronagraphic mask can obscure part of a planet's orbit. 
-As illustrated in \cref{fig:annotated-orbit}, obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
-The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and eccentricity of the exoplanet's orbit.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Circular, edge-on orbits}
-
-For the non-eccentric case, the minimum ($\alpha_\mathrm{min}$) and maximum ($\alpha_\mathrm{max}$) accessible phase angles are symmetric about \qty{90}{\degree}. 
-We define the angle $\Delta \alpha$ such that: 
-\begin{equation}
-    \label{eq:Delta_phi}
-    \Delta \alpha 
-    = \alpha_\mathrm{max} - \alpha_\mathrm{min}
-    =  2 \cdot (\alpha_\mathrm{max} - \qty{90}{\degree}) 
-    =  2 \cdot (\qty{90}{\degree} - \alpha_\mathrm{min} )  \,.
-\end{equation} 
-The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer.
-%
-The scattering angle, $\varphi$, is given by
-\begin{equation}
-    \label{eq:sinphi}
-    \sin(\varphi) = - \sin(\alpha) = \frac{r_\mathrm{proj}}{a} \,,
-\end{equation}
-assuming small angles and $r_\mathrm{proj} = \delta d_*$, where $\delta$ is the on-sky angular separation.
-%
-We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
-%
-Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
-\begin{equation}
-    \label{eq:scattering_angle}
-    \cos\left(\dfrac{\Delta \alpha}{2}\right) = \frac{\mathrm{IWA} 
-    \cdot d_*}{a} \,.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 3
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-    \centering
-    \includestandalone{tikz/scattering-angle}
-    \caption{
-        The angular separation of a planet from its star as seen by the observer, $\delta$, is related to the scattering angle $\varphi$. 
-        The star is at a distance $d_*$ from the observer, the planet orbits a distance $a$ from the star, and  the projected distance between the star and planet is  $r_\mathrm{proj}$. 
-        Not to scale. 
-    }
-    \label{fig:scattering-angle}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas}.
-%
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle ($\alpha\sim\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 4
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-4-scatterplot.pdf}
-    \script{create-figure-4-scatterplot.py}
-    \caption{
-        Scatter plot for the \HWO target stars, showing stellar effective temperatures and distances.
-        %
-        %The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
-        %
-        The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
-        %
-        The colour bar indicates the optical phenomena that can in principle be detected: all the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
-        %
-        For example, dark blue circles indicate planets for which most phenomena would be observable: planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
-    }
-    \label{fig:scatterplot}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Circular, inclined orbits}
-
-Most planets are not in edge-on orbits as seen from Earth.
-For an orbit with an inclination angle $i$ (where $i=\qty{0}{\degree}$ is face-on), we can identify the following two regimes:
-\begin{enumerate}
-    \item The coronagraph does not obscure any part of the sky-projected orbit: 
-    the maximum phase angle coverage depends solely on $i$. 
-        This holds for orbits that are relatively face-on (small $i$). 
-    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$, following \cref{eq:scattering_angle}. 
-\end{enumerate}
-
-The range of accessible phase angles is then 
-\begin{equation}
-\label{eq:Delta_phi_max}
-    \Delta \alpha = 
-    \begin{cases}
-        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA} \cdot d_* }{a} \\ 
-        2 \cos^{-1}\left(\dfrac{\mathrm{IWA} \cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA} \cdot d_* }{a}
-    \end{cases} \,
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monte Carlo simulations of accessible orbits}
-\label{subsec:2.3}
-
-\subsubsection{Circular, randomly inclined orbits}
-\label{sec:circular}
-
-For each of the 164 stars on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
-The orbital inclination angles $i$ are randomly drawn from a distribution that is uniform in $\cos i$. 
-We then use \cref{eq:Delta_phi_max} to determine the range of accessible phase angles for each hypothetical planet.
-The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible phase angles normalised to the total number of systems in the \HWO target list for different values of the \IWA.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 5
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}%[t]
-    \centering
-    \includegraphics{figures/figure-5-n-over-phase.pdf}
-    \script{create-figure-5-n-over-phase.py}
-    \caption{
-        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
-        %
-        The top and bottom $x$-axes indicate the minimum and maximum accessible phase angles, respectively.
-        %
-        These angles are symmetric about quadrature (\qty{90}{\degree}).
-        %
-        The $y$-axis on the left indicates the number of planetary systems divided by the number of Monte Carlo samples, and is thus normalised to the number of systems on the target list.
-        %
-        The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
-        %{\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
-    }
-    \label{fig:betaallofit}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 6: Yarn ball
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-6-yarnball.pdf}  
-    \script{create-figure-6-yarnball.py}
-    \caption{
-        Examples of the eccentric orbits generated for the stellar sample. 
-        Each plot shows 10 examples of orbits randomly drawn for each star.
-        The orbits are scaled to ensure an Earth-like incident flux. 
-        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
-        Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
-        It is clear that the \IWA can significantly affect the range of observable phases angles for each orbit.
-    }
-    \label{fig:ball-o-yarn}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Eccentric, randomly inclined orbits}
-\label{sec:eccentric}
-
-We repeat the Monte Carlo simulations from \cref{sec:circular}, assuming the orbital eccentricity follows a beta-distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
-%
-We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
-%
-Orbital eccentricities are drawn from the beta distribution, and inclination angles are uniform in $\cos i$. 
-%
-\Cref{fig:ball-o-yarn} shows a sample of the generated orbits.
-
-We then calculate the minimum and maximum accessible phase angles for \IWA{}s of $\qtylist{21; 41; 62; 83}{\mas}$.
-%
-The dashed lines in \cref{fig:betaallofit} show the normalised cumulative distribution of the accessible phase angles for each value of the \IWA for eccentric orbits. 
-%
-We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
-\looseness=-1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simulating contrast curves}
-\label{subsec:2.4}
-
-A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the planet/star flux ratio. 
-%
-While contrast is not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
-%
-For the contrast calculations, we start with the NASA Exoplanet Exploration Program’s Mission Star List for \HWO, which includes for every star an estimated contrast for an Earth-twin in the habitable zone.
-%
-The contrast is given by \citep[e.g.,][]{2023A&A...671A.165M}
-\begin{equation}
-    C(\alpha) 
-    = F_p(\alpha) / F_* 
-    = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
-    \label{eq:contrast}
-\end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
-%
-The contrast estimations for \HWO assumed circular orbits and a Lambertian phase function  \citep{Russell1916} with $A_\mathrm{G}=0.2$. 
-%
-For a more accurate representation of the contrast as a function of separation, we can update the phase function with the models shown in Fig. \Cref{fig:bottplot} from \citet{treesandstam2019}. 
-%
-To update the phase function, we will first correct for the reflected flux at quadrature and then map the normalised total flux and polarized flux as a function of separation.
-
-%
-We assume that the calculated contrasts in the Star List for \HWO correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
-%
-In addition, we calculate the reflected light fraction at quadrature with the updated model of \citet{treesandstam2019}, which assumes an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\qty{670}{\nano\meter}$.
-%
-We correct the star list contrast by dividing by the \qty{6.37}{\percent} reflected light fraction of the Lambertian phase function and multiplying by the reflected light fraction of the updated model.
-%
-We map the orbital phase for circular orbits to the phase angle and on-sky separation, given an inclination $i$ of \qty{90}{\degree}. 
-%
-Together, they give the reflected light contrast as a function of separation for a full orbit.
-%
-We multiply this flux with the degree of polarization to obtain the polarized contrast.
-%
-\Cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A.
-%
-These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 7
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-7-contrast-over-separation.pdf}
-    \script{create-figure-7-contrast-over-separation.py}
-    \caption{
-        Planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit  around Lalande~21185 (M2V), $\epsilon$~Eri (K2V), and $\alpha$~Cen~A (G2V) when assuming an inclination of $i = \qty{90}{\degree}$.
-        %
-        The solid black lines indicate the contrast in unpolarised light, with the contrast at quadrature marked by a black dot.
-        %
-        The contrast for linearly polarised light at $\lambda = \qty{670}{\nano\meter}$ is indicated by the coloured lines, with the colour encoding the scattering angle.
-        %
-        The small grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate \IWA's of $\qtylist{23; 46; 69}{\mas}$.
-    }
-    \label{fig:contrasts}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:3}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{The effect of eccentricity}
-\label{sec:result_eccentricity}
-
-As shown in \cref{fig:betaallofit}, a realistic eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
-%
-This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
-%
-The eccentricity does not significantly affect the number of systems for \IWA smaller than that used in the selection of the \HWO target list (yellow and blue lines).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Accessible scattering phenomena}
-%\label{sec:results_scattering_phenomena}
-
-\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be observable for different \IWA.
-%
-The Rayleigh peak, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be observable in a significant fraction of systems.
-%
-The position and shape of the rainbow peak is sensitive to the properties of the cloud droplets, allowing their characterization if the peak of the rainbow feature can be reached.
-%
-In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-    \centering
-    \caption{
-        The expected number of targets for which the peak phase angle of each phenomenon would be detectable when assuming randomly inclined, eccentric orbits (see \cref{sec:eccentric}).
-        For reference, the \HWO target list consists of 164 targets in total.
-    }
-    \label{tab:nstars_detect}
-    \input{tables/table-3-expected-number}
-    \script{create-figure-8-n-over-iwa.py}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Contrast curves}
-\label{sec:results_contrast}
-
-The contrast of the planet is also an important factor in assessing the detectability of these scattering phenomena. 
-%
-\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list, assuming an Earth-like planet on an edge-on orbit.
-%
-The overall contrast varies significantly for each system and throughout the orbit. In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
-%
-The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
-%
-It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph if these scattering features are to be observed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figure 8
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-    \centering
-    \includegraphics{figures/figure-8-n-over-iwa.pdf}
-    \script{create-figure-8-n-over-iwa.py}
-    \caption{
-        The number of systems for which key optical phenomena (Rayleigh peak, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
-        %
-        Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
-        %
-        The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        %
-        The $y$-axis on the right indicates the number of systems where all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
-    }
-    \label{fig:accessible_phase_angles}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sec:4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040s. 
-%
-The instrument design and capabilities are still under development and should be driven by the mission's science goals.
-%
-In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized and unpolarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
-%
-We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
-%
-While we do not consider the detect-ability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
-%
-A benefit of observing the polarized reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
-
-The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets around different stellar spectral types, will drive the choice of coronagraphic \IWA for \HWO.
-%
-Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarization from Rayleigh scattering.
-%
-The rainbow of light that is scattered in droplets of water clouds and ocean glint, both of which indicate the presence of a water cycle which provides additional context to the habitability of an exoplanet, would be accessible in $\sim\qty{28}{\percent}$ and $\sim\qty{10}{\percent}$ of systems respectively.
-%
-The presence of a water cycle on Earth-like exoplanets in the remaining systems would have to be established through less direct methods: spectropolarimetry or rotational mapping.
-%
-To observe more scattering features in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronagraph and potentially a larger primary mirror.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Software/Data Availability}
-
-The NASA ExEP target list containing the data used for these simulations is  available online.%
-\footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
-%
-This work has made use of \textsf{numpy} \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy}\footnote{\url{https://www.astropy.org}}, a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
-The colourblind-friendly colour schemes used in this paper were taken from \citet{Petroff_2021}.
-
-This paper has been formatted in the \texttt{showyourwork!} framework \citep{Luger2021}, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible.%
-\footnote{\url{https://github.com/mkenworthy/HWObows/}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgements}
-
-The 2023 \emph{Optimal Exoplanet Imagers} workshop, which sparked the work presented in this manuscript, was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, The Netherlands.
-%
-The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~866001 - EXACT).
-%
-This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
-%
-SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program under grant agreement \textnumero~805445.
-%
-SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
-%
-TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
-%
-KB acknowledges support from NASA Habitable Worlds grant \textnumero 80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement \textnumero~NNA13AA93A.
-%
-IL acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
-%
-PB, IL, and YG were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
-%
-EHP is supported by the NASA Hubble Fellowship grant \#HST-HF2-51467.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
-%
-LA and EC acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement \textnumero~101044152).
-%
-OA and LK acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~819155).
-%
-SYH was funded by the generous support of the Heising-Simons Foundation.
-%
-RJP is supported by NASA under award number 80GSFC21M0002.
-%
-OHS acknowledges funding from the Direction Scientifique Générale de l'ONERA (ARE Alioth).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% References
-\bibliographystyle{mnras}
-\bibliography{bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Author affiliations
-% DO NOT EDIT THIS MANUALLY
-% ALL ENTRIES ARE GENERATED BY A SCRIPT
-% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
-\section*{Author affiliations}
-\begingroup
-    \itshape
-    \raggedright
-    \affdest{1}Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK \\[0.5mm]
+\affdest{1}Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK \\[0.5mm]
 \affdest{2}Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\[0.5mm]
@@
 \affdest{12}Department of Physics, University of California, Santa Barbara, CA 93106, USA \\[0.5mm]
-    \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
-    \affdest{14}Aix Marseille Université, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
-    \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
-    \affdest{16}Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\[0.5mm]
-    \affdest{17}Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\[0.5mm]
-    \affdest{18}University of California, Santa Cruz, USA \\[0.5mm]
-    \affdest{19}DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\[0.5mm]
-    \affdest{20}DOTA, ONERA, 92322 Châtillon, France \\[0.5mm]
-    \affdest{21}Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85719, USA \\[0.5mm]
-    \affdest{22}Subaru Telescope, NAOJ, USA \\[0.5mm]
-    \affdest{23}College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\[0.5mm]
-    \affdest{24}Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\[0.5mm]
-    \affdest{25}NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\[0.5mm]
-    \affdest{26}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
-    \affdest{27}Faculty of Aerospace Engineering, TU Delft, Building 62, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
-    \affdest{28}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
-    \affdest{29}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
-    \affdest{30}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
-    \affdest{31}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
-    \affdest{32}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
-    \affdest{33}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
-    \affdest{34}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
-\endgroup
-
-\end{document}
+\affdest{13}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
+\affdest{14}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
+\affdest{15}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
+\affdest{16}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
+\affdest{17}Aix Marseille Université, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
+\affdest{18}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
+\affdest{19}Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\[0.5mm]
+\affdest{20}Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\[0.5mm]
+\affdest{21}University of California, Santa Cruz, USA \\[0.5mm]
+\affdest{22}Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85719, USA \\[0.5mm]
+\affdest{23}DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\[0.5mm]
+\affdest{24}DOTA, ONERA, 92322 Châtillon, France \\[0.5mm]
+\affdest{25}Subaru Telescope, NAOJ, USA \\[0.5mm]
+\affdest{26}College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\[0.5mm]
+\affdest{27}Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\[0.5mm]
+\affdest{28}NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\[0.5mm]
+\affdest{29}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
+\affdest{30}Faculty of Aerospace Engineering, TU Delft, Building 62, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
+\affdest{31}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
+\affdest{32}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
+\affdest{33}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
+\affdest{34}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]

### 2023-05-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \urlstyle{tt}
+
+\usepackage{datetime2}
 
@@
 \newcommand{\HWO}{HabWorlds\xspace}
+\newcommand{\Decadal}{Astro2020}  % no numbers in macro names
 
@@
 
-\newcommand{\rainbows}{% MAK: KEEP this formatting for MNRAS submission. It deserves this!
+% MAK: KEEP this formatting for MNRAS submission. It deserves this!
+\newcommand{\rainbows}{
     \textcolor{C0}{r}%
@@
 % Full title of the paper (MNRAS has no running title?)
-\title{Chasing \rainbows and \oceanglints:\\ Inner working angle constraints for the Habitable Worlds Observatory}
+\title{Chasing \rainbows and ocean glints:\\ Inner working angle constraints for the Habitable Worlds Observatory}
 
-% Authors are in alphabetical order:
-% DO NOT EDIT THIS MANUALLY
-% ALL ENTRIES ARE GENERATED BY A SCRIPT
-% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
+% Define authors (load from file)
 \author[Sophia R. Vaughan et al.]{%
-    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
-    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
-    Sarah L. Casewell\textsuperscript{\afflink{5}},
-    Nicolas B. Cowan\textsuperscript{\afflink{6}},
-    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
+    \input{authors}
     \newauthor 
-    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
-    Matthew Kenworthy\textsuperscript{\afflink{7}},
-    Johan Mazoyer\textsuperscript{\afflink{11}},
-    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
-    \newauthor
-    Olivier Absil\textsuperscript{\afflink{13}},
-    Lisa Altinier\textsuperscript{\afflink{14}},
-    Pierre Baudoz\textsuperscript{\afflink{11}},
-    Ruslan Belikov\textsuperscript{\afflink{15}},
-    Alexis Bidot\textsuperscript{\afflink{16}},
-    Jayne Birkby\textsuperscript{\afflink{1}},
-    \newauthor
-    Markus J. Bonse\textsuperscript{\afflink{10}},
-    Bernhard Brandl\textsuperscript{\afflink{7}},
-    Alexis Carlotti\textsuperscript{\afflink{16}},
-    Elodie Choquet\textsuperscript{\afflink{14}},
-    Niyati Desai\textsuperscript{\afflink{17}},
-    \newauthor
-    Kevin Fogarty\textsuperscript{\afflink{15}},
-    Jules Fowler\textsuperscript{\afflink{18}},
-    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
-    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
-    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
-    \newauthor
-    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
-    Adrien Hours\textsuperscript{\afflink{16}},
-    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
-    Elina Kleisioti\textsuperscript{\afflink{7},\afflink{27}},
-    \newauthor
-    Lorenzo König\textsuperscript{\afflink{13}},
-    Mariya Krasteva\textsuperscript{\afflink{28}},
-    Iva Laginja\textsuperscript{\afflink{11}},
-    Rico Landman\textsuperscript{\afflink{7}},
-    Lucie Leboulleux\textsuperscript{\afflink{16}},
-    \newauthor
-    David Mouillet\textsuperscript{\afflink{16}},
-    Mamadou N’Diaye\textsuperscript{\afflink{29}},
-    Emiel H. Por\textsuperscript{\afflink{30}},
-    Laurent Pueyo\textsuperscript{\afflink{30}},
-    Frans Snik\textsuperscript{\afflink{7}},
-    \newauthor
-    Daphne M. Stam\textsuperscript{\afflink{31}},
-    Victor Trees\textsuperscript{\afflink{32},\afflink{33}},
-    Dirk van Dam\textsuperscript{\afflink{7}},
-    Kyle van Gorkom\textsuperscript{\afflink{21}},
-    Maaike van Kooten\textsuperscript{\afflink{34}}
-    \newauthor \\%
+    \\%
     Affiliations are listed at the end of the paper
-    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
-    % https://academic.oup.com/mnras/article/498/2/2354/5900562
 }
-\date{Draft version: \today}
+\date{Draft version: \DTMnow}
 
@@
 %
-The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet may have the right temperature to maintain liquid surface water \citep{kasting93}.
+The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star for which a rocky planet may have the right temperature to maintain liquid surface water \citep{kasting93}.
 %
@@
 %
-Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[][]{2008Icar..195..927W}.
+Variations of the flux and degree of polarisation as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[][]{2008Icar..195..927W}.
 %
@@
 %
-In some plots, we report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.
+In some plots, we also report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.
 
@@
 %
-The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
-{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
+The maximum degree of polarisation of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
+{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.
+%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
 %
@@
 %
-The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
+The angular width of the glint in total and polarised flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
 In general, the glint on a planet will be strongest at red wavelengths because there, Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}. 
@@
 %
-This confusion can be resolved by measuring polarized phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
+This confusion can be resolved by measuring polarised phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
 the rainbow, see below.
@@
 
-Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
+Rainbows, cloudbows, glories, and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
 \footnote{
@@
 %
-In the Solar System, the polarized back-scatter from Titan indicated that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, Venus's polarized phase curve at three wavelengths showed that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture and constrained the droplet size distribution \citep{hansenhovenier1974}.
+In the Solar System, the polarised back-scatter from Titan indicated that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, and Venus's polarised phase curve at three wavelengths showed that its clouds were a $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture and constrained the droplet size distribution \citep{hansenhovenier1974}.
 %
@@
 %
-Glories and rainbows are technically visible both in total and polarized flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarized nature coupled with the suppression of other polarized features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
+Glories and rainbows are technically visible both in total and polarised flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarised nature coupled with the suppression of other polarised features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
 \looseness=-1
@@
 
-The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[Astro2020;][]{decadal} recommended a telescope with the capability to detect signatures of habitability on approximately 25 HZ planets in a new \enquote{Great Observatories} program.
+The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep[][hereafter \Decadal]{decadal} recommended a telescope with the capability to detect signatures of habitability on approximately 25 HZ planets in a new \enquote{Great Observatories} program.
 %
@@
 %
-The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
+The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarised ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
 %
@@
 %
-The performance and characteristics (e.g.
-on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
+The performance and characteristics (e.g., on- or off-axis, diameter, type of segmentation, polarimetry or photometry, coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
 %
@@
 %
-For generality, in this work, we report angular separations in milliarcseconds; however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
+For generality, in this work, we report angular separations in milliarcseconds; however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in \Decadal, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.
 
@@
     \caption{
-        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter}, \qty{670}{\nano\meter}, and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of \qty{6}{\meter}.
+        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milliarcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of \qty{6}{\meter}.
     }
     \label{tab:IWA_OWA}
-    \script{create-table-1-iwa-values.py}
+    \script{create-table-1-iwa-owa.py}
     \input{tables/table-1-iwa-owa}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{Outline}
-
-% In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
-%
-% In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.
-%
-% We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.
 
@@
 
-A preliminary target list of nearby stars (maximum distance \qty{25}{\parsec}) around which \HWO may search for habitable rocky planets, assuming a \qty{6}{\meter} telescope, has been compiled: the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
+A preliminary target list of nearby stars (maximum distance \qty{25}{\parsec}) around which \HWO may search for habitable rocky planets (assuming a \qty{6}{\meter} telescope), has been compiled: the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
 \footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
@@
     \caption{
-        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. \cref{fig:bottplot}).
+        Approximate ranges of the phase angle $\alpha$ for the optical phenomena considered in this paper (cf. \cref{fig:bottplot}). 
+        The scattering feature is present between $\alpha_\mathrm{min}$ and $\alpha_\mathrm{max}$; $\alpha_\mathrm{peak}$ is the phase angle where the scattered flux peaks.
+        \looseness=-1
     }
@@
 As illustrated in \cref{fig:annotated-orbit}, obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
-The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and eccentricity of the exoplanet's orbit.
+The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the orbital inclination and eccentricity.
 
@@
 For the non-eccentric case, the minimum ($\alpha_\mathrm{min}$) and maximum ($\alpha_\mathrm{max}$) accessible phase angles are symmetric about \qty{90}{\degree}. 
-We define the angle $\Delta \alpha$ such that: 
+We define the angle $\Delta \alpha$ (degrees from quadrature) such that: 
 \begin{equation}
@@
 
-\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas}.
+\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{61.9}{\mas}.
 %
-As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle ($\alpha\sim\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
+As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarisation due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle ($\alpha\sim\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.
 
@@
         %
-        The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
+        The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{61.9}{\mas}.
         %
@@
     \caption{
-        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
+        Expected number (computed as the average over \num{1000} simulations) of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
         %
@@
         %
-        The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
-        %{\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
+        The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$ (as in \Decadal).
     }
@@
         Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
-        It is clear that the \IWA can significantly affect the range of observable phases angles for each orbit.
+        The plot clearly shows that the \IWA can significantly affect the range of observable phases angles for each orbit.
     }
@@
 
-A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the planet/star flux ratio. 
+A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, that is, the planet-to-star flux ratio. 
 %
-While contrast is not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
+While contrast is not the main focus of this paper, we note that the total and polarised flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
 %
-For the contrast calculations, we start with the NASA Exoplanet Exploration Program’s Mission Star List for \HWO, which includes for every star an estimated contrast for an Earth-twin in the habitable zone.
+For the contrast calculations, we start with the \HWO target list (see \cref{subsec:2.1}), which, for every star, includes an estimated contrast for an Earth-twin in the habitable zone.
 %
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
 %
@@
 %
-For a more accurate representation of the contrast as a function of separation, we can update the phase function with the models shown in Fig. \Cref{fig:bottplot} from \citet{treesandstam2019}. 
+For a more accurate representation of the contrast as a function of separation, we can update the phase function with the models shown in \cref{fig:bottplot} from \citet{treesandstam2019}. 
 %
-To update the phase function, we will first correct for the reflected flux at quadrature and then map the normalised total flux and polarized flux as a function of separation.
+To update the phase function, we first correct for the reflected flux at quadrature and then map the normalised total flux and polarised flux as a function of separation.
 
 %
-We assume that the calculated contrasts in the Star List for \HWO correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
+We assume that the calculated contrasts in the \HWO target list correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
 %
-In addition, we calculate the reflected light fraction at quadrature with the updated model of \citet{treesandstam2019}, which assumes an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\qty{670}{\nano\meter}$.
+In addition, we calculate the reflected light fraction at quadrature with the updated model of \citet{treesandstam2019}, which assumes an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\lambda = \qty{670}{\nano\meter}$.
 %
@@
 %
-We map the orbital phase for circular orbits to the phase angle and on-sky separation, given an inclination $i$ of \qty{90}{\degree}. 
+We then map the orbital phase for circular orbits to the phase angle and on-sky separation, assuming an inclination $i$ of \qty{90}{\degree}. 
 %
-Together, they give the reflected light contrast as a function of separation for a full orbit.
+Together, this gives the reflected light contrast as a function of separation for a full orbit.
 %
-We multiply this flux with the degree of polarization to obtain the polarized contrast.
+We multiply this flux with the degree of polarisation to obtain the polarised contrast.
 %
@@
 %
-These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
+These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarised fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.
 
@@
     \caption{
-        Planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit  around Lalande~21185 (M2V), $\epsilon$~Eri (K2V), and $\alpha$~Cen~A (G2V) when assuming an inclination of $i = \qty{90}{\degree}$.
+        Planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its circular orbit  around Lalande~21185 (M2V), $\epsilon$~Eri (K2V), and $\alpha$~Cen~A (G2V) when assuming an inclination of $i = \qty{90}{\degree}$.
         %
@@
         %
-        The small grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate \IWA's of $\qtylist{23; 46; 69}{\mas}$.
+        The small grey dots show the contrasts at quadrature of the other targets in the star list, and the dashed lines indicate \IWA's of $\qtylist{23; 46; 69}{\mas}$ (corresponding to 1, 2, 3 $\lambda /D $ at $\lambda = \qty{670}{\nano\meter}$).
     }
@@
 %
-This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
 %
@@
 %
-The overall contrast varies significantly for each system and throughout the orbit. In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
+The overall contrast varies significantly for each system and throughout the orbit. 
+%
+In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
 %
@@
 It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph if these scattering features are to be observed.
+\looseness=-1
 
@@
         The number of systems for which key optical phenomena (Rayleigh peak, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
-        %
         Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
-        %
+        For Rayleigh scattering, the start of the feature (at $\alpha=\qty{70}{\degree}$) and the peak (at $\alpha=\qty{110}{\degree}$) coincide, as their respective phase angles are at the same distance from quadrature.
         The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        %
-        The $y$-axis on the right indicates the number of systems where all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
+        The $y$-axis on the right indicates the number of systems where all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ (as assumed in \Decadal).
     }
@@
 %
-In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized and unpolarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
+In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarised and unpolarised reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
 %
@@
 %
-While we do not consider the detect-ability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
+While we do not consider the detectability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light, but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
 %
-A benefit of observing the polarized reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.
+A benefit of observing the polarised reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarised \citep{kemp1987}.
 
@@
 %
-Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarization from Rayleigh scattering.
+Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarisation from Rayleigh scattering.
 %
@@
 %
-To observe more scattering features in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronagraph and potentially a larger primary mirror.
+To observe more scattering features in more systems, \HWO would need a smaller \IWA by use of smaller coronagraph and potentially a larger primary mirror.
 
@@
 
-The NASA ExEP target list containing the data used for these simulations is  available online.%
+The NASA ExEP target list containing the data used for our simulations is  available online.%
 \footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
 %
-This work has made use of \textsf{numpy} \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy}\footnote{\url{https://www.astropy.org}}, a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
-The colourblind-friendly colour schemes used in this paper were taken from \citet{Petroff_2021}.
-
-This paper has been formatted in the \texttt{showyourwork!} framework \citep{Luger2021}, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible.%
+This paper has been prepared using the \texttt{showyourwork!} framework \citep{Luger2021}, which makes the entire Python code to create all figures and the associated data available in an online repository to facilitate the reproducibility of our work.%
 \footnote{\url{https://github.com/mkenworthy/HWObows/}}
+\looseness=-1
 
@@
 %
-The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~866001 - EXACT).
-%
-This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France. 
-%
+The research presented in this paper was initiated at a workshop held in Leiden in February 2023 and was partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement \textnumero~866001---EXACT).
+
 SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation program under grant agreement \textnumero~805445.
@@
 %
-KB acknowledges support from NASA Habitable Worlds grant \textnumero 80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement \textnumero~NNA13AA93A.
+KB acknowledges support from NASA Habitable Worlds grant \textnumero~80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement \textnumero~NNA13AA93A.
 %
@@
 
+This work has made use of \textsf{numpy} \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy}\footnote{\url{https://www.astropy.org}}, a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.
+The colourblind-friendly colour schemes used in this paper were taken from \citet{Petroff_2021}.
+This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Author affiliations
-% DO NOT EDIT THIS MANUALLY
-% ALL ENTRIES ARE GENERATED BY A SCRIPT
-% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
+% Author affiliations (load from file)
 \section*{Author affiliations}
@@
     \raggedright
-    \affdest{1}Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK \\[0.5mm]
-    \affdest{2}Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\[0.5mm]
-    \affdest{3}NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\[0.5mm]
-    \affdest{4}NASA Nexus for Exoplanet System Science, Terrestrial Polarization Team, Orlando, FL 32826, USA \\[0.5mm]
-    \affdest{5}Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK \\[0.5mm]
-    \affdest{6}Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada \\[0.5mm]
-    \affdest{7}Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands \\[0.5mm]
-    \affdest{8}SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\[0.5mm]
-    \affdest{9}Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 Tübingen, Germany \\[0.5mm]
-    \affdest{10}ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland \\[0.5mm]
-    \affdest{11}LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, 92195 Meudon, France \\[0.5mm]
-    \affdest{12}Department of Physics, University of California, Santa Barbara, CA 93106, USA \\[0.5mm]
-    \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
-    \affdest{14}Aix Marseille Université, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
-    \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
-    \affdest{16}Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\[0.5mm]
-    \affdest{17}Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\[0.5mm]
-    \affdest{18}University of California, Santa Cruz, USA \\[0.5mm]
-    \affdest{19}DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\[0.5mm]
-    \affdest{20}DOTA, ONERA, 92322 Châtillon, France \\[0.5mm]
-    \affdest{21}Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85719, USA \\[0.5mm]
-    \affdest{22}Subaru Telescope, NAOJ, USA \\[0.5mm]
-    \affdest{23}College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\[0.5mm]
-    \affdest{24}Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\[0.5mm]
-    \affdest{25}NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\[0.5mm]
-    \affdest{26}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
-    \affdest{27}Faculty of Aerospace Engineering, TU Delft, Building 62, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
-    \affdest{28}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
-    \affdest{29}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
-    \affdest{30}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
-    \affdest{31}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
-    \affdest{32}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
-    \affdest{33}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
-    \affdest{34}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
+    \input{affiliations}
 \endgroup

### 2023-05-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-For planets in near edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
-%
 We use \HWO'~catalogue of 164 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
 %
-We find that most exo-Earths planetary systems in this catalogue will be observable at phase angles at which the Rayleigh scattering peak occurs in their polarimetric phase curve. 
+We find that the polarimetric Rayleigh scattering peak is accessible in most of the exo-Earths planetary systems.
 %
@@
 Improving the \IWA to $2\,\lambda/D$ increases the number of systems with observable rainbows and glints by factors of approximately 2 and 3, respectively.
-%
-The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.
 %
@@
         Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
-        For Rayleigh scattering, the start of the feature (at $\alpha=\qty{70}{\degree}$) and the peak (at $\alpha=\qty{110}{\degree}$) coincide, as their respective phase angles are at the same distance from quadrature.
+        For Rayleigh scattering, the line for the start of the feature (at $\alpha=\qty{70}{\degree}$) and the line for the peak (at $\alpha=\qty{110}{\degree}$) coincide, as their respective phase angles are at the same distance from quadrature.
         The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.

### 2023-06-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-NASA recently announced the Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
+NASA {\bf is engaged} in planning for a Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
 %
@@
 %
-We use \HWO'~catalogue of 164 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
+We {\bf use a recently-created catalog relevant to HabWorlds} of 164 stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and {\bf polarization effects due to Rayleigh scattering}.
 %
@@
 %
-The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be observable with \HWO in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be observable in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+Assuming a fixed Earth-like semi-major axis for each system, the rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible with \HWO in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
-Improving the \IWA to $2\,\lambda/D$ increases the number of systems with observable rainbows and glints by factors of approximately 2 and 3, respectively.
+Improving the \IWA to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of approximately 2 and 3, respectively.
 %
@@
 %
-Variations of the flux and degree of polarisation as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[][]{2008Icar..195..927W}.
+Variations of the flux and degree of polarisation as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection (i.e., glint) from water oceans \citep[e.g.][]{McCullough_2006,2008Icar..195..927W}.
 %
@@
 The angular width of the glint in total and polarised flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.
-In general, the glint on a planet will be strongest at red wavelengths because there, Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}. 
+In general, the glint on a planet will be strongest at red wavelengths because there, Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Robinson_2010,Zugger_2011}. 
 %
@@
 At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
+{\bf Recently, spectral PCA has been proposed as a tool to separate glint from atmospheric effects on Earth-like worlds \citep{Ryan_Robinson_2022}.}
 %
-This confusion can be resolved by measuring polarised phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
-the rainbow, see below.
+This confusion can be resolved by measuring polarised phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing the rainbow, see below.
 
@@
 
-We first define the phase angle ranges over which the various phenomena---glint, Rayleigh peak, rainbow, and glory---manifest themselves in \cref{tab:phase_ranges}, also shown in \cref{fig:bottplot}. 
+We first define the phase angle ranges over which the various phenomena---glint, Rayleigh scattering peak, rainbow, and glory---manifest themselves in \cref{tab:phase_ranges}, also shown in \cref{fig:bottplot}.
 %
@@
 As illustrated in \cref{fig:annotated-orbit}, obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
-The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the orbital inclination and eccentricity.
+The range of accessible phase angles will depend on the \IWA of the coronagraph, as well as on the orbital inclination and eccentricity.
 
@@
 
-For the non-eccentric case, the minimum ($\alpha_\mathrm{min}$) and maximum ($\alpha_\mathrm{max}$) accessible phase angles are symmetric about \qty{90}{\degree}. 
-We define the angle $\Delta \alpha$ (degrees from quadrature) such that: 
-\begin{equation}
-    \label{eq:Delta_phi}
-    \Delta \alpha 
-    = \alpha_\mathrm{max} - \alpha_\mathrm{min}
-    =  2 \cdot (\alpha_\mathrm{max} - \qty{90}{\degree}) 
-    =  2 \cdot (\qty{90}{\degree} - \alpha_\mathrm{min} )  \,.
-\end{equation} 
+For the non-eccentric case, the minimum and maximum accessible phase angles are symmetric about \qty{90}{\degree}. 
+%
 The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer.
@@
 %
-Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
+We can rewrite \cref{eq:sinphi} as:
 \begin{equation}
@@
 \end{equation}
+{\bf where we define $\Delta \alpha$ as the difference between the maximum and 
+minimum accessible phase angles}.
 
@@
         %
-        For example, dark blue circles indicate planets for which most phenomena would be observable: planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
+        For example, dark blue circles indicate planets for which most phenomena would be accessible: planets that can be observed at the rainbow angle can also be observed at angles exhibiting {\bf the polarization peak due to Rayleigh scattering}.
     }
@@
         Each plot shows 10 examples of orbits randomly drawn for each star.
-        The orbits are scaled to ensure an Earth-like incident flux. 
+        {\bf The orbits are scaled to ensure an Earth-like incident flux.} 
         The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
         Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
-        The plot clearly shows that the \IWA can significantly affect the range of observable phases angles for each orbit.
+        The plot clearly shows that the \IWA can significantly affect the range of accessible phases angles for each orbit.
     }
@@
 %
+The semi-major axis is kept the same for the circular and eccentric orbits - we do not renormalise the semi-major axis to keep the time averaged flux the same for all systems.
+%
 We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
@@
 %
-We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.
+We also calculate which target planets would be accessible at the phase angles where ocean glint, the rainbow, or the Rayleigh scattering peak occur.
 \looseness=-1
@@
 \end{equation}
-where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
+where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1} at full phase, thus $\phi(0^\circ)= 1.0$, and $a$ is the distance between the planet and its host star.
 %
-The contrast estimations for \HWO assumed circular orbits and a Lambertian phase function  \citep{Russell1916} with $A_\mathrm{G}=0.2$. 
+
+{\color{red}{The contrast estimations in the \HWO target list were 
+calculated assuming circular planetary orbits, and Lambertian reflecting planets \citep{Russell1916} with a geometric albedo $A_{\rm G}$ equal to 0.2.
+This albedo value was adopted from the VPL models in \citet{Robinson2011}, and the assumption of a Lambert phase function is justifiable given the apparent albedo values derived from Earthshine observations by \citet{Qiu2003} and \citet{Palle2003}.
+For a more accurate representation of the contrast as a function of the separation, we replace the Lambertian phase function with the models shown in \cref{fig:bottplot} from \citet{treesandstam2019} which assume an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\lambda = \qty{670}{\nano\meter}$.
+
+We assume that the contrasts in the \HWO target list represent the
+reflected flux at quadrature ($\alpha=90^\circ$) where a Lambertian
+phase function equals $1/\pi$, and  
+\cref{eq:contrast} to 
+}}
+%
+
+The contrast estimations for \HWO assumed circular planetary orbits and Lambertian phase functions \citep{Russell1916} with $A_\mathrm{G}=0.2$. 
 %
@@
 
+We assume that the calculated contrasts in the \HWO target list correspond to the reflected flux at quadrature, and so we use \cref{eq:contrast} to calculate the reflected light fraction.
+We assume a Lambertian reflecting planet at $\alpha=\qty{90}{\degree}$ with a geometric albedo of $0.2$ therefore $C(90) = \qty{6.37}{\percent} \left( R_p / a(90) \right)^2$.
 %
-We assume that the calculated contrasts in the \HWO target list correspond to the reflected flux at quadrature and use \cref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$ as $C(90) = \qty{6.37}{\percent}$.
+The geometric albedo is adopted from the VPL models in \citet{Robinson2011}, and the assumption of a Lambert phase function is justifiable given the apparent albedo values derived from Earthshine in \citet{Qiu2003} and \citet{Palle2003}.
 %
@@
         %
-        The solid black lines indicate the contrast in unpolarised light, with the contrast at quadrature marked by a black dot.
+        The solid black lines indicate the contrast in unpolarised light {\bf at $\lambda = \qty{670}{\nano\meter}$}, with the contrast at quadrature marked by a black dot.
         %
@@
 
-As shown in \cref{fig:betaallofit}, a realistic eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
+As shown in \cref{fig:betaallofit}, {\bf an empirical} eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
 %
-This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+This is because eccentric planets reach orbital separations greater than their semi-major axis, and hence are more likely {\bf to "peek out"} from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
 %
@@
 
-\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be observable for different \IWA.
+\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be accessible for different \IWA.
 %
-The Rayleigh peak, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
-The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be observable in a significant fraction of systems.
+The Rayleigh scattering peak {\bf in polarization}, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore accessible in the most systems.
+The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be accessible in a significant fraction of systems.
 %
@@
 %
-In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability.
+In most systems, only the first half of the glint feature would be accessible, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability.
 
@@
 %
-The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
+The accessibility of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
 %
@@
     \caption{
-        The number of systems for which key optical phenomena (Rayleigh peak, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
-        Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
+        The number of systems for which key optical phenomena (Rayleigh scattering peak, rainbow, ocean glint, glory) would be accessible as functions of the \IWA.
+        Each subplot contains three lines showing the number of systems where the start of each the phenomenon is accessible (dashed lines), where the peak is accessible (solid lines), and where the end is accessible (dotted lines).
         For Rayleigh scattering, the line for the start of the feature (at $\alpha=\qty{70}{\degree}$) and the line for the peak (at $\alpha=\qty{110}{\degree}$) coincide, as their respective phase angles are at the same distance from quadrature.
         The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
-        The $y$-axis on the right indicates the number of systems where all phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ (as assumed in \Decadal).
+        The $y$-axis on the right indicates the number of systems where all phenomena would be accessible, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ (as assumed in \Decadal).
     }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+\label{sec:4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this work, we consider each orbit to have a semi-major axis for which it would receives the same flux as Earth assuming a circular orbit.
+%
+As Earth is on the inner edge of the Habitable Zone, our estimates for the number of systems with accessible scattering features is conservative.
+%
+However, as we are only considering the accessibility of scattering features, it is possible to scale the results presented here for different semi-major axis. 
+%
+This is because the range of accessible features is determined only by the amount of the orbit covered by the mask, therefore increasing the semi-major axis of an orbit (e.g.
+double) is equivalent to decreasing the apparent mask size by the same factor (e.g.
+half).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
-\label{sec:4}
+\label{sec:5}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles with maximal polarisation from Rayleigh scattering.
+Assuming a \qty{62}{\mas} \IWA, $\sim\qty{76}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be accessible at phases angles with maximal polarisation from Rayleigh scattering.
 %
@@
 %
-To observe more scattering features in more systems, \HWO would need a smaller \IWA by use of smaller coronagraph and potentially a larger primary mirror.
+To observe more scattering features in more systems, \HWO would need a smaller \IWA by use of {\bf a better-controlled coronagraph} and potentially a larger primary mirror.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Software/Data Availability}
+\section*{Data Availability}
 

### 2023-06-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-Assuming a fixed Earth-like semi-major axis for each system, the rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible with \HWO in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible with \HWO\ {\bf for a planet with an Earth equivalent instellation} in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
 %
@@
 At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.  
-{\bf Recently, spectral PCA has been proposed as a tool to separate glint from atmospheric effects on Earth-like worlds \citep{Ryan_Robinson_2022}.}
 %
 This confusion can be resolved by measuring polarised phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing the rainbow, see below.
+%
+{\bf More recently, \citet{Ryan_Robinson_2022} have also proposed spectral PCA as a tool to separate glint from atmospheric effects on Earth-like worlds.}
 
@@
 
-For the non-eccentric case, the minimum and maximum accessible phase angles are symmetric about \qty{90}{\degree}. 
+%For the non-eccentric case, the minimum and maximum accessible phase angles are symmetric about \qty{90}{\degree}. 
 %
@@
 %
-We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
+We access the minimum{\bf /maximum} phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
 %
@@
 {\bf where we define $\Delta \alpha$ as the difference between the maximum and 
-minimum accessible phase angles}.
+minimum accessible phase angles.
+For non-eccentric orbits the minimum and maximum phase angles will be equal sides of quadrature and so $\Delta \alpha / 2$ represents the maximum angle away from quadrature that can be accessed}.
 
@@
 
-\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{61.9}{\mas}.
+\Cref{fig:scatterplot} shows {\bf $\Delta \alpha / 2$} for the systems in the \HWO target list, assuming an IWA of \qty{61.9}{\mas}.
 %
@@
 %
-The semi-major axis is kept the same for the circular and eccentric orbits - we do not renormalise the semi-major axis to keep the time averaged flux the same for all systems.
+The semi-major axis is kept the same for the circular and eccentric orbits---we do not renormalise the semi-major axis to keep the time-averaged flux the same for all systems.
 %
@@
 This albedo value was adopted from the VPL models in \citet{Robinson2011}, and the assumption of a Lambert phase function is justifiable given the apparent albedo values derived from Earthshine observations by \citet{Qiu2003} and \citet{Palle2003}.
-For a more accurate representation of the contrast as a function of the separation, we replace the Lambertian phase function with the models shown in \cref{fig:bottplot} from \citet{treesandstam2019} which assume an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\lambda = \qty{670}{\nano\meter}$.
+For a more accurate representation of the contrast as a function of the separation, we replace the Lambertian phase function with the models shown in \cref{fig:bottplot} from \citet{treesandstam2019} which assume an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds observed at $\lambda = \qty{670}{\nano\meter}$.\looseness=-1
 
-We assume that the contrasts in the \HWO target list represent the
-reflected flux at quadrature ($\alpha=90^\circ$) where a Lambertian
-phase function equals $1/\pi$, and  
-\cref{eq:contrast} to 
+We assume that the contrasts in the \HWO target list represent the reflected flux at quadrature ($\alpha = \qty{90}{\degree}$), where a Lambertian phase function equals $1/\pi$. 
+Using \cref{eq:contrast}, we correct the contrasts in the \HWO target list by dividing them with $0.2/\pi$ (=\,\qty{6.37}{\percent}) and multiplying them by the  reflected light fractions of the updated models of \citet{treesandstam2019}.
 }}
-%
-
-The contrast estimations for \HWO assumed circular planetary orbits and Lambertian phase functions \citep{Russell1916} with $A_\mathrm{G}=0.2$. 
-%
-For a more accurate representation of the contrast as a function of separation, we can update the phase function with the models shown in \cref{fig:bottplot} from \citet{treesandstam2019}.
-%
-To update the phase function, we first correct for the reflected flux at quadrature and then map the normalised total flux and polarised flux as a function of separation.
-
-We assume that the calculated contrasts in the \HWO target list correspond to the reflected flux at quadrature, and so we use \cref{eq:contrast} to calculate the reflected light fraction.
-We assume a Lambertian reflecting planet at $\alpha=\qty{90}{\degree}$ with a geometric albedo of $0.2$ therefore $C(90) = \qty{6.37}{\percent} \left( R_p / a(90) \right)^2$.
-%
-The geometric albedo is adopted from the VPL models in \citet{Robinson2011}, and the assumption of a Lambert phase function is justifiable given the apparent albedo values derived from Earthshine in \citet{Qiu2003} and \citet{Palle2003}.
-%
-In addition, we calculate the reflected light fraction at quadrature with the updated model of \citet{treesandstam2019}, which assumes an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds, observed at $\lambda = \qty{670}{\nano\meter}$.
-%
-We correct the star list contrast by dividing by the \qty{6.37}{\percent} reflected light fraction of the Lambertian phase function and multiplying by the reflected light fraction of the updated model.
-%
 We then map the orbital phase for circular orbits to the phase angle and on-sky separation, assuming an inclination $i$ of \qty{90}{\degree}. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace{-1mm}
 \section{Results}
@@
 %
-This is because eccentric planets reach orbital separations greater than their semi-major axis, and hence are more likely {\bf to "peek out"} from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+This is because eccentric planets reach orbital separations greater than their semi-major axis, and hence are more likely {\bf to \enquote{peek out}} from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-2mm}
 \section{Discussion}
@@
 
-In this work, we consider each orbit to have a semi-major axis for which it would receives the same flux as Earth assuming a circular orbit.
+In this work, we consider each orbit to have a semi-major axis for which it would receive the same flux as Earth assuming a circular orbit.
 %
-As Earth is on the inner edge of the Habitable Zone, our estimates for the number of systems with accessible scattering features is conservative.
+As Earth is at the inner edge of the Habitable Zone, our estimates for the number of systems with accessible scattering features is conservative.
 %
@@
 %
-This is because the range of accessible features is determined only by the amount of the orbit covered by the mask, therefore increasing the semi-major axis of an orbit (e.g.
-double) is equivalent to decreasing the apparent mask size by the same factor (e.g.
-half).
+This is because the range of accessible features is determined only by the amount of the orbit covered by the mask, therefore increasing the semi-major axis of an orbit (e.g., double) is equivalent to decreasing the apparent mask size by the same factor (e.g., half).
+\looseness=-1
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-2mm}
 \section{Conclusions}
@@
 %
-A benefit of observing the polarised reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarised \citep{kemp1987}.
+%A benefit of observing the polarised reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarised \citep{kemp1987}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace{-2mm}
 \section*{Data Availability}
@@
 %
+TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
+%
 SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
-%
-TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
 %

### 2023-07-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\newcommand{\todo}[1]{\textcolor{red}{[#1]}}
-\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}} % amazing
-
 % Define macros
@@
 
-% MAK: KEEP this formatting for MNRAS submission. It deserves this!
 \newcommand{\rainbows}{
@@
 \begin{abstract}
-NASA {\bf is engaged} in planning for a Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures.
+NASA is engaged in planning for a Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones and establish their habitability. 
 %
@@
 %
-We {\bf use a recently-created catalog relevant to HabWorlds} of 164 stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and {\bf polarization effects due to Rayleigh scattering}.
+We use a recently-created catalog relevant to HabWorlds of 164 stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and polarization effects due to Rayleigh scattering.
 %
@@
 %
-The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible with \HWO\ {\bf for a planet with an Earth equivalent instellation} in $\sim\num{46}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{16}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
+The rainbow due to water clouds at phase angles of ${\sim}\qtyrange{20}{60}{\degree}$ would be accessible with \HWO\ for a planet with an Earth equivalent instellation in ${\sim}\num{46}$ systems, while the ocean glint signature at phase angles of ${\sim}\qtyrange{130}{170}{\degree}$ would be accessible in ${\sim}\num{16}$ systems, assuming an \IWA${=}\qty{62}{\mas}$ ($3\lambda/D$).
 %
-Improving the \IWA to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of approximately 2 and 3, respectively.
+Improving the \IWA${=}\qty{41}{\mas}$ ($2\lambda/D$) increases accessibility to rainbows and glints by factors of approximately 2 and 3, respectively.
 %
@@
 %
-The maximum degree of polarisation of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
-{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.
-%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
+The maximum degree of polarisation of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.
 %
@@
 %
-{\bf More recently, \citet{Ryan_Robinson_2022} have also proposed spectral PCA as a tool to separate glint from atmospheric effects on Earth-like worlds.}
+More recently, \citet{Ryan_Robinson_2022} have also proposed spectral PCA as a tool to separate glint from atmospheric effects on Earth-like worlds.
 
@@
 
-%For the non-eccentric case, the minimum and maximum accessible phase angles are symmetric about \qty{90}{\degree}. 
-%
 The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer.
@@
 %
-We access the minimum{\bf /maximum} phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
+We access the minimum/maximum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
 %
@@
 \end{equation}
-{\bf where we define $\Delta \alpha$ as the difference between the maximum and 
+where we define $\Delta \alpha$ as the difference between the maximum and 
 minimum accessible phase angles.
-For non-eccentric orbits the minimum and maximum phase angles will be equal sides of quadrature and so $\Delta \alpha / 2$ represents the maximum angle away from quadrature that can be accessed}.
+For non-eccentric orbits the maximum and minimum phase angles will be equal sides of quadrature and so $\Delta \alpha / 2$ represents the maximum angle away from quadrature that can be accessed.
 
@@
         %
-        %The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
-        %
         The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{61.9}{\mas}.
@@
         %
-        For example, dark blue circles indicate planets for which most phenomena would be accessible: planets that can be observed at the rainbow angle can also be observed at angles exhibiting {\bf the polarization peak due to Rayleigh scattering}.
+        For example, dark blue circles indicate planets for which most phenomena would be accessible: planets that can be observed at the rainbow angle can also be observed at angles exhibiting the polarization peak due to Rayleigh scattering.
     }
@@
         Each plot shows 10 examples of orbits randomly drawn for each star.
-        {\bf The orbits are scaled to ensure an Earth-like incident flux.} 
+        The orbits are scaled to ensure an Earth-like incident flux.
         The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
@@
 
-{\color{red}{The contrast estimations in the \HWO target list were 
+The contrast estimations in the \HWO target list were 
 calculated assuming circular planetary orbits, and Lambertian reflecting planets \citep{Russell1916} with a geometric albedo $A_{\rm G}$ equal to 0.2.
@@
 We assume that the contrasts in the \HWO target list represent the reflected flux at quadrature ($\alpha = \qty{90}{\degree}$), where a Lambertian phase function equals $1/\pi$. 
+%
 Using \cref{eq:contrast}, we correct the contrasts in the \HWO target list by dividing them with $0.2/\pi$ (=\,\qty{6.37}{\percent}) and multiplying them by the  reflected light fractions of the updated models of \citet{treesandstam2019}.
-}}
+%
 We then map the orbital phase for circular orbits to the phase angle and on-sky separation, assuming an inclination $i$ of \qty{90}{\degree}. 
@@
         %
-        The solid black lines indicate the contrast in unpolarised light {\bf at $\lambda = \qty{670}{\nano\meter}$}, with the contrast at quadrature marked by a black dot.
+        The solid black lines indicate the contrast in unpolarised light at $\lambda = \qty{670}{\nano\meter}$, with the contrast at quadrature marked by a black dot.
         %
@@
 
-As shown in \cref{fig:betaallofit}, {\bf an empirical} eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
+As shown in \cref{fig:betaallofit}, an empirical eccentricity distribution increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
 %
-This is because eccentric planets reach orbital separations greater than their semi-major axis, and hence are more likely {\bf to \enquote{peek out}} from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
+This is because eccentric planets reach orbital separations greater than their semi-major axis, and hence are more likely to \enquote{peek out} from behind the coronagraph \IWA at extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
 %
@@
 \subsection{Accessible scattering phenomena}
-%\label{sec:results_scattering_phenomena}
 
@@
 %
-The Rayleigh scattering peak {\bf in polarization}, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore accessible in the most systems.
+The Rayleigh scattering peak in polarization, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore accessible in the most systems.
 The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be accessible in a significant fraction of systems.
@@
 While we do not consider the detectability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light, but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
-%
-%A benefit of observing the polarised reflected light is that a coronagraph is not explicitly required for quiescent stars, as the light of these stars themselves can be considered to be unpolarised \citep{kemp1987}.
 
@@
 %
-To observe more scattering features in more systems, \HWO would need a smaller \IWA by use of {\bf a better-controlled coronagraph} and potentially a larger primary mirror.
+To observe more scattering features in more systems, \HWO would need a smaller \IWA by use of a better-controlled coronagraph and potentially a larger primary mirror.
 

### 2023-07-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 }
-\date{Draft version: \DTMnow}
+\date{Accepted 2023 July 07. Submitted 2023 May 26}
 
@@
 %
-The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarised ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
+The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarised ocean glint on terrestrial planets} as a priority capability \citep[Table E.1 in][]{decadal} for ground and space-based observatories.
 %

### 2023-07-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %
-We find that the polarimetric Rayleigh scattering peak is accessible in most of the exo-Earths planetary systems.
+We find that the polarimetric Rayleigh scattering peak is accessible in most of the exo-Earth planetary systems.
 %
@@
 %
-However, as we are only considering the accessibility of scattering features, it is possible to scale the results presented here for different semi-major axis. 
+However, as we are only considering the accessibility of scattering features, it is possible to scale the results presented here for different semi-major axes. 
 %
@@
 %
-In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarised and unpolarised reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability.
+In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarised and unpolarised reflected light, such as rainbows, ocean glint, and glories, which could establish a planet's habitability.
 %
